### 2020-09-12 thesis proposal .Rmd
proposal for monday meeting

---
title: "Thesis Project Proposal Meeting"
author: "Maggie Slein"
date: "9/13/20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goals for today's meeting 
1. Discuss project proposal below and the different ways to go
2. 
3. Outline timeline for regular meetings and when milestones for this project should accomplished
4. Discuss appropriate workflow for this project

# Main Objectives: 
What is the single main objective of the research? This needs to relate to some kind of general ecological issue, and be articulated in a single concise statement.

How is environmental variation (temperature, pH, nutrient composition) systematically studied and manipulated across organismal, population, and community ecology and how does study design influence conclusions about environmental variation compare across these bodies of literature?

Does environmental variation affect how rates scale across levels of biological organization?

# Testable hypothesis: A hypothesis or set of hypotheses must:
clearly contribute to achieving the main objective
be able to be definitively tested and rejected
ideally, the set of hypotheses should together support some inference about the main objective.
Generating hypotheses is a challenge. Make sure there are not logical gaps between the
hypotheses, objective and methods used to test them, or if gaps exist, address them.

H0: There is no difference in how variation is studied at different levels of biological organization and thus the conclusions are the same across all levels.

Ha: Short term responses to simple environmental variation manipulation is prioritized at the organismal level, long term and more complex environmental variation manipulated is prioritized at the population and community level, owing to differential experimental design and conclusions about how variation type will interact in the ecosystem.

H02: Environmental variation does affect how rate scale across levels of biological organization?

# Methods:

1. I will conduct a series of ISI literature searches to get a hollistic set of studies referencing environmental variation with selected keywords

How will hypotheses be tested? General categories include experiments, observations (surveys),
meta-analyses, analysis of theoretical models. Will inferences be drawn from more than one type
of evidence?
- How will data be collected and analyzed? What statistical tests and software will you use?
- What resources are required? Budget for time (e.g., timeline) and for costs.
#Results: Consider, in advance, the 2-3 key figures or tables you will produce for your final
paper. Sketch them, including axes, and even the data that would reject or fail your hypotheses or
support your predictions.
#Inference and Next Steps: What logical steps will you use to achieve your main objectives,
and draw conclusions? How exactly will your figures (sketched above) answer your original
questions? Consider different possible versions of the results, and how you might interpret them.
Then revisit your questions – are you happy with how your hypotheses are stated, given the
possible results?
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-09-13 thesis proposal .Rmd
updated thesis meeting document

@@
 1. Discuss project proposal below and the different ways to go
-2. 
+2. Discuss how to design proper search term parameters that provides big picture but managable study count
 3. Outline timeline for regular meetings and when milestones for this project should accomplished
@@
 
-# Main Objectives: 
-What is the single main objective of the research? This needs to relate to some kind of general ecological issue, and be articulated in a single concise statement.
-
+# Main Objectives 
 How is environmental variation (temperature, pH, nutrient composition) systematically studied and manipulated across organismal, population, and community ecology and how does study design influence conclusions about environmental variation compare across these bodies of literature?
@@
 
-# Testable hypothesis: A hypothesis or set of hypotheses must:
-clearly contribute to achieving the main objective
-be able to be definitively tested and rejected
-ideally, the set of hypotheses should together support some inference about the main objective.
-Generating hypotheses is a challenge. Make sure there are not logical gaps between the
-hypotheses, objective and methods used to test them, or if gaps exist, address them.
-
+# Testable hypothesis
 H0: There is no difference in how variation is studied at different levels of biological organization and thus the conclusions are the same across all levels.
@@
 
-H02: Environmental variation does affect how rate scale across levels of biological organization?
+H02: Environmental variation does not affect how rate scale across levels of biological organization
 
-# Methods:
+Ha2: Environmental variation alters how rates scale across levels of biological organization by influencing metabolic rate
 
-1. I will conduct a series of ISI literature searches to get a hollistic set of studies referencing environmental variation with selected keywords
+# Methods
+1.
+I will conduct a series of ISI literature searches to get a hollistic set of studies referencing environmental variation with selected keywords.
+I will then select studis based on a stricy set of criteria and enter those studies into a spreadsheet, recording relevant information (study organism, biological organization level, variation type, response variable, ).
 
-How will hypotheses be tested? General categories include experiments, observations (surveys),
-meta-analyses, analysis of theoretical models. Will inferences be drawn from more than one type
-of evidence?
-- How will data be collected and analyzed? What statistical tests and software will you use?
-- What resources are required? Budget for time (e.g., timeline) and for costs.
-#Results: Consider, in advance, the 2-3 key figures or tables you will produce for your final
-paper. Sketch them, including axes, and even the data that would reject or fail your hypotheses or
-support your predictions.
-#Inference and Next Steps: What logical steps will you use to achieve your main objectives,
-and draw conclusions? How exactly will your figures (sketched above) answer your original
-questions? Consider different possible versions of the results, and how you might interpret them.
-Then revisit your questions – are you happy with how your hypotheses are stated, given the
-possible results?
-```{r cars}
-summary(cars)
+Current search term ideas:
+* "environmental variability" AND ecology AND temperature (1,281 results)
+* "environmental variability" AND ecology (4049 results)
+* "environmental variability" AND rate AND ecology (1,062 results)
+* "environmental variability" AND ecology AND organism (802 results)
+* "environmental variability" AND ecology AND population (2,458 results)
+* "environmental variability" AND ecology AND community (1,275 results)
+* "environmental variation" AND ecology (4,983 results)
+
+2.
+I will then synthesize the patterns across studies to address my main objectives.
+This will culminate in written document summarizing that language used in different sub disciplines as well as the similarities and differences in environmental variability.
+
+3. If there are studies that feature the same type of experimental design, I would like to conduct a meta-analaysis on that particular body of literature in R.
+
+4.
+If not, I would like to model environmental variability parameters from the different studies using R.
+(Less sure of how this could be achieved but I'd like to incorporate modelling components to this project)
+
+# Potential Results
+```{r, echo = FALSE}
+knitr::include_graphics("images/fig 1.png")
+```
+Figure 1.
+Quantitative representation of the relatedness of environmental variability study work across different levels of biological organization (modified from http://environmentalcomputing.net/multidimensional-scaling/)
+
+```{r, echo = FALSE}
+knitr::include_graphics("images/fig 2.png")
 ```
 
-## Including Plots
+Figure 2. Scaling of responses to environmental variability to animal body size (modified from Schramski et al 2015, PNAS)
 
-You can also embed plots, for example:
+# Timeline, Inference, and Next Steps
+To accomplish this work I'd like to hit the following milestones:
 
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
+* Narrow in on search terms (end of September)
+* Decide on filter to select studies by (end of September)
+* Filter through studies (end of fall semester)
+* Theoretical foundation of modelling work in the spring (end of October)
+* Premilinary literature review summary (end of fall semester)
+* Data extraction complete (early March)
+* Modelling work (end of March)
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Figure 1 above will answer my questions about potential dissonance and gaps in environmental variability across different levels of organization and their respective field of study.
+An NMDS-like plot would be a quanitative way to display similarities and potential distinctions.
+
+Figure 2 is a simple first pass at potentially looking about how MTE and environmental variability could interact.
+I'd be really interested in looking into the math for how environmental variability could alter metabolic rate and looking the relationship between body size.
+I think this sample figure will answer the question of whether or not environmental variability has the ability to scale with rates across levels of organization.

### 2020-09-13 thesis proposal .Rmd
final proposal updates

@@
 # Goals for today's meeting 
-1. Discuss project proposal below and the different ways to go
-2. Discuss how to design proper search term parameters that provides big picture but managable study count
+1. Discuss the project proposal below and the different ways to proceed
+2. Discuss how to design proper search term parameters that provide a representative but managable study count
 3. Outline timeline for regular meetings and when milestones for this project should accomplished
@@
 # Main Objectives 
+1.
 How is environmental variation (temperature, pH, nutrient composition) systematically studied and manipulated across organismal, population, and community ecology and how does study design influence conclusions about environmental variation compare across these bodies of literature?
 
-Does environmental variation affect how rates scale across levels of biological organization?
+2. Does environmental variation affect how rates scale across levels of biological organization?
 
@@
 
-Ha: Short term responses to simple environmental variation manipulation is prioritized at the organismal level, long term and more complex environmental variation manipulated is prioritized at the population and community level, owing to differential experimental design and conclusions about how variation type will interact in the ecosystem.
+Ha: Short term responses to simple environmental variation manipulation are prioritized at the organismal level, long term responses and more complex environmental variation manipulation is prioritized at the population and community level, owing to differential experimental design and conclusions about how variation type will interact in the ecosystem.
 
@@
 
-Ha2: Environmental variation alters how rates scale across levels of biological organization by influencing metabolic rate
+Ha2: Environmental variation alters how rates scale across levels of biological organization by influencing metabolic rate.
 
@@
 I will conduct a series of ISI literature searches to get a hollistic set of studies referencing environmental variation with selected keywords.
-I will then select studis based on a stricy set of criteria and enter those studies into a spreadsheet, recording relevant information (study organism, biological organization level, variation type, response variable, ).
+I will then select studies based on a strict set of criteria and enter those studies into a spreadsheet, recording relevant information (study organism, biological organization level, variation type, response variable, etc.).
 
@@
 I will then synthesize the patterns across studies to address my main objectives.
-This will culminate in written document summarizing that language used in different sub disciplines as well as the similarities and differences in environmental variability.
+This will culminate in a written document summarizing the language used in different sub-disciplines as well as the similarities and differences in conclusions about environmental variability.
 
@@
 Figure 2 is a simple first pass at potentially looking about how MTE and environmental variability could interact.
-I'd be really interested in looking into the math for how environmental variability could alter metabolic rate and looking the relationship between body size.
-I think this sample figure will answer the question of whether or not environmental variability has the ability to scale with rates across levels of organization.
+I'd be really interested in looking into the math for how environmental variability could alter metabolic rate and the relationship between body size.
+I think this sample figure could answer the question of whether or not environmental variability has the ability to scale with rates across levels of organization.

### 2020-09-20 thesis meeting 9:21:20.Rmd
rmd for thesis meeting on 9/21/20

---
title: "thesis meeting 9/21/20"
author: "Maggie Slein"
date: "9/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goals for today's meeting 
1. Updates to project questions and goals
2. Updates about ISI search results and terms 
3. Next steps for implementing methods

# Project updates

Last week we discussed how the second question listed below would be a better thematic guide to approaching this review but less able to address a quantitative question:

*How is environmental variation (temperature, pH, nutrient composition) systematically studied and manipulated (duration and periodicity) across organismal, population, and community ecology and how does study design influence conclusions about environmental variation compare across these bodies of literature?*

I spent some time reading through the papers that Mary sent as well as looking through Web of Science for any inspiration. I have centered on the following directions for a quantitative project that is a more detailed version of the question I posed last week:

*Does environmental variation affect how rates scale across levels of biological organization?*

## Overarching goal:
Conduct a meta-analysis on some subsection of the ecological literature with respect to environmental variation across different levels of organizations

Option 1.	Look exclusively at temperature manipulations across all levels of organization

*Driving question:*

  + How does the magnitude of temperature response connect across different levels of organization when exposed to different types of temperature variation (diurnal, stochastic, autocorrelated etc.)
  +	Potential response variables: 
    + Individual: metabolic rate, TPC 
    +	Population: population growth rate, population density, carrying capacity
    + Community: biomass?, NPP? Other rate based metrics?
  + Potential outcomes: a more nuanced understanding how the type of variation manipulation can influence differential responses across different levels of organization

Option 2: Look exclusively at diurnal fluctuations of 3 different environmental variables (temp, pH, salinity) across all levels of organization

*Driving question(s):*

  + Is there a differential response at different levels of organization to variation in each of these variables 
  +	How do non-linear and linear response variables respond to stochastic vs. predictable variation?
      + look at a couple different environmental variables and how different manipulations of those variables coincides with either linear or non-linear response variables
  +	Potential response variables:
      + Individual: metabolic rate, growth rate, respiration rate (MMR)?
      + Population: population growth rate, population density, carrying capacity
      + Community: biomass?, NPP? Other rate based metrics?
  + Potential outcomes: a better understanding of these metrics that covary may not necessarily covary in their responses to variation across different levels of organization?
3.	Looking at the trade-off between spatial and temporal variation in temperature in marine and aquatic systems

##Methods

I spent some time reading the papers Mary sent as well as several other meta-analyses to draw ideas for how to conduct this meta-analysis.

I really like the study design from Hackett et al 2013, specifically how they excluded factors that have been demonstrated as important (the effect of earthworms on soil biomass) to isolate the nuanced effects of micro and mega fauna outside of those influential factors. This inspired more attempts to devise a more detailed set of questions and outcomes from my original proposal.

I was mainly inspired by a review by Kroeker et al 2020 that address the themes in studying environmental variability from mainly an organismal perspective. This paper really addresses a lot of the thoughts Joey's Life paper addressed and I really liked the organization of this paper (really focusing on the importance and motivation behind drivers at different time scales). I would be really interested in asking similar questions with a quantitative lens. This paper focused mainly on predictable variation and wonder how these conclusion might different when compared with unpredictable variation.

This figure somewhat address what I wanted to accomplish with an NMDS plot, but I think it a good jumping off place for thinking about how to design search terms and contextualize environmental variability bodies of literature
```{r}
knitr::include_graphics("images/kroeker_fig1")
```

### 2020-09-20 thesis meeting sept21.Rmd
thesis meeting prohect updates and next steps for sept 21

---
title: "Thesis Meeting"
author: "Maggie Slein"
date: "9/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goals for today's meeting 
1. Updates to project questions and goals
2. Updates about ISI search results and terms 
3. Next steps for implementing methods

# Project updates

Last week we discussed how the second question listed below would be a better thematic guide to approaching this review but less able to address a quantitative question:

*How is environmental variation (temperature, pH, nutrient composition) systematically studied and manipulated (duration and periodicity) across organismal, population, and community ecology and how does study design influence conclusions about environmental variation compare across these bodies of literature?*

I spent some time reading through the papers that Mary sent as well as looking through Web of Science for any inspiration. As a more detailed version of the question I posed last week:

*Does environmental variation affect how rates scale across levels of biological organization?*

I have centered on the following directions for a quantitative project:

## Overarching goal:
Conduct a meta-analysis on some subsection of the ecological literature with respect to environmental variation across different levels of organizations

**Option 1.**	Look exclusively at how temperature manipulations alter ecological rates across all levels of organization

* Driving question:
  + How does the magnitude of temperature response compare across different levels of organization when exposed to different types of temperature variation (diurnal, stochastic, autocorrelated etc.)

* Potential response variables: 
    + Individual: metabolic rate, TPC 
    +	Population: population growth rate, population density, carrying capacity
    + Community: biomass?, NPP? Other rate based metrics?
* Potential outcome: a more nuanced understanding of how the type of variation manipulation can drive differential responses across different levels of organization

```{r, echo=FALSE}
knitr::include_graphics("images/kroeker_fig1.png")
```

**Option 2:** Look exclusively at diurnal fluctuations of 3 different environmental variables (temp, pH, salinity, DO, light, etc.) across all levels of organization

* Driving question(s):
  + Is there a differential response at different levels of organization to variation in each of these variables 
  +	How do non-linear and linear response variables respond to stochastic vs. predictable variation?
      + look at a couple different environmental variables and how different manipulations of those variables coincides with either linear or non-linear response variables
* Potential response variables:
    + Individual: metabolic rate, growth rate, respiration rate (MMR)?
    + Population: population growth rate, population density, carrying capacity
    + Community: biomass?, NPP? Other rate based metrics?
* Potential outcome: a better understanding of these metrics that covary may not necessarily covary in their responses to variation across different levels of organization?
```{r, echo=FALSE}
knitr::include_graphics("images/kroeker_fig1.png")
```

# Methods

I spent some time reading the papers Mary sent as well as several other meta-analyses to draw ideas for how to conduct this meta-analysis.

I really like the study design from Hackett et al 2013, specifically how they excluded factors that have been demonstrated as important (the effect of earthworms on soil biomass) to isolate the nuanced effects of micro and mega-fauna outside of those influential factors. This inspired attempts to devise a more detailed set of questions and outcomes from my original proposal.

I was mainly inspired by a review by Kroeker et al 2020 that addressed themes in studying environmental variability from mainly an organismal perspective. This paper really referenced a lot of the thoughts Joey's Life paper addressed and I really liked the organization of this paper (really focusing on the importance and motivation behind drivers at different time scales). I would be really interested in asking similar questions with a quantitative lens. This paper focused mainly on predictable variation and I wonder how these conclusions might differ when compared with unpredictable variation.

Figure 1 somewhat addresses what I wanted to accomplish with an NMDS plot, but I think it a good jumping off place for thinking about how to design search terms and contextualize environmental variability bodies of literature

```{r, fig.width=5, fig.height=5, echo=FALSE}
knitr::include_graphics("images/kroeker_fig1.png")
```

I conducted several advanced ISI searches to no avail, ending up with many more hits than my basic search, even when restricting by year and three specific journals:

1. TS=(“temporal variation” OR “spatial variation” AND temperature AND rates) OR SO=(JOURNAL OF ANIMAL ECOLOGY OR JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY) OR SO=(FRESHWATER ECOLOGY) OR SO=(ECOLOGY LETTERS OR ECOLOGY LETTERS) AND PY=(1970-2020) *(71,597)*

2. TS=( "environmental variation" AND temperature AND rates) OR SO=(JOURNAL OF ANIMAL ECOLOGY OR JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY) OR SO=(FRESHWATER ECOLOGY) OR SO=(ECOLOGY LETTERS OR ECOLOGY LETTERS) AND PY=(1970-2020) *(10,030)*

3. TS=( "environmental variation" AND temperature AND rates) OR SO=(JOURNAL OF ANIMAL ECOLOGY OR JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY) OR SO=(FRESHWATER ECOLOGY) OR SO=(ECOLOGY LETTERS OR ECOLOGY LETTERS) AND PY=(1970-2020) OR AK=("temporal variation" OR "spatial variation" OR "variable environments")  *(14,206)*

4. TS=( "environmental variation" AND temperature AND rates) OR SO=(JOURNAL OF ANIMAL ECOLOGY OR JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY) OR SO=(FRESHWATER ECOLOGY) OR SO=(ECOLOGY LETTERS OR ECOLOGY LETTERS) AND PY=(1970-2020) OR AK=("temporal variation" OR "spatial variation" OR "variable environments") 
Databases= WOS, BCI, KJD, RSCI, SCIELO, ZOOREC 
Timespan=1970-2018
Search language=Auto  *(11,242)*

5. TS=( "environmental variation" AND temperature AND rates AND diurnal OR circadian) OR SO=(JOURNAL OF ANIMAL ECOLOGY OR JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY OR THE JOURNAL OF ANIMAL ECOLOGY) OR SO=(FRESHWATER ECOLOGY) OR SO=(ECOLOGY LETTERS OR ECOLOGY LETTERS) AND PY=(1970-2020) OR AK=("variable environments")
*(276,317)*

# Next steps and questions 
* Is the world predicted to become more variable or more autocorrelated? 
    + And if the answer is both, how do they interact? 
    + Is it that while the world will become more like its surrounding conditions, those conditions will increase in amplitude or become more severe?

* How do I properly use ISI advanced searches to get narrower results?
* How do I accumulate the right amount of keywords to accomplish this?
* What is the right scope for the two proposed detailed questions for this thesis project?

### 2020-09-20 thesis meeting sept21.Rmd
including graph sketches for the two potential figures from different project directions

@@
 
+\newpage
+
 **Option 1.**	Look exclusively at how temperature manipulations alter ecological rates across all levels of organization
@@
 ```{r, echo=FALSE}
-knitr::include_graphics("images/kroeker_fig1.png")
+knitr::include_graphics("images/option1_projprop.png")
 ```
+
+\newpage
 
@@
 * Potential outcome: a better understanding of these metrics that covary may not necessarily covary in their responses to variation across different levels of organization?
+
 ```{r, echo=FALSE}
-knitr::include_graphics("images/kroeker_fig1.png")
+knitr::include_graphics("images/option2_projprop.png")
 ```
@@
 
-Figure 1 somewhat addresses what I wanted to accomplish with an NMDS plot, but I think it a good jumping off place for thinking about how to design search terms and contextualize environmental variability bodies of literature
-
-```{r, fig.width=5, fig.height=5, echo=FALSE}
-knitr::include_graphics("images/kroeker_fig1.png")
-```
+Figure 1 in Kroeker et al 2020 somewhat addresses what I wanted to accomplish with an NMDS plot, but I think it is a good jumping off place for thinking about how to design search terms and contextualize environmental variability bodies of literature
 
@@
 
+\newpage
 # Next steps and questions 
@@
     + Is it that while the world will become more like its surrounding conditions, those conditions will increase in amplitude or become more severe?
-
 * How do I properly use ISI advanced searches to get narrower results?

### 2020-09-28 meeting materials/thesis_meetingsept28.Rmd
meeting materials for sept 28th thesis meeting, including rmd summary document and importing extracted data

---
title: "Thesis"
author: "Maggie Slein"
date: "9/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Goals for today's meeting
1. Discuss figure(s) from preliminary extraction 
2. Outline extract methods and binning
3. Next steps for conducting hollistic search terms

###Study selection

Khelifa et al 2019

I really enjoy this paper and feel that its figures and methods are very straightforward. I'd like my final figure to ultimately look like this

```{r}
knitr::include_graphics("khelifa-fig2.png")
```

```{r, echo= FALSE}
library(tidyverse)
extract<-read_csv("prelim_extact.csv")
names(extract)

ggplot(extract, aes(x=org_level, y=cohen_d, color=fluctuation_pattern))+
  geom_boxplot()+
  theme_classic()+
  ggtitle("Cohen D values for effect of flux patterns across different levels of organization")

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-09-28 thesis_meetingsept28.Rmd
meeting materials for sept 28th thesis meeting, including rmd summary document and importing extracted data

---
title: "Thesis meeting"
author: "Maggie Slein"
date: "9/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goals for today's meeting
1. Discuss figure(s) from preliminary extraction 
2. Outline extract methods and binning
3. Next steps for conducting hollistic search terms

## Study selection

### Khelifa et al 2019

I really enjoy this paper and feel that its figures and methods are very straightforward. I'd like my final figure to ultimately look like this (Figure 2). I extracted all the data from this figure in my composite figure beyond.

```{r, echo=FALSE}
knitr::include_graphics("meeting materials/images/khelifa-fig2.png")
```
\newpage

### Petchey et al 2002

I wanted to include a study that feature community level responses to colored environmental variation and I settled on this one. Admittedly the papers I looked at this summer had fewer color maniupulations of temperature than I anticipated. They often manipulated other independent variables (resource concentration, dispersal etc).

I extracted that data for this study from Figure 3a, though I am not quite sure how I should have binned it. I extracted each of the data points for each treatment (fast--white, slow--red, and constant). However, I wasn't sure how to make sense of the x-axis which seems to be species richness, I assume it means how many organisms were in a given community since they varied community composition from 2 to 8 species. I ended up just taking the average of the data points and excluding community composition (this doesn't seem quite right) since I wasn't sure how to delineate this in the figure.

```{r, echo= FALSE}
knitr::include_graphics("meeting materials/images/petchey2002_fig3a.png")
```

\newpage

### Joshi et al 1996

I also seemed to struggle to find individual level studies that seemed easily wrangle-able. I landed on Joshi et al 1996 from my pre-existing library. I wasn't quite sure how to bin the sex differentiation between flux treatments and response variables, so I just ended up including them. I wasn't sure if I should just average the response and SE for both?

```{r, echo = FALSE }
knitr::include_graphics("meeting materials/images/joshi_tab1.png")
```
\newpage

## Preliminary Results

```{r, echo= FALSE, message= FALSE, results='hide'}
library(tidyverse)
extract<-read_csv("prelim_extact.csv")
names(extract)

ggplot(extract, aes(x=org_level, y=cohen_d, color=fluctuation_pattern))+
  geom_boxplot()+
  theme_classic()+
  ggtitle("Cohen D values for effect of flux patterns across different levels of organization")
```

## Current concerns and questions about selecting studies:

1. Hard to find studies that are very explicit about their variation regime and I’m not sure if I should be binning diurnal cycles that are sinusoidal or square in their pattern into different groups 
2. How should I bin diurnal cycles that vary in amplitude?
3. How should I be calculating effect size?
    * Should I split Petchey et al Figure by community composition number?
4. SD vs SE? Should these be treated differently and if so, how?
5. I’m now thinking about whether it would be good to section off time series data (like population density over time, or biomass over time) and mean effect sizes (an aggregate of all time series as well as just effect sizes listed in study)

### 2021-02-22 shared analysis files /data_wrangling_plot22feb.Rmd
salavaged rmd from rstudio server updates ish?

---
title: "data wrangling and plots"
author: "Maggie Slein"
date: "2/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}

## loading libraries 
pacman::p_load(metafor, tidyverse)
#devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
#library(orchaRd)

#load data
dat_MA_0<- read_csv("maggie_MA_2_8_21.csv")
```

```{r}
##wrangling calculations for metafor

#make standardized SD columns based on N and SE vs SD
dat_MA <-dat_MA_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_MA_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_MA, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% filter (!is.na(yi))

#fit initial MA model, abs (note structure )
mod1 <- rma.mv(abs(yi), vi, data=dat_MA_ES, 
                 method="REML", random = ~1 | response_id/ experiment_id/ study_id) 
summary(mod1)

#fit dirrectional model
mod2<-rma.mv(yi, vi, data=dat_MA_ES, 
             method="REML", random = ~1 | response_id/ experiment_id/ study_id) 
summary(mod2)
```

```{r}
# data wrangling for plots
count(dat_MA_ES, flux_range) #46 rows w/ range of 10, 51 w/ range of 15, 36 w/range of 6
count(dat_MA_ES, resp_units)
count(dat_MA_ES, mean_temp_constant, study_id) #24 C common accross 5 studies
count(dat_MA_ES, resp_def, sort = TRUE)
count(dat_MA_ES, genus)
count(dat_MA_ES, genus, study_id)
count(dat_MA_ES, resp_units) # body mass, avg. cum. num of eggs laid/fm, avg. num of eggs laid/fm, survival, rate of change, are the the most common reponses

normalized <- dat_MA_ES %>%
  filter(yi > -20)

common_unit_C <-dat_MA_ES %>% 
  filter(resp_units == "C")

common_unit_g <-dat_MA_ES %>% 
filter(resp_units == "g")
#filtering for genus to only be drosophila 
drosophila <- dat_MA_ES %>% 
  filter(genus == "Drosophila")

#filtering for just flux ranges of 10 C 
common_range <- dat_MA_ES %>% 
  filter(flux_range == "10")
#filtering for just studies with mean temp of 24
common_temp <- dat_MA_ES %>% 
  filter(mean_temp_constant == "24")

count(dat_MA_ES, larger_group)
count(dat_MA_ES, genus)
count(dat_MA_ES, genus, study_id)

#Drosophila used across 4 studies, potential to aggregate based on that?

#16 genuses to aggregate across 
count(dat_MA_ES, mean_temp_constant)

```

```{r}
#SMD across all studies
ggplot(normalized, aes(x=study_id, y=yi))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45))+
  ggtitle("SMD across all studies")
```

```{r}
# scatterplot of standardized mean response faceted by exp age
ggplot(dat_MA_ES, aes(x=log(vi), y=log(yi), color = study_id))+
  geom_point()+
  theme_bw()+
  facet_wrap(~exp_age)+
  ggtitle("Log transformed SMD across log transformed variance 
          colored by study and faceted by experimental age")
```

```{r}
# scatterplot of standardized mean response vs mean temp faceted by org level and colored by flux_range
ggplot(dat_MA_ES, aes(x=mean_temp_constant, y=yi, color = as.factor(flux_range)))+
  geom_point()+
  facet_wrap(~org_level)+
  theme_bw()+
  ggtitle("SMD across mean temperature colored by fluctation range 
          and faceted by organization level")
```

```{r}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(boxplot, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  ggtitle("SMD across fluctuation ranges colored by organization level 
          and fit with linear model")
```

```{r}
ggplot(boxplot, aes(x=flux_range, y=yi, color = study_id))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_minimal()+
  ggtitle("SMD across fluctuation ranges colored by 
          studies and fit with linear model by study")
```

```{r}
#trying to look at drosophila response across studies 
ggplot(drosophila, aes(y=yi, x=study_id, color = study_id))+
  geom_boxplot()+
  theme_minimal()+
  ggtitle("SMD across studies with the same study genus (Drosophila)")
```

```{r}
ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
  geom_jitter()+
  theme_minimal()+
  ggtitle("SMD across fluctuation ranges colored by genus")
```

```{r}
ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
  geom_jitter()+
  theme_minimal()+
  facet_wrap(~size)+
  ggtitle("SMD across fluctuation ranges colored by genus")
```

```{r}
#looking at yi across studies with the same temperature range (10 C)
ggplot(common_range, aes(y=yi, x=study_id, color = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
                                  ggtitle("SMD across studies with the same 
                                          temperature fluctuation range (10 C)")
```

```{r}
#looking at yi across studies with the same mean temperature (24 C)
ggplot(common_temp, aes(y=yi, x=study_id, color = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 7, angle = 45))+
  ggtitle("SMD across studies with the same mean temperature (24 C)")
```

```{r}
#trying to look at how size trends yi across increasing flux ranges 
ggplot(normalized, aes(y=abs(yi), x=flux_range, color = as.factor(size)))+
  geom_point(alpha = 0.5)+
  geom_smooth(formula = y~x)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 7, angle = 45))+
  ggtitle("Absolute SMD across fluctuation ranges colored 
          by size of study organism and fitted with model curve")
```

```{r}
ggplot(common_unit_C, aes(y=yi, x=study_id, color = as.factor(flux_range)))+
  geom_point(alpha = 0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 7, angle = 45))+
  ggtitle("SMD across studies with the common unit C (CT assays)")
```

```{r}
#plot across studies with common unit gram
ggplot(common_unit_g, aes(y=yi, x=flux_range, color = study_id))+
  geom_point(alpha =0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
  ggtitle(SMD across fluctuation ranges colored by studies with the common unit "grams")
```

### 2021-02-22 shared analysis files /data_wrangling_plot22feb.Rmd
updated plots

@@
 
-```{r}
+```{r, echo= FALSE}
 ##wrangling calculations for metafor
@@
 
-```{r}
-# data wrangling for plots
+```{r, echo= FALSE}
+## data wrangling for plots
+#count and propotions for the different variables
 count(dat_MA_ES, flux_range) #46 rows w/ range of 10, 51 w/ range of 15, 36 w/range of 6
-count(dat_MA_ES, resp_units)
+count(dat_MA_ES, resp_units, sort = TRUE) 
 count(dat_MA_ES, mean_temp_constant, study_id) #24 C common accross 5 studies
-count(dat_MA_ES, resp_def, sort = TRUE)
-count(dat_MA_ES, genus)
-count(dat_MA_ES, genus, study_id)
-count(dat_MA_ES, resp_units) # body mass, avg. cum. num of eggs laid/fm, avg. num of eggs laid/fm, survival, rate of change, are the the most common reponses
+count(dat_MA_ES, resp_def, sort = TRUE) # body mass, avg. cum. num of eggs laid/fm, avg. num of eggs laid/fm, survival, rate of change, are the the most common reponses
+count(dat_MA_ES, genus, study_id) # 3 studies featuring drosophila
+count(dat_MA_ES, genus)# 16 genuses to bin by 
+count(dat_MA_ES, size)# good spread for body size
+count(dat_MA_ES, larger_group) # pretty much mostly animal studies
 
+#removing outlier point to get better picture of data structure
 normalized <- dat_MA_ES %>%
   filter(yi > -20) 
-
+#filtering for studies with common variable C (CT assays)
 common_unit_C <-dat_MA_ES %>% 
   filter(resp_units == "C")
-
+#filtering for studies with common variable g 
 common_unit_g <-dat_MA_ES %>% 
@@
   filter(genus == "Drosophila")
-
 #filtering for just flux ranges of 10 C 
@@
   filter(mean_temp_constant == "24")
-
-count(dat_MA_ES, larger_group)
-count(dat_MA_ES, genus)
-count(dat_MA_ES, genus, study_id)
-
-#Drosophila used across 4 studies, potential to aggregate based on that?
-
-#16 genuses to aggregate across 
-count(dat_MA_ES, mean_temp_constant)
-
 ```
 
-```{r}
+```{r, echo= FALSE}
 #SMD across all studies
@@
 
-```{r}
+```{r, echo= FALSE}
 # scatterplot of standardized mean response faceted by exp age
@@
 
-```{r}
+```{r, echo= FALSE}
 # scatterplot of standardized mean response vs mean temp faceted by org level and colored by flux_range
@@
 
-```{r}
+```{r, echo= FALSE}
 # scatterplot of standardized mean response vs flux range colored and lm fit by org level
-ggplot(boxplot, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
+ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
   geom_point()+
@@
 
-```{r}
-ggplot(boxplot, aes(x=flux_range, y=yi, color = study_id))+
+```{r, echo= FALSE}
+ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
   geom_point()+
@@
 
-```{r}
+```{r, echo= FALSE}
 #trying to look at drosophila response across studies 
@@
 
-```{r}
+```{r, echo= FALSE}
 ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
-  geom_jitter()+
+  geom_jitter(alpha = 0.5)+
   theme_minimal()+
@@
 
-```{r}
-ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
-  geom_jitter()+
-  theme_minimal()+
-  facet_wrap(~size)+
-  ggtitle("SMD across fluctuation ranges colored by genus")
-```
-
-```{r}
+```{r, echo= FALSE}
 #looking at yi across studies with the same temperature range (10 C)
@@
 
-```{r}
+```{r, echo= FALSE}
 #looking at yi across studies with the same mean temperature (24 C)
@@
 
-```{r}
+```{r, echo= FALSE}
 #trying to look at how size trends yi across increasing flux ranges 
@@
 
-```{r}
+```{r, echo= FALSE}
 ggplot(common_unit_C, aes(y=yi, x=study_id, color = as.factor(flux_range)))+
@@
 
-```{r}
+```{r, echo= FALSE}
 #plot across studies with common unit gram
@@
                                   size = 7, angle = 45))+
-  ggtitle(SMD across fluctuation ranges colored by studies with the common unit "grams")
+  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
 ```

### 2021-02-22 shared analysis files /data_wrangling_plot22feb.Rmd
additional changes, adjusting size of points

@@
 ggplot(dat_MA_ES, aes(x=mean_temp_constant, y=yi, color = as.factor(flux_range)))+
-  geom_point()+
+  geom_point(alpha=0.5, size =3)+
   facet_wrap(~org_level)+
@@
 ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
-  geom_jitter(alpha = 0.5)+
+  geom_jitter(alpha = 0.5, size =3)+
   theme_minimal()+
@@
 ggplot(common_unit_C, aes(y=yi, x=study_id, color = as.factor(flux_range)))+
-  geom_point(alpha = 0.5)+
+  geom_point(alpha = 0.5, size=7)+
   theme_minimal()+
@@
 ggplot(common_unit_g, aes(y=yi, x=flux_range, color = study_id))+
-  geom_point(alpha =0.5)+
+  geom_point(alpha =0.5, size =6)+
   theme_minimal()+

### 2021-02-28 shared analysis files /data_wrangling_plot22feb.Rmd
revamping and reinventing initial figures

@@
 ## loading libraries 
-pacman::p_load(metafor, tidyverse)
+pacman::p_load(metafor, tidyverse, viridis, ggsci, wesanderson)
 #devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
@@
 #SMD across all studies
-ggplot(normalized, aes(x=study_id, y=yi))+
+glimpse(normalized)
+ggplot(normalized, aes(x=reorder(study_id, -yi), y=yi, color = as.factor(experiment_id)))+
   geom_boxplot()+
+  scale_color_tron()+
   theme_minimal()+
   theme(axis.text.x = element_text(face = "bold", 
-                                   size = 10, angle = 45))+
+                                   size = 10, angle = 45,hjust = 1), 
+        legend.position = "bottom")+
   ggtitle("SMD across all studies")
+
+#corresponding model
+fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
+               random = ~1 | response_id/ experiment_id/ study_id,
+                 method="REML")
+
+fig1
 ```
@@
 # scatterplot of standardized mean response faceted by exp age
-ggplot(dat_MA_ES, aes(x=log(vi), y=log(yi), color = study_id))+
+ggplot(dat_MA_ES, aes(x=vi, y=yi, color = study_id))+
   geom_point()+
   theme_bw()+
-  facet_wrap(~exp_age)+
-  ggtitle("Log transformed SMD across log transformed variance 
-          colored by study and faceted by experimental age")
+  ggtitle("=SMD across variance 
+          colored by study")
+```
+
+```{r}
+ggplot(normalized, aes(x=exp_age, y=yi, color = as.factor(exp_age)))+
+  geom_boxplot()+
+  theme_bw()+
+  ggtitle("SMD across different experimental ages")
 ```

### 2021-02-28 shared analysis files /data_wrangling_plot22feb.Rmd
half the updates to plots and models

--- 2021-02-22 shared analysis files /data_wrangling_plot22feb.Rmd
@@
 
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r}
-
+```{r, echo=FALSE, results = 'hide', message=FALSE}
 ## loading libraries 
-pacman::p_load(metafor, tidyverse)
+pacman::p_load(metafor, tidyverse, viridis, ggsci, wesanderson)
 #devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
@@
 
-```{r, echo= FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE}
 ##wrangling calculations for metafor
@@
 
-```{r, echo= FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE}
 ## data wrangling for plots
@@
 normalized <- dat_MA_ES %>%
-  filter(yi > -20) 
-#filtering for studies with common variable C (CT assays)
-common_unit_C <-dat_MA_ES %>% 
-  filter(resp_units == "C")
+  filter(yi > -20, resp_units != "C") 
 #filtering for studies with common variable g 
@@
 
-```{r, echo= FALSE}
+Figure 1.
+```{r, message=FALSE}
 #SMD across all studies
-ggplot(normalized, aes(x=study_id, y=yi))+
+ggplot(normalized, aes(x=reorder(study_id, -yi), y=yi, color = as.factor(experiment_id)))+
   geom_boxplot()+
+  scale_color_tron()+
   theme_minimal()+
   theme(axis.text.x = element_text(face = "bold", 
-                                   size = 10, angle = 45))+
+                                   size = 10, angle = 45,hjust = 1), 
+        legend.position = "bottom")+
   ggtitle("SMD across all studies")
+
+#corresponding model
+fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig1
+#basic linear model 
+simple1<-lm(yi ~ study_id, data = dat_MA_ES)
+summary(fig1)
 ```
 
+Figure 2. 
 ```{r, echo= FALSE}
 # scatterplot of standardized mean response faceted by exp age
-ggplot(dat_MA_ES, aes(x=log(vi), y=log(yi), color = study_id))+
+ggplot(normalized, aes(x=vi, y=yi, color = study_id))+
   geom_point()+
   theme_bw()+
-  facet_wrap(~exp_age)+
-  ggtitle("Log transformed SMD across log transformed variance 
-          colored by study and faceted by experimental age")
+  ggtitle("SMD across variance 
+          colored by study")
+
+#basic linear model looking at how vi interacts with yi
+simple2<-lm(yi ~ vi, data = dat_MA_ES)
+summary(simple2)
 ```
 
-```{r, echo= FALSE}
-# scatterplot of standardized mean response vs mean temp faceted by org level and colored by flux_range
-ggplot(dat_MA_ES, aes(x=mean_temp_constant, y=yi, color = as.factor(flux_range)))+
-  geom_point()+
-  facet_wrap(~org_level)+
+Figure 3. 
+```{r, message=FALSE}
+#boxplot
+ggplot(normalized, aes(x=exp_age, y=yi, color = as.factor(exp_age)))+
+  geom_boxplot()+
   theme_bw()+
-  ggtitle("SMD across mean temperature colored by fluctation range 
-          and faceted by organization level")
+  ggtitle("SMD across different experimental ages")
+
+#random effects model including exp_age as a random variable
+fig3 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~exp_age, 
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig3
+#simple linear model looking at how exp_age affects yi
+simple3<-lm(yi~exp_age, data =dat_MA_ES)
+summary(simple3)
 ```
 
-```{r, echo= FALSE}
+Figure 4. 
+```{r, message=FALSE}
+ggplot(normalized, aes(x=exp_age, y=vi, color = as.factor(exp_age)))+
+  geom_boxplot()+
+  theme_bw()+
+  ggtitle("Variance across different experimental ages")
+
+count(normalized, exp_age)
+
+#simple linear model looking at how exp_age affects vi
+simple4<-lm(vi~ exp_age, data =dat_MA_ES)
+summary(simple4)
+```
+
+Figure 5. 
+```{r, message=FALSE}
+#boxplot of SMD across levels of organization
+ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
+  geom_boxplot(alpha =0.7)+
+  scale_fill_tron()+
+  theme_bw()+
+  theme(legend.position = "bottom")+
+  ggtitle("SMD across organization level")
+
+#random effects model including org_level as a random variable
+fig5 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~org_level,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig5
+#simple linear model looking at how org_level affects yi
+simple5<-lm(yi~org_level, data =dat_MA_ES)
+summary(simple5)
+```
+
+Figure 6.
+```{r, message=FALSE}
+# boxplots of how mean temperature influences SMD
+ggplot(normalized, aes(x=as.factor(mean_temp_constant), y=yi))+
+  geom_boxplot()+
+  theme_bw()+
+  ggtitle("SMD across mean temperature")
+
+#random effects model including org_level as a random variable
+fig6 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~mean_temp_constant,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig6
+#simple linear model looking at how mean_temp_constant affects yi
+simple6<-lm(yi~mean_temp_constant, data =dat_MA_ES)
+summary(simple6)
+```
+
+Figure 7.
+```{r, message=FALSE}
+# boxplots of how fluctuation range influences SMD
+ggplot(normalized, aes(x=as.factor(flux_range), y=yi))+
+  geom_boxplot()+
+  theme_bw()+
+  ggtitle("SMD across flux_range")
+
+#random effects model including org_level as a random variable
+fig7 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig7 
+#simple linear model looking at how flux_range affects yi
+simple7<-lm(yi~flux_range, data =dat_MA_ES)
+summary(simple7)
+```
+
+Figure 8.
+```{r, message=FALSE}
 # scatterplot of standardized mean response vs flux range colored and lm fit by org level
@@
           and fit with linear model")
+
+#random effects model including org_level as a random variable
+fig8 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + org_level,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig8
+#simple linear model looking at how flux_range and org_level affect yi
+simple8<-lm(yi~flux_range*org_level, data =dat_MA_ES)
+summary(simple8)
 ```
 
-```{r, echo= FALSE}
+Figure 9.
+```{r, message=FALSE}
+#scatterplot of standardized mean difference across flux range colored by study fit with lm 
 ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
@@
           studies and fit with linear model by study")
+
+#random effects model including org_level as a random variable
+fig9 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + study_id,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig9
+#simple linear model looking at how flux_range and study_id affect yi
+simple9<-lm(yi~flux_range*study_id, data =dat_MA_ES)
+summary(simple9)
 ```
 
-```{r, echo= FALSE}
+Figure 10. 
+```{r, message=FALSE}
 #trying to look at drosophila response across studies 
-ggplot(drosophila, aes(y=yi, x=study_id, color = study_id))+
+ggplot(drosophila, aes(y=yi, x=as.factor(flux_range), color = study_id))+
   geom_boxplot()+
-  theme_minimal()+
+  scale_color_tron()+
+  theme_bw()+
   ggtitle("SMD across studies with the same study genus (Drosophila)")
+#random effects model including org_level as a random variable
+fig10 <- rma.mv(yi, vi, data=drosophila, mods = ~flux_range + study_id,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig10
+#simple linear model looking at how flux_range and study_id affect yi in drosphila specific studies
+simple10<-lm(yi~flux_range + study_id, data =drosophila)
+summary(simple10)
 ```
 
-```{r, echo= FALSE}
+```{r, message=FALSE}
 ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
-  geom_jitter(alpha = 0.5)+
+  geom_jitter(alpha = 0.5, size =3)+
   theme_minimal()+
@@
 
-```{r, echo= FALSE}
+```{r, message=FALSE}
 #looking at yi across studies with the same temperature range (10 C)
@@
 
-```{r, echo= FALSE}
+```{r, message=FALSE}
 #looking at yi across studies with the same mean temperature (24 C)
@@
 
-```{r, echo= FALSE}
+```{r, message=FALSE}
 #trying to look at how size trends yi across increasing flux ranges 
@@
 
-```{r, echo= FALSE}
-ggplot(common_unit_C, aes(y=yi, x=study_id, color = as.factor(flux_range)))+
-  geom_point(alpha = 0.5)+
-  theme_minimal()+
-  theme(axis.text.x = element_text(face = "bold", 
-                                   size = 7, angle = 45))+
-  ggtitle("SMD across studies with the common unit C (CT assays)")
-```
-
-```{r, echo= FALSE}
+```{r, message=FALSE}
 #plot across studies with common unit gram
 ggplot(common_unit_g, aes(y=yi, x=flux_range, color = study_id))+
-  geom_point(alpha =0.5)+
+  geom_point(alpha =0.5, size =6)+
   theme_minimal()+

### 2021-03-01 shared analysis files /data_wrangling_plot22feb.Rmd
re-wranging and rethinking plots for thesis meeting

@@
 
-#removing outlier point to get better picture of data structure
+#removing outlier point to get better picture of data structure AND excluding CT assays 
 normalized <- dat_MA_ES %>%
@@
 common_unit_g <-dat_MA_ES %>% 
-filter(resp_units == "g")
+filter(resp_def == "body mass")
 #filtering for genus to only be drosophila 
@@
 
-Figure 1.
+# Plots 
+## Question: How does response compare across studies and experiments?
+### Figure 1.
 ```{r, message=FALSE}
@@
   ggtitle("SMD across all studies")
-
-#corresponding model
+#corresponding random effects model 
 fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
@@
 fig1
+#trying out mixed effects model 
+fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
+                 method="FE")
+fig1me
 #basic linear model 
@@
 
-Figure 2. 
+### Figure 2. 
 ```{r, echo= FALSE}
@@
 
-Figure 3. 
+## Question: How does fluctuation amplitude affect response variables?
+
+### Figure 3.
+```{r, message=FALSE}
+# boxplots of how fluctuation range influences SMD
+ggplot(normalized, aes(x=as.factor(flux_range), y=yi))+
+  geom_boxplot()+
+  theme_bw()+
+  ggtitle("SMD across flux_range")
+
+#random effects model including org_level as a random variable
+fig7 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig7 
+#simple linear model looking at how flux_range affects yi
+simple7<-lm(yi~flux_range, data =dat_MA_ES)
+summary(simple7)
+```
+
+### Figure 4.
+```{r, message=FALSE}
+# scatterplot of standardized mean response vs flux range colored and lm fit by org level
+ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
+  geom_point()+
+  geom_smooth(method="lm", formula = y~x)+
+  ggtitle("SMD across fluctuation ranges colored by organization level 
+          and fit with linear model")
+
+#random effects model including org_level as a random variable
+fig8 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + org_level,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig8
+#simple linear model looking at how flux_range and org_level affect yi
+simple8<-lm(yi~flux_range*org_level, data =dat_MA_ES)
+summary(simple8)
+```
+
+### Figure 5.
+```{r, message=FALSE}
+#scatterplot of standardized mean difference across flux range colored by study fit with lm 
+ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
+  geom_point()+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_minimal()+
+  ggtitle("SMD across fluctuation ranges colored by 
+          studies and fit with linear model by study")
+
+#random effects model including org_level as a random variable
+fig9 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + study_id,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig9
+#simple linear model looking at how flux_range and study_id affect yi
+simple9<-lm(yi~flux_range*study_id, data =dat_MA_ES)
+summary(simple9)
+```
+
+### Figure 6.
+```{r, message=FALSE}
+#looking at yi across studies with the same temperature range (10 C)
+ggplot(common_range, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
+  geom_boxplot()+
+  theme_minimal()+
+  theme(axis.text.x = element_text(face = "bold", 
+                                  size = 7, angle = 45))+
+                                  ggtitle("SMD across studies with the same 
+                                          temperature fluctuation range (10 C)")
+#random effects model within subset data
+fig12 <- rma.mv(yi, vi, data=common_range,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig12
+#simple linear model looking at how flux_range affects yi within subset data
+simple12<-lm(yi~flux_range, data =common_range)
+summary(simple12)
+```
+
+## Question how does mean temperature affect response variables?
+
+### Figure 7.
+```{r, message=FALSE}
+# boxplots of how mean temperature influences SMD
+ggplot(normalized, aes(x=as.factor(mean_temp_constant), y=yi))+
+  geom_boxplot()+
+  theme_bw()+
+  ggtitle("SMD across mean temperature")
+
+#random effects model including org_level as a random variable
+fig6 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~mean_temp_constant,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig6
+#simple linear model looking at how mean_temp_constant affects yi
+simple6<-lm(yi~mean_temp_constant, data =dat_MA_ES)
+summary(simple6)
+```
+
+### Figure 8.
+```{r, message=FALSE}
+#looking at yi across studies with the same mean temperature (24 C)
+ggplot(common_temp, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
+  geom_boxplot()+
+  theme_minimal()+
+  theme(axis.text.x = element_text(face = "bold", 
+                                   size = 7, angle = 45))+
+  ggtitle("SMD across studies with the same mean temperature (24 C)")
+#random effects model within subset data
+fig13 <- rma.mv(yi, vi, data=common_temp,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig13
+#simple linear model looking at how flux_range affects yi within subset data
+simple13<-lm(yi~mean_temp_constant, data =common_temp)
+summary(simple13)
+```
+
+## Question: How does genus affect response?
+
+### Figure 9. 
+```{r, message=FALSE}
+#trying to look at drosophila response across studies 
+ggplot(drosophila, aes(y=yi, x=as.factor(flux_range), color = study_id))+
+  geom_boxplot()+
+  scale_color_tron()+
+  theme_bw()+
+  ggtitle("SMD across studies with the same study genus (Drosophila)")
+#random effects model including org_level as a random variable
+fig10 <- rma.mv(yi, vi, data=drosophila, mods = ~flux_range + study_id,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig10
+#simple linear model looking at how flux_range and study_id affect yi in drosphila specific studies
+simple10<-lm(yi~flux_range + study_id, data =drosophila)
+summary(simple10)
+```
+
+### Figure 10.
+```{r, message=FALSE}
+#boxplots of SMD across different genuses colored by body size
+ggplot(normalized, aes(y=yi, x= reorder(genus, -yi), color = as.factor(size)))+
+  geom_boxplot()+
+  theme_minimal()+
+  ggtitle("SMD across genuses colored by body size")+
+  theme(axis.text.x = element_text(face = "bold", 
+                                   size = 10, angle = 45,hjust = 1), 
+        legend.position = "bottom")
+#random effects model including genus as a random variable
+fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~genus,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig11
+#simple linear model looking at how genus affects yi
+simple11<-lm(yi~genus, data =dat_MA_ES)
+summary(simple11)
+```
+
+## Question how does scale (life or org level) affect response?
+### Figure 11. 
 ```{r, message=FALSE}
@@
 
-Figure 4. 
+### Figure 12. 
 ```{r, message=FALSE}
@@
 
-Figure 5. 
+### Figure 13. 
 ```{r, message=FALSE}
@@
 
-Figure 6.
+## Question: How does response variable affect response?
+### Figure 14.
 ```{r, message=FALSE}
-# boxplots of how mean temperature influences SMD
-ggplot(normalized, aes(x=as.factor(mean_temp_constant), y=yi))+
-  geom_boxplot()+
-  theme_bw()+
-  ggtitle("SMD across mean temperature")
-
-#random effects model including org_level as a random variable
-fig6 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~mean_temp_constant,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig6
-#simple linear model looking at how mean_temp_constant affects yi
-simple6<-lm(yi~mean_temp_constant, data =dat_MA_ES)
-summary(simple6)
-```
-
-Figure 7.
-```{r, message=FALSE}
-# boxplots of how fluctuation range influences SMD
-ggplot(normalized, aes(x=as.factor(flux_range), y=yi))+
-  geom_boxplot()+
-  theme_bw()+
-  ggtitle("SMD across flux_range")
-
-#random effects model including org_level as a random variable
-fig7 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig7 
-#simple linear model looking at how flux_range affects yi
-simple7<-lm(yi~flux_range, data =dat_MA_ES)
-summary(simple7)
-```
-
-Figure 8.
-```{r, message=FALSE}
-# scatterplot of standardized mean response vs flux range colored and lm fit by org level
-ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
-  geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  ggtitle("SMD across fluctuation ranges colored by organization level 
-          and fit with linear model")
-
-#random effects model including org_level as a random variable
-fig8 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + org_level,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig8
-#simple linear model looking at how flux_range and org_level affect yi
-simple8<-lm(yi~flux_range*org_level, data =dat_MA_ES)
-summary(simple8)
-```
-
-Figure 9.
-```{r, message=FALSE}
-#scatterplot of standardized mean difference across flux range colored by study fit with lm 
-ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
-  geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  theme_minimal()+
-  ggtitle("SMD across fluctuation ranges colored by 
-          studies and fit with linear model by study")
-
-#random effects model including org_level as a random variable
-fig9 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + study_id,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig9
-#simple linear model looking at how flux_range and study_id affect yi
-simple9<-lm(yi~flux_range*study_id, data =dat_MA_ES)
-summary(simple9)
-```
-
-Figure 10. 
-```{r, message=FALSE}
-#trying to look at drosophila response across studies 
-ggplot(drosophila, aes(y=yi, x=as.factor(flux_range), color = study_id))+
-  geom_boxplot()+
-  scale_color_tron()+
-  theme_bw()+
-  ggtitle("SMD across studies with the same study genus (Drosophila)")
-#random effects model including org_level as a random variable
-fig10 <- rma.mv(yi, vi, data=drosophila, mods = ~flux_range + study_id,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig10
-#simple linear model looking at how flux_range and study_id affect yi in drosphila specific studies
-simple10<-lm(yi~flux_range + study_id, data =drosophila)
-summary(simple10)
-```
-
-```{r, message=FALSE}
-ggplot(normalized, aes(y=yi, x=flux_range, color = genus))+
-  geom_jitter(alpha = 0.5, size =3)+
-  theme_minimal()+
-  ggtitle("SMD across fluctuation ranges colored by genus")
-```
-
-```{r, message=FALSE}
-#looking at yi across studies with the same temperature range (10 C)
-ggplot(common_range, aes(y=yi, x=study_id, color = study_id))+
+#plot across studies with common response variable body mass
+ggplot(common_unit_g, aes(y=yi, color=as.factor(flux_range), x = study_id))+
   geom_boxplot()+
@@
                                   size = 7, angle = 45))+
-                                  ggtitle("SMD across studies with the same 
-                                          temperature fluctuation range (10 C)")
+  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
+#random effects model within subset data
+fig14 <- rma.mv(yi, vi, data=common_unit_g,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig14
+#simple linear model looking at how flux_range affects yi within subset data
+simple14<-lm(yi~study_id, data =common_unit_g)
+summary(simple14)
 ```
-
-```{r, message=FALSE}
-#looking at yi across studies with the same mean temperature (24 C)
-ggplot(common_temp, aes(y=yi, x=study_id, color = study_id))+
-  geom_boxplot()+
-  theme_minimal()+
-  theme(axis.text.x = element_text(face = "bold", 
-                                   size = 7, angle = 45))+
-  ggtitle("SMD across studies with the same mean temperature (24 C)")
-```
-
-```{r, message=FALSE}
-#trying to look at how size trends yi across increasing flux ranges 
-ggplot(normalized, aes(y=abs(yi), x=flux_range, color = as.factor(size)))+
-  geom_point(alpha = 0.5)+
-  geom_smooth(formula = y~x)+
-  theme_minimal()+
-  theme(axis.text.x = element_text(face = "bold", 
-                                   size = 7, angle = 45))+
-  ggtitle("Absolute SMD across fluctuation ranges colored 
-          by size of study organism and fitted with model curve")
-```
-
-```{r, message=FALSE}
-#plot across studies with common unit gram
-ggplot(common_unit_g, aes(y=yi, x=flux_range, color = study_id))+
-  geom_point(alpha =0.5, size =6)+
-  theme_minimal()+
-  theme(axis.text.x = element_text(face = "bold", 
-                                  size = 7, angle = 45))+
-  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
-```

### 2021-03-03 shared analysis files /data_wrangling_plot22feb.Rmd
adding in full model with response variable as a mod, officially removing CT assays from dataset

@@
 # Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
-dat_MA_ES<-dat_MA_ES %>% filter (!is.na(yi))
+dat_MA_ES<-dat_MA_ES %>% 
+  filter (!is.na(yi)) %>%
+  filter(resp_units != "C")
 
@@
 normalized <- dat_MA_ES %>%
-  filter(yi > -20, resp_units != "C") 
+  filter(yi > -20) 
 #filtering for studies with common variable g 
@@
 ### Figure 11.
+
 ```{r, message=FALSE}
@@
 ### Figure 14.
+```{r}
+#full random effects model looking at all the responses in the entire dataset
+response_variable_mod <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~resp_def, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+response_variable_mod
+
+#full random effects model looking at all the response units in the entire dataset
+response_units_mod <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~resp_units, 
+               random = ~1 | experiment_id/ study_id /response_id,
+                 method="REML") 
+response_units_mod
+```
+
 ```{r, message=FALSE}

### 2021-03-07 gbif_datawranging_7march.R
r file for gbif data

@@ -1,0 +1,118 @@
+#grabbing data from rgbif 
+#extracting occurences for each species as available in gbif
+pacman::p_load(rvest, rgbif, tidyverse)
+c_reevesii <- occ_search(scientificName = "Chinemys reevesii")
+a <- c_reevesii$data %>% 
+  as.data.frame()%>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
+x_arvicola <- occ_search(scientificName = "Xylotrechus arvicola")
+b <- x_arvicola$data %>% 
+  as.data.frame()%>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
+g_firmus <- occ_search(scientificName = "Gryllus firmus")
+c <- g_firmus$data %>% 
+  as.data.frame()%>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
+l_peronii <- occ_search(scientificName = "Limnodynastes	peronii")
+d <- l_peronii$data %>% 
+  as.data.frame() %>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
+l_tasmaniensis <- occ_search(scientificName = "Limnodynastes	tasmaniensis")
+e <- l_tasmaniensis$data %>% 
+  as.data.frame()%>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
+d_simulans <- occ_search(scientificName = "Drosophila simulans")
+f <- d_simulans$data %>%
+  as.data.frame()%>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
+r_prolixus <- occ_search(scientificName = "Rhodnius	prolixus")
+g <- r_prolixus$data %>%
+  as.data.frame()%>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
+c_latirostris <- occ_search(scientificName = "Caiman latirostris")
+h <- c_latirostris$data %>%
+  as.data.frame()%>% 
+  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  drop_na()
+
[…]

### 2021-03-08 shared analysis files /data_wrangling_plot22feb.Rmd
gbif data wrangling file, thesis meeting explanation of wrangling + questions, modified datawrangling plots

@@
 #load data
-dat_MA_0<- read_csv("maggie_MA_2_8_21.csv")
+dat_MA_0<- read_csv("metafor7march.csv")
+regions <- read_csv("gbif_regiondata.csv")
 ```
@@
 count(dat_MA_ES, genus, study_id) # 3 studies featuring drosophila
-count(dat_MA_ES, genus)# 16 genuses to bin by 
+count(dat_MA_ES, genus, species)# 16 genuses to bin by 
 count(dat_MA_ES, size)# good spread for body size

### 2021-03-08 thesis_meeting7march.Rmd
gbif data wrangling file, thesis meeting explanation of wrangling + questions, modified datawrangling plots

---
title: "Thesis meeting"
author: "Maggie Slein"
date: "3/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#loading libraries 
pacman::p_load(rvest, rgbif, tidyverse, rnaturalearth, rnaturalearthdata, rgeos)
```

## Pulling down data from GBIF

Notes:

- Excluded genuses from the final meta analysis dataset that did not have greater than 8 sample size (included Platyplectum ornatum (6), Natrix natrix (5), Trachemys scripta (2), Leptopilinia boulardii (6), Clematis vitalba (1))

- Any of the remaining organisms have a sample size of at least 8 when aggregated by genus

- S. enterica was the only organism with no data on GBIF and was therefor leftout

```{r}
#extracting occurences for each species as available in gbif
c_reevesii <- occ_search(scientificName = "Chinemys reevesii")
a <- c_reevesii$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

x_arvicola <- occ_search(scientificName = "Xylotrechus arvicola")
b <- x_arvicola$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

g_firmus <- occ_search(scientificName = "Gryllus firmus")
c <- g_firmus$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

l_peronii <- occ_search(scientificName = "Limnodynastes	peronii")
d <- l_peronii$data %>% 
  as.data.frame() %>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

l_tasmaniensis <- occ_search(scientificName = "Limnodynastes	tasmaniensis")
e <- l_tasmaniensis$data %>% 
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

d_simulans <- occ_search(scientificName = "Drosophila simulans")
f <- d_simulans$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

r_prolixus <- occ_search(scientificName = "Rhodnius	prolixus")
g <- r_prolixus$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

c_latirostris <- occ_search(scientificName = "Caiman latirostris")
h <- c_latirostris$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

d_melanogaster <- occ_search(scientificName = "Drosophila melanogaster")
i <- d_melanogaster$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

p_chinensis <- occ_search(scientificName = "Plestiodon chinensis")
j <- p_chinensis$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

#no country column 
e_coli <- occ_search(scientificName = "Escherichia 	coli")
k <- e_coli$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()

#zero observations 
s_enetrica <- occ_search(scientificName = "Salmonella	enetrica")$data %>%
  as.data.frame()

d_birchii <- occ_search(scientificName = "Drosophila birchii")
l <- d_birchii$data %>%
  as.data.frame()%>% 
  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
  drop_na()
#binding all the data together
gbif_data <- bind_rows(a,b,c,d,e,f,g,h,i,j,k,l) %>%
  drop_na()
```

```{r}
#binning data into different temperature zones
regions <- gbif_data %>% 
  mutate(thermal_zone = case_when(between(decimalLatitude, -20, 20.9) ~ "tropical", 
                                  between(decimalLatitude, 21, 40.9) ~ "subtropical",
                                  between(decimalLatitude, -21, -40.9) ~ "subtropical",
                                  between(decimalLatitude, -41, -60.9) ~ "temperate",
                                  between(decimalLatitude, 41, 60.9) ~ "temperate", 
                                  between(decimalLatitude, 61, 90) ~ "arctic", 
                                  between(decimalLatitude, -61, -90) ~ "arctic")) %>%
  mutate(genus = case_when(species == "Caiman latirostris" ~ "Caiman", 
         species %in% c("Drosophila birchii", "Drosophila melanogaster", 
                        "Drosophila simulans") ~ "Drosophila", 
         species == "Escherichia coli " ~ "Escherichia", 
         species == "Gryllus firmus " ~ "Gryllus",
         species %in% c("Limnodynastes peronii", 
                        "Limnodynastes tasmaniensis") ~ "Limnodynastes",
         species == "Mauremys reevesii" ~ "Mauremys",
         species == "Plestiodon chinensis " ~ "Plestiodon",
         species == "Rhodnius prolixus" ~ "Rhodnius",
         species == "Xylotrechus arvicola" ~ "Xylotrechus")) %>%
  drop_na()
```

## Plotting preliminary plot from GBIF data

- The climate reference bands were pulle from Devinei et al 2015 with the following groups:

Tropical: 20 N - 20 S
Subtropical: 20-40 N, 20-40 S
Extratropical: 40-60 N, 40-60 S

Added in my own category of polar: anything above 60 N or below 60 S

```{r}
#generating data for map overlay
world <- ne_countries(scale = "medium", returnclass = "sf")
#plotting up data 
ggplot(data = world)+
  geom_sf()+
  geom_point(data = regions, aes(x=decimalLongitude, y=decimalLatitude, color=genus), alpha = 0.5)+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -20, ymax = 20, alpha = 0.2, fill = "lightcoral")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 20, ymax = 40, alpha = 0.2, fill = "lightgoldenrod")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -40, ymax = -20, alpha = 0.2, fill = "lightgoldenrod")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 40, ymax = 60, alpha = 0.2, fill = "yellowgreen")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -60, ymax = -40, alpha = 0.2, fill = "yellowgreen")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = 60, ymax = 90, alpha = 0.2, fill = "skyblue")+
  annotate("rect",xmin = -Inf, xmax = Inf, ymin = -90, ymax = -60, alpha = 0.2, fill = "skyblue")+
  geom_hline(yintercept = 0, alpha =0.5, linetype= "dashed")+
  theme_classic()+
  theme(legend.position="bottom")+
  ylab("Latitude")+
  xlab("Longitude")
```

```{r}
#writing to csv 
write_csv(regions, "gbif_regiondata.csv")
```

## Running Questions:

- How should thermal regimes be designated? By latitude or country code? Losing some occurence data because country is listed but latitude/longitude are not.

- How can I successfully join this dataset with the meta analysis dataset to run a model that takes into account location? How should I handle some species that span multiple zones (i.e. Drosophila )

-Should I be grouping by genus or by species?

### 2021-03-08 shared analysis files /data_wrangling_plot22feb.Rmd
saving gbif wrangling, trying to compile pdf for thesis meeting and edits to preliminary plots

@@
 #load data
-dat_MA_0<- read_csv("metafor7march.csv")
+dat_MA_0<- read_csv("metafor.csv")
 regions <- read_csv("gbif_regiondata.csv")
@@
   filter (!is.na(yi)) %>%
-  filter(resp_units != "C")
+  filter(resp_units != "C") %>%
+  filter(study_id != "piccau2017")
 
@@
 # boxplots of how fluctuation range influences SMD
-ggplot(normalized, aes(x=as.factor(flux_range), y=yi))+
-  geom_boxplot()+
+ggplot(normalized, aes(x=flux_range, y=yi))+
+  geom_point()+
   theme_bw()+
+  geom_smooth(method="lm", formula = y~x)+
   ggtitle("SMD across flux_range")
@@
   geom_boxplot()+
+  geom_point(alpha = 0.5)+
   theme_minimal()+
@@
 ```{r, message=FALSE}
-# boxplots of how mean temperature influences SMD
-ggplot(normalized, aes(x=as.factor(mean_temp_constant), y=yi))+
-  geom_boxplot()+
+# scatterplot of how mean temperature influences SMD
+ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
+  geom_point()+
   theme_bw()+
@@
 ggplot(normalized, aes(y=yi, x= reorder(genus, -yi), color = as.factor(size)))+
-  geom_boxplot()+
+  geom_point()+
   theme_minimal()+
@@
 #random effects model including genus as a random variable
-fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~genus,
+fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~size,
                random = ~1 | experiment_id/ study_id,
@@
 
-```{r, message=FALSE}
+___________________________________
+Leftover/excluded plots
+
+```{r, message=FALSE, results = 'hide'}
 #plot across studies with common response variable body mass

### 2021-03-08 thesis_meeting7march.Rmd
saving gbif wrangling, trying to compile pdf for thesis meeting and edits to preliminary plots

@@
 
-- S. enterica was the only organism with no data on GBIF and was therefor leftout
-
-```{r}
-#extracting occurences for each species as available in gbif
-c_reevesii <- occ_search(scientificName = "Chinemys reevesii")
-a <- c_reevesii$data %>% 
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-x_arvicola <- occ_search(scientificName = "Xylotrechus arvicola")
-b <- x_arvicola$data %>% 
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-g_firmus <- occ_search(scientificName = "Gryllus firmus")
-c <- g_firmus$data %>% 
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-l_peronii <- occ_search(scientificName = "Limnodynastes	peronii")
-d <- l_peronii$data %>% 
-  as.data.frame() %>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-l_tasmaniensis <- occ_search(scientificName = "Limnodynastes	tasmaniensis")
-e <- l_tasmaniensis$data %>% 
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-d_simulans <- occ_search(scientificName = "Drosophila simulans")
-f <- d_simulans$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-r_prolixus <- occ_search(scientificName = "Rhodnius	prolixus")
-g <- r_prolixus$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-c_latirostris <- occ_search(scientificName = "Caiman latirostris")
-h <- c_latirostris$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-d_melanogaster <- occ_search(scientificName = "Drosophila melanogaster")
-i <- d_melanogaster$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-p_chinensis <- occ_search(scientificName = "Plestiodon chinensis")
-j <- p_chinensis$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-#no country column 
-e_coli <- occ_search(scientificName = "Escherichia 	coli")
-k <- e_coli$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-
-#zero observations 
-s_enetrica <- occ_search(scientificName = "Salmonella	enetrica")$data %>%
-  as.data.frame()
-
-d_birchii <- occ_search(scientificName = "Drosophila birchii")
-l <- d_birchii$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
-#binding all the data together
-gbif_data <- bind_rows(a,b,c,d,e,f,g,h,i,j,k,l) %>%
-  drop_na()
-```
-
-```{r}
-#binning data into different temperature zones
-regions <- gbif_data %>% 
-  mutate(thermal_zone = case_when(between(decimalLatitude, -20, 20.9) ~ "tropical", 
-                                  between(decimalLatitude, 21, 40.9) ~ "subtropical",
-                                  between(decimalLatitude, -21, -40.9) ~ "subtropical",
-                                  between(decimalLatitude, -41, -60.9) ~ "temperate",
-                                  between(decimalLatitude, 41, 60.9) ~ "temperate", 
-                                  between(decimalLatitude, 61, 90) ~ "arctic", 
-                                  between(decimalLatitude, -61, -90) ~ "arctic")) %>%
-  mutate(genus = case_when(species == "Caiman latirostris" ~ "Caiman", 
-         species %in% c("Drosophila birchii", "Drosophila melanogaster", 
-                        "Drosophila simulans") ~ "Drosophila", 
-         species == "Escherichia coli " ~ "Escherichia", 
-         species == "Gryllus firmus " ~ "Gryllus",
-         species %in% c("Limnodynastes peronii", 
-                        "Limnodynastes tasmaniensis") ~ "Limnodynastes",
-         species == "Mauremys reevesii" ~ "Mauremys",
-         species == "Plestiodon chinensis " ~ "Plestiodon",
-         species == "Rhodnius prolixus" ~ "Rhodnius",
-         species == "Xylotrechus arvicola" ~ "Xylotrechus")) %>%
-  drop_na()
-```
+- S. enterica and E. coli had no data on GBIF and was therefore leftout.
 
@@
 #generating data for map overlay
+regions <- read_csv("gbif_regiondata.csv")
 world <- ne_countries(scale = "medium", returnclass = "sf")
@@
 ```{r}
-#writing to csv 
-write_csv(regions, "gbif_regiondata.csv")
+#attempting to pull thermal performance data from devRate
+
+myDf <- exTropicalMoth$raw$egg
+devRateModelAll(dfData = myDf)
+
+devRateFind(species = "Drosophila melanogaster")
+devRateFind(species = "Drosophila simulans")
 ```
@@
 -Should I be grouping by genus or by species?
+
+________
+
+Additional notes/questions from figures last week
+
+- Semenov 2007 and Kern et al 2015 were the two studies with the strangely smooth curves
+
+  - Kern et al 2015--I changed the sample sizes but not sure they are right?
+    - Looks less curved after sample size change
+  -Not sure about the sample size for Semenov: sample size of 3 for 3 replicates, should I be multiplying that by the original CFU count?
+
+- Removed Piccau et al (the big negative yi)--only study without a comparable constant and fluctuating mean
+
+- How to best determine whether or not the temperatures each species is experiencing are heat stress temperatures?
+  -It looks like in the final processes meta-analysis file there are quite a few species with range of temperatures
+
+-Would it be useful to pull in and extract the thermal performance acclimation studies?

### 2021-03-08 gbif_datawranging_7march.R

@@ -1,8 +1,9 @@
-#grabbing data from rgbif 
+#loading libraries 
+pacman::p_load(rgbif, tidyverse)
+
 #extracting occurences for each species as available in gbif
-pacman::p_load(rvest, rgbif, tidyverse, rnaturalearth, rnaturalearthdata, rgeos)
-c_reevesii <- occ_search(scientificName = "Chinemys reevesii")
-a <- c_reevesii$data %>% 
+m_reevesii <- occ_search(scientificName = "Mauremys reevesii")
+a <- m_reevesii$data %>% 
   as.data.frame()%>% 
   select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
   drop_na()
@@ -61,16 +62,16 @@ j <- p_chinensis$data %>%
   select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
   drop_na()
 
-#no country column 
-e_coli <- occ_search(scientificName = "Escherichia 	coli")
-k <- e_coli$data %>%
-  as.data.frame()%>% 
-  select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
-  drop_na()
+#zero observations 
+#e_coli <- occ_search(scientificName = "Escherichia coli")
+#k <- e_coli$data %>%
+  #as.data.frame()%>% 
+  #select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
+  #drop_na()
 
 #zero observations 
-s_enetrica <- occ_search(scientificName = "Salmonella	enetrica")$data %>%
-  as.data.frame()
+#s_enetrica <- occ_search(scientificName = "Salmonella	enetrica")$data %>%
+#as.data.frame()
 
 d_birchii <- occ_search(scientificName = "Drosophila birchii")
 l <- d_birchii$data %>%
@@ -78,30 +79,8 @@ l <- d_birchii$data %>%
   select(countryCode, genus, species,  decimalLongitude, decimalLatitude) %>%
   drop_na()
 #binding all the data together
-gbif_data <- bind_rows(a,b,c,d,e,f,g,h,i,j,k,l) %>%
-  drop_na
-
-#generating data for map overlay
-world <- ne_countries(scale = "medium", returnclass = "sf")
-
-?ne_countries
-#plotting up data 
-ggplot(data = world)+
-  geom_sf()+
-  geom_point(data = gbif_data, aes(x=decimalLongitude, y=decimalLatitude, color=species), alpha = 0.5)+
[…]

### 2021-03-08 thesis_meeting7march.Rmd
updating gbif R script and meeting file

@@
 #loading libraries 
-pacman::p_load(rvest, rgbif, tidyverse, rnaturalearth, rnaturalearthdata, rgeos)
+pacman::p_load(rvest, rgbif, tidyverse, rnaturalearth, rnaturalearthdata, rgeos, devRate)
 ```
@@
 #generating data for map overlay
-regions <- read_csv("gbif_regiondata.csv")
+regions <- read_csv("gbif_region_data.csv")
 world <- ne_countries(scale = "medium", returnclass = "sf")
@@
 
-```{r}
+```{r, results='hide'}
 #attempting to pull thermal performance data from devRate
@@
 devRateModelAll(dfData = myDf)
-
 devRateFind(species = "Drosophila melanogaster")

### 2021-03-08 gbif_datawranging_7march.R

@@ -82,7 +82,7 @@ l <- d_birchii$data %>%
 gbif_data <- bind_rows(a,b,c,d,e,f,g,h,i,j,l) %>%
   drop_na()
 
-#binning data into different temperature zones
+#binning data into different temperature zones and creating a genus column
 regions <- gbif_data %>% 
   mutate(thermal_zone = case_when(between(decimalLatitude, -20, 20.9) ~ "tropical", 
                                   between(decimalLatitude, 21, 40.9) ~ "subtropical",
@@ -103,8 +103,8 @@ regions <- gbif_data %>%
                            species == "Rhodnius prolixus" ~ "Rhodnius",
                            species == "Xylotrechus arvicola" ~ "Xylotrechus")) %>%
   drop_na()
-
-write_csv(regions, "gbif_regiondata.csv")
-
+#writing all this data to a csv 
+write_csv(regions, "gbif_region_data.csv")
+#spread of geographic data
 count<-count(regions, countryCode, thermal_zone, species)
 glimpse(regions)

### 2021-03-08 shared analysis files /data_wrangling_plot22feb.Rmd
updated rmd and pdf for first pass plots
updated and re-arranged questions and plots

--- 2021-03-08 shared analysis files /data_wrangling_plot22feb.Rmd
@@
 author: "Maggie Slein"
-date: "2/22/2021"
+date: "3/8/21"
 output: pdf_document
@@
 #load data
-dat_MA_0<- read_csv("metafor7march.csv")
-regions <- read_csv("gbif_regiondata.csv")
+dat_MA_0<- read_csv("metafor.csv")
 ```
@@
   filter (!is.na(yi)) %>%
-  filter(resp_units != "C")
-
-#fit initial MA model, abs (note structure )
-mod1 <- rma.mv(abs(yi), vi, data=dat_MA_ES, 
-                 method="REML", random = ~1 | response_id/ experiment_id/ study_id) 
-summary(mod1)
-
-#fit dirrectional model
-mod2<-rma.mv(yi, vi, data=dat_MA_ES, 
-             method="REML", random = ~1 | response_id/ experiment_id/ study_id) 
-summary(mod2)
+  filter(resp_units != "C") %>%
+  filter(study_id != "piccau2017")
 ```
@@
 # Plots
+
+## Question: How does response vary with study covariates?
+
+```{r}
+full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range +
+                        exp_age + size + org_level + mean_temp_constant, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+full_rf_model
+```
+
+## Relevant plots
+### Figure 1.
+```{r}
+# boxplots of how fluctuation range influences SMD
+ggplot(normalized, aes(x=flux_range, y=yi))+
+  geom_point(alpha = 0.5)+
+  theme_bw()+
+  geom_smooth(method="lm", formula = y~x)+
+  ggtitle("SMD across flux_range")
+```
+
+### Figure 2.
+```{r}
+# scatterplot of standardized mean response vs flux range colored and lm fit by org level
+ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
+  geom_point()+
+  geom_smooth(method="lm", formula = y~x)+
+  ggtitle("SMD across fluctuation ranges colored by organization level 
+          and fit with linear model")
+```
+
+### Figure 3. 
+```{r}
+#boxplot of SMD across levels of organization
+ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
+  geom_boxplot(alpha =0.7)+
+  geom_point(alpha = 0.3)+
+  scale_fill_tron()+
+  theme_bw()+
+  theme(legend.position = "bottom")+
+  ggtitle("SMD across organization level")
+```
+
+### Figure 4.
+```{r}
+#scatterplot of standardized mean difference across flux range colored by study fit with lm 
+ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
+  geom_point()+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_minimal()+
+  ggtitle("SMD across fluctuation ranges colored by 
+          studies and fit with linear model by study")
+```
+
 ## Question: How does response compare across studies and experiments?
-### Figure 1.
 ```{r, message=FALSE}
+#corresponding random effects model 
+fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig1
+#trying out mixed effects model 
+fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
+                 method="FE")
+fig1me
+```
+
+## Relevant Plots 
+### Figure 5.
+```{r}
 #SMD across all studies
@@
   ggtitle("SMD across all studies")
-#corresponding random effects model 
-fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig1
-#trying out mixed effects model 
-fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
-                 method="FE")
-fig1me
-#basic linear model 
-simple1<-lm(yi ~ study_id, data = dat_MA_ES)
-summary(fig1)
 ```
 
-### Figure 2. 
+### Figure 6. 
 ```{r, echo= FALSE}
@@
           colored by study")
-
-#basic linear model looking at how vi interacts with yi
-simple2<-lm(yi ~ vi, data = dat_MA_ES)
-summary(simple2)
 ```
 
-## Question: How does fluctuation amplitude affect response variables?
-
-### Figure 3.
-```{r, message=FALSE}
-# boxplots of how fluctuation range influences SMD
-ggplot(normalized, aes(x=as.factor(flux_range), y=yi))+
-  geom_boxplot()+
-  theme_bw()+
-  ggtitle("SMD across flux_range")
-
-#random effects model including org_level as a random variable
-fig7 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig7 
-#simple linear model looking at how flux_range affects yi
-simple7<-lm(yi~flux_range, data =dat_MA_ES)
-summary(simple7)
-```
-
-### Figure 4.
-```{r, message=FALSE}
-# scatterplot of standardized mean response vs flux range colored and lm fit by org level
-ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
-  geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  ggtitle("SMD across fluctuation ranges colored by organization level 
-          and fit with linear model")
-
-#random effects model including org_level as a random variable
-fig8 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + org_level,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig8
-#simple linear model looking at how flux_range and org_level affect yi
-simple8<-lm(yi~flux_range*org_level, data =dat_MA_ES)
-summary(simple8)
-```
-
-### Figure 5.
-```{r, message=FALSE}
-#scatterplot of standardized mean difference across flux range colored by study fit with lm 
-ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
-  geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  theme_minimal()+
-  ggtitle("SMD across fluctuation ranges colored by 
-          studies and fit with linear model by study")
-
-#random effects model including org_level as a random variable
-fig9 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + study_id,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig9
-#simple linear model looking at how flux_range and study_id affect yi
-simple9<-lm(yi~flux_range*study_id, data =dat_MA_ES)
-summary(simple9)
-```
-
-### Figure 6.
-```{r, message=FALSE}
+### Figure 7.
+```{r}
 #looking at yi across studies with the same temperature range (10 C)
@@
                                           temperature fluctuation range (10 C)")
-#random effects model within subset data
-fig12 <- rma.mv(yi, vi, data=common_range,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig12
-#simple linear model looking at how flux_range affects yi within subset data
-simple12<-lm(yi~flux_range, data =common_range)
-summary(simple12)
-```
-
-## Question how does mean temperature affect response variables?
-
-### Figure 7.
-```{r, message=FALSE}
-# boxplots of how mean temperature influences SMD
-ggplot(normalized, aes(x=as.factor(mean_temp_constant), y=yi))+
-  geom_boxplot()+
-  theme_bw()+
-  ggtitle("SMD across mean temperature")
-
-#random effects model including org_level as a random variable
-fig6 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~mean_temp_constant,
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig6
-#simple linear model looking at how mean_temp_constant affects yi
-simple6<-lm(yi~mean_temp_constant, data =dat_MA_ES)
-summary(simple6)
 ```
@@
 ### Figure 8.
-```{r, message=FALSE}
+```{r}
+# scatterplot of how mean temperature influences SMD
+ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
+  geom_point()+
+  theme_bw()+
+  ggtitle("SMD across mean temperature")
+```
+
+_________________________________
+Supplementary Plots/Code
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+
 #looking at yi across studies with the same mean temperature (24 C)
 ggplot(common_temp, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
+  geom_point()+
   geom_boxplot()+
@@
 
-## Question: How does genus affect response?
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#boxplots of SMD across different genuses colored by body size
+ggplot(normalized, aes(y=yi, x= reorder(genus, -yi), color = as.factor(size)))+
+  geom_point()+
+  theme_minimal()+
+  ggtitle("SMD across genuses colored by body size")+
+  theme(axis.text.x = element_text(face = "bold", 
+                                   size = 10, angle = 45,hjust = 1), 
+        legend.position = "bottom")
+#random effects model including genus as a random variable
+fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~size,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig11
+#simple linear model looking at how genus affects yi
+simple11<-lm(yi~genus, data =dat_MA_ES)
+summary(simple11)
+```
 
-### Figure 9. 
-```{r, message=FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
 #trying to look at drosophila response across studies 
@@
   geom_boxplot()+
+  geom_point()+
   scale_color_tron()+
@@
 
-### Figure 10.
-```{r, message=FALSE}
-#boxplots of SMD across different genuses colored by body size
-ggplot(normalized, aes(y=yi, x= reorder(genus, -yi), color = as.factor(size)))+
-  geom_boxplot()+
-  theme_minimal()+
-  ggtitle("SMD across genuses colored by body size")+
-  theme(axis.text.x = element_text(face = "bold", 
-                                   size = 10, angle = 45,hjust = 1), 
-        legend.position = "bottom")
-#random effects model including genus as a random variable
-fig11 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~genus,
+```{r, echo= FALSE, results = 'hide', message=FALSE}
+#random effects model including org_level as a random variable
+fig6 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~mean_temp_constant,
                random = ~1 | experiment_id/ study_id,
                  method="REML") 
-fig11
-#simple linear model looking at how genus affects yi
-simple11<-lm(yi~genus, data =dat_MA_ES)
-summary(simple11)
+fig6
+#simple linear model looking at how mean_temp_constant affects yi
+simple6<-lm(yi~mean_temp_constant, data =dat_MA_ES)
+summary(simple6)
 ```
 
-## Question how does scale (life or org level) affect response?
-### Figure 11.
+```{r, echo= FALSE, results = 'hide', message=FALSE}
+#basic linear model for study_id
+simple1<-lm(yi ~ study_id, data = dat_MA_ES)
+summary(fig1)
+#basic linear model looking at how vi interacts with yi
+simple2<-lm(yi ~ vi, data = dat_MA_ES)
+summary(simple2)
+```
 
-```{r, message=FALSE}
-#boxplot
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#boxplot of experimental ages and how their yi's stack up 
 ggplot(normalized, aes(x=exp_age, y=yi, color = as.factor(exp_age)))+
@@
 
-### Figure 12. 
-```{r, message=FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#boxplot of different experimental ages and how they vi's stack up 
 ggplot(normalized, aes(x=exp_age, y=vi, color = as.factor(exp_age)))+
@@
 
-### Figure 13. 
-```{r, message=FALSE}
-#boxplot of SMD across levels of organization
-ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
-  geom_boxplot(alpha =0.7)+
-  scale_fill_tron()+
-  theme_bw()+
-  theme(legend.position = "bottom")+
-  ggtitle("SMD across organization level")
-
+```{r, echo= FALSE, results = 'hide', message=FALSE}
 #random effects model including org_level as a random variable
@@
 
-## Question: How does response variable affect response?
-### Figure 14.
-```{r}
+```{r, echo= FALSE, results = 'hide', message=FALSE}
 #full random effects model looking at all the responses in the entire dataset
@@
 
-```{r, message=FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE}
+#random effects model within subset data
+fig12 <- rma.mv(yi, vi, data=common_range,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig12
+#simple linear model looking at how flux_range affects yi within subset data
+simple12<-lm(yi~flux_range, data =common_range)
+summary(simple12)
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE}
+#random effects model including org_level as a random variable
+fig9 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + study_id,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig9
+#simple linear model looking at how flux_range and study_id affect yi
+simple9<-lm(yi~flux_range*study_id, data =dat_MA_ES)
+summary(simple9)
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE}
+#random effects model including org_level as a random variable
+fig8 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + org_level,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig8
+#simple linear model looking at how flux_range and org_level affect yi
+simple8<-lm(yi~flux_range*org_level, data =dat_MA_ES)
+summary(simple8)
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE}
+#random effects model including org_level as a random variable
+fig7 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range,
+               random = ~1 | experiment_id/ study_id,
+                 method="REML") 
+fig7 
+#simple linear model looking at how flux_range affects yi
+simple7<-lm(yi~flux_range, data =dat_MA_ES)
+summary(simple7)
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+rf_model<-rma.mv(abs(yi), vi, data=dat_MA_ES, mods = ~flux_range +
+                        exp_age + size + org_level + mean_temp_constant, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML")
+rf_model
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
 #plot across studies with common response variable body mass

### 2021-03-08 shared analysis files /data_wrangling_plot22feb.Rmd
adding in hypothesis and initial summary conclusions for the two questions

@@
 
+Hypothesis: SMD between constant and fluctuating environments is most affected by temperature parameters (range and mean) but is also likely affected by demographic parameters (age, size, organization level)
+
+Initial conclusions: Flux_range and mean temperature appear to be the most important contributers to variation in yi, though organization also contributes.
+
 ```{r}
@@
   geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
   ggtitle("SMD across fluctuation ranges colored by organization level 
@@
   geom_smooth(method="lm", formula = y~x)+
-  theme_minimal()+
+  theme_bw()+
   ggtitle("SMD across fluctuation ranges colored by 
@@
 ## Question: How does response compare across studies and experiments?
+Hypothesis: There should be differences across studies because of differences in experimental designs and study organism that would mean different magnitudes of response.
+
+Initial conclusions: The studies overall are different in their responses but not owing to study_id or experiment.
+However, the mixed effects model does suggest some differences when you include study _id as a mod.
+
 ```{r, message=FALSE}
@@
   geom_point()+
-  theme_bw()+
+  theme_minimal()+
   ggtitle("SMD across mean temperature")

### 2021-03-15 shared analysis files /data_wrangling_plot8march.Rmd
updates to the rmd to remove unnecessary EDA and updating metafor file

--- 2021-03-08 shared analysis files /data_wrangling_plot22feb.Rmd
@@
 dat_MA_ES<-dat_MA_ES %>% 
-  filter (!is.na(yi)) %>%
-  filter(resp_units != "C") %>%
-  filter(study_id != "piccau2017")
+  filter (!is.na(yi)) 
 ```
@@
   filter(yi > -20) 
-#filtering for studies with common variable g 
-common_unit_g <-dat_MA_ES %>% 
-filter(resp_def == "body mass")
-#filtering for genus to only be drosophila 
-drosophila <- dat_MA_ES %>% 
-  filter(genus == "Drosophila")
 #filtering for just flux ranges of 10 C 

### 2021-03-15 shared analysis files /data_wrangling_plot8march.Rmd
updated with additional columns and models

@@
 ## loading libraries 
-pacman::p_load(metafor, tidyverse, viridis, ggsci, wesanderson)
+pacman::p_load(metafor, tidyverse, viridis, ggsci, plotly)
 #devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
@@
 #load data
-dat_MA_0<- read_csv("metafor.csv")
+dat_MA_0<- read_csv("metafor.csv") %>%
+  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))
+
+glimpse(dat_MA_0)
+
+count(dat_MA_0, study_id, genus, species, mean_temp_constant, min_temp, max_temp)
 ```
@@
   filter (!is.na(yi))
+
+#removing outlier point to get better picture of data structure AND excluding CT assays 
+normalized <- dat_MA_ES %>%
+  filter(yi > -20) 
+#filtering for just flux ranges of 10 C 
+common_range <- dat_MA_ES %>% 
+  filter(flux_range == "10")
+#filtering for just studies with mean temp of 24
+common_temp <- dat_MA_ES %>% 
+  filter(mean_temp_constant == "24")
 ```
@@
 count(dat_MA_ES, larger_group) # pretty much mostly animal studies
-
-#removing outlier point to get better picture of data structure AND excluding CT assays 
-normalized <- dat_MA_ES %>%
-  filter(yi > -20) 
-#filtering for just flux ranges of 10 C 
-common_range <- dat_MA_ES %>% 
-  filter(flux_range == "10")
-#filtering for just studies with mean temp of 24
-common_temp <- dat_MA_ES %>% 
-  filter(mean_temp_constant == "24")
 ```
@@
 ```{r}
-full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range +
-                        exp_age + size + org_level + mean_temp_constant, 
+#with interaction term 
+full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
+                        exp_age + size + org_level, 
                random = ~1 | experiment_id/ study_id/ response_id,
@@
 full_rf_model
+
+#without interaction term
+full_rf_model2<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + mean_temp_constant +
+                        exp_age + size + org_level, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+full_rf_model2
+#full model split out by stocks/rates 
+full_rf_stocks<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + mean_temp_constant +
+                        exp_age + size + org_level + resp_type, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+full_rf_stocks
+
+#full model split out by life history outcomes
+full_rf_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + mean_temp_constant +
+                        exp_age + size + org_level + stressful , 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+full_rf_thermal
 ```
@@
 # scatterplot of standardized mean response faceted by exp age
-ggplot(normalized, aes(x=vi, y=yi, color = study_id))+
-  geom_point()+
+ggplot(normalized, aes(x=yi, y=vi, color = study_id))+
+  geom_point(alpha = 0.5)+
   theme_bw()+
@@
   theme_minimal()+
+  geom_smooth(formula = y~x, method = "lm")+
   ggtitle("SMD across mean temperature")
+```
+
+```{r}
+#contour plot with flux_range on the x axis, mean temp of the y axis and colored by yi
+fig <- plot_ly(
+  x = dat_MA_ES$flux_range, 
+  y = dat_MA_ES$mean_temp_constant, 
+  z = dat_MA_ES$yi, 
+  type = "contour" 
+)
+
+#funnel plot 
+### set up 2x2 array for plotting
+par(mfrow=c(2,2))
+
+funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
+                        exp_age + size + org_level, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+### draw funnel plots
+#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
+funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+
+?funnel
+
 ```

### 2021-03-16 shared analysis files /data_wrangling_plot8march.Rmd
changing the contour plot to excluded big negative yi

@@
 fig <- plot_ly(
-  x = dat_MA_ES$flux_range, 
-  y = dat_MA_ES$mean_temp_constant, 
-  z = dat_MA_ES$yi, 
-  type = "contour" 
-)
+  x = normalized$flux_range, 
+  y = normalized$mean_temp_constant, 
+  z = normalized$yi, 
+  type = "contour")
+fig
 

### 2021-03-21 thesis draft/thesis draft.Rmd
trying to use thesisdown very inefficiently

---
title: 'My Final College Paper'
author: 'Margaret Anne Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
# If you have more two advisors, un-silence line 7
#altadvisor: 'Your Other Advisor'
department: 'Biology'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True

# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
  thesisdown::thesis_pdf: default 
#  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  I want to thank a few people.
dedication: |
  You can have a dedication here if you wish. 
preface: |
  This is an example of a thesis setup to use the reed thesis document class 
  (for LaTeX) and the R bookdown package, in general.

# Specify the location of the bibliography below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line).

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

Welcome to the _R Markdown_ thesis template. This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before.  Using _R Markdown_ will also allow you to easily keep track of your analyses in **R** chunks of code, with the resulting plots and output included as well.  The hope is this _R Markdown_ template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.

Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis.  The use of LaTeX in combination with _Markdown_ is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file. While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities.  After working with _Markdown_ and **R** together for a few weeks, we are confident this will be your reporting style of choice going forward.

<!-- 
If you're still on the fence about using _R Markdown_, check out the resource
for newbies available at <https://rbasics.netlify.com>. 
-->

**Why use it?**

_R Markdown_ creates a simple and straightforward way to interface with the beauty of LaTeX.  Packages have been written in **R** to work directly with LaTeX to produce nicely formatting tables and paragraphs. In addition to creating a user friendly interface to LaTeX, _R Markdown_ also allows you to read in your data, to analyze it and to visualize it using **R** functions, and also to provide the documentation and commentary on the results of your project.  Further, it allows for **R** results to be passed inline to the commentary of your results.  You'll see more on this later.

<!-- 
Having your code and commentary all together in one place has a plethora of 
benefits!
-->

**Who should use it?**

Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use _R Markdown_. Of particular use should be anyone in the sciences, but the user-friendly nature of _Markdown_ and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc. should make it of great benefit to nearly anyone writing a thesis project.

**For additional help with bookdown**

Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).

### 2021-03-21 thesis draft/thesis.Rmd
trying to make an r file for this thesis draft?

<!--
This is for including Chapter 1.  Notice that it's also good practice to name your chunk.  This will help you debug potential issues as you knit.  The chunk above is called intro and the one below is called chapter1.  Feel free to change the name of the Rmd file as you wish, but don't forget to change it here from chap1.Rmd.
-->

<!--
The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of Chapter 1.  These labels will automatically be generated (if not specified) by changing the spaces to hyphens and capital letters to lowercase.  Look for the reference to this label at the beginning of Chapter 2.
-->

# R Markdown Basics {#rmd-basics}

Here is a brief introduction into using _R Markdown_. _Markdown_ is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. _R Markdown_ provides the flexibility of _Markdown_ with the implementation of **R** input and output.  For more details on using _R Markdown_ see <https://rmarkdown.rstudio.com>.

Be careful with your spacing in _Markdown_ documents.  While whitespace largely is ignored, it does at times give _Markdown_ signals as to how to proceed.  As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph.  In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).

## Lists

It's easy to create a list.  It can be unordered like

* Item 1
* Item 2

or it can be ordered like

1. Item 1
4. Item 2

Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a tab).  (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
    - Item 3a
    - Item 3b

## Line breaks

Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
This should be a new paragraph.

*Now for the correct way:*

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.

This should be a new paragraph.

## R chunks

When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document. You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):

```{r cars}
summary(cars)
```

## Inline code

If you'd like to put the results of your analysis directly into your discussion, add inline code like this:

> The `cos` of $2 \pi$ is `r cos(2*pi)`.

Another example would be the direct calculation of the standard deviation:

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.

One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be indented.

As you see with `$2 \pi$` above, mathematics can be added by surrounding the mathematical text with dollar signs.  More examples of this are in [Mathematics and Science] if you uncomment the code in [Math].

## Including plots

You can also embed plots. For example, here is a way to use the base **R** graphics package to produce a plot using the built-in `pressure` dataset:

```{r pressure, echo=FALSE, cache=TRUE, fig.height=3, fig.width=5}
plot(pressure)
```

Note that the `echo=FALSE` parameter was added to the code chunk to prevent printing of the **R** code that generated the plot. There are plenty of other ways to add chunk options (like `fig.height` and `fig.width` in the chunk above).  More information is available at <https://yihui.org/knitr/options/>.

Another useful chunk option is the setting of `cache=TRUE` as you see here.  If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.  Later in this file, you'll see a way to reference plots created in **R** or external figures.

## Loading and exploring data

Included in this template is a file called `flights.csv`.  This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014. More information about this dataset and its **R** package is available at <https://github.com/ismayc/pnwflights14>. This subset includes only Portland flights and only rows that were complete with no missing values. Merges were also done with the `airports` and `airlines` data sets in the `pnwflights14` package to get more descriptive airport and airline names.

We can load in this data set using the following commands:

```{r load_data}
# flights.csv is in the data directory
flights_path <- here::here("data", "flights.csv")
# string columns will be read in as strings and not factors now
flights <- read.csv(flights_path, stringsAsFactors = FALSE)
```

The data is now stored in the data frame called `flights` in **R**.  To get a better feel for the variables included in this dataset we can use a variety of functions. Here we can see the dimensions (rows by columns) and also the names of the columns.

```{r str}
dim(flights)
names(flights)
```

Another good idea is to take a look at the dataset in table form.  With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here. I recommend you enter the command into the Console **_after_** you have run the **R** chunks above to load the data into **R**.

```{r view_flights, eval=FALSE}
View(flights)
```

While not required, it is highly recommended you use the `dplyr` package to manipulate and summarize your data set as needed.  It uses a syntax that is easy to understand using chaining operations.  Below I've created a few examples of using `dplyr` to get information about the Portland flights in 2014.  You will also see the use of the `ggplot2` package, which produces beautiful, high-quality academic visuals.

We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:

```{r load_pkgs, message=FALSE}
# List of packages required for this analysis
pkg <- c("dplyr", "ggplot2", "knitr", "bookdown")
# Check if packages are not installed and assign the
# names of the packages not installed to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]
# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg)) {
  install.packages(new.pkg, repos = "https://cran.rstudio.com")
}
# Load packages
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
```

\clearpage

The example we show here does the following:

- Selects only the `carrier_name` and `arr_delay` from the `flights` dataset and then assigns this subset to a new variable called `flights2`.

- Using `flights2`, we determine the largest arrival delay for each of the carriers.

```{r max_delays}
flights2 <- flights %>%
  select(carrier_name, arr_delay)
max_delays <- flights2 %>%
  group_by(carrier_name) %>%
  summarize(max_arr_delay = max(arr_delay, na.rm = TRUE))
```

A useful function in the `knitr` package for making nice tables in _R Markdown_ is called `kable`.  It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.  This again goes to show how nice reproducible documents can be! (Note the use of `results="asis"`, which will produce the table instead of the code to create the table.)  The `caption.short` argument is used to include a shorter title to appear in the List of Tables.

```{r maxdelays, results="asis"}
kable(max_delays,
  col.names = c("Airline", "Max Arrival Delay"),
  caption = "Maximum Delays by Airline",
  caption.short = "Max Delays by Airline",
  longtable = TRUE,
  booktabs = TRUE
)
```

The last two options make the table a little easier-to-read.

We can further look into the properties of the largest value here for American Airlines Inc.  To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original `flights` dataset.

```{r max_props}
flights %>%
  filter(
    arr_delay == 1539,
    carrier_name == "American Airlines Inc."
  ) %>%
  select(-c(
    month, day, carrier, dest_name, hour,
    minute, carrier_name, arr_delay
  ))
```

We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth.  Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.

```{r march3plot, fig.height=3, fig.width=6}
flights %>%
  filter(month == 3, day == 3) %>%
  ggplot(aes(x = dep_time, y = arr_delay)) +
  geom_point()
```

## Additional resources

- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>

- _R Markdown_
    - Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf>

- _RStudio IDE_
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf>
    - Official website - <https://rstudio.com/products/rstudio/>

- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html>

- `ggplot2`
    - Documentation - <https://ggplot2.tidyverse.org/>
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf>

<!--chapter:end:01-chap1.Rmd-->

# Mathematics and Science {#math-sci}

<!-- Required to number equations in HTML files -->
<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { equationNumbers: { autoNumber: "AMS" } }
});
</script>

## Math

\TeX\ is the best way to typeset mathematics. Donald Knuth designed \TeX\ when he got frustrated at how long it was taking the typesetters to finish his book, which contained a lot of mathematics.  One nice feature of _R Markdown_ is its ability to read LaTeX code directly.

If you are doing a thesis that will involve lots of math, you will want to read the following section which has been commented out. If you're not going to use math, skip over or delete this next commented section.

<!-- MATH and PHYSICS majors: Uncomment the following section -->
<!--
$$\sum_{j=1}^n (\delta\theta_j)^2 \leq {{\beta_i^2}\over{\delta_i^2 + \rho_i^2}}
\left[ 2\rho_i^2 + {\delta_i^2\beta_i^2\over{\delta_i^2 + \rho_i^2}} \right] \equiv \omega_i^2
$$

From Informational Dynamics, we have the following (Dave Braden):

After _n_ such encounters the posterior density for $\theta$ is

$$
\pi(\theta|X_1< y_1,\dots,X_n<y_n) \varpropto \pi(\theta) \prod_{i=1}^n\int_{-\infty}^{y_i}
   \exp\left(-{(x-\theta)^2\over{2\sigma^2}}\right)\ dx
$$

Another equation:

$$\det\left|\,\begin{matrix}%
c_0&c_1\hfill&c_2\hfill&\ldots&c_n\hfill\cr
c_1&c_2\hfill&c_3\hfill&\ldots&c_{n+1}\hfill\cr
c_2&c_3\hfill&c_4\hfill&\ldots&c_{n+2}\hfill\cr
\,\vdots\hfill&\,\vdots\hfill&
  \,\vdots\hfill&&\,\vdots\hfill\cr
c_n&c_{n+1}\hfill&c_{n+2}\hfill&\ldots&c_{2n}\hfill\cr
\end{matrix}\right|>0$$

Lapidus and Pindar, Numerical Solution of Partial Differential Equations in Science and
Engineering.  Page 54

$$
\int_t\left\{\sum_{j=1}^3 T_j \left({d\phi_j\over dt}+k\phi_j\right)-kT_e\right\}w_i(t)\ dt=0,
   \qquad\quad i=1,2,3.
$$

L\&P  Galerkin method weighting functions.  Page 55

$$
\sum_{j=1}^3 T_j\int_0^1\left\{{d\phi_j\over dt} + k\phi_j\right\} \phi_i\ dt
   = \int_{0}^1k\,T_e\phi_idt, \qquad i=1,2,3 $$

Another L\&P (p145)

$$
\int_{-1}^1\!\int_{-1}^1\!\int_{-1}^1 f\big(\xi,\eta,\zeta\big)
   = \sum_{k=1}^n\sum_{j=1}^n\sum_{i=1}^n w_i w_j w_k f\big( \xi,\eta,\zeta\big).
$$

Another L\&P (p126)

$$
\int_{A_e} (\,\cdot\,) dx dy = \int_{-1}^1\!\int_{-1}^1 (\,\cdot\,) \det[J] d\xi d\eta.
$$
-->

## Chemistry 101: Symbols

Chemical formulas will look best if they are not italicized. Get around math mode's automatic italicizing in LaTeX by using the argument `$\mathrm{formula here}$`, with your formula inside the curly brackets.  (Notice the use of the backticks here which enclose text that acts as code.)

So, $\mathrm{Fe_2^{2+}Cr_2O_4}$ is written `$\mathrm{Fe_2^{2+}Cr_2O_4}$`.

<!--
The \noindent command below does what you'd expect:  it forces the current line/paragraph to not indent. This was done here to match the format of the LaTeX thesis PDF.
-->

\noindent Exponent or Superscript: $\mathrm{O^-}$

\noindent Subscript: $\mathrm{CH_4}$

To stack numbers or letters as in $\mathrm{Fe_2^{2+}}$, the subscript is defined first, and then the superscript is defined.

\noindent Bullet: CuCl $\bullet$ $\mathrm{7H_{2}O}$

\noindent Delta: $\Delta$

\noindent Reaction Arrows: $\longrightarrow$ or  $\xrightarrow{solution}$

\noindent Resonance Arrows: $\leftrightarrow$

\noindent Reversible Reaction Arrows: $\rightleftharpoons$

### Typesetting reactions

You may wish to put your reaction in an equation environment, which means that LaTeX will place the reaction where it fits and will number the equations for you.

\begin{equation}
  \mathrm{C_6H_{12}O_6  + 6O_2} \longrightarrow \mathrm{6CO_2 + 6H_2O}
  (\#eq:reaction)
\end{equation}

We can reference this combustion of glucose reaction via Equation \@ref(eq:reaction).

### Other examples of reactions

$\mathrm{NH_4Cl_{(s)}}$ $\rightleftharpoons$ $\mathrm{NH_{3(g)}+HCl_{(g)}}$

\noindent $\mathrm{MeCH_2Br + Mg}$ $\xrightarrow[below]{above}$ $\mathrm{MeCH_2\bullet Mg \bullet Br}$

## Physics

Many of the symbols you will need can be found on the math page <https://web.reed.edu/cis/help/latex/math.html> and the Comprehensive LaTeX Symbol Guide (<https://mirror.utexas.edu/ctan/info/symbols/comprehensive/symbols-letter.pdf>).

## Biology

You will probably find the resources at <https://www.lecb.ncifcrf.gov/~toms/latex.html> helpful, particularly the links to bsts for various journals. You may also be interested in TeXShade for nucleotide typesetting (<https://homepages.uni-tuebingen.de/beitz/txe.html>).  Be sure to read the proceeding chapter on graphics and tables.

<!--chapter:end:02-chap2.Rmd-->

```{r include_packages_2, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis and also two functions
# used for labeling and referencing
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("remotes")',
        "first in the Console."
      )
    )
  }
}
if (!require(dplyr)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("dplyr", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("dplyr")',
        "first in the Console."
      )
    )
  }
}
if (!require(ggplot2)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("ggplot2", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("ggplot2")',
        "first in the Console."
      )
    )
  }
}
if (!require(bookdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("bookdown", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("bookdown")',
        "first in the Console."
      )
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
flights <- read.csv("data/flights.csv", stringsAsFactors = FALSE)
```

# Graphics, References, and Labels {#ref-labels}

## Figures

If your thesis has a lot of figures, _R Markdown_ might behave better for you than that other word processor.  One perk is that it will automatically number the figures accordingly in each chapter.    You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier.  If you label your figures, you can move the figures around and _R Markdown_ will automatically adjust the numbering for you.  No need for you to remember!  So that you don't have to get too far into LaTeX to do this, a couple **R** functions have been created for you to assist.  You'll see their use below.

<!--
One thing that may be annoying is the way _R Markdown_ handles "floats" like tables and figures (it's really \LaTeX's fault). \LaTeX\ will try to find the best place to put your object based on the text around it and until you're really, truly done writing you should just leave it where it lies. There are some optional arguments specified in the options parameter of the `label` function.  If you need to shift your figure around, it might be good to look here on tweaking the options argument:  <https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions>

If you need a graphic or tabular material to be part of the text, you can just put it inline. If you need it to appear in the list of figures or tables, it should be placed in a code chunk.
-->

In the **R** chunk below, we will load in a picture stored as `reed.jpg` in our main directory.  We then give it the caption of "Reed logo", the label of "reedlogo", and specify that this is a figure.  Make note of the different **R** chunk options that are given in the R Markdown file (not shown in the knitted document).

```{r reedlogo, fig.cap="Reed logo", out.width="0.2\\linewidth", fig.align="center"}
include_graphics(path = "figure/reed.jpg")
```

<!-- Note the use of `out.width` as a chunk option here. The resulting
image is 20% of what the linewidth is in LaTeX. You can also center
the image using `fig.align="center"` as shown.-->

Here is a reference to the Reed logo: Figure \@ref(fig:reedlogo).  Note the use of the `fig:` code here.  By naming the **R** chunk that contains the figure, we can then reference that figure later as done in the first sentence here.  We can also specify the caption for the figure via the R chunk option `fig.cap`.

\clearpage

<!-- clearpage ends the page, and also dumps out all floats.
  Floats are things like tables and figures. -->

Below we will investigate how to save the output of an **R** plot and label it in a way similar to that done above.  Recall the `flights` dataset from Chapter \@ref(rmd-basics).  (Note that we've shown a different way to reference a section or chapter here.)  We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.

```{r delaysboxplot, warnings=FALSE, messages=FALSE, fig.cap="Mean Delays by Airline", fig.width=6, fig.height=5}
mean_delay_by_carrier <- flights %>%
  group_by(carrier) %>%
  summarize(mean_dep_delay = mean(dep_delay))
ggplot(mean_delay_by_carrier, aes(x = carrier, y = mean_dep_delay)) +
  geom_bar(position = "identity", stat = "identity", fill = "red")
```

Here is a reference to this image: Figure \@ref(fig:delaysboxplot).

A table linking these carrier codes to airline names is available at <https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv>.

\clearpage

Next, we will explore the use of the `out.extra` chunk option, which can be used to shrink or expand an image loaded from a file by specifying `"scale= "`. Here we use the mathematical graph stored in the "subdivision.pdf" file.

```{r subd, results="asis", echo=FALSE, fig.cap="Subdiv. graph", out.extra="scale=0.75"}
include_graphics("figure/subdivision.pdf")
```

Here is a reference to this image: Figure \@ref(fig:subd).  Note that `echo=FALSE` is specified so that the **R** code is hidden in the document.

**More Figure Stuff**

Lastly, we will explore how to rotate and enlarge figures using the `out.extra` chunk option.  (Currently this only works in the PDF version of the book.)

```{r subd2, results="asis", echo=FALSE, out.extra="angle=180, scale=1.1", fig.cap="A Larger Figure, Flipped Upside Down"}
include_graphics("figure/subdivision.pdf")
```

As another example, here is a reference: Figure \@ref(fig:subd2).

## Footnotes and Endnotes

You might want to footnote something. ^[footnote text] The footnote will be in a smaller font and placed appropriately. Endnotes work in much the same way. More information can be found about both on the CUS site or feel free to reach out to <data@reed.edu>.

## Bibliographies

Of course you will need to cite things, and you will probably accumulate an armful of sources. There are a variety of tools available for creating a bibliography database (stored with the .bib extension).  In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero.  The Reed librarians have created Zotero documentation at <https://libguides.reed.edu/citation/zotero>.  In addition, a tutorial is available from Middlebury College at <https://sites.middlebury.edu/zoteromiddlebury/>.

_R Markdown_ uses _pandoc_ (<https://pandoc.org/>) to build its bibliographies.  One nice caveat of this is that you won't have to do a second compile to load in references as standard LaTeX requires. To cite references in your thesis (after creating your bibliography database), place the reference name inside square brackets and precede it by the "at" symbol.  For example, here's a reference to a book about worrying:  [@Molina1994].  This `Molina1994` entry appears in a file called `thesis.bib` in the `bib` folder.  This bibliography database file was created by a program called BibTeX.  You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the `bib` folder.

For more information about BibTeX and bibliographies, see our CUS site (<https://web.reed.edu/cis/help/latex/index.html>)^[@reedweb2007]. There are three pages on this topic:  _bibtex_ (which talks about using BibTeX, at <https://web.reed.edu/cis/help/latex/bibtex.html>), _bibtexstyles_ (about how to find and use the bibliography style that best suits your needs, at <https://web.reed.edu/cis/help/latex/bibtexstyles.html>) and _bibman_ (which covers how to make and maintain a bibliography by hand, without BibTeX, at <https://web.reed.edu/cis/help/latex/bibman.html>). The last page will not be useful unless you have only a few sources.

If you look at the YAML header at the top of the main .Rmd file you can see that we can specify the style of the bibliography by referencing the appropriate csl file.  You can download a variety of different style files at <https://www.zotero.org/styles>.  Make sure to download the file into the csl folder.

<!-- Fill the rest of the page with the content below for the PDF version. -->

\vfill

**Tips for Bibliographies**

- Like with thesis formatting, the sooner you start compiling your bibliography for something as large as thesis, the better. Typing in source after source is mind-numbing enough; do you really want to do it for hours on end in late April? Think of it as procrastination.
- The cite key (a citation's label) needs to be unique from the other entries.
- When you have more than one author or editor, you need to separate each author's name by the word "and" e.g. `Author = {Noble, Sam and Youngberg, Jessica},`.
- Bibliographies made using BibTeX (whether manually or using a manager) accept LaTeX markup, so you can italicize and add symbols as necessary.
- To force capitalization in an article title or where all lowercase is generally used, bracket the capital letter in curly braces.
- You can add a Reed Thesis citation^[@noble2002] option. The best way to do this is to use the phdthesis type of citation, and use the optional "type" field to enter "Reed thesis" or "Undergraduate thesis."

## Anything else?

If you'd like to see examples of other things in this template, please contact the Data @ Reed team (email <data@reed.edu>) with your suggestions. We love to see people using _R Markdown_ for their theses, and are happy to help.

<!--chapter:end:03-chap3.Rmd-->

# Conclusion {-}

If we don't want Conclusion to have a chapter number next to it, we can add the `{-}` attribute.

**More info**

And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph. Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.

<!--chapter:end:04-conclusion.Rmd-->

`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

# The First Appendix

<!--
If you feel it necessary to include an appendix, it goes here.
-->

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.

**In the main Rmd file**

```{r ref.label='include_packages', results='hide', echo = TRUE}
```

**In Chapter \@ref(ref-labels):**

```{r ref.label='include_packages_2', results='hide', echo = TRUE}
```

# The Second Appendix, for Fun

<!--chapter:end:05-appendix.Rmd-->

<!--
The bib chunk below must go last in this document according to how R Markdown renders.  More info is at http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
-->

\backmatter

<!-- 
If you'd like to change the name of the bibliography to something else,
delete "References" and replace it.
-->

# References {-}
<!--
This manually sets the header for this unnumbered chapter.
-->
\markboth{References}{References}
<!--
To remove the indentation of the first entry.
-->
\noindent

<!--
To create a hanging indent and spacing between entries.  These three lines may need to be removed for styles that don't require the hanging indent.
-->

\setlength{\parindent}{-0.20in}
\setlength{\leftskip}{0.20in}
\setlength{\parskip}{8pt}

<!--
This is just for testing with more citations for the bibliography at the end.  Add other entries into the list here if you'd like them to appear in the bibliography even if they weren't explicitly cited in the document.
-->

---
nocite: | 
  @angel2000, @angel2001, @angel2002a
...

<!--chapter:end:99-references.Rmd-->

---
title: 'My Final College Paper'
author: 'Margaret Anne Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
# If you have more two advisors, un-silence line 7
#altadvisor: 'Your Other Advisor'
department: 'Biology'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True

# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
  thesisdown::thesis_pdf: default 
#  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  I want to thank a few people.
dedication: |
  You can have a dedication here if you wish. 
preface: |
  This is an example of a thesis setup to use the reed thesis document class 
  (for LaTeX) and the R bookdown package, in general.

# Specify the location of the bibliography below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line).

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

Welcome to the _R Markdown_ thesis template. This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before.  Using _R Markdown_ will also allow you to easily keep track of your analyses in **R** chunks of code, with the resulting plots and output included as well.  The hope is this _R Markdown_ template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.

Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis.  The use of LaTeX in combination with _Markdown_ is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file. While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities.  After working with _Markdown_ and **R** together for a few weeks, we are confident this will be your reporting style of choice going forward.

<!-- 
If you're still on the fence about using _R Markdown_, check out the resource
for newbies available at <https://rbasics.netlify.com>. 
-->

**Why use it?**

_R Markdown_ creates a simple and straightforward way to interface with the beauty of LaTeX.  Packages have been written in **R** to work directly with LaTeX to produce nicely formatting tables and paragraphs. In addition to creating a user friendly interface to LaTeX, _R Markdown_ also allows you to read in your data, to analyze it and to visualize it using **R** functions, and also to provide the documentation and commentary on the results of your project.  Further, it allows for **R** results to be passed inline to the commentary of your results.  You'll see more on this later.

<!-- 
Having your code and commentary all together in one place has a plethora of 
benefits!
-->

**Who should use it?**

Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use _R Markdown_. Of particular use should be anyone in the sciences, but the user-friendly nature of _Markdown_ and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc. should make it of great benefit to nearly anyone writing a thesis project.

**For additional help with bookdown**

Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).

<!--chapter:end:thesis-draft.Rmd-->

### 2021-04-02 shared analysis files /data_wrangling_plot.Rmd
fixing up final plots to potentially include in results and discussion

--- 2021-03-15 shared analysis files /data_wrangling_plot8march.Rmd
@@
 author: "Maggie Slein"
-date: "3/8/21"
+date: "4/2/21"
 output: pdf_document
+
 ---
@@
   mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))
-
-glimpse(dat_MA_0)
-
-count(dat_MA_0, study_id, genus, species, mean_temp_constant, min_temp, max_temp)
 ```
@@
   filter (!is.na(yi)) 
+count(dat_MA_ES, study_id)
 
@@
 common_range <- dat_MA_ES %>% 
-  filter(flux_range == "10")
+  filter(flux_range %in% c(10))
 #filtering for just studies with mean temp of 24
@@
 ## Relevant plots
-### Figure 1.
-```{r}
+
+```{r, echo= FALSE}
 # boxplots of how fluctuation range influences SMD
@@
 ```
+**Figure 1.**
+Data visualization supports model findings that there is a differene between shorter fluctuation ranges and longer fluctuation ranges.
 
-### Figure 2.
-```{r}
+```{r, echo= FALSE}
 # scatterplot of standardized mean response vs flux range colored and lm fit by org level
@@
 ```
+**Figure 2.**
+We can also see that there a difference in effect size in organization level responses across fluctuation range.
+Population level reponses trended positively across fluctuation ranges, while organism level responses trended negatively across fluctuation ranges.
 
-### Figure 3. 
-```{r}
-#boxplot of SMD across levels of organization
-ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
-  geom_boxplot(alpha =0.7)+
-  geom_point(alpha = 0.3)+
-  scale_fill_tron()+
-  theme_bw()+
-  theme(legend.position = "bottom")+
-  ggtitle("SMD across organization level")
-```
-
-### Figure 4.
-```{r}
+```{r, echo= FALSE}
 #scatterplot of standardized mean difference across flux range colored by study fit with lm 
@@
 ```
+**Figure 3.**
+When broken out by study, we can see that there is disagreement in the effect of fluctuation range on effect size.
+
+```{r, echo= FALSE}
+# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
+normalized %>%
+  drop_na(stressful) %>%
+ggplot(aes(x=flux_range, y=yi, color = stressful))+
+  geom_point()+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  ggtitle("SMD across fluctuation ranges colored by 
+          whether or mean temperature was thermally stressful")
+```
+**Figure 4.** 
+When we consider where the temperatures in the fluctuation range reach thermally stressful levels, there is a small difference between responses to stressful  temperatures (negative) and non-stressful temperatures (positive).
+
+```{r, echo= FALSE}
+# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
+normalized %>%
+ggplot(aes(x=flux_range, y=yi, color = resp_type))+
+  geom_point()+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  ggtitle("SMD across fluctuation ranges colored by response type")
+```
+**Figure 5.**
 
@@
 
-```{r, message=FALSE}
+```{r, echo= FALSE, message=FALSE}
 #corresponding random effects model 
@@
 fig1
-#trying out mixed effects model 
-fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
-                 method="FE")
-fig1me
 ```
@@
 ## Relevant Plots
-### Figure 5.
-```{r}
+
+```{r, echo = FALSE}
 #SMD across all studies
@@
 
-### Figure 6. 
-```{r, echo= FALSE}
-# scatterplot of standardized mean response faceted by exp age
-ggplot(normalized, aes(x=yi, y=vi, color = study_id))+
-  geom_point(alpha = 0.5)+
-  theme_bw()+
-  ggtitle("SMD across variance 
-          colored by study")
+**Figure 6.**
+When we look at experiment and study, we can see across the 15 studies included in this meta-analysis there are differences across and within experiments.
+
+```{r, echo = FALSE}
+#forest plot that shows the same thing?
+forest <- normalized %>%
+  group_by(study_id) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  arrange(desc(mean_yi))
+forest(forest$mean_yi, forest$mean_vi, slab=forest$study_id)
 ```
 
-### Figure 7.
-```{r}
+**Figure 7.**
+I kind of like this figure better than the boxplot in Figure 5 as an overall figure?
+
+```{r, echo = FALSE}
 #looking at yi across studies with the same temperature range (10 C)
-ggplot(common_range, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
+#excluded wang because there were too few points to do a boxplot
+common_range %>%
+  filter(study_id != "wang2015") %>%
+ggplot(aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
   geom_boxplot()+
@@
 
-### Figure 8.
-```{r}
+**Figure 8.**
+
+```{r, echo = FALSE}
 # scatterplot of how mean temperature influences SMD
@@
 
-```{r}
-#contour plot with flux_range on the x axis, mean temp of the y axis and colored by yi
-fig <- plot_ly(
-  x = dat_MA_ES$flux_range, 
-  y = dat_MA_ES$mean_temp_constant, 
-  z = dat_MA_ES$yi, 
-  type = "contour" 
-)
+```{r, echo = FALSE}
 
@@
                  method="REML") 
+```
+
+**Figure 9.**
+
+```{r, echo = FALSE, fig.height=8, fig.width=12}
 ### draw funnel plots
@@
 funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+funnel(funnel_model, yaxis="vi", main="Sampling Variance")
+```
 
-?funnel
-
-```
+**Figure 10.** 
+Well, from what I can gather, a majority of the effect sizes from our meta-analysis have a non-signficant effect size value.
+However, there is a population of a couple influential effect sizes that are significantly important.
 
@@
 Supplementary Plots/Code
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#trying out mixed effects model 
+fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
+                 method="FE")
+fig1me
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#boxplot of SMD across levels of organization
+ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
+  geom_boxplot(alpha =0.7)+
+  geom_point(alpha = 0.3)+
+  scale_fill_tron()+
+  theme_bw()+
+  theme(legend.position = "bottom")+
+  ggtitle("SMD across organization level")
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+# scatterplot of standardized mean response faceted by exp age
+ggplot(normalized, aes(x=yi, y=vi, color = study_id))+
+  geom_point(alpha = 0.5)+
+  theme_bw()+
+  ggtitle("SMD across variance 
+          colored by study")
+```
 
@@
 
-```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+```{r, echo = FALSE, eval= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
 #trying to look at drosophila response across studies 
@@
 
-```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+```{r, echo = FALSE, eval= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
 #plot across studies with common response variable body mass

### 2021-04-06 shared analysis files /data_wrangling_plot.Rmd
contour plot craziness

@@
 ## loading libraries 
-pacman::p_load(metafor, tidyverse, viridis, ggsci, plotly)
+pacman::p_load(metafor, tidyverse, viridis, ggsci, plotly,  mgcv)
 #devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
@@
                         exp_age + size + org_level, 
-               random = ~1 | experiment_id/ study_id/ response_id,
+               random = ~1 |  study_id/ experiment_id/ response_id,
                  method="REML") 
@@
 However, there is a population of a couple influential effect sizes that are significantly important.
+
+```{r}
+fig <- plot_ly(
+  x = normalized$mean_temp_constant, 
+  y = normalized$flux_range, 
+  z = normalized$yi, 
+  type = "contour", 
+  contours = list(
+    start = -19.42062,
+    end = 5.286697,
+    size = 1, 
+    coloring = 'heatmap'
+))
+fig
+
+#GAM APPROACH
+#fit GAM
+gam1<-gam(yi~te(mean_temp_constant,flux_range,k=6),data=normalized) # note, lots of decisions that I blew by here
+summary(gam1)
+#estimate values via gam
+vals<-expand.grid(flux_range=seq(0,20,length.out = 100), mean_temp_constant=seq(0,35,length.out = 100))
+pd.gam<-predict(gam1,newdata=vals)
+surf<-data.frame(vals,value=pd.gam)
+#plot values as a surface
+# setting up custom color palette for surfaces, or anything besides the default green/ yellow one
+or2<-c('#fdd49e','#fdbb84','#fc8d59','#e34a33','#b30000')
+or2.bl2.pal<-c(colorRampPalette(colors=or2)(100))
+#plot gam surface
+ggplot(surf,aes(x=mean_temp_constant,y=flux_range,z=value))+
+  geom_raster(aes(fill=value))+
+  #geom_abline(intercept=0,slope=1,linetype=2)+
+  geom_contour(colour=gray(0.2),binwidth=0.5,size=0.2)+
+  #geom_contour(colour='black',breaks=c(mean(surf$value)),size=2)+
+  #geom_point(data=sample.locs,alpha=0.5)+
+  geom_contour(colour="black",breaks=c(0),size=2)+
+  scale_x_continuous('Mean Temp (°C)',expand=c(0,0))+
+  scale_y_continuous('Temp Range (°C)',expand=c(0,0))+
+  scale_fill_gradientn('Size',
+                       colors= or2.bl2.pal,
+                       breaks=seq(-2,5,2),limits=c(-2,5),
+                       guide = guide_colourbar(ticks.colour = gray(.2),
+                                               ticks.linewidth = 1.,
+                                               frame.colour = "black"))+
+  coord_cartesian(xlim=c(0,35),ylim=c(0,20))+
+  theme_bw()
+
+max(normalized$yi)
+
+```
 

### 2021-04-08 shared analysis files /data_wrangling_plot.Rmd
updates for plots for inclusion in results and discussion

--- 2021-04-02 shared analysis files /data_wrangling_plot.Rmd
@@
 ---
-title: "data wrangling and plots"
+title: "final figures for inclusion"
 author: "Maggie Slein"
-date: "4/2/21"
+date: "4/7/21"
 output: pdf_document
@@
 
-# Plots
+## Question: How does response compare across studies and experiments?
+Hypothesis: There should be differences across studies because of differences in experimental designs and study organism that would mean different magnitudes of response.
 
+Initial conclusions: The studies overall are different in their responses but not owing to study_id or experiment.
+However, the mixed effects model does suggest some differences when you include study _id as a mod.
+
+```{r, echo= FALSE, message=FALSE}
+#simple random effects model 
+simple_model <- rma.mv(yi, vi, data=dat_MA_ES, 
+               random = ~1 | study_id / experiment_id/ response_id,
+                 method="REML") 
+simple_model
+```
+
+## Relevant Plots
+
+```{r, echo = FALSE}
+#forest plot
+forest <- normalized %>%
+  group_by(study_id, experiment_id) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
+forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2")
+```
+**Figure 1.**
+Forest plot broken out by study and experiment and their observed effected sizes and sampling variances.
+
+```{r, echo = FALSE}
+#funnel plot 
+### set up 2x2 array for plotting
+par(mfrow=c(2,2))
+funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
+                        exp_age + size + org_level, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+```
+
+```{r, echo = FALSE, fig.height=8, fig.width=12}
+### draw funnel plots
+#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
+funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+```
+
+**Figure 2.** 
+Well, from what I can gather, a majority of the effect sizes from our meta-analysis have a non-signficant effect size value.
+However, there is a population of a couple influential effect sizes that are significantly important.
+
+```{r, echo= FALSE, message= FALSE}
+#GAM APPROACH
+#fit GAM
+gam1<-gam(yi~te(mean_temp_constant,flux_range,k=7),data=normalized) # note, lots of decisions that I blew by here
+summary(gam1)
+#estimate values via gam
+vals<-expand.grid(flux_range=seq(0,20,length.out = 100), mean_temp_constant=seq(0,35,length.out = 100))
+pd.gam<-predict(gam1,newdata=vals)
+surf<-data.frame(vals,value=pd.gam)
+vertices <- data.frame(
+  x = c(0, 0, 20),
+  y = c(10, 0, 0), 
+  id = c("a", "a", "a")
+)
+#plot values as a surface
+```
+
+```{r, echo= FALSE, message= FALSE}
+#plot gam surface
+ggplot()+
+  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range, fill=value))+
+  geom_contour(data = surf, 
+               aes(x=mean_temp_constant,y=flux_range,z=value), colour=gray(0.2),binwidth=0.5,size=0.2)+
+  geom_contour(data = surf,  aes(x=mean_temp_constant,y=flux_range,z=value), 
+               colour="black", breaks=c(0),size=2)+
+  #geom_contour_filled()
+  scale_x_continuous('Mean Temp (°C)',expand=c(0,0))+
+  scale_y_continuous('Temp Range (°C)',expand=c(0,0))+
+  scale_fill_gradientn('Effect Size',
+                       colors= viridis(100),
+                       breaks=seq(-2,5,2),limits=c(-2,5),
+                       guide = guide_colourbar(ticks.colour = gray(.2),
+                                               ticks.linewidth = 1.,
+                                               frame.colour = "black"))+
+  coord_cartesian(xlim=c(0,35),ylim=c(0,20))+
+  geom_jitter(data = normalized, mapping = aes(x=mean_temp_constant,y=flux_range,z= yi,fill = yi),
+              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
+  scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")+
+  theme_bw()+ 
+  theme(legend.position="bottom")+
+  geom_polygon(data = vertices, mapping = aes(x = x, y = y), fill = "gray")
+```
+**Figure 3.** Contour plot of GAM generated effects sizes overlayed with raw effect sizes from the extracted dataset.
+Black lines denote the boundary from negative to positive effect sizes.
+Grey triange denotes excluded contour that lacks any raw data.
+
+\newpage
 ## Question: How does response vary with study covariates?
@@
 full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
-                        exp_age + size + org_level, 
-               random = ~1 | experiment_id/ study_id/ response_id,
+                        exp_age + size + org_level + resp_type, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
                  method="REML") 
@@
 
+#separate model with thermal history trait as a mod 
+full_rf_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = 
+                          ~flux_range * mean_temp_constant + stressful, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+full_rf_thermal
+#absolute yi model 
+absolute_model<-rma.mv(abs(yi), vi, data=dat_MA_ES, mods = 
+                          ~flux_range * mean_temp_constant +
+                        exp_age + size + org_level + resp_type, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+absolute_model
+```
+
+## Relevant plots
+
+```{r, echo= FALSE}
+# scatterplot of standardized mean response vs flux range colored and lm fit by org level
+ggplot(normalized, aes(x=flux_range, y=yi, color = factor(org_level)))+
+  geom_point()+
+  scale_color_viridis(discrete = TRUE, option = "D")+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  theme(legend.position="bottom")+
+  ggtitle("SMD across fluctuation ranges colored by organization level 
+          and fit with linear model")
+```
+**Figure 4.**
+We can also see that there a difference in effect size in organization level responses across fluctuation range.
+Population level reponses trended positively across fluctuation ranges, while organism level responses trended negatively across fluctuation ranges.
+
+```{r, echo= FALSE}
+# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
+normalized %>%
+  drop_na(stressful) %>%
+ggplot(aes(x=flux_range, y=yi, color = factor(stressful)))+
+  geom_jitter()+
+  scale_color_viridis(discrete = TRUE, option = "D")+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  theme(legend.position="bottom")+
+  ggtitle("SMD across fluctuation ranges colored by 
+          whether or mean temperature was thermally stressful")
+```
+**Figure 5.** 
+When we consider where the temperatures in the fluctuation range reach thermally stressful levels, there is a small difference between responses to stressful  temperatures (negative) and non-stressful temperatures (positive).
+
+```{r, echo= FALSE}
+# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
+normalized %>%
+ggplot(aes(x=flux_range, y=yi, color = resp_type))+
+  geom_jitter()+
+  scale_color_viridis(discrete = TRUE, option = "D")+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  theme(legend.position="bottom")+
+  ggtitle("SMD across fluctuation ranges colored by response type")
+
+```
+**Figure 6.** We cab see that the ranges at which traits are measured is much wider than rates, however, there appears to a minorly negative trend in effect sizes as traits increase in fluctuation range.
+The opposite is true for rates.
+Great way to show that there is a lack of information on rates in the collected dataset.
+
+\newpage
+
+Running questions:
+
+- How do all the different models sound? Does it make sense to have these different models?
+- Do the figures match up with the story/questions associated with them?
+- Jittering versus standard plotting for the regression plots?
+_________________________________
+Supplementary Plots/Code
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#scatterplot of standardized mean difference across flux range colored by study fit with lm 
+ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
+  geom_point()+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  ggtitle("SMD across fluctuation ranges colored by 
+          studies and fit with linear model by study")
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+# boxplots of how fluctuation range influences SMD
+ggplot(normalized, aes(x=flux_range, y=yi))+
+  geom_point(alpha = 0.5)+
+  theme_bw()+
+  geom_smooth(method="lm", formula = y~x)+
+  ggtitle("SMD across flux_range")
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
 #without interaction term
@@
 full_rf_stocks
-
-#full model split out by life history outcomes
-full_rf_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range + mean_temp_constant +
-                        exp_age + size + org_level + stressful , 
-               random = ~1 | experiment_id/ study_id/ response_id,
-                 method="REML") 
-full_rf_thermal
 ```
 
-## Relevant plots
-
-```{r, echo= FALSE}
-# boxplots of how fluctuation range influences SMD
-ggplot(normalized, aes(x=flux_range, y=yi))+
-  geom_point(alpha = 0.5)+
-  theme_bw()+
-  geom_smooth(method="lm", formula = y~x)+
-  ggtitle("SMD across flux_range")
-```
-**Figure 1.**
-Data visualization supports model findings that there is a differene between shorter fluctuation ranges and longer fluctuation ranges.
-
-```{r, echo= FALSE}
-# scatterplot of standardized mean response vs flux range colored and lm fit by org level
-ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+# scatterplot of how mean temperature influences SMD
+ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
   geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  theme_bw()+
-  ggtitle("SMD across fluctuation ranges colored by organization level 
-          and fit with linear model")
-```
-**Figure 2.**
-We can also see that there a difference in effect size in organization level responses across fluctuation range.
-Population level reponses trended positively across fluctuation ranges, while organism level responses trended negatively across fluctuation ranges.
-
-```{r, echo= FALSE}
-#scatterplot of standardized mean difference across flux range colored by study fit with lm 
-ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
-  geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  theme_bw()+
-  ggtitle("SMD across fluctuation ranges colored by 
-          studies and fit with linear model by study")
-```
-**Figure 3.**
-When broken out by study, we can see that there is disagreement in the effect of fluctuation range on effect size.
-
-```{r, echo= FALSE}
-# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
-normalized %>%
-  drop_na(stressful) %>%
-ggplot(aes(x=flux_range, y=yi, color = stressful))+
-  geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  theme_bw()+
-  ggtitle("SMD across fluctuation ranges colored by 
-          whether or mean temperature was thermally stressful")
-```
-**Figure 4.** 
-When we consider where the temperatures in the fluctuation range reach thermally stressful levels, there is a small difference between responses to stressful  temperatures (negative) and non-stressful temperatures (positive).
-
-```{r, echo= FALSE}
-# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
-normalized %>%
-ggplot(aes(x=flux_range, y=yi, color = resp_type))+
-  geom_point()+
-  geom_smooth(method="lm", formula = y~x)+
-  theme_bw()+
-  ggtitle("SMD across fluctuation ranges colored by response type")
-```
-**Figure 5.**
-
-## Question: How does response compare across studies and experiments?
-Hypothesis: There should be differences across studies because of differences in experimental designs and study organism that would mean different magnitudes of response.
-
-Initial conclusions: The studies overall are different in their responses but not owing to study_id or experiment.
-However, the mixed effects model does suggest some differences when you include study _id as a mod.
-
-```{r, echo= FALSE, message=FALSE}
-#corresponding random effects model 
-fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
-               random = ~1 | experiment_id/ study_id,
-                 method="REML") 
-fig1
+  theme_minimal()+
+  geom_smooth(formula = y~x, method = "lm")+
+  ggtitle("SMD across mean temperature")
 ```
 
-## Relevant Plots
-
-```{r, echo = FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
 #SMD across all studies
@@
 
-**Figure 6.**
-When we look at experiment and study, we can see across the 15 studies included in this meta-analysis there are differences across and within experiments.
-
-```{r, echo = FALSE}
-#forest plot that shows the same thing?
-forest <- normalized %>%
-  group_by(study_id) %>%
-  summarise(mean_yi = mean(yi),
-            mean_vi = mean(vi)) %>%
-  arrange(desc(mean_yi))
-forest(forest$mean_yi, forest$mean_vi, slab=forest$study_id)
-```
-
-**Figure 7.**
-I kind of like this figure better than the boxplot in Figure 5 as an overall figure?
-
-```{r, echo = FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
 #looking at yi across studies with the same temperature range (10 C)
@@
 
-**Figure 8.**
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+fig <- plot_ly(
+  x = normalized$mean_temp_constant, 
+  y = normalized$flux_range, 
+  z = normalized$yi, 
+  type = "contour", 
+  contours = list(
+    start = -19.42062,
+    end = 5.286697,
+    size = 1, 
+    coloring = 'heatmap'
+))
+fig
 
-```{r, echo = FALSE}
-# scatterplot of how mean temperature influences SMD
-ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
-  geom_point()+
-  theme_minimal()+
-  geom_smooth(formula = y~x, method = "lm")+
-  ggtitle("SMD across mean temperature")
 ```
-
-```{r, echo = FALSE}
-
-#funnel plot 
-### set up 2x2 array for plotting
-par(mfrow=c(2,2))
-
-funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
-                        exp_age + size + org_level, 
-               random = ~1 | experiment_id/ study_id/ response_id,
-                 method="REML") 
-```
-
-**Figure 9.**
-
-```{r, echo = FALSE, fig.height=8, fig.width=12}
-### draw funnel plots
-#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
-funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
-funnel(funnel_model, yaxis="vi", main="Sampling Variance")
-```
-
-**Figure 10.** 
-Well, from what I can gather, a majority of the effect sizes from our meta-analysis have a non-signficant effect size value.
-However, there is a population of a couple influential effect sizes that are significantly important.
-
-_________________________________
-Supplementary Plots/Code
 
@@
 
-```{r, echo= FALSE, results = 'hide', message=FALSE}
+```{r, echo= FALSE, results = 'hide', message=FALSE, eval = FALSE}
 #basic linear model for study_id

### 2021-04-14 shared analysis files /data_wrangling_plot.Rmd
rethinking contour plot

@@
 #fit GAM
-gam1<-gam(yi~te(mean_temp_constant,flux_range,k=7),data=normalized) # note, lots of decisions that I blew by here
+gam1<-gam(yi~te(mean_temp_constant + te(as.factor(flux_range), k=7)),data=normalized) # note, lots of decisions that I blew by here
 summary(gam1)
@@
 ggplot()+
-  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range, fill=value))+
+  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range))+
   geom_contour(data = surf, 
@@
   scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")+
-  theme_bw()+ 
+  theme_classic()+ 
   theme(legend.position="bottom")+
   geom_polygon(data = vertices, mapping = aes(x = x, y = y), fill = "gray")
+
+normalized %>%
+ggplot(aes(x=mean_temp_constant, y=flux_range, fill = yi))+
+  geom_jitter(data = normalized, mapping = aes(x=mean_temp_constant,y=flux_range,z= yi,fill = yi),
+              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
+  scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")
 ```
@@
 - Jittering versus standard plotting for the regression plots?
+
+```{r}
+normalized %>%
+  ggplot(aes(y=yi, x=mean_temp_constant, color = flux_range))+
+  geom_point(size=4)+
+  scale_color_viridis(option= "B")+
+  theme_classic()
+
+normalized %>%
+  ggplot(aes(y=flux_range, x=mean_temp_constant, color = yi))+
+  geom_jitter(size=4)+
+  scale_color_viridis(option= "B")+
+  theme_classic()
+
+```
+
 _________________________________

### 2021-04-20 shared analysis files /data_wrangling_plot.Rmd
trying to reboot my thesis formatting and also updating final results models and figures

--- 2021-04-08 shared analysis files /data_wrangling_plot.Rmd
@@
 
+```{r}
+dat_MA_ES %>%
+  count(study_id)
+```
+
 ```{r, echo=FALSE, results = 'hide', message=FALSE}
 ## loading libraries 
-pacman::p_load(metafor, tidyverse, viridis, ggsci, plotly,  mgcv)
+pacman::p_load(metafor, tidyverse, viridis, visreg)
+#pacman::p_load(ggsci, plotly,  mgcv)
 #devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
@@
 normalized <- dat_MA_ES %>%
-  filter(yi > -20) 
+  filter(yi > -4) 
 #filtering for just flux ranges of 10 C 
@@
 simple_model
+
+absolute_model <- rma.mv(abs(yi), vi, data=dat_MA_ES, 
+               random = ~1 | study_id / experiment_id/ response_id,
+                 method="REML") 
+absolute_model
 ```
@@
 #forest plot
-forest <- normalized %>%
+forest <- dat_MA_ES %>%
   group_by(study_id, experiment_id) %>%
@@
 ```{r, echo= FALSE, message= FALSE}
-#GAM APPROACH
-#fit GAM
-gam1<-gam(yi~te(mean_temp_constant,flux_range,k=7),data=normalized) # note, lots of decisions that I blew by here
-summary(gam1)
-#estimate values via gam
-vals<-expand.grid(flux_range=seq(0,20,length.out = 100), mean_temp_constant=seq(0,35,length.out = 100))
-pd.gam<-predict(gam1,newdata=vals)
-surf<-data.frame(vals,value=pd.gam)
-vertices <- data.frame(
-  x = c(0, 0, 20),
-  y = c(10, 0, 0), 
-  id = c("a", "a", "a")
-)
-#plot values as a surface
+normalized %>%
+ggplot(aes(fill=mean_temp_constant, y = yi, x=flux_range))+
+  geom_jitter( alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
+  scale_fill_viridis('Mean Temperature (C)', option = "C")+
+  theme_classic()+
+  xlab("Temperature range")+
+  ylab("Effect Size")+
+  geom_line(y=0, linetype = "dashed")+
+  theme(legend.position="bottom")
 ```
 
-```{r, echo= FALSE, message= FALSE}
-#plot gam surface
-ggplot()+
-  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range, fill=value))+
-  geom_contour(data = surf, 
-               aes(x=mean_temp_constant,y=flux_range,z=value), colour=gray(0.2),binwidth=0.5,size=0.2)+
-  geom_contour(data = surf,  aes(x=mean_temp_constant,y=flux_range,z=value), 
-               colour="black", breaks=c(0),size=2)+
-  #geom_contour_filled()
-  scale_x_continuous('Mean Temp (°C)',expand=c(0,0))+
-  scale_y_continuous('Temp Range (°C)',expand=c(0,0))+
-  scale_fill_gradientn('Effect Size',
-                       colors= viridis(100),
-                       breaks=seq(-2,5,2),limits=c(-2,5),
-                       guide = guide_colourbar(ticks.colour = gray(.2),
-                                               ticks.linewidth = 1.,
-                                               frame.colour = "black"))+
-  coord_cartesian(xlim=c(0,35),ylim=c(0,20))+
-  geom_jitter(data = normalized, mapping = aes(x=mean_temp_constant,y=flux_range,z= yi,fill = yi),
-              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
-  scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")+
-  theme_bw()+ 
-  theme(legend.position="bottom")+
-  geom_polygon(data = vertices, mapping = aes(x = x, y = y), fill = "gray")
+```{r}
+# keep this to justify range of (-4:5)
+dat_MA_ES %>%
+filter(yi > -20) %>%
+  ggplot(aes(x=yi))+
+  geom_histogram(binwidth = 1)
 ```
@@
 full_rf_model
+```
+
+```{r}
+#plot to include to demonstrate model estimates
+full_model_coefficients<- as.data.frame(coef(full_rf_model)) %>%
+  tibble::rownames_to_column("covariates")
+
+full_model_coefficients %>%
+  ggplot(aes(x = fct_reorder(covariates, coef(full_rf_model)),
+             y = coef(full_rf_model),
+             fill = coef(full_rf_model) > 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  labs(x = NULL, y = "Coefficient estimate") +
+  scale_fill_viridis(discrete = TRUE)+
+  theme_bw()
+```
+
+```{r}
+#thermal history only bar graph
 
 #separate model with thermal history trait as a mod 
-full_rf_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = 
-                          ~flux_range * mean_temp_constant + stressful, 
+only_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = 
+                          ~ stressful, 
                random = ~1 | experiment_id/ study_id/ response_id,
                  method="REML") 
-full_rf_thermal
-#absolute yi model 
-absolute_model<-rma.mv(abs(yi), vi, data=dat_MA_ES, mods = 
-                          ~flux_range * mean_temp_constant +
-                        exp_age + size + org_level + resp_type, 
-               random = ~1 | experiment_id/ study_id/ response_id,
-                 method="REML") 
-absolute_model
+only_thermal
+
+thermal_only_coefficients<- as.data.frame(coef(only_thermal)) %>%
+  tibble::rownames_to_column("covariates")
+
 ```
@@
 ## Relevant plots
+
+```{r}
+ggplot(normalized, aes(x=flux_range, y=yi, color = factor(exp_age)))+
+  geom_point()+
+  labs(title = "SMD across fluctuation ranges colored by organism size and fit with linear model", 
+       x = "Temperature fluctuation range (C)",
+       y ="SMD",
+       color = "Experimental age")+
+  scale_color_viridis(discrete = TRUE, option = "D", labels = c("larval", "juvenile", "adult"))+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  theme(legend.position="bottom")
+```
+
+```{r}
+# scatterplot of standardized mean response vs flux range colored and lm fit by organism size
+ggplot(normalized, aes(x=flux_range, y=yi, color = factor(size)))+
+  geom_point()+
+  labs(title = "SMD across fluctuation ranges colored by organism size and fit with linear model", 
+       x = "Temperature fluctuation range (C)",
+       y ="SMD",
+       color = "Size")+
+  scale_color_viridis(discrete = TRUE, option = "D", labels = c("extra-small", "small", "medium", "large"))+
+  geom_smooth(method="lm", formula = y~x)+
+  theme_bw()+
+  theme(legend.position="bottom")
+```
 
@@
   geom_point()+
-  scale_color_viridis(discrete = TRUE, option = "D")+
+  labs(title = "SMD across fluctuation ranges colored by organization level and fit with linear model", 
+       x = "Temperature fluctuation range (C)",
+       y ="SMD",
+       color = "Biological organization level")+
+  scale_color_viridis(discrete = TRUE, option = "D", labels = c("organismal", "population"))+
   geom_smooth(method="lm", formula = y~x)+
   theme_bw()+
-  theme(legend.position="bottom")+
-  ggtitle("SMD across fluctuation ranges colored by organization level 
-          and fit with linear model")
+  theme(legend.position="bottom") 
 ```
@@
   geom_jitter()+
-  scale_color_viridis(discrete = TRUE, option = "D")+
+  labs(title = "SMD across fluctuation ranges colored by thermal stress experienced", 
+       x = "Temperature fluctuation range (C)",
+       y ="SMD",
+       color = "Thermally stressfull")+
+  scale_color_viridis(discrete = TRUE, option = "D", labels = c("no", "yes"))+
   geom_smooth(method="lm", formula = y~x)+
   theme_bw()+
-  theme(legend.position="bottom")+
-  ggtitle("SMD across fluctuation ranges colored by 
-          whether or mean temperature was thermally stressful")
+  theme(legend.position="bottom")
 ```
@@
   geom_jitter()+
-  scale_color_viridis(discrete = TRUE, option = "D")+
+  labs(title = "SMD across fluctuation ranges colored by response type", 
+       x = "Temperature fluctuation range (C)",
+       y ="SMD",
+       color = "Life history metric")+
+  scale_color_viridis(discrete = TRUE, option = "D", labels = c("rate", "trait"))+
   geom_smooth(method="lm", formula = y~x)+
   theme_bw()+
-  theme(legend.position="bottom")+
-  ggtitle("SMD across fluctuation ranges colored by response type")
-
+  theme(legend.position="bottom")
 ```
@@
 - Jittering versus standard plotting for the regression plots?
+
 _________________________________
 Supplementary Plots/Code
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+za <- dat_MA_ES %>%
+  filter(study_id != "garcaruiz2011")
+
+full_rf_thermal_2 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
+                        exp_age + size + org_level + resp_type + stressful, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+full_rf_thermal_2
+za_model <- rma.mv(yi, vi, data=za, mods = ~flux_range * mean_temp_constant +
+                        exp_age + size + org_level + resp_type, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+za_model
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#GAM APPROACH
+#fit GAM
+gam1<-gam(yi~te(mean_temp_constant + te(as.factor(flux_range), k=7)),data=normalized) # note, lots of decisions that I blew by here
+summary(gam1)
+#estimate values via gam
+vals<-expand.grid(flux_range=seq(0,20,length.out = 100), mean_temp_constant=seq(0,35,length.out = 100))
+pd.gam<-predict(gam1,newdata=vals)
+surf<-data.frame(vals,value=pd.gam)
+vertices <- data.frame(
+  x = c(0, 0, 20),
+  y = c(10, 0, 0), 
+  id = c("a", "a", "a")
+)
+#plot values as a surface
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+#plot gam surface
+ggplot()+
+  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range))+
+  geom_contour(data = surf, 
+               aes(x=mean_temp_constant,y=flux_range,z=value), colour=gray(0.2),binwidth=0.5,size=0.2)+
+  geom_contour(data = surf,  aes(x=mean_temp_constant,y=flux_range,z=value), 
+               colour="black", breaks=c(0),size=2)+
+  #geom_contour_filled()
+  scale_x_continuous('Mean Temp (°C)',expand=c(0,0))+
+  scale_y_continuous('Temp Range (°C)',expand=c(0,0))+
+  scale_fill_gradientn('Effect Size',
+                       colors= viridis(100),
+                       breaks=seq(-2,5,2),limits=c(-2,5),
+                       guide = guide_colourbar(ticks.colour = gray(.2),
+                                               ticks.linewidth = 1.,
+                                               frame.colour = "black"))+
+  coord_cartesian(xlim=c(0,35),ylim=c(0,20))+
+  geom_jitter(data = normalized, mapping = aes(x=mean_temp_constant,y=flux_range,z= yi,fill = yi),
+              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
+  scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")+
+  theme_classic()+ 
+  theme(legend.position="bottom")+
+  geom_polygon(data = vertices, mapping = aes(x = x, y = y), fill = "gray")
+
+```
+
+```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
+yi_direcionality <- normalized %>%
+  mutate(yi_direction = case_when(
+    yi > 0 ~ "positive", 
+    TRUE~ "negative"
+  ))
+
+yi_direcionality %>%
+ggplot(aes(x=mean_temp_constant, y= flux_range, fill= as.factor(yi_direction)))+
+  geom_jitter(data = yi_direcionality, 
+              mapping = aes(x=mean_temp_constant,y=flux_range,fill = yi_direction),
+              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
+  geom_smooth(method = "lm", color = "black")
+```
 

### 2021-04-20 writing /thesis_draft/_book/thesis.tex
i somewhat fixed the knitting problem but the formatting is still off

% This is the Reed College LaTeX thesis template. Most of the work
% for the document class was done by Sam Noble (SN), as well as this
% template. Later comments etc. by Ben Salzberg (BTS). Additional
% restructuring and APA support by Jess Youngberg (JY).
% Your comments and suggestions are more than welcome; please email
% them to cus@reed.edu
%
% See https://www.reed.edu/cis/help/LaTeX/index.html for help. There are a
% great bunch of help pages there, with notes on
% getting started, bibtex, etc. Go there and read it if you're not
% already familiar with LaTeX.
%
% Any line that starts with a percent symbol is a comment.
% They won't show up in the document, and are useful for notes
% to yourself and explaining commands.
% Commenting also removes a line from the document;
% very handy for troubleshooting problems. -BTS

% As far as I know, this follows the requirements laid out in
% the 2002-2003 Senior Handbook. Ask a librarian to check the
% document before binding. -SN

%%
%% Preamble
%%
% \documentclass{<something>} must begin each LaTeX document
\documentclass[12pt,twoside]{reedthesis}
% Packages are extensions to the basic LaTeX functions. Whatever you
% want to typeset, there is probably a package out there for it.
% Chemistry (chemtex), screenplays, you name it.
% Check out CTAN to see: https://www.ctan.org/
%%
\usepackage{graphicx,latexsym}
\usepackage{amsmath}
\usepackage{amssymb,amsthm}
\usepackage{longtable,booktabs,setspace}
\usepackage{chemarr} %% Useful for one reaction arrow, useless if you're not a chem major
\usepackage[hyphens]{url}
% Added by CII
\usepackage{hyperref}
\usepackage{lmodern}
\usepackage{float}
\floatplacement{figure}{H}
% Thanks, @Xyv
\usepackage{calc}
% End of CII addition
\usepackage{rotating}

% Next line commented out by CII
%%% \usepackage{natbib}
% Comment out the natbib line above and uncomment the following two lines to use the new
% biblatex-chicago style, for Chicago A. Also make some changes at the end where the
% bibliography is included.
%\usepackage{biblatex-chicago}
%\bibliography{thesis}

% Added by CII (Thanks, Hadley!)
% Use ref for internal links
\renewcommand{\hyperref}[2][???]{\autoref{#1}}
\def\chapterautorefname{Chapter}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Subsection}
% End of CII addition

% Added by CII
\usepackage{caption}
\captionsetup{width=5in}
% End of CII addition

% \usepackage{times} % other fonts are available like times, bookman, charter, palatino

% Syntax highlighting #22
  \usepackage{color}
  \usepackage{fancyvrb}
  \newcommand{\VerbBar}{|}
  \newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
  % Add ',fontsize=\small' for more characters per line
  \usepackage{framed}
  \definecolor{shadecolor}{RGB}{248,248,248}
  \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
  \newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
  \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
  \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
  \newcommand{\BuiltInTok}[1]{#1}
  \newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
  \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
  \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
  \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
  \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
  \newcommand{\ExtensionTok}[1]{#1}
  \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
  \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\ImportTok}[1]{#1}
  \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
  \newcommand{\NormalTok}[1]{#1}
  \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
  \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
  \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
  \newcommand{\RegionMarkerTok}[1]{#1}
  \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% To pass between YAML and LaTeX the dollar signs are added by CII
\title{My Final College Paper}
\author{Margaret A. Slein}
% The month and year that you submit your FINAL draft TO THE LIBRARY (May or December)
\date{May 2021}
\division{Mathematics and Natural Sciences}
\advisor{Sam Fey}
\institution{Reed College}
\degree{Bachelor of Arts}
%If you have two advisors for some reason, you can use the following
% Uncommented out by CII
% End of CII addition

%%% Remember to use the correct department!
\department{Mathematics}
% if you're writing a thesis in an interdisciplinary major,
% uncomment the line below and change the text as appropriate.
% check the Senior Handbook if unsure.
%\thedivisionof{The Established Interdisciplinary Committee for}
% if you want the approval page to say "Approved for the Committee",
% uncomment the next line
%\approvedforthe{Committee}

% Added by CII
%%% Copied from knitr
%% maxwidth is the original width if it's less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

% From {rticles}

\renewcommand{\contentsname}{Table of Contents}
% End of CII addition

\setlength{\parskip}{0pt}

% Added by CII

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\Acknowledgements{
The future is bright, the future is dark

Time stands still, yet flies like a lark

Memories endure and memories fade

People are rough, yet smooth like suede

The days seem long, the days seem short

It rains all year, until the sunshine retorts

Reed is an experience, Reed is a place

Reed is a journey of learning how to keep pace

Reed is a individual, Reed is a kollective

Reed is space for being reflective

Reed is the present, Reed is the past

Reed is the net for our future that we cast

To those from Reed

To those from before

To those from the outside

To those at the core

To those I still talk to

To those I talk to no more

To those still here today

To those who now soar

A few words of dear thanks\ldots{}.

Thank you for the laughs, the tears, the screams

Gratitude for the late nights, the dancing, the dreams

Thank you for the walks, the talks, and meals together

Many thanks for the shenanigans I'll cherish forever

Thank you for the compassion, for the coffee and tea

Appreciation for the love, the hate, and learning to be

Thank you for the support, and most ardently, for embracing me
}

\Dedication{
You can have a dedication here if you wish.
}

\Preface{
This is an example of a thesis setup to use the reed thesis document class
(for LaTeX) and the R bookdown package, in general.
}

\Abstract{
The preface pretty much says it all.

\par

Second paragraph of abstract starts here.
}

% End of CII addition
%%
%% End Preamble
%%
%
\begin{document}

% Everything below added by CII
  \maketitle

\frontmatter % this stuff will be roman-numbered
\pagestyle{empty} % this removes page numbers from the frontmatter
  \begin{acknowledgements}
    The future is bright, the future is dark

    Time stands still, yet flies like a lark

    Memories endure and memories fade

    People are rough, yet smooth like suede

    The days seem long, the days seem short

    It rains all year, until the sunshine retorts

    Reed is an experience, Reed is a place

    Reed is a journey of learning how to keep pace

    Reed is a individual, Reed is a kollective

    Reed is space for being reflective

    Reed is the present, Reed is the past

    Reed is the net for our future that we cast

    To those from Reed

    To those from before

    To those from the outside

    To those at the core

    To those I still talk to

    To those I talk to no more

    To those still here today

    To those who now soar

    A few words of dear thanks\ldots{}.

    Thank you for the laughs, the tears, the screams

    Gratitude for the late nights, the dancing, the dreams

    Thank you for the walks, the talks, and meals together

    Many thanks for the shenanigans I'll cherish forever

    Thank you for the compassion, for the coffee and tea

    Appreciation for the love, the hate, and learning to be

    Thank you for the support, and most ardently, for embracing me
  \end{acknowledgements}
  \begin{preface}
    This is an example of a thesis setup to use the reed thesis document class
    (for LaTeX) and the R bookdown package, in general.
  \end{preface}
  \hypersetup{linkcolor=black}
  \setcounter{secnumdepth}{2}
  \setcounter{tocdepth}{2}
  \tableofcontents

  \listoftables

  \listoffigures
  \begin{abstract}
    The preface pretty much says it all.

    \par

    Second paragraph of abstract starts here.
  \end{abstract}
  \begin{dedication}
    You can have a dedication here if you wish.
  \end{dedication}
\mainmatter % here the regular arabic numbering starts
\pagestyle{fancyplain} % turns page numbering back on

\hypertarget{introduction}{%
\chapter*{Introduction}\label{introduction}}
\addcontentsline{toc}{chapter}{Introduction}

Welcome to the \emph{R Markdown} thesis template. This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before. Using \emph{R Markdown} will also allow you to easily keep track of your analyses in \textbf{R} chunks of code, with the resulting plots and output included as well. The hope is this \emph{R Markdown} template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.

Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis. The use of LaTeX in combination with \emph{Markdown} is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file. While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities. After working with \emph{Markdown} and \textbf{R} together for a few weeks, we are confident this will be your reporting style of choice going forward.

\textbf{Why use it?}

\emph{R Markdown} creates a simple and straightforward way to interface with the beauty of LaTeX. Packages have been written in \textbf{R} to work directly with LaTeX to produce nicely formatting tables and paragraphs. In addition to creating a user friendly interface to LaTeX, \emph{R Markdown} also allows you to read in your data, to analyze it and to visualize it using \textbf{R} functions, and also to provide the documentation and commentary on the results of your project. Further, it allows for \textbf{R} results to be passed inline to the commentary of your results. You'll see more on this later.

\textbf{Who should use it?}

Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use \emph{R Markdown}. Of particular use should be anyone in the sciences, but the user-friendly nature of \emph{Markdown} and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc. should make it of great benefit to nearly anyone writing a thesis project.

\textbf{For additional help with bookdown}

Please visit \href{https://bookdown.org/yihui/bookdown/}{the free online bookdown reference guide}.

\hypertarget{rmd-basics}{%
\chapter{R Markdown Basics}\label{rmd-basics}}

Here is a brief introduction into using \emph{R Markdown}. \emph{Markdown} is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. \emph{R Markdown} provides the flexibility of \emph{Markdown} with the implementation of \textbf{R} input and output. For more details on using \emph{R Markdown} see \url{https://rmarkdown.rstudio.com}.

Be careful with your spacing in \emph{Markdown} documents. While whitespace largely is ignored, it does at times give \emph{Markdown} signals as to how to proceed. As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph. In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).

\hypertarget{lists}{%
\section{Lists}\label{lists}}

It's easy to create a list. It can be unordered like
\begin{itemize}
\tightlist
\item
  Item 1
\item
  Item 2
\end{itemize}
or it can be ordered like
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Item 1
\item
  Item 2
\end{enumerate}
Notice that I intentionally mislabeled Item 2 as number 4. \emph{Markdown} automatically figures this out! You can put any numbers in the list and it will create the list. Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a tab). (Here's one case where indentation is key!)
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Item 1
\item
  Item 2
\item
  Item 3
  \begin{itemize}
  \tightlist
  \item
    Item 3a
  \item
    Item 3b
  \end{itemize}
\end{enumerate}
\hypertarget{line-breaks}{%
\section{Line breaks}\label{line-breaks}}

Make sure to add white space between lines if you'd like to start a new paragraph. Look at what happens below in the outputted document if you don't:

Here is the first sentence. Here is another sentence. Here is the last sentence to end the paragraph.
This should be a new paragraph.

\emph{Now for the correct way:}

Here is the first sentence. Here is another sentence. Here is the last sentence to end the paragraph.

This should be a new paragraph.

\hypertarget{r-chunks}{%
\section{R chunks}\label{r-chunks}}

When you click the \textbf{Knit} button above a document will be generated that includes both content as well as the output of any embedded \textbf{R} code chunks within the document. You can embed an \textbf{R} code chunk like this (\texttt{cars} is a built-in \textbf{R} dataset):
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(cars)}
\end{Highlighting}
\end{Shaded}
\begin{verbatim}
     speed           dist       
 Min.   : 4.0   Min.   :  2.00  
 1st Qu.:12.0   1st Qu.: 26.00  
 Median :15.0   Median : 36.00  
 Mean   :15.4   Mean   : 42.98  
 3rd Qu.:19.0   3rd Qu.: 56.00  
 Max.   :25.0   Max.   :120.00  
\end{verbatim}
\hypertarget{inline-code}{%
\section{Inline code}\label{inline-code}}

If you'd like to put the results of your analysis directly into your discussion, add inline code like this:
\begin{quote}
The \texttt{cos} of \(2 \pi\) is 1.
\end{quote}
Another example would be the direct calculation of the standard deviation:
\begin{quote}
The standard deviation of \texttt{speed} in \texttt{cars} is 5.2876444.
\end{quote}
One last neat feature is the use of the \texttt{ifelse} conditional statement which can be used to output text depending on the result of an \textbf{R} calculation:
\begin{quote}
The standard deviation is less than 6.
\end{quote}
Note the use of \texttt{\textgreater{}} here, which signifies a quotation environment that will be indented.

As you see with \texttt{\$2\ \textbackslash{}pi\$} above, mathematics can be added by surrounding the mathematical text with dollar signs. More examples of this are in \protect\hyperlink{math-sci}{Mathematics and Science} if you uncomment the code in \protect\hyperlink{math}{Math}.

\hypertarget{including-plots}{%
\section{Including plots}\label{including-plots}}

You can also embed plots. For example, here is a way to use the base \textbf{R} graphics package to produce a plot using the built-in \texttt{pressure} dataset:

\includegraphics{thesis_files/figure-latex/pressure-1.pdf}

Note that the \texttt{echo=FALSE} parameter was added to the code chunk to prevent printing of the \textbf{R} code that generated the plot. There are plenty of other ways to add chunk options (like \texttt{fig.height} and \texttt{fig.width} in the chunk above). More information is available at \url{https://yihui.org/knitr/options/}.

Another useful chunk option is the setting of \texttt{cache=TRUE} as you see here. If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance. Later in this file, you'll see a way to reference plots created in \textbf{R} or external figures.

\hypertarget{loading-and-exploring-data}{%
\section{Loading and exploring data}\label{loading-and-exploring-data}}

Included in this template is a file called \texttt{flights.csv}. This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014. More information about this dataset and its \textbf{R} package is available at \url{https://github.com/ismayc/pnwflights14}. This subset includes only Portland flights and only rows that were complete with no missing values. Merges were also done with the \texttt{airports} and \texttt{airlines} data sets in the \texttt{pnwflights14} package to get more descriptive airport and airline names.

We can load in this data set using the following commands:
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# flights.csv is in the data directory}
\NormalTok{flights_path <-}\StringTok{ }\NormalTok{here}\OperatorTok{::}\KeywordTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"flights.csv"}\NormalTok{)}
\CommentTok{# string columns will be read in as strings and not factors now}
\NormalTok{flights <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(flights_path, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
The data is now stored in the data frame called \texttt{flights} in \textbf{R}. To get a better feel for the variables included in this dataset we can use a variety of functions. Here we can see the dimensions (rows by columns) and also the names of the columns.
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dim}\NormalTok{(flights)}
\end{Highlighting}
\end{Shaded}
\begin{verbatim}
[1] 12649    16
\end{verbatim}
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{names}\NormalTok{(flights)}
\end{Highlighting}
\end{Shaded}
\begin{verbatim}
 [1] "month"        "day"          "dep_time"     "dep_delay"   
 [5] "arr_time"     "arr_delay"    "carrier"      "tailnum"     
 [9] "flight"       "dest"         "air_time"     "distance"    
[13] "hour"         "minute"       "carrier_name" "dest_name"   
\end{verbatim}
Another good idea is to take a look at the dataset in table form. With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here. I recommend you enter the command into the Console \textbf{\emph{after}} you have run the \textbf{R} chunks above to load the data into \textbf{R}.
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{View}\NormalTok{(flights)}
\end{Highlighting}
\end{Shaded}
While not required, it is highly recommended you use the \texttt{dplyr} package to manipulate and summarize your data set as needed. It uses a syntax that is easy to understand using chaining operations. Below I've created a few examples of using \texttt{dplyr} to get information about the Portland flights in 2014. You will also see the use of the \texttt{ggplot2} package, which produces beautiful, high-quality academic visuals.

We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# List of packages required for this analysis}
\NormalTok{pkg <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{, }\StringTok{"ggplot2"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"bookdown"}\NormalTok{)}
\CommentTok{# Check if packages are not installed and assign the}
\CommentTok{# names of the packages not installed to the variable new.pkg}
\NormalTok{new.pkg <-}\StringTok{ }\NormalTok{pkg[}\OperatorTok{!}\NormalTok{(pkg }\OperatorTok{%in%}\StringTok{ }\KeywordTok{installed.packages}\NormalTok{())]}
\CommentTok{# If there are any packages in the list that aren't installed,}
\CommentTok{# install them}
\ControlFlowTok{if}\NormalTok{ (}\KeywordTok{length}\NormalTok{(new.pkg)) \{}
  \KeywordTok{install.packages}\NormalTok{(new.pkg, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
\NormalTok{\}}
\CommentTok{# Load packages}
\KeywordTok{library}\NormalTok{(thesisdown)}
\KeywordTok{library}\NormalTok{(dplyr)}
\end{Highlighting}
\end{Shaded}
\begin{verbatim}
Warning: package 'dplyr' was built under R version 4.0.2
\end{verbatim}
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(ggplot2)}
\end{Highlighting}
\end{Shaded}
\begin{verbatim}
Warning: package 'ggplot2' was built under R version 4.0.2
\end{verbatim}
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(knitr)}
\end{Highlighting}
\end{Shaded}
\begin{verbatim}
Warning: package 'knitr' was built under R version 4.0.2
\end{verbatim}
\clearpage

The example we show here does the following:
\begin{itemize}
\item
  Selects only the \texttt{carrier\_name} and \texttt{arr\_delay} from the \texttt{flights} dataset and then assigns this subset to a new variable called \texttt{flights2}.
\item
  Using \texttt{flights2}, we determine the largest arrival delay for each of the carriers.
\end{itemize}
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{flights2 <-}\StringTok{ }\NormalTok{flights }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(carrier_name, arr_delay)}
\NormalTok{max_delays <-}\StringTok{ }\NormalTok{flights2 }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(carrier_name) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{max_arr_delay =} \KeywordTok{max}\NormalTok{(arr_delay, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}
A useful function in the \texttt{knitr} package for making nice tables in \emph{R Markdown} is called \texttt{kable}. It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX. This again goes to show how nice reproducible documents can be! (Note the use of \texttt{results="asis"}, which will produce the table instead of the code to create the table.) The \texttt{caption.short} argument is used to include a shorter title to appear in the List of Tables.
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{kable}\NormalTok{(max_delays,}
  \DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"Airline"}\NormalTok{, }\StringTok{"Max Arrival Delay"}\NormalTok{),}
  \DataTypeTok{caption =} \StringTok{"Maximum Delays by Airline"}\NormalTok{,}
  \DataTypeTok{caption.short =} \StringTok{"Max Delays by Airline"}\NormalTok{,}
  \DataTypeTok{longtable =} \OtherTok{TRUE}\NormalTok{,}
  \DataTypeTok{booktabs =} \OtherTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\begin{longtable}[t]{lr}
\caption[Max Delays by Airline]{\label{tab:maxdelays}Maximum Delays by Airline}\\
\toprule
Airline & Max Arrival Delay\\
\midrule
Alaska Airlines Inc. & 338\\
American Airlines Inc. & 1539\\
Delta Air Lines Inc. & 371\\
Frontier Airlines Inc. & 166\\
Hawaiian Airlines Inc. & 116\\
\addlinespace
JetBlue Airways & 256\\
SkyWest Airlines Inc. & 321\\
Southwest Airlines Co. & 315\\
United Air Lines Inc. & 319\\
US Airways Inc. & 347\\
\addlinespace
Virgin America & 366\\
\bottomrule
\end{longtable}
The last two options make the table a little easier-to-read.

We can further look into the properties of the largest value here for American Airlines Inc.~To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original \texttt{flights} dataset.
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{flights }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}
\NormalTok{    arr_delay }\OperatorTok{==}\StringTok{ }\DecValTok{1539}\NormalTok{,}
\NormalTok{    carrier_name }\OperatorTok{==}\StringTok{ "American Airlines Inc."}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}
\NormalTok{    month, day, carrier, dest_name, hour,}
\NormalTok{    minute, carrier_name, arr_delay}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}
\begin{verbatim}
  dep_time dep_delay arr_time tailnum flight dest air_time distance
1     1403      1553     1934  N595AA   1568  DFW      182     1616
\end{verbatim}
We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth. Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{flights }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(month }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{, day }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ dep_time, }\DataTypeTok{y =}\NormalTok{ arr_delay)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{()}
\end{Highlighting}
\end{Shaded}
\includegraphics{thesis_files/figure-latex/march3plot-1.pdf}

\hypertarget{additional-resources}{%
\section{Additional resources}\label{additional-resources}}
\begin{itemize}
\item
  \emph{Markdown} Cheatsheet - \url{https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet}
\item
  \emph{R Markdown}
  \begin{itemize}
  \tightlist
  \item
    Reference Guide - \url{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}
  \item
    Cheatsheet - \url{https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf}
  \end{itemize}
\item
  \emph{RStudio IDE}
  \begin{itemize}
  \tightlist
  \item
    Cheatsheet - \url{https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf}
  \item
    Official website - \url{https://rstudio.com/products/rstudio/}
  \end{itemize}
\item
  Introduction to \texttt{dplyr} - \url{https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html}
\item
  \texttt{ggplot2}
  \begin{itemize}
  \tightlist
  \item
    Documentation - \url{https://ggplot2.tidyverse.org/}
  \item
    Cheatsheet - \url{https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf}
  \end{itemize}
\end{itemize}
\hypertarget{math-sci}{%
\chapter{Mathematics and Science}\label{math-sci}}

\hypertarget{math}{%
\section{Math}\label{math}}

\TeX~is the best way to typeset mathematics. Donald Knuth designed \TeX~when he got frustrated at how long it was taking the typesetters to finish his book, which contained a lot of mathematics. One nice feature of \emph{R Markdown} is its ability to read LaTeX code directly.

If you are doing a thesis that will involve lots of math, you will want to read the following section which has been commented out. If you're not going to use math, skip over or delete this next commented section.

\hypertarget{chemistry-101-symbols}{%
\section{Chemistry 101: Symbols}\label{chemistry-101-symbols}}

Chemical formulas will look best if they are not italicized. Get around math mode's automatic italicizing in LaTeX by using the argument \texttt{\$\textbackslash{}mathrm\{formula\ here\}\$}, with your formula inside the curly brackets. (Notice the use of the backticks here which enclose text that acts as code.)

So, \(\mathrm{Fe_2^{2+}Cr_2O_4}\) is written \texttt{\$\textbackslash{}mathrm\{Fe\_2\^{}\{2+\}Cr\_2O\_4\}\$}.

\noindent Exponent or Superscript: \(\mathrm{O^-}\)

\noindent Subscript: \(\mathrm{CH_4}\)

To stack numbers or letters as in \(\mathrm{Fe_2^{2+}}\), the subscript is defined first, and then the superscript is defined.

\noindent Bullet: CuCl \(\bullet\) \(\mathrm{7H_{2}O}\)

\noindent Delta: \(\Delta\)

\noindent Reaction Arrows: \(\longrightarrow\) or \(\xrightarrow{solution}\)

\noindent Resonance Arrows: \(\leftrightarrow\)

\noindent Reversible Reaction Arrows: \(\rightleftharpoons\)

\hypertarget{typesetting-reactions}{%
\subsection{Typesetting reactions}\label{typesetting-reactions}}

You may wish to put your reaction in an equation environment, which means that LaTeX will place the reaction where it fits and will number the equations for you.
\begin{equation}
  \mathrm{C_6H_{12}O_6  + 6O_2} \longrightarrow \mathrm{6CO_2 + 6H_2O}
  \label{eq:reaction}
\end{equation}
We can reference this combustion of glucose reaction via Equation \eqref{eq:reaction}.

\hypertarget{other-examples-of-reactions}{%
\subsection{Other examples of reactions}\label{other-examples-of-reactions}}

\(\mathrm{NH_4Cl_{(s)}}\) \(\rightleftharpoons\) \(\mathrm{NH_{3(g)}+HCl_{(g)}}\)

\noindent \(\mathrm{MeCH_2Br + Mg}\) \(\xrightarrow[below]{above}\) \(\mathrm{MeCH_2\bullet Mg \bullet Br}\)

\hypertarget{physics}{%
\section{Physics}\label{physics}}

Many of the symbols you will need can be found on the math page \url{https://web.reed.edu/cis/help/latex/math.html} and the Comprehensive LaTeX Symbol Guide (\url{https://mirror.utexas.edu/ctan/info/symbols/comprehensive/symbols-letter.pdf}).

\hypertarget{biology}{%
\section{Biology}\label{biology}}

You will probably find the resources at \url{https://www.lecb.ncifcrf.gov/~toms/latex.html} helpful, particularly the links to bsts for various journals. You may also be interested in TeXShade for nucleotide typesetting (\url{https://homepages.uni-tuebingen.de/beitz/txe.html}). Be sure to read the proceeding chapter on graphics and tables.

\hypertarget{ref-labels}{%
\chapter{Graphics, References, and Labels}\label{ref-labels}}

\hypertarget{figures}{%
\section{Figures}\label{figures}}

If your thesis has a lot of figures, \emph{R Markdown} might behave better for you than that other word processor. One perk is that it will automatically number the figures accordingly in each chapter. You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier. If you label your figures, you can move the figures around and \emph{R Markdown} will automatically adjust the numbering for you. No need for you to remember! So that you don't have to get too far into LaTeX to do this, a couple \textbf{R} functions have been created for you to assist. You'll see their use below.

In the \textbf{R} chunk below, we will load in a picture stored as \texttt{reed.jpg} in our main directory. We then give it the caption of ``Reed logo'', the label of ``reedlogo'', and specify that this is a figure. Make note of the different \textbf{R} chunk options that are given in the R Markdown file (not shown in the knitted document).
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{include_graphics}\NormalTok{(}\DataTypeTok{path =} \StringTok{"figure/reed.jpg"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\begin{figure}

{\centering \includegraphics[width=0.2\linewidth]{figure/reed}

}

\caption{Reed logo}\label{fig:reedlogo}
\end{figure}
Here is a reference to the Reed logo: Figure \ref{fig:reedlogo}. Note the use of the \texttt{fig:} code here. By naming the \textbf{R} chunk that contains the figure, we can then reference that figure later as done in the first sentence here. We can also specify the caption for the figure via the R chunk option \texttt{fig.cap}.

\clearpage

Below we will investigate how to save the output of an \textbf{R} plot and label it in a way similar to that done above. Recall the \texttt{flights} dataset from Chapter \ref{rmd-basics}. (Note that we've shown a different way to reference a section or chapter here.) We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mean_delay_by_carrier <-}\StringTok{ }\NormalTok{flights }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(carrier) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{mean_dep_delay =} \KeywordTok{mean}\NormalTok{(dep_delay))}
\KeywordTok{ggplot}\NormalTok{(mean_delay_by_carrier, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ carrier, }\DataTypeTok{y =}\NormalTok{ mean_dep_delay)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{position =} \StringTok{"identity"}\NormalTok{, }\DataTypeTok{stat =} \StringTok{"identity"}\NormalTok{, }\DataTypeTok{fill =} \StringTok{"red"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\begin{figure}
\centering
\includegraphics{thesis_files/figure-latex/delaysboxplot-1.pdf}
\caption{\label{fig:delaysboxplot}Mean Delays by Airline}
\end{figure}
Here is a reference to this image: Figure \ref{fig:delaysboxplot}.

A table linking these carrier codes to airline names is available at \url{https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv}.

\clearpage

Next, we will explore the use of the \texttt{out.extra} chunk option, which can be used to shrink or expand an image loaded from a file by specifying \texttt{"scale=\ "}. Here we use the mathematical graph stored in the ``subdivision.pdf'' file.
\begin{figure}
\includegraphics[scale=0.75]{figure/subdivision} \caption{Subdiv. graph}\label{fig:subd}
\end{figure}
Here is a reference to this image: Figure \ref{fig:subd}. Note that \texttt{echo=FALSE} is specified so that the \textbf{R} code is hidden in the document.

\textbf{More Figure Stuff}

Lastly, we will explore how to rotate and enlarge figures using the \texttt{out.extra} chunk option. (Currently this only works in the PDF version of the book.)
\begin{figure}
\includegraphics[angle=180, scale=1.1]{figure/subdivision} \caption{A Larger Figure, Flipped Upside Down}\label{fig:subd2}
\end{figure}
As another example, here is a reference: Figure \ref{fig:subd2}.

\hypertarget{footnotes-and-endnotes}{%
\section{Footnotes and Endnotes}\label{footnotes-and-endnotes}}

You might want to footnote something.\footnote{footnote text} The footnote will be in a smaller font and placed appropriately. Endnotes work in much the same way. More information can be found about both on the CUS site or feel free to reach out to \href{mailto:data@reed.edu}{\nolinkurl{data@reed.edu}}.

\hypertarget{bibliographies}{%
\section{Bibliographies}\label{bibliographies}}

Of course you will need to cite things, and you will probably accumulate an armful of sources. There are a variety of tools available for creating a bibliography database (stored with the .bib extension). In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero. The Reed librarians have created Zotero documentation at \url{https://libguides.reed.edu/citation/zotero}. In addition, a tutorial is available from Middlebury College at \url{https://sites.middlebury.edu/zoteromiddlebury/}.

\emph{R Markdown} uses \emph{pandoc} (\url{https://pandoc.org/}) to build its bibliographies. One nice caveat of this is that you won't have to do a second compile to load in references as standard LaTeX requires. To cite references in your thesis (after creating your bibliography database), place the reference name inside square brackets and precede it by the ``at'' symbol. For example, here's a reference to a book about worrying: (Molina \& Borkovec, 1994). This \texttt{Molina1994} entry appears in a file called \texttt{thesis.bib} in the \texttt{bib} folder. This bibliography database file was created by a program called BibTeX. You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the \texttt{bib} folder.

For more information about BibTeX and bibliographies, see our CUS site (\url{https://web.reed.edu/cis/help/latex/index.html})\footnote{Reed~College (2007)}. There are three pages on this topic: \emph{bibtex} (which talks about using BibTeX, at \url{https://web.reed.edu/cis/help/latex/bibtex.html}), \emph{bibtexstyles} (about how to find and use the bibliography style that best suits your needs, at \url{https://web.reed.edu/cis/help/latex/bibtexstyles.html}) and \emph{bibman} (which covers how to make and maintain a bibliography by hand, without BibTeX, at \url{https://web.reed.edu/cis/help/latex/bibman.html}). The last page will not be useful unless you have only a few sources.

If you look at the YAML header at the top of the main .Rmd file you can see that we can specify the style of the bibliography by referencing the appropriate csl file. You can download a variety of different style files at \url{https://www.zotero.org/styles}. Make sure to download the file into the csl folder.

\vfill

\textbf{Tips for Bibliographies}
\begin{itemize}
\tightlist
\item
  Like with thesis formatting, the sooner you start compiling your bibliography for something as large as thesis, the better. Typing in source after source is mind-numbing enough; do you really want to do it for hours on end in late April? Think of it as procrastination.
\item
  The cite key (a citation's label) needs to be unique from the other entries.
\item
  When you have more than one author or editor, you need to separate each author's name by the word ``and'' e.g. \texttt{Author\ =\ \{Noble,\ Sam\ and\ Youngberg,\ Jessica\},}.
\item
  Bibliographies made using BibTeX (whether manually or using a manager) accept LaTeX markup, so you can italicize and add symbols as necessary.
\item
  To force capitalization in an article title or where all lowercase is generally used, bracket the capital letter in curly braces.
\item
  You can add a Reed Thesis citation\footnote{Noble (2002)} option. The best way to do this is to use the phdthesis type of citation, and use the optional ``type'' field to enter ``Reed thesis'' or ``Undergraduate thesis.''
\end{itemize}
\hypertarget{anything-else}{%
\section{Anything else?}\label{anything-else}}

If you'd like to see examples of other things in this template, please contact the Data @ Reed team (email \href{mailto:data@reed.edu}{\nolinkurl{data@reed.edu}}) with your suggestions. We love to see people using \emph{R Markdown} for their theses, and are happy to help.

\hypertarget{conclusion}{%
\chapter*{Conclusion}\label{conclusion}}
\addcontentsline{toc}{chapter}{Conclusion}

If we don't want Conclusion to have a chapter number next to it, we can add the \texttt{\{-\}} attribute.

\textbf{More info}

And here's some other random info: the first paragraph after a chapter title or section head \emph{shouldn't be} indented, because indents are to tell the reader that you're starting a new paragraph. Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.

\appendix

\hypertarget{the-first-appendix}{%
\chapter{The First Appendix}\label{the-first-appendix}}

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the \texttt{include\ =\ FALSE} chunk tag) to help with readibility and/or setup.

\textbf{In the main Rmd file}
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# This chunk ensures that the thesisdown package is}
\CommentTok{# installed and loaded. This thesisdown package includes}
\CommentTok{# the template files for the thesis.}
\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \KeywordTok{stop}\NormalTok{(}
      \KeywordTok{paste}\NormalTok{(}\StringTok{'You need to run install.packages("remotes")",}
\StringTok{            "first in the Console.'}\NormalTok{)}
\NormalTok{    )}
\NormalTok{  \}}
\NormalTok{\}}
\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \KeywordTok{stop}\NormalTok{(}
      \KeywordTok{paste}\NormalTok{(}
        \StringTok{"You need to run"}\NormalTok{,}
        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
        \StringTok{"first in the Console."}
\NormalTok{      )}
\NormalTok{    )}
\NormalTok{  \}}
\NormalTok{\}}
\KeywordTok{library}\NormalTok{(thesisdown)}
\CommentTok{# Set how wide the R output will go}
\KeywordTok{options}\NormalTok{(}\DataTypeTok{width =} \DecValTok{70}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\textbf{In Chapter \ref{ref-labels}:}
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# This chunk ensures that the thesisdown package is}
\CommentTok{# installed and loaded. This thesisdown package includes}
\CommentTok{# the template files for the thesis and also two functions}
\CommentTok{# used for labeling and referencing}
\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \KeywordTok{stop}\NormalTok{(}
      \KeywordTok{paste}\NormalTok{(}
        \StringTok{'You need to run install.packages("remotes")'}\NormalTok{,}
        \StringTok{"first in the Console."}
\NormalTok{      )}
\NormalTok{    )}
\NormalTok{  \}}
\NormalTok{\}}
\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(dplyr)) \{}
  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \KeywordTok{stop}\NormalTok{(}
      \KeywordTok{paste}\NormalTok{(}
        \StringTok{'You need to run install.packages("dplyr")'}\NormalTok{,}
        \StringTok{"first in the Console."}
\NormalTok{      )}
\NormalTok{    )}
\NormalTok{  \}}
\NormalTok{\}}
\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(ggplot2)) \{}
  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \KeywordTok{stop}\NormalTok{(}
      \KeywordTok{paste}\NormalTok{(}
        \StringTok{'You need to run install.packages("ggplot2")'}\NormalTok{,}
        \StringTok{"first in the Console."}
\NormalTok{      )}
\NormalTok{    )}
\NormalTok{  \}}
\NormalTok{\}}
\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(bookdown)) \{}
  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"bookdown"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \KeywordTok{stop}\NormalTok{(}
      \KeywordTok{paste}\NormalTok{(}
        \StringTok{'You need to run install.packages("bookdown")'}\NormalTok{,}
        \StringTok{"first in the Console."}
\NormalTok{      )}
\NormalTok{    )}
\NormalTok{  \}}
\NormalTok{\}}
\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \KeywordTok{stop}\NormalTok{(}
      \KeywordTok{paste}\NormalTok{(}
        \StringTok{"You need to run"}\NormalTok{,}
        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
        \StringTok{"first in the Console."}
\NormalTok{      )}
\NormalTok{    )}
\NormalTok{  \}}
\NormalTok{\}}
\KeywordTok{library}\NormalTok{(thesisdown)}
\KeywordTok{library}\NormalTok{(dplyr)}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(knitr)}
\NormalTok{flights <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data/flights.csv"}\NormalTok{, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\hypertarget{the-second-appendix-for-fun}{%
\chapter{The Second Appendix, for Fun}\label{the-second-appendix-for-fun}}

\backmatter

\hypertarget{references}{%
\chapter*{References}\label{references}}
\addcontentsline{toc}{chapter}{References}

\markboth{References}{References}

\noindent

\setlength{\parindent}{-0.20in}
\setlength{\leftskip}{0.20in}
\setlength{\parskip}{8pt}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-angel2000}{}%
Angel, E. (2000). \emph{Interactive computer graphics : A top-down approach with opengl}. Boston, MA: Addison Wesley Longman.

\leavevmode\hypertarget{ref-angel2001}{}%
Angel, E. (2001a). \emph{Batch-file computer graphics : A bottom-up approach with quicktime}. Boston, MA: Wesley Addison Longman.

\leavevmode\hypertarget{ref-angel2002a}{}%
Angel, E. (2001b). \emph{Test second book by angel}. Boston, MA: Wesley Addison Longman.

\leavevmode\hypertarget{ref-Molina1994}{}%
Molina, S. T., \& Borkovec, T. D. (1994). The Penn State worry questionnaire: Psychometric properties and associated characteristics. In G. C. L. Davey \& F. Tallis (Eds.), \emph{Worrying: Perspectives on theory, assessment and treatment} (pp. 265--283). New York: Wiley.

\leavevmode\hypertarget{ref-noble2002}{}%
Noble, S. G. (2002). \emph{Turning images into simple line-art} (Undergraduate thesis). Reed College.

\leavevmode\hypertarget{ref-reedweb2007}{}%
Reed~College. (2007). LaTeX your document. Retrieved from \url{https://web.reed.edu/cis/help/LaTeX/index.html}

% Index?

\end{document}

### 2021-04-20 writing /thesis_draft/index.Rmd
i somewhat fixed the knitting problem but the formatting is still off

--- 2021-03-21 thesis draft/thesis.Rmd
@@
-<!--
-This is for including Chapter 1.
-Notice that it's also good practice to name your chunk.
-This will help you debug potential issues as you knit.
-The chunk above is called intro and the one below is called chapter1.
-Feel free to change the name of the Rmd file as you wish, but don't forget to change it here from chap1.Rmd.
--->
-
-<!--
-The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of Chapter 1.
-These labels will automatically be generated (if not specified) by changing the spaces to hyphens and capital letters to lowercase.
-Look for the reference to this label at the beginning of Chapter 2.
--->
-
-# R Markdown Basics {#rmd-basics}
-
-Here is a brief introduction into using _R Markdown_.
-_Markdown_ is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
-_R Markdown_ provides the flexibility of _Markdown_ with the implementation of **R** input and output.
-For more details on using _R Markdown_ see <https://rmarkdown.rstudio.com>.
-
-Be careful with your spacing in _Markdown_ documents.
-While whitespace largely is ignored, it does at times give _Markdown_ signals as to how to proceed.
-As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph.
-In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).
-
-## Lists
-
-It's easy to create a list.  It can be unordered like
-
-* Item 1
-* Item 2
-
-or it can be ordered like
-
-1. Item 1
-4. Item 2
-
-Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.
-
-To create a sublist, just indent the values a bit (at least four spaces or a tab).  (Here's one case where indentation is key!)
-
-1. Item 1
-1. Item 2
-1. Item 3
-    - Item 3a
-    - Item 3b
-
-## Line breaks
-
-Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:
-
-Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
-This should be a new paragraph.
-
-*Now for the correct way:*
-
-Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
-
-This should be a new paragraph.
-
-## R chunks
-
-When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document.
-You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):
-
-```{r cars}
-summary(cars)
-```
-
-## Inline code
-
-If you'd like to put the results of your analysis directly into your discussion, add inline code like this:
-
-> The `cos` of $2 \pi$ is `r cos(2*pi)`.
-
-Another example would be the direct calculation of the standard deviation:
-
-> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.
-
-One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:
-
-> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`
-
-Note the use of `>` here, which signifies a quotation environment that will be indented.
-
-As you see with `$2 \pi$` above, mathematics can be added by surrounding the mathematical text with dollar signs.
-More examples of this are in [Mathematics and Science] if you uncomment the code in [Math].
-
-## Including plots
-
-You can also embed plots. For example, here is a way to use the base **R** graphics package to produce a plot using the built-in `pressure` dataset:
-
-```{r pressure, echo=FALSE, cache=TRUE, fig.height=3, fig.width=5}
-plot(pressure)
-```
-
-Note that the `echo=FALSE` parameter was added to the code chunk to prevent printing of the **R** code that generated the plot.
-There are plenty of other ways to add chunk options (like `fig.height` and `fig.width` in the chunk above).
-More information is available at <https://yihui.org/knitr/options/>.
-
-Another useful chunk option is the setting of `cache=TRUE` as you see here.
-If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.
-Later in this file, you'll see a way to reference plots created in **R** or external figures.
-
-## Loading and exploring data
-
-Included in this template is a file called `flights.csv`.
-This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014.
-More information about this dataset and its **R** package is available at <https://github.com/ismayc/pnwflights14>.
-This subset includes only Portland flights and only rows that were complete with no missing values.
-Merges were also done with the `airports` and `airlines` data sets in the `pnwflights14` package to get more descriptive airport and airline names.
-
-We can load in this data set using the following commands:
-
-```{r load_data}
-# flights.csv is in the data directory
-flights_path <- here::here("data", "flights.csv")
-# string columns will be read in as strings and not factors now
-flights <- read.csv(flights_path, stringsAsFactors = FALSE)
-```
-
-The data is now stored in the data frame called `flights` in **R**.
-To get a better feel for the variables included in this dataset we can use a variety of functions.
-Here we can see the dimensions (rows by columns) and also the names of the columns.
-
-```{r str}
-dim(flights)
-names(flights)
-```
-
-Another good idea is to take a look at the dataset in table form.
-With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here.
-I recommend you enter the command into the Console **_after_** you have run the **R** chunks above to load the data into **R**.
-
-```{r view_flights, eval=FALSE}
-View(flights)
-```
-
-While not required, it is highly recommended you use the `dplyr` package to manipulate and summarize your data set as needed.
-It uses a syntax that is easy to understand using chaining operations.
-Below I've created a few examples of using `dplyr` to get information about the Portland flights in 2014.
-You will also see the use of the `ggplot2` package, which produces beautiful, high-quality academic visuals.
-
-We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:
-
-```{r load_pkgs, message=FALSE}
-# List of packages required for this analysis
-pkg <- c("dplyr", "ggplot2", "knitr", "bookdown")
-# Check if packages are not installed and assign the
-# names of the packages not installed to the variable new.pkg
-new.pkg <- pkg[!(pkg %in% installed.packages())]
-# If there are any packages in the list that aren't installed,
-# install them
-if (length(new.pkg)) {
-  install.packages(new.pkg, repos = "https://cran.rstudio.com")
-}
-# Load packages
-library(thesisdown)
-library(dplyr)
-library(ggplot2)
-library(knitr)
-```
-
-\clearpage
-
-The example we show here does the following:
-
-- Selects only the `carrier_name` and `arr_delay` from the `flights` dataset and then assigns this subset to a new variable called `flights2`.
-
-- Using `flights2`, we determine the largest arrival delay for each of the carriers.
-
-```{r max_delays}
-flights2 <- flights %>%
-  select(carrier_name, arr_delay)
-max_delays <- flights2 %>%
-  group_by(carrier_name) %>%
-  summarize(max_arr_delay = max(arr_delay, na.rm = TRUE))
-```
-
-A useful function in the `knitr` package for making nice tables in _R Markdown_ is called `kable`.
-It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.
-This again goes to show how nice reproducible documents can be!
-(Note the use of `results="asis"`, which will produce the table instead of the code to create the table.)  The `caption.short` argument is used to include a shorter title to appear in the List of Tables.
-
-```{r maxdelays, results="asis"}
-kable(max_delays,
-  col.names = c("Airline", "Max Arrival Delay"),
-  caption = "Maximum Delays by Airline",
-  caption.short = "Max Delays by Airline",
-  longtable = TRUE,
-  booktabs = TRUE
-)
-```
-
-The last two options make the table a little easier-to-read.
-
-We can further look into the properties of the largest value here for American Airlines Inc.
-To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original `flights` dataset.
-
-```{r max_props}
-flights %>%
-  filter(
-    arr_delay == 1539,
-    carrier_name == "American Airlines Inc."
-  ) %>%
-  select(-c(
-    month, day, carrier, dest_name, hour,
-    minute, carrier_name, arr_delay
-  ))
-```
-
-We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth.
-Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.
-
-```{r march3plot, fig.height=3, fig.width=6}
-flights %>%
-  filter(month == 3, day == 3) %>%
-  ggplot(aes(x = dep_time, y = arr_delay)) +
-  geom_point()
-```
-
-## Additional resources
-
-- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>
-
-- _R Markdown_
-    - Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf>
-
-- _RStudio IDE_
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf>
-    - Official website - <https://rstudio.com/products/rstudio/>
-
-- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html>
-
-- `ggplot2`
-    - Documentation - <https://ggplot2.tidyverse.org/>
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf>
-
-<!--chapter:end:01-chap1.Rmd-->
-
-# Mathematics and Science {#math-sci}
-
-<!-- Required to number equations in HTML files -->
-<script type="text/x-mathjax-config">
-MathJax.Hub.Config({
-  TeX: { equationNumbers: { autoNumber: "AMS" } }
-});
-</script>
-
-## Math
-
-\TeX\ is the best way to typeset mathematics.
-Donald Knuth designed \TeX\ when he got frustrated at how long it was taking the typesetters to finish his book, which contained a lot of mathematics.
-One nice feature of _R Markdown_ is its ability to read LaTeX code directly.
-
-If you are doing a thesis that will involve lots of math, you will want to read the following section which has been commented out.
-If you're not going to use math, skip over or delete this next commented section.
-
-<!-- MATH and PHYSICS majors: Uncomment the following section -->
-<!--
-$$\sum_{j=1}^n (\delta\theta_j)^2 \leq {{\beta_i^2}\over{\delta_i^2 + \rho_i^2}}
-\left[ 2\rho_i^2 + {\delta_i^2\beta_i^2\over{\delta_i^2 + \rho_i^2}} \right] \equiv \omega_i^2
-$$
-
-From Informational Dynamics, we have the following (Dave Braden):
-
-After _n_ such encounters the posterior density for $\theta$ is
-
-$$
-\pi(\theta|X_1< y_1,\dots,X_n<y_n) \varpropto \pi(\theta) \prod_{i=1}^n\int_{-\infty}^{y_i}
-   \exp\left(-{(x-\theta)^2\over{2\sigma^2}}\right)\ dx
-$$
-
-Another equation:
-
-$$\det\left|\,\begin{matrix}%
-c_0&c_1\hfill&c_2\hfill&\ldots&c_n\hfill\cr
-c_1&c_2\hfill&c_3\hfill&\ldots&c_{n+1}\hfill\cr
-c_2&c_3\hfill&c_4\hfill&\ldots&c_{n+2}\hfill\cr
-\,\vdots\hfill&\,\vdots\hfill&
-  \,\vdots\hfill&&\,\vdots\hfill\cr
-c_n&c_{n+1}\hfill&c_{n+2}\hfill&\ldots&c_{2n}\hfill\cr
-\end{matrix}\right|>0$$
-
-Lapidus and Pindar, Numerical Solution of Partial Differential Equations in Science and
-Engineering.  Page 54
-
-$$
-\int_t\left\{\sum_{j=1}^3 T_j \left({d\phi_j\over dt}+k\phi_j\right)-kT_e\right\}w_i(t)\ dt=0,
-   \qquad\quad i=1,2,3.
-$$
-
-L\&P  Galerkin method weighting functions.  Page 55
-
-$$
-\sum_{j=1}^3 T_j\int_0^1\left\{{d\phi_j\over dt} + k\phi_j\right\} \phi_i\ dt
-   = \int_{0}^1k\,T_e\phi_idt, \qquad i=1,2,3 $$
-
-Another L\&P (p145)
-
-$$
-\int_{-1}^1\!\int_{-1}^1\!\int_{-1}^1 f\big(\xi,\eta,\zeta\big)
-   = \sum_{k=1}^n\sum_{j=1}^n\sum_{i=1}^n w_i w_j w_k f\big( \xi,\eta,\zeta\big).
-$$
-
-Another L\&P (p126)
-
-$$
-\int_{A_e} (\,\cdot\,) dx dy = \int_{-1}^1\!\int_{-1}^1 (\,\cdot\,) \det[J] d\xi d\eta.
-$$
--->
-
-## Chemistry 101: Symbols
-
-Chemical formulas will look best if they are not italicized.
-Get around math mode's automatic italicizing in LaTeX by using the argument `$\mathrm{formula here}$`, with your formula inside the curly brackets.
-(Notice the use of the backticks here which enclose text that acts as code.)
-
-So, $\mathrm{Fe_2^{2+}Cr_2O_4}$ is written `$\mathrm{Fe_2^{2+}Cr_2O_4}$`.
-
-<!--
-The \noindent command below does what you'd expect:  it forces the current line/paragraph to not indent. This was done here to match the format of the LaTeX thesis PDF.
--->
-
-\noindent Exponent or Superscript: $\mathrm{O^-}$
-
-\noindent Subscript: $\mathrm{CH_4}$
-
-To stack numbers or letters as in $\mathrm{Fe_2^{2+}}$, the subscript is defined first, and then the superscript is defined.
-
-\noindent Bullet: CuCl $\bullet$ $\mathrm{7H_{2}O}$
-
-\noindent Delta: $\Delta$
-
-\noindent Reaction Arrows: $\longrightarrow$ or  $\xrightarrow{solution}$
-
-\noindent Resonance Arrows: $\leftrightarrow$
-
-\noindent Reversible Reaction Arrows: $\rightleftharpoons$
-
-### Typesetting reactions
-
-You may wish to put your reaction in an equation environment, which means that LaTeX will place the reaction where it fits and will number the equations for you.
-
-\begin{equation}
-  \mathrm{C_6H_{12}O_6  + 6O_2} \longrightarrow \mathrm{6CO_2 + 6H_2O}
-  (\#eq:reaction)
-\end{equation}
-
-We can reference this combustion of glucose reaction via Equation \@ref(eq:reaction).
-
-### Other examples of reactions
-
-$\mathrm{NH_4Cl_{(s)}}$ $\rightleftharpoons$ $\mathrm{NH_{3(g)}+HCl_{(g)}}$
-
-\noindent $\mathrm{MeCH_2Br + Mg}$ $\xrightarrow[below]{above}$ $\mathrm{MeCH_2\bullet Mg \bullet Br}$
-
-## Physics
-
-Many of the symbols you will need can be found on the math page <https://web.reed.edu/cis/help/latex/math.html> and the Comprehensive LaTeX Symbol Guide (<https://mirror.utexas.edu/ctan/info/symbols/comprehensive/symbols-letter.pdf>).
-
-## Biology
-
-You will probably find the resources at <https://www.lecb.ncifcrf.gov/~toms/latex.html> helpful, particularly the links to bsts for various journals.
-You may also be interested in TeXShade for nucleotide typesetting (<https://homepages.uni-tuebingen.de/beitz/txe.html>).
-Be sure to read the proceeding chapter on graphics and tables.
-
-<!--chapter:end:02-chap2.Rmd-->
-
-```{r include_packages_2, include = FALSE}
-# This chunk ensures that the thesisdown package is
-# installed and loaded. This thesisdown package includes
-# the template files for the thesis and also two functions
-# used for labeling and referencing
-if (!require(remotes)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("remotes", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("remotes")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(dplyr)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("dplyr", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("dplyr")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(ggplot2)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("ggplot2", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("ggplot2")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(bookdown)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("bookdown", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("bookdown")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(thesisdown)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    remotes::install_github("ismayc/thesisdown")
-  } else {
-    stop(
-      paste(
-        "You need to run",
-        'remotes::install_github("ismayc/thesisdown")',
-        "first in the Console."
-      )
-    )
-  }
-}
-library(thesisdown)
-library(dplyr)
-library(ggplot2)
-library(knitr)
-flights <- read.csv("data/flights.csv", stringsAsFactors = FALSE)
-```
-
-# Graphics, References, and Labels {#ref-labels}
-
-## Figures
-
-If your thesis has a lot of figures, _R Markdown_ might behave better for you than that other word processor.
-One perk is that it will automatically number the figures accordingly in each chapter.
-You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier.
-If you label your figures, you can move the figures around and _R Markdown_ will automatically adjust the numbering for you.
-No need for you to remember!
-So that you don't have to get too far into LaTeX to do this, a couple **R** functions have been created for you to assist.
-You'll see their use below.
-
-<!--
-One thing that may be annoying is the way _R Markdown_ handles "floats" like tables and figures (it's really \LaTeX's fault).
-\LaTeX\ will try to find the best place to put your object based on the text around it and until you're really, truly done writing you should just leave it where it lies.
-There are some optional arguments specified in the options parameter of the `label` function.
-If you need to shift your figure around, it might be good to look here on tweaking the options argument:  <https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions>
-
-If you need a graphic or tabular material to be part of the text, you can just put it inline. If you need it to appear in the list of figures or tables, it should be placed in a code chunk.
--->
-
-In the **R** chunk below, we will load in a picture stored as `reed.jpg` in our main directory.
-We then give it the caption of "Reed logo", the label of "reedlogo", and specify that this is a figure.
-Make note of the different **R** chunk options that are given in the R Markdown file (not shown in the knitted document).
-
-```{r reedlogo, fig.cap="Reed logo", out.width="0.2\\linewidth", fig.align="center"}
-include_graphics(path = "figure/reed.jpg")
-```
-
-<!-- Note the use of `out.width` as a chunk option here. The resulting
-image is 20% of what the linewidth is in LaTeX. You can also center
-the image using `fig.align="center"` as shown.-->
-
-Here is a reference to the Reed logo: Figure \@ref(fig:reedlogo).
-Note the use of the `fig:` code here.
-By naming the **R** chunk that contains the figure, we can then reference that figure later as done in the first sentence here.
-We can also specify the caption for the figure via the R chunk option `fig.cap`.
-
-\clearpage
-
-<!-- clearpage ends the page, and also dumps out all floats.
-  Floats are things like tables and figures. -->
-
-Below we will investigate how to save the output of an **R** plot and label it in a way similar to that done above.
-Recall the `flights` dataset from Chapter \@ref(rmd-basics).
-(Note that we've shown a different way to reference a section or chapter here.)  We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.
-
-```{r delaysboxplot, warnings=FALSE, messages=FALSE, fig.cap="Mean Delays by Airline", fig.width=6, fig.height=5}
-mean_delay_by_carrier <- flights %>%
-  group_by(carrier) %>%
-  summarize(mean_dep_delay = mean(dep_delay))
-ggplot(mean_delay_by_carrier, aes(x = carrier, y = mean_dep_delay)) +
-  geom_bar(position = "identity", stat = "identity", fill = "red")
-```
-
-Here is a reference to this image: Figure \@ref(fig:delaysboxplot).
-
-A table linking these carrier codes to airline names is available at <https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv>.
-
-\clearpage
-
-Next, we will explore the use of the `out.extra` chunk option, which can be used to shrink or expand an image loaded from a file by specifying `"scale= "`.
-Here we use the mathematical graph stored in the "subdivision.pdf" file.
-
-```{r subd, results="asis", echo=FALSE, fig.cap="Subdiv. graph", out.extra="scale=0.75"}
-include_graphics("figure/subdivision.pdf")
-```
-
-Here is a reference to this image: Figure \@ref(fig:subd).  Note that `echo=FALSE` is specified so that the **R** code is hidden in the document.
-
-**More Figure Stuff**
-
-Lastly, we will explore how to rotate and enlarge figures using the `out.extra` chunk option.  (Currently this only works in the PDF version of the book.)
-
-```{r subd2, results="asis", echo=FALSE, out.extra="angle=180, scale=1.1", fig.cap="A Larger Figure, Flipped Upside Down"}
-include_graphics("figure/subdivision.pdf")
-```
-
-As another example, here is a reference: Figure \@ref(fig:subd2).
-
-## Footnotes and Endnotes
-
-You might want to footnote something.
-^[footnote text] The footnote will be in a smaller font and placed appropriately.
-Endnotes work in much the same way.
-More information can be found about both on the CUS site or feel free to reach out to <data@reed.edu>.
-
-## Bibliographies
-
-Of course you will need to cite things, and you will probably accumulate an armful of sources.
-There are a variety of tools available for creating a bibliography database (stored with the .bib extension).
-In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero.
-The Reed librarians have created Zotero documentation at <https://libguides.reed.edu/citation/zotero>.
-In addition, a tutorial is available from Middlebury College at <https://sites.middlebury.edu/zoteromiddlebury/>.
-
-_R Markdown_ uses _pandoc_ (<https://pandoc.org/>) to build its bibliographies.
-One nice caveat of this is that you won't have to do a second compile to load in references as standard LaTeX requires.
-To cite references in your thesis (after creating your bibliography database), place the reference name inside square brackets and precede it by the "at" symbol.
-For example, here's a reference to a book about worrying:  [@Molina1994].
-This `Molina1994` entry appears in a file called `thesis.bib` in the `bib` folder.
-This bibliography database file was created by a program called BibTeX.
-You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the `bib` folder.
-
-For more information about BibTeX and bibliographies, see our CUS site (<https://web.reed.edu/cis/help/latex/index.html>)^[@reedweb2007].
-There are three pages on this topic:  _bibtex_ (which talks about using BibTeX, at <https://web.reed.edu/cis/help/latex/bibtex.html>), _bibtexstyles_ (about how to find and use the bibliography style that best suits your needs, at <https://web.reed.edu/cis/help/latex/bibtexstyles.html>) and _bibman_ (which covers how to make and maintain a bibliography by hand, without BibTeX, at <https://web.reed.edu/cis/help/latex/bibman.html>).
-The last page will not be useful unless you have only a few sources.
-
-If you look at the YAML header at the top of the main .Rmd file you can see that we can specify the style of the bibliography by referencing the appropriate csl file.
-You can download a variety of different style files at <https://www.zotero.org/styles>.
-Make sure to download the file into the csl folder.
-
-<!-- Fill the rest of the page with the content below for the PDF version. -->
-
-\vfill
-
-**Tips for Bibliographies**
-
-- Like with thesis formatting, the sooner you start compiling your bibliography for something as large as thesis, the better.
-Typing in source after source is mind-numbing enough; do you really want to do it for hours on end in late April?
-Think of it as procrastination.
-- The cite key (a citation's label) needs to be unique from the other entries.
-- When you have more than one author or editor, you need to separate each author's name by the word "and" e.g. `Author = {Noble, Sam and Youngberg, Jessica},`.
-- Bibliographies made using BibTeX (whether manually or using a manager) accept LaTeX markup, so you can italicize and add symbols as necessary.
-- To force capitalization in an article title or where all lowercase is generally used, bracket the capital letter in curly braces.
-- You can add a Reed Thesis citation^[@noble2002] option.
-The best way to do this is to use the phdthesis type of citation, and use the optional "type" field to enter "Reed thesis" or "Undergraduate thesis."
-
-## Anything else?
-
-If you'd like to see examples of other things in this template, please contact the Data @ Reed team (email <data@reed.edu>) with your suggestions.
-We love to see people using _R Markdown_ for their theses, and are happy to help.
-
-<!--chapter:end:03-chap3.Rmd-->
-
-# Conclusion {-}
-
-If we don't want Conclusion to have a chapter number next to it, we can add the `{-}` attribute.
-
-**More info**
-
-And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph.
-Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.
-
-<!--chapter:end:04-conclusion.Rmd-->
-
-`r if(knitr:::is_latex_output()) '\\appendix'`
-
-`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`
-
-# The First Appendix
-
-<!--
-If you feel it necessary to include an appendix, it goes here.
--->
-
-This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.
-
-**In the main Rmd file**
-
-```{r ref.label='include_packages', results='hide', echo = TRUE}
-```
-
-**In Chapter \@ref(ref-labels):**
-
-```{r ref.label='include_packages_2', results='hide', echo = TRUE}
-```
-
-# The Second Appendix, for Fun
-
-<!--chapter:end:05-appendix.Rmd-->
-
-<!--
-The bib chunk below must go last in this document according to how R Markdown renders.  More info is at http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
--->
-
-\backmatter
-
-<!-- 
-If you'd like to change the name of the bibliography to something else,
-delete "References" and replace it.
--->
-
-# References {-}
-<!--
-This manually sets the header for this unnumbered chapter.
--->
-\markboth{References}{References}
-<!--
-To remove the indentation of the first entry.
--->
-\noindent
-
-<!--
-To create a hanging indent and spacing between entries.  These three lines may need to be removed for styles that don't require the hanging indent.
--->
-
-\setlength{\parindent}{-0.20in}
-\setlength{\leftskip}{0.20in}
-\setlength{\parskip}{8pt}
-
-<!--
-This is just for testing with more citations for the bibliography at the end.
-Add other entries into the list here if you'd like them to appear in the bibliography even if they weren't explicitly cited in the document.
--->
-
----
-nocite: | 
-  @angel2000, @angel2001, @angel2002a
-...
-
-<!--chapter:end:99-references.Rmd-->
-
 ---
 title: 'My Final College Paper'
-author: 'Margaret Anne Slein'
+author: 'Margaret A. Slein'
 date: 'May 2021'
@@
 #altadvisor: 'Your Other Advisor'
-department: 'Biology'
+department: 'Mathematics'
 degree: 'Bachelor of Arts'
@@
 acknowledgements: |
-  I want to thank a few people.
+  The future is bright, the future is dark
+
+  Time stands still, yet flies like a lark
+
+  Memories endure and memories fade
+
+  People are rough, yet smooth like suede
+
+  The days seem long, the days seem short
+
+  It rains all year, until the sunshine retorts
+
+  Reed is an experience, Reed is a place
+
+  Reed is a journey of learning how to keep pace
+
+  Reed is a individual, Reed is a kollective
+
+  Reed is space for being reflective
+
+  Reed is the present, Reed is the past
+
+  Reed is the net for our future that we cast
+
+  To those from Reed
+
+  To those from before
+
+  To those from the outside
+
+  To those at the core
+
+  To those I still talk to
+
+  To those I talk to no more
+
+  To those still here today
+
+  To those who now soar
+
+  A few words of dear thanks….
+
+  Thank you for the laughs, the tears, the screams
+
+  Gratitude for the late nights, the dancing, the dreams
+
+  Thank you for the walks, the talks, and meals together
+
+  Many thanks for the shenanigans I’ll cherish forever
+
+  Thank you for the compassion, for the coffee and tea
+
+  Appreciation for the love, the hate, and learning to be
+
+  Thank you for the support, and most ardently, for embracing me
+
 dedication: |
@@
 Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).
-
-<!--chapter:end:thesis-draft.Rmd-->

### 2021-04-24 writing /thesis_draft/01-chap1.Rmd
starting get my act together to format this thesis

<!--
This is for including Chapter 1.  Notice that it's also good practice to name your chunk.  This will help you debug potential issues as you knit.  The chunk above is called intro and the one below is called chapter1.  Feel free to change the name of the Rmd file as you wish, but don't forget to change it here from chap1.Rmd.
-->

<!--
The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of Chapter 1.  These labels will automatically be generated (if not specified) by changing the spaces to hyphens and capital letters to lowercase.  Look for the reference to this label at the beginning of Chapter 2.
-->

# Background 
## Environmental variability
Environmental variability describes the dynamic nature of environmental conditions in the natural world and how organisms respond (cite). Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure X).

```{r, echo=FALSE, fig.cap="Figure 1. An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability. Shown is a comparison of how duration of variability compares across the same environmental factor", out.width = '90%'}
knitr::include_graphics("figures/figure1.png")
```

### Mean
Aggregating values together, whether it be temperature, light availability, precipitation, allows for a better understanding of trends and patterns as opposed to random events. The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time. It provides a departure point for what we expect the average value to be for all time. We can then expand additional information adjacent to the mean (i.e. range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk than just simply considering increases in global temperature (2014).

### Amplitude/Range 
We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience. This range can be short or long, for example, geographic location drastically influences the amount of sunlight an ecosystem receives. The Amazon Rainforest experiences less variation in light availability during the year than Denali National Park in Alaska, which experiences high variation in light availability between the summer and winter seasons. The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare et al 2013, 2012), as well as additional life history traits (size, speed, etc.).

### Duration 
The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability. Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure X). Short term fluctuations, like Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time, have differing effects on performance (Colinet et al 2018). While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e. heat waves, cold fronts etc.).

Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with organism, population, and community level processes. For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during that specific window of development (cite). Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (cite). When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (cite).

### Predictability 
FTR can also describe the predictability of the fluctuation period. Variability may not have evenly spaced intervals (i.e. heatwaves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time. Predictability can also attempt to describe stochastic vs. autocorrelated variation, however, that will be discussed in a different context below (Figure 4).

Link between temporal and spatial variability 
While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact. Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (cite). Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e. Diel vertical migration (DVM). In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (cite). However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e. an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (cite).

Though environmental variability can include many different forms, from fluctuations in resource availability (cite) to light (cite), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (cite).

## Thermal variability

### What is temperature variability?
Changes in temperature can be grouped into a couple different orientations, the first being the difference between the terms “fluctuating” and “variation” with respect to temperature. Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as “a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis” (2018). In any case, temperature variability describes changes in environmental temperature through time (cite).

### Why is it important to ecological processes?
Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (cite). Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of  ecology (MTE), though highly debated (Clarke et al 2006), which serves as an origin point for understanding how organisms respond to environmental conditions. MTE is described by the following equation:

$$
B = b_0M^{3/4}e^{-E/kT}
$$
Where $B$ is organismal metabolic rate, $b_0$ is a normalization constant unrelated to temperate or body size, $M$ is organismal mass, and $e^-E/kT$ is referred to as the “Boltzman factor”, where $E$ is activation energy, $k$ is the Boltzman constant and $T$ is temperature. Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate. We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates should be the same. However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clark and Johnston 1999, Clarke et al 2006). Others believe that MTE makes valid assumptions and is more flexible than opponents argue (Gillooly et al 2006).

MTE supports temperature impacting metabolic rate, which can have downstream effects on fitness characteristics. This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 2).

```{r, echo=FALSE, fig.cap="Figure 2. Schematic of thermal performance curve and it’s important components for understanding thermal performance. Adapted from (cite, cite)", out.width = '90%'}
knitr::include_graphics("figures/figure2.png")
```

The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range. CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms’ performance is negatively impacted (cite). Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.

### Jensen’s Inequality and its impacts on metabolic rate
TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (cite, Bernhardt et al 2018). However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e. 15-35 C in 5 C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance). These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.

Given the characteristically non-linear shape of TPCs, Jensen’s inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions. Jensen’s inequality describes the phenomenon that environmental variance can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999). Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time. It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019.

Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms. For instance, Amarasekare et al (year) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (cite 2012). Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (cite 2012).

### The relevancy of temperature variability in the modern context

```{r, echo=FALSE, fig.cap="Figure 3. DON’T FORGET THIS IS LOG SCALED!! Plot of generation time in days against fluctuation period in days, with yellow, red, and grey horizontal panels representing diurnal, autocorrelated, and stochastic fluctuation patterns. The blue, green, purple, and pink horizontal lines represent different examples of variety in organismal generation time. *Figure out how to overlap pngs of plankton for 1, drosophila for 10, beetle for 100, and fish for 4000*", out.width = '90%'}
knitr::include_graphics("figures/figure3.png")
```

Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate. Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance. Khelifa et al (2019) demonstrated that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.

```{r, echo=FALSE, fig.cap="Figure 4. Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F). Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.", out.width = '90%'}
knitr::include_graphics("figures/figure4.png")
```

However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 4). Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (2019). Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature. Vasseur and Yodzis (2004) outline how the frequency (1/period) of environmental fluctuations describes the “color of environmental noise.” In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods. Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (cite). Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.

Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level (cite):

* At the individual level, this usually comes in the form of reduced fitness or increased development rates (cite)
* At the population level, this usually comes in the form of increases to population saturation
* At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration

It however, less understood, is how variation patterns influence these characteristic responses at each level of organization. Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (cite).

## Objectives, driving questions, hypotheses

There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (cite). To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions:

1. Does variability pattern drive differences in response magnitude? (i.e. does autocorrelated variation have a larger impact on response than stochastic or diurnal?)
2. Does temperature variability drive differential response levels at different levels of biological organization?
3. How can the magnitude of response be influenced by additional factors (i.e. periodicity, age, size, thermal history, etc)?

We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.

## Lists

It's easy to create a list.  It can be unordered like

* Item 1
* Item 2

or it can be ordered like

1. Item 1
4. Item 2

Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a tab).  (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
    - Item 3a
    - Item 3b

## Line breaks

Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
This should be a new paragraph.

*Now for the correct way:*

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.

This should be a new paragraph.

## R chunks

When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document. You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):

```{r cars}
summary(cars)
```

## Inline code

If you'd like to put the results of your analysis directly into your discussion, add inline code like this:

> The `cos` of $2 \pi$ is `r cos(2*pi)`.

Another example would be the direct calculation of the standard deviation:

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.

One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be indented.

As you see with `$2 \pi$` above, mathematics can be added by surrounding the mathematical text with dollar signs.  More examples of this are in [Mathematics and Science] if you uncomment the code in [Math].

## Including plots

You can also embed plots. For example, here is a way to use the base **R** graphics package to produce a plot using the built-in `pressure` dataset:

```{r pressure, echo=FALSE, cache=TRUE, fig.height=3, fig.width=5}
plot(pressure)
```

Note that the `echo=FALSE` parameter was added to the code chunk to prevent printing of the **R** code that generated the plot. There are plenty of other ways to add chunk options (like `fig.height` and `fig.width` in the chunk above).  More information is available at <https://yihui.org/knitr/options/>.

Another useful chunk option is the setting of `cache=TRUE` as you see here.  If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.  Later in this file, you'll see a way to reference plots created in **R** or external figures.

## Loading and exploring data

Included in this template is a file called `flights.csv`.  This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014. More information about this dataset and its **R** package is available at <https://github.com/ismayc/pnwflights14>. This subset includes only Portland flights and only rows that were complete with no missing values. Merges were also done with the `airports` and `airlines` data sets in the `pnwflights14` package to get more descriptive airport and airline names.

We can load in this data set using the following commands:

```{r load_data}
# flights.csv is in the data directory
flights_path <- here::here("data", "flights.csv")
# string columns will be read in as strings and not factors now
flights <- read.csv(flights_path, stringsAsFactors = FALSE)
```

The data is now stored in the data frame called `flights` in **R**.  To get a better feel for the variables included in this dataset we can use a variety of functions. Here we can see the dimensions (rows by columns) and also the names of the columns.

```{r str}
dim(flights)
names(flights)
```

Another good idea is to take a look at the dataset in table form.  With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here. I recommend you enter the command into the Console **_after_** you have run the **R** chunks above to load the data into **R**.

```{r view_flights, eval=FALSE}
View(flights)
```

While not required, it is highly recommended you use the `dplyr` package to manipulate and summarize your data set as needed.  It uses a syntax that is easy to understand using chaining operations.  Below I've created a few examples of using `dplyr` to get information about the Portland flights in 2014.  You will also see the use of the `ggplot2` package, which produces beautiful, high-quality academic visuals.

We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:

```{r load_pkgs, message=FALSE}
# List of packages required for this analysis
pkg <- c("dplyr", "ggplot2", "knitr", "bookdown")
# Check if packages are not installed and assign the
# names of the packages not installed to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]
# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg)) {
  install.packages(new.pkg, repos = "https://cran.rstudio.com")
}
# Load packages
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
```

\clearpage

The example we show here does the following:

- Selects only the `carrier_name` and `arr_delay` from the `flights` dataset and then assigns this subset to a new variable called `flights2`.

- Using `flights2`, we determine the largest arrival delay for each of the carriers.

```{r max_delays}
flights2 <- flights %>%
  select(carrier_name, arr_delay)
max_delays <- flights2 %>%
  group_by(carrier_name) %>%
  summarize(max_arr_delay = max(arr_delay, na.rm = TRUE))
```

A useful function in the `knitr` package for making nice tables in _R Markdown_ is called `kable`.  It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.  This again goes to show how nice reproducible documents can be! (Note the use of `results="asis"`, which will produce the table instead of the code to create the table.)  The `caption.short` argument is used to include a shorter title to appear in the List of Tables.

```{r maxdelays, results="asis"}
kable(max_delays,
  col.names = c("Airline", "Max Arrival Delay"),
  caption = "Maximum Delays by Airline",
  caption.short = "Max Delays by Airline",
  longtable = TRUE,
  booktabs = TRUE
)
```

The last two options make the table a little easier-to-read.

We can further look into the properties of the largest value here for American Airlines Inc.  To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original `flights` dataset.

```{r max_props}
flights %>%
  filter(
    arr_delay == 1539,
    carrier_name == "American Airlines Inc."
  ) %>%
  select(-c(
    month, day, carrier, dest_name, hour,
    minute, carrier_name, arr_delay
  ))
```

We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth.  Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.

```{r march3plot, fig.height=3, fig.width=6}
flights %>%
  filter(month == 3, day == 3) %>%
  ggplot(aes(x = dep_time, y = arr_delay)) +
  geom_point()
```

## Additional resources

- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>

- _R Markdown_
    - Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf>

- _RStudio IDE_
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf>
    - Official website - <https://rstudio.com/products/rstudio/>

- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html>

- `ggplot2`
    - Documentation - <https://ggplot2.tidyverse.org/>
    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf>

### 2021-04-24 writing /thesis_draft/02-chap2.Rmd
starting get my act together to format this thesis

--- 2021-03-21 thesis draft/thesis.Rmd
@@
-<!--
-This is for including Chapter 1.
-Notice that it's also good practice to name your chunk.
-This will help you debug potential issues as you knit.
-The chunk above is called intro and the one below is called chapter1.
-Feel free to change the name of the Rmd file as you wish, but don't forget to change it here from chap1.Rmd.
--->
-
-<!--
-The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of Chapter 1.
-These labels will automatically be generated (if not specified) by changing the spaces to hyphens and capital letters to lowercase.
-Look for the reference to this label at the beginning of Chapter 2.
--->
-
-# R Markdown Basics {#rmd-basics}
-
-Here is a brief introduction into using _R Markdown_.
-_Markdown_ is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
-_R Markdown_ provides the flexibility of _Markdown_ with the implementation of **R** input and output.
-For more details on using _R Markdown_ see <https://rmarkdown.rstudio.com>.
-
-Be careful with your spacing in _Markdown_ documents.
-While whitespace largely is ignored, it does at times give _Markdown_ signals as to how to proceed.
-As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph.
-In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).
-
-## Lists
-
-It's easy to create a list.  It can be unordered like
-
-* Item 1
-* Item 2
-
-or it can be ordered like
-
-1. Item 1
-4. Item 2
-
-Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.
-
-To create a sublist, just indent the values a bit (at least four spaces or a tab).  (Here's one case where indentation is key!)
-
-1. Item 1
-1. Item 2
-1. Item 3
-    - Item 3a
-    - Item 3b
-
-## Line breaks
-
-Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:
-
-Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
-This should be a new paragraph.
-
-*Now for the correct way:*
-
-Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
-
-This should be a new paragraph.
-
-## R chunks
-
-When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document.
-You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):
-
-```{r cars}
-summary(cars)
-```
-
-## Inline code
-
-If you'd like to put the results of your analysis directly into your discussion, add inline code like this:
-
-> The `cos` of $2 \pi$ is `r cos(2*pi)`.
-
-Another example would be the direct calculation of the standard deviation:
-
-> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.
-
-One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:
-
-> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`
-
-Note the use of `>` here, which signifies a quotation environment that will be indented.
-
-As you see with `$2 \pi$` above, mathematics can be added by surrounding the mathematical text with dollar signs.
-More examples of this are in [Mathematics and Science] if you uncomment the code in [Math].
-
-## Including plots
-
-You can also embed plots. For example, here is a way to use the base **R** graphics package to produce a plot using the built-in `pressure` dataset:
-
-```{r pressure, echo=FALSE, cache=TRUE, fig.height=3, fig.width=5}
-plot(pressure)
-```
-
-Note that the `echo=FALSE` parameter was added to the code chunk to prevent printing of the **R** code that generated the plot.
-There are plenty of other ways to add chunk options (like `fig.height` and `fig.width` in the chunk above).
-More information is available at <https://yihui.org/knitr/options/>.
-
-Another useful chunk option is the setting of `cache=TRUE` as you see here.
-If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.
-Later in this file, you'll see a way to reference plots created in **R** or external figures.
-
-## Loading and exploring data
-
-Included in this template is a file called `flights.csv`.
-This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014.
-More information about this dataset and its **R** package is available at <https://github.com/ismayc/pnwflights14>.
-This subset includes only Portland flights and only rows that were complete with no missing values.
-Merges were also done with the `airports` and `airlines` data sets in the `pnwflights14` package to get more descriptive airport and airline names.
-
-We can load in this data set using the following commands:
-
-```{r load_data}
-# flights.csv is in the data directory
-flights_path <- here::here("data", "flights.csv")
-# string columns will be read in as strings and not factors now
-flights <- read.csv(flights_path, stringsAsFactors = FALSE)
-```
-
-The data is now stored in the data frame called `flights` in **R**.
-To get a better feel for the variables included in this dataset we can use a variety of functions.
-Here we can see the dimensions (rows by columns) and also the names of the columns.
-
-```{r str}
-dim(flights)
-names(flights)
-```
-
-Another good idea is to take a look at the dataset in table form.
-With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here.
-I recommend you enter the command into the Console **_after_** you have run the **R** chunks above to load the data into **R**.
-
-```{r view_flights, eval=FALSE}
-View(flights)
-```
-
-While not required, it is highly recommended you use the `dplyr` package to manipulate and summarize your data set as needed.
-It uses a syntax that is easy to understand using chaining operations.
-Below I've created a few examples of using `dplyr` to get information about the Portland flights in 2014.
-You will also see the use of the `ggplot2` package, which produces beautiful, high-quality academic visuals.
-
-We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:
-
-```{r load_pkgs, message=FALSE}
-# List of packages required for this analysis
-pkg <- c("dplyr", "ggplot2", "knitr", "bookdown")
-# Check if packages are not installed and assign the
-# names of the packages not installed to the variable new.pkg
-new.pkg <- pkg[!(pkg %in% installed.packages())]
-# If there are any packages in the list that aren't installed,
-# install them
-if (length(new.pkg)) {
-  install.packages(new.pkg, repos = "https://cran.rstudio.com")
-}
-# Load packages
-library(thesisdown)
-library(dplyr)
-library(ggplot2)
-library(knitr)
-```
-
-\clearpage
-
-The example we show here does the following:
-
-- Selects only the `carrier_name` and `arr_delay` from the `flights` dataset and then assigns this subset to a new variable called `flights2`.
-
-- Using `flights2`, we determine the largest arrival delay for each of the carriers.
-
-```{r max_delays}
-flights2 <- flights %>%
-  select(carrier_name, arr_delay)
-max_delays <- flights2 %>%
-  group_by(carrier_name) %>%
-  summarize(max_arr_delay = max(arr_delay, na.rm = TRUE))
-```
-
-A useful function in the `knitr` package for making nice tables in _R Markdown_ is called `kable`.
-It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.
-This again goes to show how nice reproducible documents can be!
-(Note the use of `results="asis"`, which will produce the table instead of the code to create the table.)  The `caption.short` argument is used to include a shorter title to appear in the List of Tables.
-
-```{r maxdelays, results="asis"}
-kable(max_delays,
-  col.names = c("Airline", "Max Arrival Delay"),
-  caption = "Maximum Delays by Airline",
-  caption.short = "Max Delays by Airline",
-  longtable = TRUE,
-  booktabs = TRUE
-)
-```
-
-The last two options make the table a little easier-to-read.
-
-We can further look into the properties of the largest value here for American Airlines Inc.
-To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original `flights` dataset.
-
-```{r max_props}
-flights %>%
-  filter(
-    arr_delay == 1539,
-    carrier_name == "American Airlines Inc."
-  ) %>%
-  select(-c(
-    month, day, carrier, dest_name, hour,
-    minute, carrier_name, arr_delay
-  ))
-```
-
-We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth.
-Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.
-
-```{r march3plot, fig.height=3, fig.width=6}
-flights %>%
-  filter(month == 3, day == 3) %>%
-  ggplot(aes(x = dep_time, y = arr_delay)) +
-  geom_point()
-```
-
-## Additional resources
-
-- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>
-
-- _R Markdown_
-    - Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf>
-
-- _RStudio IDE_
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf>
-    - Official website - <https://rstudio.com/products/rstudio/>
-
-- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html>
-
-- `ggplot2`
-    - Documentation - <https://ggplot2.tidyverse.org/>
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf>
-
-<!--chapter:end:01-chap1.Rmd-->
-
-# Mathematics and Science {#math-sci}
-
-<!-- Required to number equations in HTML files -->
-<script type="text/x-mathjax-config">
-MathJax.Hub.Config({
-  TeX: { equationNumbers: { autoNumber: "AMS" } }
-});
-</script>
+# Methods
 
@@
 Be sure to read the proceeding chapter on graphics and tables.
-
-<!--chapter:end:02-chap2.Rmd-->
-
-```{r include_packages_2, include = FALSE}
-# This chunk ensures that the thesisdown package is
-# installed and loaded. This thesisdown package includes
-# the template files for the thesis and also two functions
-# used for labeling and referencing
-if (!require(remotes)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("remotes", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("remotes")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(dplyr)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("dplyr", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("dplyr")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(ggplot2)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("ggplot2", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("ggplot2")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(bookdown)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("bookdown", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste(
-        'You need to run install.packages("bookdown")',
-        "first in the Console."
-      )
-    )
-  }
-}
-if (!require(thesisdown)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    remotes::install_github("ismayc/thesisdown")
-  } else {
-    stop(
-      paste(
-        "You need to run",
-        'remotes::install_github("ismayc/thesisdown")',
-        "first in the Console."
-      )
-    )
-  }
-}
-library(thesisdown)
-library(dplyr)
-library(ggplot2)
-library(knitr)
-flights <- read.csv("data/flights.csv", stringsAsFactors = FALSE)
-```
-
-# Graphics, References, and Labels {#ref-labels}
-
-## Figures
-
-If your thesis has a lot of figures, _R Markdown_ might behave better for you than that other word processor.
-One perk is that it will automatically number the figures accordingly in each chapter.
-You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier.
-If you label your figures, you can move the figures around and _R Markdown_ will automatically adjust the numbering for you.
-No need for you to remember!
-So that you don't have to get too far into LaTeX to do this, a couple **R** functions have been created for you to assist.
-You'll see their use below.
-
-<!--
-One thing that may be annoying is the way _R Markdown_ handles "floats" like tables and figures (it's really \LaTeX's fault).
-\LaTeX\ will try to find the best place to put your object based on the text around it and until you're really, truly done writing you should just leave it where it lies.
-There are some optional arguments specified in the options parameter of the `label` function.
-If you need to shift your figure around, it might be good to look here on tweaking the options argument:  <https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions>
-
-If you need a graphic or tabular material to be part of the text, you can just put it inline. If you need it to appear in the list of figures or tables, it should be placed in a code chunk.
--->
-
-In the **R** chunk below, we will load in a picture stored as `reed.jpg` in our main directory.
-We then give it the caption of "Reed logo", the label of "reedlogo", and specify that this is a figure.
-Make note of the different **R** chunk options that are given in the R Markdown file (not shown in the knitted document).
-
-```{r reedlogo, fig.cap="Reed logo", out.width="0.2\\linewidth", fig.align="center"}
-include_graphics(path = "figure/reed.jpg")
-```
-
-<!-- Note the use of `out.width` as a chunk option here. The resulting
-image is 20% of what the linewidth is in LaTeX. You can also center
-the image using `fig.align="center"` as shown.-->
-
-Here is a reference to the Reed logo: Figure \@ref(fig:reedlogo).
-Note the use of the `fig:` code here.
-By naming the **R** chunk that contains the figure, we can then reference that figure later as done in the first sentence here.
-We can also specify the caption for the figure via the R chunk option `fig.cap`.
-
-\clearpage
-
-<!-- clearpage ends the page, and also dumps out all floats.
-  Floats are things like tables and figures. -->
-
-Below we will investigate how to save the output of an **R** plot and label it in a way similar to that done above.
-Recall the `flights` dataset from Chapter \@ref(rmd-basics).
-(Note that we've shown a different way to reference a section or chapter here.)  We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.
-
-```{r delaysboxplot, warnings=FALSE, messages=FALSE, fig.cap="Mean Delays by Airline", fig.width=6, fig.height=5}
-mean_delay_by_carrier <- flights %>%
-  group_by(carrier) %>%
-  summarize(mean_dep_delay = mean(dep_delay))
-ggplot(mean_delay_by_carrier, aes(x = carrier, y = mean_dep_delay)) +
-  geom_bar(position = "identity", stat = "identity", fill = "red")
-```
-
-Here is a reference to this image: Figure \@ref(fig:delaysboxplot).
-
-A table linking these carrier codes to airline names is available at <https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv>.
-
-\clearpage
-
-Next, we will explore the use of the `out.extra` chunk option, which can be used to shrink or expand an image loaded from a file by specifying `"scale= "`.
-Here we use the mathematical graph stored in the "subdivision.pdf" file.
-
-```{r subd, results="asis", echo=FALSE, fig.cap="Subdiv. graph", out.extra="scale=0.75"}
-include_graphics("figure/subdivision.pdf")
-```
-
-Here is a reference to this image: Figure \@ref(fig:subd).  Note that `echo=FALSE` is specified so that the **R** code is hidden in the document.
-
-**More Figure Stuff**
-
-Lastly, we will explore how to rotate and enlarge figures using the `out.extra` chunk option.  (Currently this only works in the PDF version of the book.)
-
-```{r subd2, results="asis", echo=FALSE, out.extra="angle=180, scale=1.1", fig.cap="A Larger Figure, Flipped Upside Down"}
-include_graphics("figure/subdivision.pdf")
-```
-
-As another example, here is a reference: Figure \@ref(fig:subd2).
-
-## Footnotes and Endnotes
-
-You might want to footnote something.
-^[footnote text] The footnote will be in a smaller font and placed appropriately.
-Endnotes work in much the same way.
-More information can be found about both on the CUS site or feel free to reach out to <data@reed.edu>.
-
-## Bibliographies
-
-Of course you will need to cite things, and you will probably accumulate an armful of sources.
-There are a variety of tools available for creating a bibliography database (stored with the .bib extension).
-In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero.
-The Reed librarians have created Zotero documentation at <https://libguides.reed.edu/citation/zotero>.
-In addition, a tutorial is available from Middlebury College at <https://sites.middlebury.edu/zoteromiddlebury/>.
-
-_R Markdown_ uses _pandoc_ (<https://pandoc.org/>) to build its bibliographies.
-One nice caveat of this is that you won't have to do a second compile to load in references as standard LaTeX requires.
-To cite references in your thesis (after creating your bibliography database), place the reference name inside square brackets and precede it by the "at" symbol.
-For example, here's a reference to a book about worrying:  [@Molina1994].
-This `Molina1994` entry appears in a file called `thesis.bib` in the `bib` folder.
-This bibliography database file was created by a program called BibTeX.
-You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the `bib` folder.
-
-For more information about BibTeX and bibliographies, see our CUS site (<https://web.reed.edu/cis/help/latex/index.html>)^[@reedweb2007].
-There are three pages on this topic:  _bibtex_ (which talks about using BibTeX, at <https://web.reed.edu/cis/help/latex/bibtex.html>), _bibtexstyles_ (about how to find and use the bibliography style that best suits your needs, at <https://web.reed.edu/cis/help/latex/bibtexstyles.html>) and _bibman_ (which covers how to make and maintain a bibliography by hand, without BibTeX, at <https://web.reed.edu/cis/help/latex/bibman.html>).
-The last page will not be useful unless you have only a few sources.
-
-If you look at the YAML header at the top of the main .Rmd file you can see that we can specify the style of the bibliography by referencing the appropriate csl file.
-You can download a variety of different style files at <https://www.zotero.org/styles>.
-Make sure to download the file into the csl folder.
-
-<!-- Fill the rest of the page with the content below for the PDF version. -->
-
-\vfill
-
-**Tips for Bibliographies**
-
-- Like with thesis formatting, the sooner you start compiling your bibliography for something as large as thesis, the better.
-Typing in source after source is mind-numbing enough; do you really want to do it for hours on end in late April?
-Think of it as procrastination.
-- The cite key (a citation's label) needs to be unique from the other entries.
-- When you have more than one author or editor, you need to separate each author's name by the word "and" e.g. `Author = {Noble, Sam and Youngberg, Jessica},`.
-- Bibliographies made using BibTeX (whether manually or using a manager) accept LaTeX markup, so you can italicize and add symbols as necessary.
-- To force capitalization in an article title or where all lowercase is generally used, bracket the capital letter in curly braces.
-- You can add a Reed Thesis citation^[@noble2002] option.
-The best way to do this is to use the phdthesis type of citation, and use the optional "type" field to enter "Reed thesis" or "Undergraduate thesis."
-
-## Anything else?
-
-If you'd like to see examples of other things in this template, please contact the Data @ Reed team (email <data@reed.edu>) with your suggestions.
-We love to see people using _R Markdown_ for their theses, and are happy to help.
-
-<!--chapter:end:03-chap3.Rmd-->
-
-# Conclusion {-}
-
-If we don't want Conclusion to have a chapter number next to it, we can add the `{-}` attribute.
-
-**More info**
-
-And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph.
-Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.
-
-<!--chapter:end:04-conclusion.Rmd-->
-
-`r if(knitr:::is_latex_output()) '\\appendix'`
-
-`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`
-
-# The First Appendix
-
-<!--
-If you feel it necessary to include an appendix, it goes here.
--->
-
-This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.
-
-**In the main Rmd file**
-
-```{r ref.label='include_packages', results='hide', echo = TRUE}
-```
-
-**In Chapter \@ref(ref-labels):**
-
-```{r ref.label='include_packages_2', results='hide', echo = TRUE}
-```
-
-# The Second Appendix, for Fun
-
-<!--chapter:end:05-appendix.Rmd-->
-
-<!--
-The bib chunk below must go last in this document according to how R Markdown renders.  More info is at http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
--->
-
-\backmatter
-
-<!-- 
-If you'd like to change the name of the bibliography to something else,
-delete "References" and replace it.
--->
-
-# References {-}
-<!--
-This manually sets the header for this unnumbered chapter.
--->
-\markboth{References}{References}
-<!--
-To remove the indentation of the first entry.
--->
-\noindent
-
-<!--
-To create a hanging indent and spacing between entries.  These three lines may need to be removed for styles that don't require the hanging indent.
--->
-
-\setlength{\parindent}{-0.20in}
-\setlength{\leftskip}{0.20in}
-\setlength{\parskip}{8pt}
-
-<!--
-This is just for testing with more citations for the bibliography at the end.
-Add other entries into the list here if you'd like them to appear in the bibliography even if they weren't explicitly cited in the document.
--->
-
----
-nocite: | 
-  @angel2000, @angel2001, @angel2002a
-...
-
-<!--chapter:end:99-references.Rmd-->
-
----
-title: 'My Final College Paper'
-author: 'Margaret Anne Slein'
-date: 'May 2021'
-institution: 'Reed College'
-division: 'Mathematics and Natural Sciences'
-advisor: 'Sam Fey'
-# If you have more two advisors, un-silence line 7
-#altadvisor: 'Your Other Advisor'
-department: 'Biology'
-degree: 'Bachelor of Arts'
-knit: bookdown::render_book
-site: bookdown::bookdown_site
-
-# This will automatically install the {remotes} package and {thesisdown}
-# Change this to FALSE if you'd like to install them manually on your own.
-params:
-  'Install needed packages for {thesisdown}': True
-
-# Remove the hashtag to specify which version of output you would like.
-# Can only choose one at a time.
-output:
-  thesisdown::thesis_pdf: default 
-#  thesisdown::thesis_gitbook: default         
-#  thesisdown::thesis_word: default
-#  thesisdown::thesis_epub: default
-
-# If you are creating a PDF you'll need to write your preliminary content 
-# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
-# for the .RMD files. If you are NOT producing a PDF, delete or silence
-# lines 25-39 in this YAML header.
-abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
-# If you'd rather include the preliminary content in files instead of inline
-# like below, use a command like that for the abstract above.  Note that a tab 
-# is needed on the line after the `|`.
-acknowledgements: |
-  I want to thank a few people.
-dedication: |
-  You can have a dedication here if you wish. 
-preface: |
-  This is an example of a thesis setup to use the reed thesis document class 
-  (for LaTeX) and the R bookdown package, in general.
-
-# Specify the location of the bibliography below
-bibliography: bib/thesis.bib
-# Download your specific csl file and refer to it in the line below.
-csl: csl/apa.csl
-lot: true
-lof: true
----
-
-<!--
-Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
-metadata used to produce the document.  Be careful with spacing in this header!
-
-If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line).
-
-If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.
-
-If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.
-
-If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
--->
-
-```{r include_packages, include=FALSE}
-# This chunk ensures that the thesisdown package is
-# installed and loaded. This thesisdown package includes
-# the template files for the thesis.
-if (!require(remotes)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    install.packages("remotes", repos = "https://cran.rstudio.com")
-  } else {
-    stop(
-      paste('You need to run install.packages("remotes")",
-            "first in the Console.')
-    )
-  }
-}
-if (!require(thesisdown)) {
-  if (params$`Install needed packages for {thesisdown}`) {
-    remotes::install_github("ismayc/thesisdown")
-  } else {
-    stop(
-      paste(
-        "You need to run",
-        'remotes::install_github("ismayc/thesisdown")',
-        "first in the Console."
-      )
-    )
-  }
-}
-library(thesisdown)
-# Set how wide the R output will go
-options(width = 70)
-```
-
-<!--
-The acknowledgments, preface, dedication, and abstract are added into the PDF
-version automatically by inputting them in the YAML at the top of this file.
-Alternatively, you can put that content in files like 00--prelim.Rmd and
-00-abstract.Rmd like done below.
--->
-
-```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}
-
-```
-
-```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}
-
-```
-
-<!-- The {.unnumbered} option here means that the introduction will be 
-"Chapter 0." You can also use {-} for no numbers on chapters.
--->
-
-# Introduction {.unnumbered}
-
-Welcome to the _R Markdown_ thesis template.
-This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before.
-Using _R Markdown_ will also allow you to easily keep track of your analyses in **R** chunks of code, with the resulting plots and output included as well.
-The hope is this _R Markdown_ template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.
-
-Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis.
-The use of LaTeX in combination with _Markdown_ is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file.
-While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities.
-After working with _Markdown_ and **R** together for a few weeks, we are confident this will be your reporting style of choice going forward.
-
-<!-- 
-If you're still on the fence about using _R Markdown_, check out the resource
-for newbies available at <https://rbasics.netlify.com>. 
--->
-
-**Why use it?**
-
-_R Markdown_ creates a simple and straightforward way to interface with the beauty of LaTeX.
-Packages have been written in **R** to work directly with LaTeX to produce nicely formatting tables and paragraphs.
-In addition to creating a user friendly interface to LaTeX, _R Markdown_ also allows you to read in your data, to analyze it and to visualize it using **R** functions, and also to provide the documentation and commentary on the results of your project.
-Further, it allows for **R** results to be passed inline to the commentary of your results.
-You'll see more on this later.
-
-<!-- 
-Having your code and commentary all together in one place has a plethora of 
-benefits!
--->
-
-**Who should use it?**
-
-Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use _R Markdown_.
-Of particular use should be anyone in the sciences, but the user-friendly nature of _Markdown_ and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc.
-should make it of great benefit to nearly anyone writing a thesis project.
-
-**For additional help with bookdown**
-
-Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).
-
-<!--chapter:end:thesis-draft.Rmd-->

### 2021-04-24 writing /thesis_draft/_book/thesis.tex
starting get my act together to format this thesis

@@
 %%% Remember to use the correct department!
-\department{Mathematics}
+\department{Biology}
 % if you're writing a thesis in an interdisciplinary major,
@@
 \Preface{
-This is an example of a thesis setup to use the reed thesis document class
-(for LaTeX) and the R bookdown package, in general.
+
 }
@@
   \end{acknowledgements}
-  \begin{preface}
-    This is an example of a thesis setup to use the reed thesis document class
-    (for LaTeX) and the R bookdown package, in general.
-  \end{preface}
+
   \hypersetup{linkcolor=black}
@@
 
-Welcome to the \emph{R Markdown} thesis template.
-This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before.
-Using \emph{R Markdown} will also allow you to easily keep track of your analyses in \textbf{R} chunks of code, with the resulting plots and output included as well.
-The hope is this \emph{R Markdown} template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.
+Placeholder
 
-Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis.
-The use of LaTeX in combination with \emph{Markdown} is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file.
-While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities.
-After working with \emph{Markdown} and \textbf{R} together for a few weeks, we are confident this will be your reporting style of choice going forward.
+\hypertarget{the-historical-significance-of-variability}{%
+\section{The historical significance of variability}\label{the-historical-significance-of-variability}}
 
-\textbf{Why use it?}
+\hypertarget{relevant-significance-of-variability-today}{%
+\section{Relevant significance of variability today}\label{relevant-significance-of-variability-today}}
 
-\emph{R Markdown} creates a simple and straightforward way to interface with the beauty of LaTeX.
-Packages have been written in \textbf{R} to work directly with LaTeX to produce nicely formatting tables and paragraphs.
-In addition to creating a user friendly interface to LaTeX, \emph{R Markdown} also allows you to read in your data, to analyze it and to visualize it using \textbf{R} functions, and also to provide the documentation and commentary on the results of your project.
-Further, it allows for \textbf{R} results to be passed inline to the commentary of your results.
-You'll see more on this later.
+\hypertarget{why-this-study-on-variability-matters}{%
+\section{Why this study on variability matters}\label{why-this-study-on-variability-matters}}
 
-\textbf{Who should use it?}
+\hypertarget{background}{%
+\chapter{Background}\label{background}}
 
-Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use \emph{R Markdown}.
-Of particular use should be anyone in the sciences, but the user-friendly nature of \emph{Markdown} and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc.
-should make it of great benefit to nearly anyone writing a thesis project.
+\hypertarget{environmental-variability}{%
+\section{Environmental variability}\label{environmental-variability}}
 
-\textbf{For additional help with bookdown}
+Environmental variability describes the dynamic nature of environmental conditions in the natural world and how organisms respond (cite).
+Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure X).
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figures/figure1} \caption{Figure 1.
+An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
+Shown is a comparison of how duration of variability compares across the same environmental factor}\label{fig:unnamed-chunk-1}
+\end{figure}
+\hypertarget{mean}{%
+\subsection{Mean}\label{mean}}
 
-Please visit \href{https://bookdown.org/yihui/bookdown/}{the free online bookdown reference guide}.
+Aggregating values together, whether it be temperature, light availability, precipitation, allows for a better understanding of trends and patterns as opposed to random events.
+The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time.
+It provides a departure point for what we expect the average value to be for all time.
+We can then expand additional information adjacent to the mean (i.e.~range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk than just simply considering increases in global temperature (2014).
 
-\hypertarget{rmd-basics}{%
-\chapter{R Markdown Basics}\label{rmd-basics}}
+\hypertarget{amplituderange}{%
+\subsection{Amplitude/Range}\label{amplituderange}}
 
-Here is a brief introduction into using \emph{R Markdown}.
-\emph{Markdown} is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
-\emph{R Markdown} provides the flexibility of \emph{Markdown} with the implementation of \textbf{R} input and output.
-For more details on using \emph{R Markdown} see \url{https://rmarkdown.rstudio.com}.
+We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
+This range can be short or long, for example, geographic location drastically influences the amount of sunlight an ecosystem receives.
+The Amazon Rainforest experiences less variation in light availability during the year than Denali National Park in Alaska, which experiences high variation in light availability between the summer and winter seasons.
+The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare et al 2013, 2012), as well as additional life history traits (size, speed, etc.).
 
-Be careful with your spacing in \emph{Markdown} documents.
-While whitespace largely is ignored, it does at times give \emph{Markdown} signals as to how to proceed.
-As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph.
-In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).
+\hypertarget{duration}{%
+\subsection{Duration}\label{duration}}
+
+The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability.
+Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure X).
+Short term fluctuations, like Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time, have differing effects on performance (Colinet et al 2018).
+While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.~heat waves, cold fronts etc.).
+
+Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with organism, population, and community level processes.
+For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during that specific window of development (cite).
+Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (cite).
+When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (cite).
+
+\hypertarget{predictability}{%
+\subsection{Predictability}\label{predictability}}
+
+FTR can also describe the predictability of the fluctuation period.
+Variability may not have evenly spaced intervals (i.e.~heatwaves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
+Predictability can also attempt to describe stochastic vs.~autocorrelated variation, however, that will be discussed in a different context below (Figure 4).
+
+Link between temporal and spatial variability
+While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact.
+Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (cite).
+Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e.~Diel vertical migration (DVM).
+In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (cite).
+However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (cite).
+
+Though environmental variability can include many different forms, from fluctuations in resource availability (cite) to light (cite), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (cite).
+
+\hypertarget{thermal-variability}{%
+\section{Thermal variability}\label{thermal-variability}}
+
+\hypertarget{what-is-temperature-variability}{%
+\subsection{What is temperature variability?}\label{what-is-temperature-variability}}
+
+Changes in temperature can be grouped into a couple different orientations, the first being the difference between the terms ``fluctuating'' and ``variation'' with respect to temperature.
+Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as ``a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis'' (2018).
+In any case, temperature variability describes changes in environmental temperature through time (cite).
+
+\hypertarget{why-is-it-important-to-ecological-processes}{%
+\subsection{Why is it important to ecological processes?}\label{why-is-it-important-to-ecological-processes}}
+
+Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (cite).
+Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE), though highly debated (Clarke et al 2006), which serves as an origin point for understanding how organisms respond to environmental conditions.
+MTE is described by the following equation:
+
+\[
+B = b_0M^{3/4}e^{-E/kT}
+\]
+Where \(B\) is organismal metabolic rate, \(b_0\) is a normalization constant unrelated to temperate or body size, \(M\) is organismal mass, and \(e^-E/kT\) is referred to as the ``Boltzman factor'', where \(E\) is activation energy, \(k\) is the Boltzman constant and \(T\) is temperature.
+Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate.
+We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates should be the same.
+However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clark and Johnston 1999, Clarke et al 2006).
+Others believe that MTE makes valid assumptions and is more flexible than opponents argue (Gillooly et al 2006).
+
+MTE supports temperature impacting metabolic rate, which can have downstream effects on fitness characteristics.
+This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 2).
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figures/figure2} \caption{Figure 2.
+Schematic of thermal performance curve and it’s important components for understanding thermal performance.
+Adapted from (cite, cite)}\label{fig:unnamed-chunk-2}
+\end{figure}
+The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms' performance is negatively impacted (cite).
+Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
+
+\hypertarget{jensens-inequality-and-its-impacts-on-metabolic-rate}{%
+\subsection{Jensen's Inequality and its impacts on metabolic rate}\label{jensens-inequality-and-its-impacts-on-metabolic-rate}}
+
+TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (cite, Bernhardt et al 2018).
+However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.~15-35 C in 5 C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
+These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.
+
+Given the characteristically non-linear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
+Jensen's inequality describes the phenomenon that environmental variance can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999).
+Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time.
+It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019.
+
+Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
+For instance, Amarasekare et al (year) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (cite 2012).
+Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (cite 2012).
+
+\hypertarget{the-relevancy-of-temperature-variability-in-the-modern-context}{%
+\subsection{The relevancy of temperature variability in the modern context}\label{the-relevancy-of-temperature-variability-in-the-modern-context}}
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figures/figure3} \caption{Figure 3.
+DON’T FORGET THIS IS LOG SCALED!!
+Plot of generation time in days against fluctuation period in days, with yellow, red, and grey horizontal panels representing diurnal, autocorrelated, and stochastic fluctuation patterns.
+The blue, green, purple, and pink horizontal lines represent different examples of variety in organismal generation time.
+*Figure out how to overlap pngs of plankton for 1, drosophila for 10, beetle for 100, and fish for 4000*}\label{fig:unnamed-chunk-3}
+\end{figure}
+Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
+Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance.
+Khelifa et al (2019) demonstrated that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figures/figure4} \caption{Figure 4.
+Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
+Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.}\label{fig:unnamed-chunk-4}
+\end{figure}
+However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 4).
+Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (2019).
+Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
+Vasseur and Yodzis (2004) outline how the frequency (1/period) of environmental fluctuations describes the ``color of environmental noise.'' In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
+Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (cite).
+Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
+
+Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level (cite):
+\begin{itemize}
+\tightlist
+\item
+  At the individual level, this usually comes in the form of reduced fitness or increased development rates (cite)
+\item
+  At the population level, this usually comes in the form of increases to population saturation
+\item
+  At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration
+\end{itemize}
+It however, less understood, is how variation patterns influence these characteristic responses at each level of organization.
+Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (cite).
+
+\hypertarget{objectives-driving-questions-hypotheses}{%
+\section{Objectives, driving questions, hypotheses}\label{objectives-driving-questions-hypotheses}}
+
+There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (cite).
+To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions:
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Does variability pattern drive differences in response magnitude? (i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal?)
+\item
+  Does temperature variability drive differential response levels at different levels of biological organization?
+\item
+  How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc)?
+\end{enumerate}
+We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
 
@@
 \begin{verbatim}
- [1] "month"        "day"          "dep_time"     "dep_delay"   
- [5] "arr_time"     "arr_delay"    "carrier"      "tailnum"     
- [9] "flight"       "dest"         "air_time"     "distance"    
-[13] "hour"         "minute"       "carrier_name" "dest_name"   
+ [1] "month"        "day"          "dep_time"     "dep_delay"    "arr_time"    
+ [6] "arr_delay"    "carrier"      "tailnum"      "flight"       "dest"        
+[11] "air_time"     "distance"     "hour"         "minute"       "carrier_name"
+[16] "dest_name"   
 \end{verbatim}
@@
 
+Placeholder
+
 \hypertarget{math}{%
 \section{Math}\label{math}}
-
-\TeX~is the best way to typeset mathematics.
-Donald Knuth designed \TeX~when he got frustrated at how long it was taking the typesetters to finish his book, which contained a lot of mathematics.
-One nice feature of \emph{R Markdown} is its ability to read LaTeX code directly.
-
-If you are doing a thesis that will involve lots of math, you will want to read the following section which has been commented out.
-If you're not going to use math, skip over or delete this next commented section.
 
@@
 
-Chemical formulas will look best if they are not italicized.
-Get around math mode's automatic italicizing in LaTeX by using the argument \texttt{\$\textbackslash{}mathrm\{formula\ here\}\$}, with your formula inside the curly brackets.
-(Notice the use of the backticks here which enclose text that acts as code.)
-
-So, \(\mathrm{Fe_2^{2+}Cr_2O_4}\) is written \texttt{\$\textbackslash{}mathrm\{Fe\_2\^{}\{2+\}Cr\_2O\_4\}\$}.
-
-\noindent Exponent or Superscript: \(\mathrm{O^-}\)
-
-\noindent Subscript: \(\mathrm{CH_4}\)
-
-To stack numbers or letters as in \(\mathrm{Fe_2^{2+}}\), the subscript is defined first, and then the superscript is defined.
-
-\noindent Bullet: CuCl \(\bullet\) \(\mathrm{7H_{2}O}\)
-
-\noindent Delta: \(\Delta\)
-
-\noindent Reaction Arrows: \(\longrightarrow\) or \(\xrightarrow{solution}\)
-
-\noindent Resonance Arrows: \(\leftrightarrow\)
-
-\noindent Reversible Reaction Arrows: \(\rightleftharpoons\)
-
 \hypertarget{typesetting-reactions}{%
 \subsection{Typesetting reactions}\label{typesetting-reactions}}
-
-You may wish to put your reaction in an equation environment, which means that LaTeX will place the reaction where it fits and will number the equations for you.
-\begin{equation}
-  \mathrm{C_6H_{12}O_6  + 6O_2} \longrightarrow \mathrm{6CO_2 + 6H_2O}
-  \label{eq:reaction}
-\end{equation}
-We can reference this combustion of glucose reaction via Equation \eqref{eq:reaction}.
 
@@
 
-\(\mathrm{NH_4Cl_{(s)}}\) \(\rightleftharpoons\) \(\mathrm{NH_{3(g)}+HCl_{(g)}}\)
-
-\noindent \(\mathrm{MeCH_2Br + Mg}\) \(\xrightarrow[below]{above}\) \(\mathrm{MeCH_2\bullet Mg \bullet Br}\)
-
 \hypertarget{physics}{%
 \section{Physics}\label{physics}}
-
-Many of the symbols you will need can be found on the math page \url{https://web.reed.edu/cis/help/latex/math.html} and the Comprehensive LaTeX Symbol Guide (\url{https://mirror.utexas.edu/ctan/info/symbols/comprehensive/symbols-letter.pdf}).
 
@@
 
-You will probably find the resources at \url{https://www.lecb.ncifcrf.gov/~toms/latex.html} helpful, particularly the links to bsts for various journals.
-You may also be interested in TeXShade for nucleotide typesetting (\url{https://homepages.uni-tuebingen.de/beitz/txe.html}).
-Be sure to read the proceeding chapter on graphics and tables.
-
 \hypertarget{ref-labels}{%
 \chapter{Graphics, References, and Labels}\label{ref-labels}}
+
+Placeholder
 
@@
 
-If your thesis has a lot of figures, \emph{R Markdown} might behave better for you than that other word processor.
-One perk is that it will automatically number the figures accordingly in each chapter.
-You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier.
-If you label your figures, you can move the figures around and \emph{R Markdown} will automatically adjust the numbering for you.
-No need for you to remember!
-So that you don't have to get too far into LaTeX to do this, a couple \textbf{R} functions have been created for you to assist.
-You'll see their use below.
-
-In the \textbf{R} chunk below, we will load in a picture stored as \texttt{reed.jpg} in our main directory.
-We then give it the caption of ``Reed logo'', the label of ``reedlogo'', and specify that this is a figure.
-Make note of the different \textbf{R} chunk options that are given in the R Markdown file (not shown in the knitted document).
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{include_graphics}\NormalTok{(}\DataTypeTok{path =} \StringTok{"figure/reed.jpg"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-\begin{figure}
-
-{\centering \includegraphics[width=0.2\linewidth]{figure/reed}
-
-}
-
-\caption{Reed logo}\label{fig:reedlogo}
-\end{figure}
-Here is a reference to the Reed logo: Figure \ref{fig:reedlogo}.
-Note the use of the \texttt{fig:} code here.
-By naming the \textbf{R} chunk that contains the figure, we can then reference that figure later as done in the first sentence here.
-We can also specify the caption for the figure via the R chunk option \texttt{fig.cap}.
-
-\clearpage
-
-Below we will investigate how to save the output of an \textbf{R} plot and label it in a way similar to that done above.
-Recall the \texttt{flights} dataset from Chapter \ref{rmd-basics}.
-(Note that we've shown a different way to reference a section or chapter here.) We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mean_delay_by_carrier <-}\StringTok{ }\NormalTok{flights }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{group_by}\NormalTok{(carrier) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{mean_dep_delay =} \KeywordTok{mean}\NormalTok{(dep_delay))}
-\KeywordTok{ggplot}\NormalTok{(mean_delay_by_carrier, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ carrier, }\DataTypeTok{y =}\NormalTok{ mean_dep_delay)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{position =} \StringTok{"identity"}\NormalTok{, }\DataTypeTok{stat =} \StringTok{"identity"}\NormalTok{, }\DataTypeTok{fill =} \StringTok{"red"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/delaysboxplot-1.pdf}
-\caption{\label{fig:delaysboxplot}Mean Delays by Airline}
-\end{figure}
-Here is a reference to this image: Figure \ref{fig:delaysboxplot}.
-
-A table linking these carrier codes to airline names is available at \url{https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv}.
-
-\clearpage
-
-Next, we will explore the use of the \texttt{out.extra} chunk option, which can be used to shrink or expand an image loaded from a file by specifying \texttt{"scale=\ "}.
-Here we use the mathematical graph stored in the ``subdivision.pdf'' file.
-\begin{figure}
-\includegraphics[scale=0.75]{figure/subdivision} \caption{Subdiv. graph}\label{fig:subd}
-\end{figure}
-Here is a reference to this image: Figure \ref{fig:subd}. Note that \texttt{echo=FALSE} is specified so that the \textbf{R} code is hidden in the document.
-
-\textbf{More Figure Stuff}
-
-Lastly, we will explore how to rotate and enlarge figures using the \texttt{out.extra} chunk option. (Currently this only works in the PDF version of the book.)
-\begin{figure}
-\includegraphics[angle=180, scale=1.1]{figure/subdivision} \caption{A Larger Figure, Flipped Upside Down}\label{fig:subd2}
-\end{figure}
-As another example, here is a reference: Figure \ref{fig:subd2}.
-
 \hypertarget{footnotes-and-endnotes}{%
 \section{Footnotes and Endnotes}\label{footnotes-and-endnotes}}
-
-You might want to footnote something.\footnote{footnote text} The footnote will be in a smaller font and placed appropriately.
-Endnotes work in much the same way.
-More information can be found about both on the CUS site or feel free to reach out to \href{mailto:data@reed.edu}{\nolinkurl{data@reed.edu}}.
 
@@
 
-Of course you will need to cite things, and you will probably accumulate an armful of sources.
-There are a variety of tools available for creating a bibliography database (stored with the .bib extension).
-In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero.
-The Reed librarians have created Zotero documentation at \url{https://libguides.reed.edu/citation/zotero}.
-In addition, a tutorial is available from Middlebury College at \url{https://sites.middlebury.edu/zoteromiddlebury/}.
-
-\emph{R Markdown} uses \emph{pandoc} (\url{https://pandoc.org/}) to build its bibliographies.
-One nice caveat of this is that you won't have to do a second compile to load in references as standard LaTeX requires.
-To cite references in your thesis (after creating your bibliography database), place the reference name inside square brackets and precede it by the ``at'' symbol.
-For example, here's a reference to a book about worrying: (Molina \& Borkovec, 1994).
-This \texttt{Molina1994} entry appears in a file called \texttt{thesis.bib} in the \texttt{bib} folder.
-This bibliography database file was created by a program called BibTeX.
-You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the \texttt{bib} folder.
-
-For more information about BibTeX and bibliographies, see our CUS site (\url{https://web.reed.edu/cis/help/latex/index.html})\footnote{Reed~College (2007)}.
-There are three pages on this topic: \emph{bibtex} (which talks about using BibTeX, at \url{https://web.reed.edu/cis/help/latex/bibtex.html}), \emph{bibtexstyles} (about how to find and use the bibliography style that best suits your needs, at \url{https://web.reed.edu/cis/help/latex/bibtexstyles.html}) and \emph{bibman} (which covers how to make and maintain a bibliography by hand, without BibTeX, at \url{https://web.reed.edu/cis/help/latex/bibman.html}).
-The last page will not be useful unless you have only a few sources.
-
-If you look at the YAML header at the top of the main .Rmd file you can see that we can specify the style of the bibliography by referencing the appropriate csl file.
-You can download a variety of different style files at \url{https://www.zotero.org/styles}.
-Make sure to download the file into the csl folder.
-
-\vfill
-
-\textbf{Tips for Bibliographies}
-\begin{itemize}
-\tightlist
-\item
-  Like with thesis formatting, the sooner you start compiling your bibliography for something as large as thesis, the better.
-Typing in source after source is mind-numbing enough; do you really want to do it for hours on end in late April?
-Think of it as procrastination.
-\item
-  The cite key (a citation's label) needs to be unique from the other entries.
-\item
-  When you have more than one author or editor, you need to separate each author's name by the word ``and'' e.g. \texttt{Author\ =\ \{Noble,\ Sam\ and\ Youngberg,\ Jessica\},}.
-\item
-  Bibliographies made using BibTeX (whether manually or using a manager) accept LaTeX markup, so you can italicize and add symbols as necessary.
-\item
-  To force capitalization in an article title or where all lowercase is generally used, bracket the capital letter in curly braces.
-\item
-  You can add a Reed Thesis citation\footnote{Noble (2002)} option.
-The best way to do this is to use the phdthesis type of citation, and use the optional ``type'' field to enter ``Reed thesis'' or ``Undergraduate thesis.''
-\end{itemize}
 \hypertarget{anything-else}{%
 \section{Anything else?}\label{anything-else}}
-
-If you'd like to see examples of other things in this template, please contact the Data @ Reed team (email \href{mailto:data@reed.edu}{\nolinkurl{data@reed.edu}}) with your suggestions.
-We love to see people using \emph{R Markdown} for their theses, and are happy to help.
 
@@
 \textbf{In the main Rmd file}
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# This chunk ensures that the thesisdown package is}
-\CommentTok{# installed and loaded. This thesisdown package includes}
-\CommentTok{# the template files for the thesis.}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}\StringTok{'You need to run install.packages("remotes")",}
-\StringTok{            "first in the Console.'}\NormalTok{)}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{"You need to run"}\NormalTok{,}
-        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\KeywordTok{library}\NormalTok{(thesisdown)}
-\CommentTok{# Set how wide the R output will go}
-\KeywordTok{options}\NormalTok{(}\DataTypeTok{width =} \DecValTok{70}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+
 \textbf{In Chapter \ref{ref-labels}:}
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# This chunk ensures that the thesisdown package is}
-\CommentTok{# installed and loaded. This thesisdown package includes}
-\CommentTok{# the template files for the thesis and also two functions}
-\CommentTok{# used for labeling and referencing}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("remotes")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(dplyr)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("dplyr")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(ggplot2)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("ggplot2")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(bookdown)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"bookdown"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("bookdown")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{"You need to run"}\NormalTok{,}
-        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\KeywordTok{library}\NormalTok{(thesisdown)}
-\KeywordTok{library}\NormalTok{(dplyr)}
-\KeywordTok{library}\NormalTok{(ggplot2)}
-\KeywordTok{library}\NormalTok{(knitr)}
-\NormalTok{flights <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data/flights.csv"}\NormalTok{, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+
 \hypertarget{the-second-appendix-for-fun}{%
 \chapter{The Second Appendix, for Fun}\label{the-second-appendix-for-fun}}
-
-\backmatter
 
@@
 
-\markboth{References}{References}
-
-\noindent
-
-\setlength{\parindent}{-0.20in}
-\setlength{\leftskip}{0.20in}
-\setlength{\parskip}{8pt}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-angel2000}{}%
-Angel, E. (2000). \emph{Interactive computer graphics : A top-down approach with opengl}. Boston, MA: Addison Wesley Longman.
-
-\leavevmode\hypertarget{ref-angel2001}{}%
-Angel, E. (2001a). \emph{Batch-file computer graphics : A bottom-up approach with quicktime}. Boston, MA: Wesley Addison Longman.
-
-\leavevmode\hypertarget{ref-angel2002a}{}%
-Angel, E. (2001b). \emph{Test second book by angel}. Boston, MA: Wesley Addison Longman.
-
-\leavevmode\hypertarget{ref-Molina1994}{}%
-Molina, S.
-T., \& Borkovec, T.
-D.
-(1994).
-The Penn State worry questionnaire: Psychometric properties and associated characteristics.
-In G.
-C.
-L.
-Davey \& F.
-Tallis (Eds.), \emph{Worrying: Perspectives on theory, assessment and treatment} (pp.
-265--283).
-New York: Wiley.
-
-\leavevmode\hypertarget{ref-noble2002}{}%
-Noble, S. G. (2002). \emph{Turning images into simple line-art} (Undergraduate thesis). Reed College.
-
-\leavevmode\hypertarget{ref-reedweb2007}{}%
-Reed~College. (2007). LaTeX your document. Retrieved from \url{https://web.reed.edu/cis/help/LaTeX/index.html}
+Placeholder
 

### 2021-04-24 writing /thesis_draft/index.Rmd
starting get my act together to format this thesis

@@
 advisor: 'Sam Fey'
-# If you have more two advisors, un-silence line 7
-#altadvisor: 'Your Other Advisor'
-department: 'Mathematics'
+department: 'Biology'
 degree: 'Bachelor of Arts'
@@
 site: bookdown::bookdown_site
-
-# This will automatically install the {remotes} package and {thesisdown}
-# Change this to FALSE if you'd like to install them manually on your own.
 params:
   'Install needed packages for {thesisdown}': True
-
-# Remove the hashtag to specify which version of output you would like.
-# Can only choose one at a time.
 output:
   thesisdown::thesis_pdf: default 
-#  thesisdown::thesis_gitbook: default         
-#  thesisdown::thesis_word: default
-#  thesisdown::thesis_epub: default
+abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
+acknowledgements: |
 
-# If you are creating a PDF you'll need to write your preliminary content 
-# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
-# for the .RMD files. If you are NOT producing a PDF, delete or silence
-# lines 25-39 in this YAML header.
-abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
-# If you'd rather include the preliminary content in files instead of inline
-# like below, use a command like that for the abstract above.  Note that a tab 
-# is needed on the line after the `|`.
-acknowledgements: |
   The future is bright, the future is dark
@@
   You can have a dedication here if you wish. 
-preface: |
-  This is an example of a thesis setup to use the reed thesis document class 
-  (for LaTeX) and the R bookdown package, in general.
-
-# Specify the location of the bibliography below
 bibliography: bib/thesis.bib
-# Download your specific csl file and refer to it in the line below.
 csl: csl/apa.csl
@@
 
-Welcome to the _R Markdown_ thesis template.
-This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before.
-Using _R Markdown_ will also allow you to easily keep track of your analyses in **R** chunks of code, with the resulting plots and output included as well.
-The hope is this _R Markdown_ template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.
+## The historical significance of variability 
+It is no secret that the natural world has relied on resilience and persistence in the face of environmental change for millenia (cite).
+Cyanobacteria can trace their origin back to a bacteria that evolved the ability to perform oxygenic photosynthesis to persist in a changing world (Soo et al 2017, science).
+Similarly, vascular plants have evolved different photosynthetic systems (C3, CAM) to manage a variety of thermal and spatial conditions (cite).
+While cyanobacteria and vascular plants evolved new strategies and structures to persist in the environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity may become even more relevant (cite).
+Understanding how organisms respond to the environment and its dynamic nature is significant for how to capitulate ecosystems currently and in the future.
 
-Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis.
-The use of LaTeX in combination with _Markdown_ is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file.
-While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities.
-After working with _Markdown_ and **R** together for a few weeks, we are confident this will be your reporting style of choice going forward.
+## Relevant significance of variability today 
+In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
+Ecosystems and organisms inhabiting those environments have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.
+ENSO).
+The scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (cite).
+As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.
+hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important (cite).
 
-<!-- 
-If you're still on the fence about using _R Markdown_, check out the resource
-for newbies available at <https://rbasics.netlify.com>. 
--->
-
-**Why use it?**
-
-_R Markdown_ creates a simple and straightforward way to interface with the beauty of LaTeX.
-Packages have been written in **R** to work directly with LaTeX to produce nicely formatting tables and paragraphs.
-In addition to creating a user friendly interface to LaTeX, _R Markdown_ also allows you to read in your data, to analyze it and to visualize it using **R** functions, and also to provide the documentation and commentary on the results of your project.
-Further, it allows for **R** results to be passed inline to the commentary of your results.
-You'll see more on this later.
-
-<!-- 
-Having your code and commentary all together in one place has a plethora of 
-benefits!
--->
-
-**Who should use it?**
-
-Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use _R Markdown_.
-Of particular use should be anyone in the sciences, but the user-friendly nature of _Markdown_ and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc.
-should make it of great benefit to nearly anyone writing a thesis project.
-
-**For additional help with bookdown**
-
-Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).
+## Why this study on variability matters 
+While there is a large body of literature on environmental variability including its relevance for X, y, and Z (include citation), there is a lack of consensus as to which scales of biological organization will be most severely affected by different shifts in environmental conditions (problem statement).
+We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as whether different patterns of variation differentially affect different levels of organization (aims).
+To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.

### 2021-04-27 shared analysis files /data_wrangling_plot.Rmd
updates and trying to finalize figures

@@
 
-```{r}
-dat_MA_ES %>%
-  count(study_id)
-```
-
 ```{r, echo=FALSE, results = 'hide', message=FALSE}
 ## loading libraries 
-pacman::p_load(metafor, tidyverse, viridis, visreg)
+pacman::p_load(metafor, tidyverse, viridis, visreg, forcats)
 #pacman::p_load(ggsci, plotly,  mgcv)
@@
   ggplot(aes(x=yi))+
-  geom_histogram(binwidth = 1)
+  geom_histogram(binwidth = 1)+
+  theme_bw()+
+  scale_fill_viridis()
 ```
@@
 full_model_coefficients<- as.data.frame(coef(full_rf_model)) %>%
-  tibble::rownames_to_column("covariates")
+  tibble::rownames_to_column("covariates") %>%
+  rename(value = "coef(full_rf_model)") %>%
+  arrange(desc(value))
 
-full_model_coefficients %>%
-  ggplot(aes(x = fct_reorder(covariates, coef(full_rf_model)),
-             y = coef(full_rf_model),
-             fill = coef(full_rf_model) > 0)) +
+label.df <- data.frame(covariates = c("flux_range", "flux_range:mean_constant_temp", "org_level"),
+                       value = c(0.3, -0.01, -0.69))
+
+ggplot(data=full_model_coefficients, aes(x = reorder(covariates, value),
+             y = value,
+             fill = value > 0)) +
   geom_col(show.legend = FALSE) +
-  coord_flip() +
   labs(x = NULL, y = "Coefficient estimate") +
   scale_fill_viridis(discrete = TRUE)+
-  theme_bw()
+  theme_bw()+
+  coord_flip()+
+  geom_text(data=label.df, label = "*")
+
+p + aes(x= reorder(covariates, value))
+
 ```
@@
        color = "Biological organization level")+
-  scale_color_viridis(discrete = TRUE, option = "D", labels = c("organismal", "population"))+
+  scale_color_viridis(discrete = TRUE, option = "D", labels = c("individual", "population"))+
   geom_smooth(method="lm", formula = y~x)+

### 2021-04-28 writing /thesis_draft/01-chap1.Rmd
an ugly but somewhat formatted draft

@@
 
-# Background 
+# Background {-}
 ## Environmental variability
-Environmental variability describes the dynamic nature of environmental conditions in the natural world and how organisms respond (cite).
-Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure X).
 
-```{r, echo=FALSE, fig.cap="Figure 1.
-An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
-Shown is a comparison of how duration of variability compares across the same environmental factor", out.width = '90%'}
+```{r, echo=FALSE, fig.cap="An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
+Shown is a comparison of how duration of variability compares across the same environmental factor", out.width = '105%'}
 knitr::include_graphics("figures/figure1.png")
 ```
+
+Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms’ response to environmental conditions (Parepa et al 2013, Gudmundson et al 2015).
+Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1).
 
@@
 It provides a departure point for what we expect the average value to be for all time.
-We can then expand additional information adjacent to the mean (i.e.
-range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk than just simply considering increases in global temperature (2014).
+In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, i.e.
+averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations.
+However, in other cases, the mean is less useful compared to the variability.
+For instance, if there are key factors contributing to population growth rates from specific locations, comparing the amount of variability in additional variables across locations (individual respiration rates, temperature, etc.) is more fruitful for the task at this scale.
+At some point however, all data of interest are averages.
+The important distinction is the scale at which we average the data and the level of detail provided at each scale.
+Means are more useful in describing variation amongst normally distributed data, however, other statistical metrics like the median or mode may be more appropriate when the data is not normally distributed.
 
 ### Amplitude/Range 
+We can expand additional information adjacent to the mean (i.e.
+range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014).
 We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
@@
 The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare et al 2013, 2012), as well as additional life history traits (size, speed, etc.).
+An organism’s ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible (Amarasekare 2012).
+If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale their adaptive evolution can manage.
 
@@
 The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability.
-Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure X).
-Short term fluctuations, like Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time, have differing effects on performance (Colinet et al 2018).
-While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.
-heat waves, cold fronts etc.).
-
-Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with organism, population, and community level processes.
-For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during that specific window of development (cite).
-Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (cite).
-When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (cite).
+Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1).
+Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet et al 2018).
+Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment (Husak and Lailvaux 2017, Amarasekare and Savage 2012).While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.
+heat waves, cold fronts etc.) in which the benefits of FTR in mitigated chill injuries may be helpful.
+Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
+For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Delmas et al 2007, Bowden and Paitz 2018).
+Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (Lawson et al 2015) depending on the magnitude of variance.
+When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Romanuk and Kolasa 2002, Dobramys and Tauber 2013, Bastille-Rousseau et al 2018).
 
 ### Predictability
+
 FTR can also describe the predictability of the fluctuation period.
 Variability may not have evenly spaced intervals (i.e.
-heatwaves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
+heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
 Predictability can also attempt to describe stochastic vs.
@@
 
-Link between temporal and spatial variability 
+### Link between temporal and spatial variability
+
 While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact.
-Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (cite).
+Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Horne and Schneider 1995, Farhang-Sardroodi et al 2019).
 Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e.
 Diel vertical migration (DVM).
-In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (cite).
+In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (Lampert 1989, Brierly 2014).
 However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.
-an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (cite).
+an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey 1974, Kearney et al 2011).
 
-Though environmental variability can include many different forms, from fluctuations in resource availability (cite) to light (cite), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (cite).
+Though environmental variability can include many different forms, from fluctuations in resource availability (Dempster and Pollard 1981, Sommer 1984) to light (Ruel and Ayres 1999, Morison 2020), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (Huey et al 1990, Colinet et al 2015).
 
@@
 ### What is temperature variability?
+
 Changes in temperature can be grouped into a couple different orientations, the first being the difference between the terms “fluctuating” and “variation” with respect to temperature.
 Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as “a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis” (2018).
-In any case, temperature variability describes changes in environmental temperature through time (cite).
+Most generally, temperature variability describes changes in environmental temperature through time (Holmes et al 2016, Colinet et al 2018).
 
 ### Why is it important to ecological processes?
-Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (cite).
+
+Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey et al 1990).
 Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of  ecology (MTE), though highly debated (Clarke et al 2006), which serves as an origin point for understanding how organisms respond to environmental conditions.
@@
 However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clark and Johnston 1999, Clarke et al 2006).
-Others believe that MTE makes valid assumptions and is more flexible than opponents argue (Gillooly et al 2006).
+Others believe that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al 2006).
+Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.
 
@@
 The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms’ performance is negatively impacted (cite).
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms’ performance is negatively impacted (Brown et al 2004, Dowd et al 2015).
 Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
 
+TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (i.e.
+metabolism) (Amarasekare et al 2012).
+If only temperature’s effect on metabolism were considered, the shape of the relationship may be an exponential one.
+However, there are several other constraining factors (i.e.
+energy acquisition, nutrient availability etc.) that derive the characteristically nonlinear pattern of TPCs.
+TPCs strive to represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
+
 ### Jensen’s Inequality and its impacts on metabolic rate
-TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (cite, Bernhardt et al 2018).
+
+TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt et al 2018).
 However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.
@@
 Given the characteristically non-linear shape of TPCs, Jensen’s inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
-Jensen’s inequality describes the phenomenon that environmental variance can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999).
+Jensen’s inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999).
+For instance, Ruel and Ayres (1999) provide the development rate of ectotherms as an example to demonstrate Jensen’s inequality.
+The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel and Ayres, 1999).
+With Jensen’s inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel and Ayres, 1999).
 Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time.
-It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019.
+It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019).
 
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al (year) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (cite 2012).
-Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (cite 2012).
+For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare et al 2012, Paajimans et al 2013).
+Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare et al 2012).
 
@@
 
-```{r, echo=FALSE, fig.cap="Figure 3.
-DON’T FORGET THIS IS LOG SCALED!!
-Plot of generation time in days against fluctuation period in days, with yellow, red, and grey horizontal panels representing diurnal, autocorrelated, and stochastic fluctuation patterns.
-The blue, green, purple, and pink horizontal lines represent different examples of variety in organismal generation time.
-*Figure out how to overlap pngs of plankton for 1, drosophila for 10, beetle for 100, and fish for 4000*", out.width = '90%'}
-knitr::include_graphics("figures/figure3.png")
-```
+Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
+Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
+Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
 
-Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
-Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance.
-Khelifa et al (2019) demonstrated that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
+However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 3).
+Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (2019).
+Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
+Vasseur and Yodzis (2004) outline how the frequency (1/period) of environmental fluctuations describes the “color of environmental noise.” In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
+Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (Wigley et al 1998, Tabari et al 2012).
+Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
 
-```{r, echo=FALSE, fig.cap="Figure 4.
-Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
+```{r, echo=FALSE, fig.cap="Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
 Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.", out.width = '90%'}
@@
 
-However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 4).
-Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (2019).
-Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
-Vasseur and Yodzis (2004) outline how the frequency (1/period) of environmental fluctuations describes the “color of environmental noise.” In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
-Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (cite).
-Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
-
-Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level (cite):
-
-* At the individual level, this usually comes in the form of reduced fitness or increased development rates (cite)
-* At the population level, this usually comes in the form of increases to population saturation
-* At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration
-
-It however, less understood, is how variation patterns influence these characteristic responses at each level of organization.
-Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (cite).
-
 ## Objectives, driving questions, hypotheses
 
-There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (cite).
+Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
+* At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve (Bartheld et al 2017) to changes in TSD ratios (Bowden and Paitz 2018)
+* At the population level, this usually comes in the form of decreasing population densities (Anders and Post 2006)
+* At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration (Descamps-Julien and Gonzalez 2005), inflationary effects (Holt et al 2003)
+It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
+Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (Wang et al 2016, Petchey et al 2000).
+
+There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al 2019, Massey et al 2020).
 To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions:
-
-1. Does variability pattern drive differences in response magnitude? (i.e. does autocorrelated variation have a larger impact on response than stochastic or diurnal?)
-2. Does temperature variability drive differential response levels at different levels of biological organization?
-3. How can the magnitude of response be influenced by additional factors (i.e. periodicity, age, size, thermal history, etc)?
-
+1) Does variability pattern drive differences in response magnitude?
+(i.e.
+does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.
+periodicity, age, size, thermal history, etc)?
 We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
 
-## Lists
-
-It's easy to create a list.  It can be unordered like
-
-* Item 1
-* Item 2
-
-or it can be ordered like
-
-1. Item 1
-4. Item 2
-
-Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.
-
-To create a sublist, just indent the values a bit (at least four spaces or a tab).  (Here's one case where indentation is key!)
-
-1. Item 1
-1. Item 2
-1. Item 3
-    - Item 3a
-    - Item 3b
-
-## Line breaks
-
-Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:
-
-Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
-This should be a new paragraph.
-
-*Now for the correct way:*
-
-Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
-
-This should be a new paragraph.
-
-## R chunks
-
-When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document.
-You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):
-
-```{r cars}
-summary(cars)
-```
-
-## Inline code
-
-If you'd like to put the results of your analysis directly into your discussion, add inline code like this:
-
-> The `cos` of $2 \pi$ is `r cos(2*pi)`.
-
-Another example would be the direct calculation of the standard deviation:
-
-> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.
-
-One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:
-
-> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`
-
-Note the use of `>` here, which signifies a quotation environment that will be indented.
-
-As you see with `$2 \pi$` above, mathematics can be added by surrounding the mathematical text with dollar signs.
-More examples of this are in [Mathematics and Science] if you uncomment the code in [Math].
-
-## Including plots
-
-You can also embed plots. For example, here is a way to use the base **R** graphics package to produce a plot using the built-in `pressure` dataset:
-
-```{r pressure, echo=FALSE, cache=TRUE, fig.height=3, fig.width=5}
-plot(pressure)
-```
-
-Note that the `echo=FALSE` parameter was added to the code chunk to prevent printing of the **R** code that generated the plot.
-There are plenty of other ways to add chunk options (like `fig.height` and `fig.width` in the chunk above).
-More information is available at <https://yihui.org/knitr/options/>.
-
-Another useful chunk option is the setting of `cache=TRUE` as you see here.
-If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.
-Later in this file, you'll see a way to reference plots created in **R** or external figures.
-
-## Loading and exploring data
-
-Included in this template is a file called `flights.csv`.
-This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014.
-More information about this dataset and its **R** package is available at <https://github.com/ismayc/pnwflights14>.
-This subset includes only Portland flights and only rows that were complete with no missing values.
-Merges were also done with the `airports` and `airlines` data sets in the `pnwflights14` package to get more descriptive airport and airline names.
-
-We can load in this data set using the following commands:
-
-```{r load_data}
-# flights.csv is in the data directory
-flights_path <- here::here("data", "flights.csv")
-# string columns will be read in as strings and not factors now
-flights <- read.csv(flights_path, stringsAsFactors = FALSE)
-```
-
-The data is now stored in the data frame called `flights` in **R**.
-To get a better feel for the variables included in this dataset we can use a variety of functions.
-Here we can see the dimensions (rows by columns) and also the names of the columns.
-
-```{r str}
-dim(flights)
-names(flights)
-```
-
-Another good idea is to take a look at the dataset in table form.
-With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here.
-I recommend you enter the command into the Console **_after_** you have run the **R** chunks above to load the data into **R**.
-
-```{r view_flights, eval=FALSE}
-View(flights)
-```
-
-While not required, it is highly recommended you use the `dplyr` package to manipulate and summarize your data set as needed.
-It uses a syntax that is easy to understand using chaining operations.
-Below I've created a few examples of using `dplyr` to get information about the Portland flights in 2014.
-You will also see the use of the `ggplot2` package, which produces beautiful, high-quality academic visuals.
-
-We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:
-
-```{r load_pkgs, message=FALSE}
-# List of packages required for this analysis
-pkg <- c("dplyr", "ggplot2", "knitr", "bookdown")
-# Check if packages are not installed and assign the
-# names of the packages not installed to the variable new.pkg
-new.pkg <- pkg[!(pkg %in% installed.packages())]
-# If there are any packages in the list that aren't installed,
-# install them
-if (length(new.pkg)) {
-  install.packages(new.pkg, repos = "https://cran.rstudio.com")
-}
-# Load packages
-library(thesisdown)
-library(dplyr)
-library(ggplot2)
-library(knitr)
-```
-
-\clearpage
-
-The example we show here does the following:
-
-- Selects only the `carrier_name` and `arr_delay` from the `flights` dataset and then assigns this subset to a new variable called `flights2`.
-
-- Using `flights2`, we determine the largest arrival delay for each of the carriers.
-
-```{r max_delays}
-flights2 <- flights %>%
-  select(carrier_name, arr_delay)
-max_delays <- flights2 %>%
-  group_by(carrier_name) %>%
-  summarize(max_arr_delay = max(arr_delay, na.rm = TRUE))
-```
-
-A useful function in the `knitr` package for making nice tables in _R Markdown_ is called `kable`.
-It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.
-This again goes to show how nice reproducible documents can be!
-(Note the use of `results="asis"`, which will produce the table instead of the code to create the table.)  The `caption.short` argument is used to include a shorter title to appear in the List of Tables.
-
-```{r maxdelays, results="asis"}
-kable(max_delays,
-  col.names = c("Airline", "Max Arrival Delay"),
-  caption = "Maximum Delays by Airline",
-  caption.short = "Max Delays by Airline",
-  longtable = TRUE,
-  booktabs = TRUE
-)
-```
-
-The last two options make the table a little easier-to-read.
-
-We can further look into the properties of the largest value here for American Airlines Inc.
-To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original `flights` dataset.
-
-```{r max_props}
-flights %>%
-  filter(
-    arr_delay == 1539,
-    carrier_name == "American Airlines Inc."
-  ) %>%
-  select(-c(
-    month, day, carrier, dest_name, hour,
-    minute, carrier_name, arr_delay
-  ))
-```
-
-We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth.
-Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.
-
-```{r march3plot, fig.height=3, fig.width=6}
-flights %>%
-  filter(month == 3, day == 3) %>%
-  ggplot(aes(x = dep_time, y = arr_delay)) +
-  geom_point()
-```
-
-## Additional resources
-
-- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>
-
-- _R Markdown_
-    - Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf>
-
-- _RStudio IDE_
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf>
-    - Official website - <https://rstudio.com/products/rstudio/>
-
-- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html>
-
-- `ggplot2`
-    - Documentation - <https://ggplot2.tidyverse.org/>
-    - Cheatsheet - <https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf>
+We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and larger organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).

### 2021-04-28 writing /thesis_draft/02-chap2.Rmd
an ugly but somewhat formatted draft

@@
-# Methods
+# Methods {-}
+```{r, include=FALSE}
+pacman::p_load(knitr, kableExtra, tidyverse)
+```
 
-## Math
+```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
+Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%'}
+knitr::include_graphics("figures/figure5.png")
+```
 
-\TeX\ is the best way to typeset mathematics.
-Donald Knuth designed \TeX\ when he got frustrated at how long it was taking the typesetters to finish his book, which contained a lot of mathematics.
-One nice feature of _R Markdown_ is its ability to read LaTeX code directly.
+We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 3).
+After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*) ) AND SU=(Life Sciences & Biomedicine).
+This search returned a total of 176 results.
+We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1).
+Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 5).
+Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
 
-If you are doing a thesis that will involve lots of math, you will want to read the following section which has been commented out.
-If you're not going to use math, skip over or delete this next commented section.
+* We excluded studies that did not feature both a constant and fluctuating treatment (i.e.
+studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
+* We excluded studies that were not completed in a laboratory (i.e. modeling predictions or results from models, mesocosm experiments, in situ experiments)
+* We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (i.e.
+organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
+* We excluded any articles not written in English 
+* We excluded studies that did not report measurements of error
+* We excluded studies in which the response variables are comparable or extrable (i.e. one study performed Principal component analysis for their responses)
 
-<!-- MATH and PHYSICS majors: Uncomment the following section -->
-<!--
-$$\sum_{j=1}^n (\delta\theta_j)^2 \leq {{\beta_i^2}\over{\delta_i^2 + \rho_i^2}}
-\left[ 2\rho_i^2 + {\delta_i^2\beta_i^2\over{\delta_i^2 + \rho_i^2}} \right] \equiv \omega_i^2
-$$
+```{r, echo = FALSE}
+table1 <- data.frame(Cat = c("Delava et al", "Du et al", 
+                         "Garcia-Ruiz et al", "Glass et al", 
+                         "Kern et al", "Kern et al", 
+                         "Lowenborg et al", "Maneti et al", 
+                         "Pendlebury et al", "Qu et al", 
+                         "Rolandi et al", "Saxon et al", 
+                         "Semenov et al", "Simonici et al", 
+                         "Triedel et al"), 
+                 Year = c("2016", "2009", "2011", "2009", "2015", "2014",
+                          "2012", "2014", "2004", "2014", "2018", "2017",
+                          "2007", "2019", "2015"), 
+                 `Extracted data points (n)` = c("16", "32", "16", "16", "24",
+                                                 "4", "10", "24", "16", "40", "4", 
+                                                 "12", "32", "30", "4"))
+kable(table1, "latex", booktabs = T, col.names = c("Study", "Year", "Extracted data points (n)"), escape = F, caption = "Studies included in analysis of thermal variability (n=15)") %>%
+   row_spec(0,bold=TRUE)
+```
 
-From Informational Dynamics, we have the following (Dave Braden):
+We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.
+treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
+We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al 2004).
+The final number of data points included in the analysis is broken down by study (Table 1).
 
-After _n_ such encounters the posterior density for $\theta$ is
+## Data extraction and analysis
 
-$$
-\pi(\theta|X_1< y_1,\dots,X_n<y_n) \varpropto \pi(\theta) \prod_{i=1}^n\int_{-\infty}^{y_i}
-   \exp\left(-{(x-\theta)^2\over{2\sigma^2}}\right)\ dx
-$$
+From the studies that met our inclusion criteria, we extracted data from tables and figures.
+Figures were extracted via png files imported into WebPlotDigitizer to obtain the raw data.
+The extracted data was then imported into a master spreadsheet of data from all studies for analysis.
+We extracted mean values for response variables, any measure of variance (Standard Deviation (SD) or Standard Error of the Mean (SEM)), and sample size.
+For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM (see code appendix).
+We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal history, and aggregated these values (Table 2).
 
-Another equation:
+```{r, echo = FALSE, include = TRUE}
+table2 <- data.frame(Variable = c("mean_temp_constant", "flux_range", 
+                         "org_level", "size", 
+                         "exp_age", "resp_type", 
+                         "stressful"),
+                     Description = c("Reported mean temperature in study",
+                                     "Reported range (difference between minimum 
+                                     and maximum) of temperature fluctuation experienced", 
+                                     "Level of biological organization, categorized
+                                     as either 0 for individual or 1 for population level
+                                     responses", 
+                                     "Size estimate of study organism, 0 for extra-small
+                                     (microscopic), 1 for small (macroscopic), 2 for medium 
+                                     (less than a foot), and 3 for large (more than a foot)", 
+                                     "Life stage at which experiment was conducted, 
+                                     categorized as 0 for larval, 1 for juvenile, and 
+                                     2 for adult",
+                                     "Response variables units classified as rates to traits, 
+                                     such that rates included any response variable units with a
+                                     “per day” clarifier were classified as rates, all else as
+                                     traits",
+                                     "Whether or not the organisms experienced thermally stressful 
+                                     temperatures throughout study duration, if temperature range 
+                                     exceeded Topt, classified as thermally stressful, else not
+                                     thermally stressful. See appendix for references in which
+                                     organisms’ thermal history was included"))
 
-$$\det\left|\,\begin{matrix}%
-c_0&c_1\hfill&c_2\hfill&\ldots&c_n\hfill\cr
-c_1&c_2\hfill&c_3\hfill&\ldots&c_{n+1}\hfill\cr
-c_2&c_3\hfill&c_4\hfill&\ldots&c_{n+2}\hfill\cr
-\,\vdots\hfill&\,\vdots\hfill&
-  \,\vdots\hfill&&\,\vdots\hfill\cr
-c_n&c_{n+1}\hfill&c_{n+2}\hfill&\ldots&c_{2n}\hfill\cr
-\end{matrix}\right|>0$$
+kable(table2, caption = "All covariates included in the model", "latex", booktabs = T) %>%
+  column_spec(2, width = "8cm") %>%
+   row_spec(0,bold=TRUE)
+```
 
-Lapidus and Pindar, Numerical Solution of Partial Differential Equations in Science and
-Engineering.  Page 54
+If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Cochrane Handbook for Systematic Reviews of Interventions, Version 5.1.0, 2011).
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
 
-$$
-\int_t\left\{\sum_{j=1}^3 T_j \left({d\phi_j\over dt}+k\phi_j\right)-kT_e\right\}w_i(t)\ dt=0,
-   \qquad\quad i=1,2,3.
-$$
+## Metafor
 
-L\&P  Galerkin method weighting functions.  Page 55
+We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
+Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer et al 2010).
+In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
 
-$$
-\sum_{j=1}^3 T_j\int_0^1\left\{{d\phi_j\over dt} + k\phi_j\right\} \phi_i\ dt
-   = \int_{0}^1k\,T_e\phi_idt, \qquad i=1,2,3 $$
+```{r, echo = FALSE}
+table3 <- data.frame(Cat = c("Mf", "SDf", 
+                         "Mc", "SDc", 
+                         "Nf", "Nc"), 
+                 Purpose = c("Mean of experimental group ", 
+                             "Standard deviation of the experimental group",
+                             "Mean of the control group",
+                             "Standard deviation of the control group",
+                             "Sample size (n) of the experimental group",
+                             "Sample size (n) of the control group"))
+kable(table3, "latex", booktabs = T, col.names = c("Column name", "Purpose"), escape = F, caption = "Column names expanded to wider format for data analysis in metafor") %>%
+   row_spec(0,bold=TRUE)
+```
 
-Another L\&P (p145)
+Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the escalc function in metafor to calculate effect size and sampling variance:
+```{r, eval= FALSE}
+escalc(measure="SMD", m1i=`constant_resp`, m2i=flux_resp, 
+            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, 
+       n2i=flux_samp, data=dat_MA, slab=paste(study_id, experiment_id, 
+                                              response_id, sep=", "))
+```
+We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head (Viechtbauer 2010).
+It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer 2010).
+Each of the subsequent arguments of the escalc function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
+The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 6).
 
-$$
-\int_{-1}^1\!\int_{-1}^1\!\int_{-1}^1 f\big(\xi,\eta,\zeta\big)
-   = \sum_{k=1}^n\sum_{j=1}^n\sum_{i=1}^n w_i w_j w_k f\big( \xi,\eta,\zeta\big).
-$$
+```{r, echo= FALSE, fig.cap= "Workflow diagram of how data were extracted, analyzed, and processed for downstream analysis in metafor.", out.width="100%"}
+knitr::include_graphics("figures/figure6.png")
+```
 
-Another L\&P (p126)
+First, it calculates the pooled standard deviation (Pooled SD) from the sample sizes of both studies.
+Our data had studies with a wide range of sample sizes, so we wanted to include separate sample sizes between the constant and fluctuating treatments.
+The Pooled SD, Cohen’s d, which is one of the most common ways to measure effect size, is used to normalize the mean differences between constant and fluctuating treatments.
+Then, using a correction factor J, which again normalizes the data further, we obtain the final SMD.
+The SMD is then ultimately used to calculate the sampling variance, which creates a unitless metric for variance.
 
-$$
-\int_{A_e} (\,\cdot\,) dx dy = \int_{-1}^1\!\int_{-1}^1 (\,\cdot\,) \det[J] d\xi d\eta.
-$$
--->
+Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response (see insert of pseudo code).
 
-## Chemistry 101: Symbols
+```{r, eval = FALSE}
+rma.mv(yi, vi, 
+       data=dat_MA_ES, 
+       random = ~1 | study_id / experiment_id/ response_id, 
+       method="REML") 
+```
 
-Chemical formulas will look best if they are not italicized.
-Get around math mode's automatic italicizing in LaTeX by using the argument `$\mathrm{formula here}$`, with your formula inside the curly brackets.
-(Notice the use of the backticks here which enclose text that acts as code.)
+The arguments of the rma.mv function include SMD (yi), sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
+We specified a nested structure to account for the colinearity of data points from within the same study, experiment, and response metric.
+We used Restricted Maximum-Likelihood (REML) to specify the fit of the model, as REML is best at handling small samples sizes compared to Maximum Likelihood (ML) (Viechtbauer 2010).
+We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population (Viechtbauer 2010).
 
-So, $\mathrm{Fe_2^{2+}Cr_2O_4}$ is written `$\mathrm{Fe_2^{2+}Cr_2O_4}$`.
+We then ran two multivariate random effects models, one with modifiers and one without (see insert above).
+This was to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies.
+The full model we ran included:
 
-<!--
-The \noindent command below does what you'd expect:  it forces the current line/paragraph to not indent. This was done here to match the format of the LaTeX thesis PDF.
--->
+```{r, eval = FALSE}
+rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant 
+       + exp_age + size + org_level + resp_type, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+```
 
-\noindent Exponent or Superscript: $\mathrm{O^-}$
+We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal history data for one study.
+That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz 2011, Table 1).
+We ran a separate model without this study and included the thermal history covariate separately from all the other modifiers:
 
-\noindent Subscript: $\mathrm{CH_4}$
-
-To stack numbers or letters as in $\mathrm{Fe_2^{2+}}$, the subscript is defined first, and then the superscript is defined.
-
-\noindent Bullet: CuCl $\bullet$ $\mathrm{7H_{2}O}$
-
-\noindent Delta: $\Delta$
-
-\noindent Reaction Arrows: $\longrightarrow$ or  $\xrightarrow{solution}$
-
-\noindent Resonance Arrows: $\leftrightarrow$
-
-\noindent Reversible Reaction Arrows: $\rightleftharpoons$
-
-### Typesetting reactions
-
-You may wish to put your reaction in an equation environment, which means that LaTeX will place the reaction where it fits and will number the equations for you.
-
-\begin{equation}
-  \mathrm{C_6H_{12}O_6  + 6O_2} \longrightarrow \mathrm{6CO_2 + 6H_2O}
-  (\#eq:reaction)
-\end{equation}
-
-We can reference this combustion of glucose reaction via Equation \@ref(eq:reaction).
-
-### Other examples of reactions
-
-$\mathrm{NH_4Cl_{(s)}}$ $\rightleftharpoons$ $\mathrm{NH_{3(g)}+HCl_{(g)}}$
-
-\noindent $\mathrm{MeCH_2Br + Mg}$ $\xrightarrow[below]{above}$ $\mathrm{MeCH_2\bullet Mg \bullet Br}$
-
-## Physics
-
-Many of the symbols you will need can be found on the math page <https://web.reed.edu/cis/help/latex/math.html> and the Comprehensive LaTeX Symbol Guide (<https://mirror.utexas.edu/ctan/info/symbols/comprehensive/symbols-letter.pdf>).
-
-## Biology
-
-You will probably find the resources at <https://www.lecb.ncifcrf.gov/~toms/latex.html> helpful, particularly the links to bsts for various journals.
-You may also be interested in TeXShade for nucleotide typesetting (<https://homepages.uni-tuebingen.de/beitz/txe.html>).
-Be sure to read the proceeding chapter on graphics and tables.
+```{r, eval =FALSE}
+rma.mv(yi, vi, data=dat_MA_ES, mods =  ~ stressful, 
+               random = ~1 | experiment_id/ study_id/ response_id,
+                 method="REML") 
+```

### 2021-04-28 writing /thesis_draft/03-chap3.Rmd
an ugly but somewhat formatted draft

```{r include_packages_2, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis and also two functions
# used for labeling and referencing
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("remotes")',
        "first in the Console."
      )
    )
  }
}
if (!require(dplyr)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("dplyr", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("dplyr")',
        "first in the Console."
      )
    )
  }
}
if (!require(ggplot2)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("ggplot2", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("ggplot2")',
        "first in the Console."
      )
    )
  }
}
if (!require(bookdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("bookdown", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste(
        'You need to run install.packages("bookdown")',
        "first in the Console."
      )
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
library(dplyr)
library(ggplot2)
library(knitr)
flights <- read.csv("data/flights.csv", stringsAsFactors = FALSE)
```

# Results {-}

```{r, echo=FALSE, results = 'hide', message=FALSE}
pacman::p_load(metafor, tidyverse, viridis, visreg, forcats, kableExtra)
dat_MA_0<- read_csv("metafor.csv") %>%
  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))

#make standardized SD columns based on N and SE vs SD
dat_MA <-dat_MA_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_MA_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_MA, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) 
count(dat_MA_ES, study_id)

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -4) 
```

Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 6. These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis. As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.

```{r, echo= FALSE, message= FALSE, fig.cap="Forest plot of effects sizes by study and experiment within a study. Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study. Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", out.width="110%"}
forest <- dat_MA_ES %>%
  group_by(study_id, experiment_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2")
```

Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal history), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p < 0.0001). Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (df = 139, p < 0.00001). Organization level was also important in the model, as a statistically significant covariate (df= 139, p < 0.05).

```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.", out.width="105%"}
full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML")

#plot to include to demonstrate model estimates
full_model_coefficients<- as.data.frame(coef(full_rf_model)) %>%
  tibble::rownames_to_column("covariates") %>%
  rename(value = "coef(full_rf_model)") %>%
  arrange(desc(value)) %>%
  mutate(coefficients = c("intercept", "response trait type", 
                          "fluctuation range", "mean temperature",
                          "fluctuation range:mean temperature",
                                       "size", "experimental age","organization level" 
                                      ))

label.df <- data.frame(coefficients = c("fluctuation range", "fluctuation range:mean temperature", "organization level"),
                       value = c(0.35, -0.07, -0.69))

ggplot(data=full_model_coefficients, aes(x = reorder(coefficients, value),
             y = value,
             fill = value > 0)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "Coefficient estimate") +
  scale_fill_viridis(discrete = TRUE)+
  theme_bw()+
  coord_flip()+
  geom_text(data=label.df, label = "*")
```

Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 7).The most negative estimate (organization level) was significantly influential in our model at a significance level of p < 0.05.  Surprisingly, the most negative estimates of our model coefficients were not statistically significant. In our analysis of thermal history via a separate random effects model, we found thermal history was not statistically significant in explaining the effect size (Table 3).

```{r, echo= FALSE}
table3 <- data.frame(Cat = c("intercept", "stressful"),
                     estimate = c(0.2050, -0.0489), 
                 se = c(0.1846, 0.0702),
                 zval = c(1.1104, -0.6975),
                 pval = c(0.2668, 0.4855), 
                 ci.lb = c(-0.1568, -0.1864), 
                 ci.ub = c(0.5669, 0.0886))
kable(table3, "latex", booktabs = T, col.names = c("Term", "estimate", "se", "zval", "pval", "ci.lb", "ci.ub"), escape = F, caption = "Model summary statistics for random effects model using thermal history as sole modifier") %>%
   row_spec(0,bold=TRUE)
```

The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 8. At lower mean temperatures, higher fluctuation ranges have a more positive effect on..., though that trend starkly ends at about 24O C. Beyond 24O C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 8).

```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature range, and effect size. Temperature fluctuation range from 0-20C, Mean temperature range from 7-33OC. Effect size range from -2 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)", out.width = '90%'}
normalized %>%
ggplot(aes(x=mean_temp_constant, y = yi, fill=flux_range))+
  geom_jitter( alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Temperature range (C)', option = "C")+
  theme_classic()+
  xlab("Mean Temperature (C)")+
  ylab("Effect Size")+
  geom_line(y=0, linetype = "dashed")+
  theme(legend.position="bottom")
```

The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 9). Interestingly, there were not any significant differences between thermal history (Figure 10), life history metric (Figure 11), body size (Figure 12), or age (Figure 13).

As noted in the methods, when we excluded one highly influential study and included thermal history (exclusion model), size became much more important in the thermal stress model (df=123, p < 0.001) as did mean temperature. Surprisingly, organization level, fluctuation range, and the interaction between range and mean temperature became insignificant in the thermal stress model (df=123, p < 0.1).

```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.", out.width = '90%'}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(org_level)))+
  geom_point()+
  labs(x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Biological organization level")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("individual", "population"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom") 
```

```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal history: stressful or not stressful.", out.width = '90%'}
normalized %>%
  drop_na(stressful) %>%
ggplot(aes(x=flux_range, y=yi, color = factor(stressful)))+
  geom_jitter()+
  labs(x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Thermally stressfull")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("no", "yes"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.", out.width = '90%'}
normalized %>%
ggplot(aes(x=flux_range, y=yi, color = resp_type))+
  geom_jitter()+
  labs(x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Life history metric")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("rate", "trait"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.", out.width = '90%'}
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(size)))+
  geom_point()+
  labs(x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Size")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("extra-small", "small", "medium", "large"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.", out.width = '90%'}
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(exp_age)))+
  geom_point()+
  labs(x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Experimental age")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("larval", "juvenile", "adult"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

# Discussion {-}

The results of this meta analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level. There were not overall differences in the direction o,

These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e. higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures. The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change (Vasseur et al 2014, Sunday et al 2012). With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed “the point of no return” (Russill and Nyssa 2009), understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades (West et al 2002, Cheung et al 2008). In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature (Vasseur et al 2014).

These conclusions further drive home the decadal conclusion that Jensen’s Inequality has demonstrated: averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance (Ruel and Ayres 1999, Bernhardt et al 2018). The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive.

Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal history, life history traits, and experimental age. Body size has been identified as important in the allometric scaling of MTE and it is interesting in these subsets of studies, body size did not explain the differences in responses across studies. This could simply be because we had a small sample size of studies (n=15), though multiple  responses in studies contributed to a large sample of effect sizes (n=139) and there were only a subset of relative body sizes included. It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms (Clark and Johnston 1999, Clarke et al 2006). This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns variability in temperature drives.

An important side note is that our full model and thermal history model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola). This study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding Garcia-Ruiz et al (2011) from our full model. We decided to include this study in our analysis because the data were extracted from a table, therefore there were no inaccuracies in obtaining the data. It provides an important counter and discussion for how and why responses from different studies may be so different from others.

While body size and thermal history emerged as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship thermal variability and MTE. If we were able to obtain additional information about the CTmax, CTmin, thermal breadth, and tolerance range for each of the species included in the analysis, we may also be able to better understand the trends and the significance of variability patterns in dictating responses.

Both the type of response (i.e. whether it was a rate or trait) and experimental age were also not important predictors in our full model. We expected rates and traits to have different effects, as traits may be more heavily affected by variability because they are not as temporally dynamic as rates. The lack of difference between rates and traits may be because variability is not differentially acting on rate or trait based responses, simply dampening their overall effects. Again, this may be an artifact of having a small sample size, with mainly trait based responses as opposed to rate based responses.

We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018). This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile. However, there have been meta analyses, both qualitative (Massey et al 2020), and quantitative (Noble et al 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al 2018). These results align with patterns of juvenile organisms (i.e. fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Portner and Farrell 2008, Moyano et al 2017).

Beyond the data included in our meta analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project. As Figure 4 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures. It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated in fact. Are these diurnal temperature fluctuations accurate for what is expected under natural conditions? This lack of variety in pattern may be due to a lack of consensus on the language used to classify and discuss thermal variability. Though there is a large body of literature that investigates temperature variation, there appears to be dissonance in the scientific literature about the language used to discuss such variation. Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures (Amarasekare et al 2012) as well as different ranges of temperature, both of which are different concepts. For instance, Amarasekare et al 2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits. Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern. Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes (Colinet et al 2018, Colinet et al 2015). Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.

Many of these studies failed to justify why certain temperatures were chosen as mean temperatures as well as the range of temperatures studied. In order to compile information on the thermal history of each of the organisms included in the studies of this meta analysis, additional resources were needed (see Appendix Table S1). Understanding where these organisms' performance falls in relation to temperature on its TPC is crucial for drawing reliable conclusions of how thermal variability affects performance. Without it, only broad conclusions can be drawn, as have been drawn here, that variability and range affects thermal performance.

Many of the studies made little to no mention of generation time with respect to duration of their experiments or study organisms used. Understanding how organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt et al 2020). This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta analysis. Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times. The converse is true for seasonal or decadal thermal fluctuations.

The results of this analysis differ from previous attempts to understand variability as they quantitatively address how variability affects organisms across taxa, not simply one group. Understanding how allometric scaling and thermal variability patterns interact is important for predicting realistic organismal responses in the face of variable conditions in the environment.

### 2021-04-28 writing /thesis_draft/04-conclusion.Rmd
an ugly but somewhat formatted draft

# Conclusions and Future Directions {-}

While there are limitations to the reach of this meta analysis and its results, it is the first cross-taxa quantitative summary of how environmental variability has been studied in the literature. Our results suggest that thermal variability influences responses across several studies, positively as lower mean temperatures with higher fluctuation ranges, and less so at high mean temperatures with lower fluctuation ranges.

However, most importantly, our results highlight the importance of learning from previous research to push how environmental variability is studied further. It has been noted for years the importance of including thermal variability in ecological research, yet there are still an overwhelming number of studies that employ the same patterns of variability with the same lack of reference to thermal performance. Going forward, experiments that explicitly account for and justify variation patterns with respect to generation time, organization level, duration, and natural relevance are needed. In order to draw conclusions that are relevant to natural conditions, patterns of variability across colors of noise (red, white, etc.) and across lengths of time must be investigated. A larger sample size of studies to draw more robust conclusions with respect to support for or against allometric scaling and MTE with respect to thermal variability would be useful. There is more to be studied on how additional experimental designs, i.e. acclimation or heat-wave simulations, and biological performance respond to environmental variability, however, this is an important first step in quantitatively summarizing progress so far.

### 2021-04-28 writing /thesis_draft/_book/thesis.tex
an ugly but somewhat formatted draft

@@
 % To pass between YAML and LaTeX the dollar signs are added by CII
-\title{My Final College Paper}
+\title{Mean temperature, fluctuation range, and biological organization affect biological responses}
 \author{Margaret A. Slein}
@@
 \Dedication{
-You can have a dedication here if you wish.
+
 }
@@
 \Abstract{
-The preface pretty much says it all.
-
-\par
-
-Second paragraph of abstract starts here.
+Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns.
+Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
+Here, we present a quantitative meta analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
+Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model.
+Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.
 }
 
+	\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 % End of CII addition
@@
   \begin{abstract}
-    The preface pretty much says it all.
+    Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns.
+Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
+Here, we present a quantitative meta analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
+Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model.
+Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.
+  \end{abstract}
 
-    \par
-
-    Second paragraph of abstract starts here.
-  \end{abstract}
-  \begin{dedication}
-    You can have a dedication here if you wish.
-  \end{dedication}
 \mainmatter % here the regular arabic numbering starts
@@
 
-Placeholder
-
 \hypertarget{the-historical-significance-of-variability}{%
 \section{The historical significance of variability}\label{the-historical-significance-of-variability}}
+
+It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bouchard 2014, Bambach 1993).
+Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Soo et al 2017, Blankenship 2017).
+Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll 2008).
+In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnik et al 2007).
+Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Reed et al 2010, Burggren 2018).
+Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.
 
@@
 
+In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
+Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Nino Southern Oscillation).
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (Khelifa et al 2019, Stewart et al 2013).
+How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
+As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Helmuth et al 2010, Bernardo 2014, Huey et al 2012).
+
 \hypertarget{why-this-study-on-variability-matters}{%
@@
 
+There is a large body of literature on environmental variability including its relevance for positive effects for populations and communities, i.e.~inflationary effects in sink-source systems (Holt et al 2003, Vasseur and Fox 2009), such that variability aids population density when organisms emigrate.
+However, there is a lack of consensus as to which scales of biological organization, i.e.~individuals, populations, or communities of organisms will be most severely affected by different shifts in environmental conditions (problem statement).
+We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response (aims).
+To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.
+
 \hypertarget{background}{%
-\chapter{Background}\label{background}}
+\chapter*{Background}\label{background}}
+\addcontentsline{toc}{chapter}{Background}
 
@@
 \section{Environmental variability}\label{environmental-variability}}
+\begin{figure}
+\includegraphics[width=1.05\linewidth]{figures/figure1} \caption{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
+Shown is a comparison of how duration of variability compares across the same environmental factor}\label{fig:unnamed-chunk-3}
+\end{figure}
+Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms' response to environmental conditions (Parepa et al 2013, Gudmundson et al 2015).
+Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1).
 
-Environmental variability describes the dynamic nature of environmental conditions in the natural world and how organisms respond (cite).
-Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure X).
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{figures/figure1} \caption{Figure 1.
-An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
-Shown is a comparison of how duration of variability compares across the same environmental factor}\label{fig:unnamed-chunk-1}
-\end{figure}
 \hypertarget{mean}{%
@@
 It provides a departure point for what we expect the average value to be for all time.
-We can then expand additional information adjacent to the mean (i.e.~range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk than just simply considering increases in global temperature (2014).
+In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, i.e.~averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations.
+However, in other cases, the mean is less useful compared to the variability.
+For instance, if there are key factors contributing to population growth rates from specific locations, comparing the amount of variability in additional variables across locations (individual respiration rates, temperature, etc.) is more fruitful for the task at this scale.
+At some point however, all data of interest are averages.
+The important distinction is the scale at which we average the data and the level of detail provided at each scale.
+Means are more useful in describing variation amongst normally distributed data, however, other statistical metrics like the median or mode may be more appropriate when the data is not normally distributed.
 
@@
 
+We can expand additional information adjacent to the mean (i.e.~range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014).
 We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
@@
 The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare et al 2013, 2012), as well as additional life history traits (size, speed, etc.).
+An organism's ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible (Amarasekare 2012).
+If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale their adaptive evolution can manage.
 
@@
 The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability.
-Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure X).
-Short term fluctuations, like Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time, have differing effects on performance (Colinet et al 2018).
-While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.~heat waves, cold fronts etc.).
-
-Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with organism, population, and community level processes.
-For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during that specific window of development (cite).
-Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (cite).
-When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (cite).
+Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1).
+Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet et al 2018).
+Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Husak and Lailvaux 2017, Amarasekare and Savage 2012).While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.~heat waves, cold fronts etc.) in which the benefits of FTR in mitigated chill injuries may be helpful.
+Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
+For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Delmas et al 2007, Bowden and Paitz 2018).
+Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (Lawson et al 2015) depending on the magnitude of variance.
+When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Romanuk and Kolasa 2002, Dobramys and Tauber 2013, Bastille-Rousseau et al 2018).
 
@@
 FTR can also describe the predictability of the fluctuation period.
-Variability may not have evenly spaced intervals (i.e.~heatwaves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
+Variability may not have evenly spaced intervals (i.e.~heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
 Predictability can also attempt to describe stochastic vs.~autocorrelated variation, however, that will be discussed in a different context below (Figure 4).
 
-Link between temporal and spatial variability
+\hypertarget{link-between-temporal-and-spatial-variability}{%
+\subsection{Link between temporal and spatial variability}\label{link-between-temporal-and-spatial-variability}}
+
 While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact.
-Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (cite).
+Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Horne and Schneider 1995, Farhang-Sardroodi et al 2019).
 Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e.~Diel vertical migration (DVM).
-In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (cite).
-However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (cite).
+In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (Lampert 1989, Brierly 2014).
+However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey 1974, Kearney et al 2011).
 
-Though environmental variability can include many different forms, from fluctuations in resource availability (cite) to light (cite), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (cite).
+Though environmental variability can include many different forms, from fluctuations in resource availability (Dempster and Pollard 1981, Sommer 1984) to light (Ruel and Ayres 1999, Morison 2020), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (Huey et al 1990, Colinet et al 2015).
 
@@
 Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as ``a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis'' (2018).
-In any case, temperature variability describes changes in environmental temperature through time (cite).
+Most generally, temperature variability describes changes in environmental temperature through time (Holmes et al 2016, Colinet et al 2018).
 
@@
 
-Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (cite).
+Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey et al 1990).
 Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE), though highly debated (Clarke et al 2006), which serves as an origin point for understanding how organisms respond to environmental conditions.
@@
 Schematic of thermal performance curve and it’s important components for understanding thermal performance.
-Adapted from (cite, cite)}\label{fig:unnamed-chunk-2}
+Adapted from (cite, cite)}\label{fig:unnamed-chunk-4}
 \end{figure}
 The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms' performance is negatively impacted (cite).
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms' performance is negatively impacted (Brown et al 2004, Dowd et al 2015).
 Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
+
+TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (i.e.~metabolism) (Amarasekare et al 2012).
+If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one.
+However, there are several other constraining factors (i.e.~energy acquisition, nutrient availability etc.) that derive the characteristically nonlinear pattern of TPCs.
+TPCs strive to represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
 
@@
 
-TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (cite, Bernhardt et al 2018).
+TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt et al 2018).
 However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.~15-35 C in 5 C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
@@
 Given the characteristically non-linear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
-Jensen's inequality describes the phenomenon that environmental variance can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999).
+Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999).
+For instance, Ruel and Ayres (1999) provide the development rate of ectotherms as an example to demonstrate Jensen's inequality.
+The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel and Ayres, 1999).
+With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel and Ayres, 1999).
 Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time.
-It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019.
+It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019).
 
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al (year) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (cite 2012).
-Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (cite 2012).
+For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare et al 2012, Paajimans et al 2013).
+Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare et al 2012).
 
@@
 \subsection{The relevancy of temperature variability in the modern context}\label{the-relevancy-of-temperature-variability-in-the-modern-context}}
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{figures/figure3} \caption{Figure 3.
-DON’T FORGET THIS IS LOG SCALED!!
-Plot of generation time in days against fluctuation period in days, with yellow, red, and grey horizontal panels representing diurnal, autocorrelated, and stochastic fluctuation patterns.
-The blue, green, purple, and pink horizontal lines represent different examples of variety in organismal generation time.
-*Figure out how to overlap pngs of plankton for 1, drosophila for 10, beetle for 100, and fish for 4000*}\label{fig:unnamed-chunk-3}
-\end{figure}
+
 Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
-Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance.
-Khelifa et al (2019) demonstrated that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{figures/figure4} \caption{Figure 4.
-Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
-Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.}\label{fig:unnamed-chunk-4}
-\end{figure}
-However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 4).
+Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
+Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
+
+However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 3).
 Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (2019).
@@
 Vasseur and Yodzis (2004) outline how the frequency (1/period) of environmental fluctuations describes the ``color of environmental noise.'' In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
-Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (cite).
+Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (Wigley et al 1998, Tabari et al 2012).
 Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figures/figure4} \caption{Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
+Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.}\label{fig:unnamed-chunk-5}
+\end{figure}
+\hypertarget{objectives-driving-questions-hypotheses}{%
+\section{Objectives, driving questions, hypotheses}\label{objectives-driving-questions-hypotheses}}
 
-Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level (cite):
+Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
+* At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve (Bartheld et al 2017) to changes in TSD ratios (Bowden and Paitz 2018)
+* At the population level, this usually comes in the form of decreasing population densities (Anders and Post 2006)
+* At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration (Descamps-Julien and Gonzalez 2005), inflationary effects (Holt et al 2003)
+It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
+Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (Wang et al 2016, Petchey et al 2000).
+
+There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al 2019, Massey et al 2020).
+To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions:
+1) Does variability pattern drive differences in response magnitude?
+(i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc)?
+We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
+
+We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and larger organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).
+
+\hypertarget{methods}{%
+\chapter*{Methods}\label{methods}}
+\addcontentsline{toc}{chapter}{Methods}
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figures/figure5} \caption{Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
+Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
+\end{figure}
+We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 3).
+After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*) ) AND SU=(Life Sciences \& Biomedicine).
+This search returned a total of 176 results.
+We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1).
+Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 5).
+Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
 \begin{itemize}
@@
 \item
-  At the individual level, this usually comes in the form of reduced fitness or increased development rates (cite)
+  We excluded studies that did not feature both a constant and fluctuating treatment (i.e.~studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
 \item
-  At the population level, this usually comes in the form of increases to population saturation
+  We excluded studies that were not completed in a laboratory (i.e.~modeling predictions or results from models, mesocosm experiments, in situ experiments)
 \item
-  At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration
+  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (i.e.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
+\item
+  We excluded any articles not written in English
+\item
+  We excluded studies that did not report measurements of error
+\item
+  We excluded studies in which the response variables are comparable or extrable (i.e.~one study performed Principal component analysis for their responses)
 \end{itemize}
-It however, less understood, is how variation patterns influence these characteristic responses at each level of organization.
-Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (cite).
+\begin{table}
 
-\hypertarget{objectives-driving-questions-hypotheses}{%
-\section{Objectives, driving questions, hypotheses}\label{objectives-driving-questions-hypotheses}}
+\caption{\label{tab:unnamed-chunk-8}Studies included in analysis of thermal variability (n=15)}
+\centering
+\begin{tabular}[t]{lll}
+\toprule
+\textbf{Study} & \textbf{Year} & \textbf{Extracted data points (n)}\\
+\midrule
+Delava et al & 2016 & 16\\
+Du et al & 2009 & 32\\
+Garcia-Ruiz et al & 2011 & 16\\
+Glass et al & 2009 & 16\\
+Kern et al & 2015 & 24\\
+\addlinespace
+Kern et al & 2014 & 4\\
+Lowenborg et al & 2012 & 10\\
+Maneti et al & 2014 & 24\\
+Pendlebury et al & 2004 & 16\\
+Qu et al & 2014 & 40\\
+\addlinespace
+Rolandi et al & 2018 & 4\\
+Saxon et al & 2017 & 12\\
+Semenov et al & 2007 & 32\\
+Simonici et al & 2019 & 30\\
+Triedel et al & 2015 & 4\\
+\bottomrule
+\end{tabular}
+\end{table}
+We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
+We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al 2004).
+The final number of data points included in the analysis is broken down by study (Table 1).
 
-There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (cite).
-To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions:
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Does variability pattern drive differences in response magnitude? (i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal?)
-\item
-  Does temperature variability drive differential response levels at different levels of biological organization?
-\item
-  How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc)?
-\end{enumerate}
-We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
+\hypertarget{data-extraction-and-analysis}{%
+\section{Data extraction and analysis}\label{data-extraction-and-analysis}}
 
-\hypertarget{lists}{%
-\section{Lists}\label{lists}}
+From the studies that met our inclusion criteria, we extracted data from tables and figures.
+Figures were extracted via png files imported into WebPlotDigitizer to obtain the raw data.
+The extracted data was then imported into a master spreadsheet of data from all studies for analysis.
+We extracted mean values for response variables, any measure of variance (Standard Deviation (SD) or Standard Error of the Mean (SEM)), and sample size.
+For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM (see code appendix).
+We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal history, and aggregated these values (Table 2).
+\begin{table}
 
-It's easy to create a list. It can be unordered like
-\begin{itemize}
-\tightlist
-\item
-  Item 1
-\item
-  Item 2
-\end{itemize}
-or it can be ordered like
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Item 1
-\item
-  Item 2
-\end{enumerate}
-Notice that I intentionally mislabeled Item 2 as number 4. \emph{Markdown} automatically figures this out! You can put any numbers in the list and it will create the list. Check it out below.
+\caption{\label{tab:unnamed-chunk-9}All covariates included in the model}
+\centering
+\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+\textbf{Variable} & \textbf{Description}\\
+\midrule
+mean\_temp\_constant & Reported mean temperature in study\\
+flux\_range & Reported range (difference between minimum 
+                                     and maximum) of temperature fluctuation experienced\\
+org\_level & Level of biological organization, categorized
+                                     as either 0 for individual or 1 for population level
+                                     responses\\
+size & Size estimate of study organism, 0 for extra-small
+                                     (microscopic), 1 for small (macroscopic), 2 for medium 
+                                     (less than a foot), and 3 for large (more than a foot)\\
+exp\_age & Life stage at which experiment was conducted, 
+                                     categorized as 0 for larval, 1 for juvenile, and 
+                                     2 for adult\\
+\addlinespace
+resp\_type & Response variables units classified as rates to traits, 
+                                     such that rates included any response variable units with a
+                                     “per day” clarifier were classified as rates, all else as
+                                     traits\\
+stressful & Whether or not the organisms experienced thermally stressful 
+                                     temperatures throughout study duration, if temperature range 
+                                     exceeded Topt, classified as thermally stressful, else not
+                                     thermally stressful. See appendix for references in which
+                                     organisms’ thermal history was included\\
+\bottomrule
+\end{tabular}
+\end{table}
+If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Cochrane Handbook for Systematic Reviews of Interventions, Version 5.1.0, 2011).
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
 
-To create a sublist, just indent the values a bit (at least four spaces or a tab). (Here's one case where indentation is key!)
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Item 1
-\item
-  Item 2
-\item
-  Item 3
-  \begin{itemize}
-  \tightlist
-  \item
-    Item 3a
-  \item
-    Item 3b
-  \end{itemize}
-\end{enumerate}
-\hypertarget{line-breaks}{%
-\section{Line breaks}\label{line-breaks}}
+\hypertarget{metafor}{%
+\section{Metafor}\label{metafor}}
 
-Make sure to add white space between lines if you'd like to start a new paragraph. Look at what happens below in the outputted document if you don't:
+We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
+Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer et al 2010).
+In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
+\begin{table}
 
-Here is the first sentence. Here is another sentence. Here is the last sentence to end the paragraph.
-This should be a new paragraph.
-
-\emph{Now for the correct way:}
-
-Here is the first sentence. Here is another sentence. Here is the last sentence to end the paragraph.
-
-This should be a new paragraph.
-
-\hypertarget{r-chunks}{%
-\section{R chunks}\label{r-chunks}}
-
-When you click the \textbf{Knit} button above a document will be generated that includes both content as well as the output of any embedded \textbf{R} code chunks within the document.
-You can embed an \textbf{R} code chunk like this (\texttt{cars} is a built-in \textbf{R} dataset):
+\caption{\label{tab:unnamed-chunk-10}Column names expanded to wider format for data analysis in metafor}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+\textbf{Column name} & \textbf{Purpose}\\
+\midrule
+Mf & Mean of experimental group\\
+SDf & Standard deviation of the experimental group\\
+Mc & Mean of the control group\\
+SDc & Standard deviation of the control group\\
+Nf & Sample size (n) of the experimental group\\
+\addlinespace
+Nc & Sample size (n) of the control group\\
+\bottomrule
+\end{tabular}
+\end{table}
+Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the escalc function in metafor to calculate effect size and sampling variance:
 \begin{Shaded}
 \begin{Highlighting}[]
-\KeywordTok{summary}\NormalTok{(cars)}
+\KeywordTok{escalc}\NormalTok{(}\DataTypeTok{measure=}\StringTok{"SMD"}\NormalTok{, }\DataTypeTok{m1i=}\StringTok{`}\DataTypeTok{constant_resp}\StringTok{`}\NormalTok{, }\DataTypeTok{m2i=}\NormalTok{flux_resp, }
+            \DataTypeTok{sd1i=}\StringTok{`}\DataTypeTok{SD_constant}\StringTok{`}\NormalTok{, }\DataTypeTok{sd2i=} \StringTok{`}\DataTypeTok{SD_variable}\StringTok{`}\NormalTok{, }\DataTypeTok{n1i=}\NormalTok{constant_samp, }
+       \DataTypeTok{n2i=}\NormalTok{flux_samp, }\DataTypeTok{data=}\NormalTok{dat_MA, }\DataTypeTok{slab=}\KeywordTok{paste}\NormalTok{(study_id, experiment_id, }
+\NormalTok{                                              response_id, }\DataTypeTok{sep=}\StringTok{", "}\NormalTok{))}
 \end{Highlighting}
 \end{Shaded}
-\begin{verbatim}
-     speed           dist       
- Min.   : 4.0   Min.   :  2.00  
- 1st Qu.:12.0   1st Qu.: 26.00  
- Median :15.0   Median : 36.00  
- Mean   :15.4   Mean   : 42.98  
- 3rd Qu.:19.0   3rd Qu.: 56.00  
- Max.   :25.0   Max.   :120.00  
-\end{verbatim}
-\hypertarget{inline-code}{%
-\section{Inline code}\label{inline-code}}
+We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head (Viechtbauer 2010).
+It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer 2010).
+Each of the subsequent arguments of the escalc function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
+The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 6).
+\begin{figure}
+\includegraphics[width=1\linewidth]{figures/figure6} \caption{Workflow diagram of how data were extracted, analyzed, and processed for downstream analysis in metafor.}\label{fig:unnamed-chunk-12}
+\end{figure}
+First, it calculates the pooled standard deviation (Pooled SD) from the sample sizes of both studies.
+Our data had studies with a wide range of sample sizes, so we wanted to include separate sample sizes between the constant and fluctuating treatments.
+The Pooled SD, Cohen's d, which is one of the most common ways to measure effect size, is used to normalize the mean differences between constant and fluctuating treatments.
+Then, using a correction factor J, which again normalizes the data further, we obtain the final SMD.
+The SMD is then ultimately used to calculate the sampling variance, which creates a unitless metric for variance.
 
-If you'd like to put the results of your analysis directly into your discussion, add inline code like this:
-\begin{quote}
-The \texttt{cos} of \(2 \pi\) is 1.
-\end{quote}
-Another example would be the direct calculation of the standard deviation:
-\begin{quote}
-The standard deviation of \texttt{speed} in \texttt{cars} is 5.2876444.
-\end{quote}
-One last neat feature is the use of the \texttt{ifelse} conditional statement which can be used to output text depending on the result of an \textbf{R} calculation:
-\begin{quote}
-The standard deviation is less than 6.
-\end{quote}
-Note the use of \texttt{\textgreater{}} here, which signifies a quotation environment that will be indented.
-
-As you see with \texttt{\$2\ \textbackslash{}pi\$} above, mathematics can be added by surrounding the mathematical text with dollar signs.
-More examples of this are in \protect\hyperlink{math-sci}{Mathematics and Science} if you uncomment the code in \protect\hyperlink{math}{Math}.
-
-\hypertarget{including-plots}{%
-\section{Including plots}\label{including-plots}}
-
-You can also embed plots. For example, here is a way to use the base \textbf{R} graphics package to produce a plot using the built-in \texttt{pressure} dataset:
-
-\includegraphics{thesis_files/figure-latex/pressure-1.pdf}
-
-Note that the \texttt{echo=FALSE} parameter was added to the code chunk to prevent printing of the \textbf{R} code that generated the plot.
-There are plenty of other ways to add chunk options (like \texttt{fig.height} and \texttt{fig.width} in the chunk above).
-More information is available at \url{https://yihui.org/knitr/options/}.
-
-Another useful chunk option is the setting of \texttt{cache=TRUE} as you see here.
-If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.
-Later in this file, you'll see a way to reference plots created in \textbf{R} or external figures.
-
-\hypertarget{loading-and-exploring-data}{%
-\section{Loading and exploring data}\label{loading-and-exploring-data}}
-
-Included in this template is a file called \texttt{flights.csv}.
-This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014.
-More information about this dataset and its \textbf{R} package is available at \url{https://github.com/ismayc/pnwflights14}.
-This subset includes only Portland flights and only rows that were complete with no missing values.
-Merges were also done with the \texttt{airports} and \texttt{airlines} data sets in the \texttt{pnwflights14} package to get more descriptive airport and airline names.
-
-We can load in this data set using the following commands:
+Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response (see insert of pseudo code).
 \begin{Shaded}
 \begin{Highlighting}[]
-\CommentTok{# flights.csv is in the data directory}
-\NormalTok{flights_path <-}\StringTok{ }\NormalTok{here}\OperatorTok{::}\KeywordTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"flights.csv"}\NormalTok{)}
-\CommentTok{# string columns will be read in as strings and not factors now}
-\NormalTok{flights <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(flights_path, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
+\KeywordTok{rma.mv}\NormalTok{(yi, vi, }
+       \DataTypeTok{data=}\NormalTok{dat_MA_ES, }
+       \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{ }\NormalTok{study_id }\OperatorTok{/}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id, }
+       \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
 \end{Highlighting}
 \end{Shaded}
-The data is now stored in the data frame called \texttt{flights} in \textbf{R}.
-To get a better feel for the variables included in this dataset we can use a variety of functions.
-Here we can see the dimensions (rows by columns) and also the names of the columns.
+The arguments of the rma.mv function include SMD (yi), sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
+We specified a nested structure to account for the colinearity of data points from within the same study, experiment, and response metric.
+We used Restricted Maximum-Likelihood (REML) to specify the fit of the model, as REML is best at handling small samples sizes compared to Maximum Likelihood (ML) (Viechtbauer 2010).
+We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population (Viechtbauer 2010).
+
+We then ran two multivariate random effects models, one with modifiers and one without (see insert above).
+This was to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies.
+The full model we ran included:
 \begin{Shaded}
 \begin{Highlighting}[]
-\KeywordTok{dim}\NormalTok{(flights)}
+\KeywordTok{rma.mv}\NormalTok{(yi, vi, }\DataTypeTok{data=}\NormalTok{dat_MA_ES, }\DataTypeTok{mods =} \OperatorTok{~}\NormalTok{flux_range }\OperatorTok{*}\StringTok{ }\NormalTok{mean_temp_constant }
+       \OperatorTok{+}\StringTok{ }\NormalTok{exp_age }\OperatorTok{+}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{org_level }\OperatorTok{+}\StringTok{ }\NormalTok{resp_type, }
+               \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{  }\NormalTok{study_id}\OperatorTok{/}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id,}
+                 \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
 \end{Highlighting}
 \end{Shaded}
-\begin{verbatim}
-[1] 12649    16
-\end{verbatim}
+We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal history data for one study.
+That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz 2011, Table 1).
+We ran a separate model without this study and included the thermal history covariate separately from all the other modifiers:
 \begin{Shaded}
 \begin{Highlighting}[]
-\KeywordTok{names}\NormalTok{(flights)}
+\KeywordTok{rma.mv}\NormalTok{(yi, vi, }\DataTypeTok{data=}\NormalTok{dat_MA_ES, }\DataTypeTok{mods =}  \OperatorTok{~}\StringTok{ }\NormalTok{stressful, }
+               \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{study_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id,}
+                 \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
 \end{Highlighting}
 \end{Shaded}
-\begin{verbatim}
- [1] "month"        "day"          "dep_time"     "dep_delay"    "arr_time"    
- [6] "arr_delay"    "carrier"      "tailnum"      "flight"       "dest"        
-[11] "air_time"     "distance"     "hour"         "minute"       "carrier_name"
-[16] "dest_name"   
-\end{verbatim}
-Another good idea is to take a look at the dataset in table form.
-With this dataset having more than 20,000 rows, we won't explicitly show the results of the command here.
-I recommend you enter the command into the Console \textbf{\emph{after}} you have run the \textbf{R} chunks above to load the data into \textbf{R}.
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{View}\NormalTok{(flights)}
-\end{Highlighting}
-\end{Shaded}
-While not required, it is highly recommended you use the \texttt{dplyr} package to manipulate and summarize your data set as needed.
-It uses a syntax that is easy to understand using chaining operations.
-Below I've created a few examples of using \texttt{dplyr} to get information about the Portland flights in 2014.
-You will also see the use of the \texttt{ggplot2} package, which produces beautiful, high-quality academic visuals.
+\hypertarget{results}{%
+\chapter*{Results}\label{results}}
+\addcontentsline{toc}{chapter}{Results}
 
-We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# List of packages required for this analysis}
-\NormalTok{pkg <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{, }\StringTok{"ggplot2"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"bookdown"}\NormalTok{)}
-\CommentTok{# Check if packages are not installed and assign the}
-\CommentTok{# names of the packages not installed to the variable new.pkg}
-\NormalTok{new.pkg <-}\StringTok{ }\NormalTok{pkg[}\OperatorTok{!}\NormalTok{(pkg }\OperatorTok{%in%}\StringTok{ }\KeywordTok{installed.packages}\NormalTok{())]}
-\CommentTok{# If there are any packages in the list that aren't installed,}
-\CommentTok{# install them}
-\ControlFlowTok{if}\NormalTok{ (}\KeywordTok{length}\NormalTok{(new.pkg)) \{}
-  \KeywordTok{install.packages}\NormalTok{(new.pkg, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{\}}
-\CommentTok{# Load packages}
-\KeywordTok{library}\NormalTok{(thesisdown)}
-\KeywordTok{library}\NormalTok{(dplyr)}
-\end{Highlighting}
-\end{Shaded}
-\begin{verbatim}
-Warning: package 'dplyr' was built under R version 4.0.2
-\end{verbatim}
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(ggplot2)}
-\end{Highlighting}
-\end{Shaded}
-\begin{verbatim}
-Warning: package 'ggplot2' was built under R version 4.0.2
-\end{verbatim}
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(knitr)}
-\end{Highlighting}
-\end{Shaded}
-\begin{verbatim}
-Warning: package 'knitr' was built under R version 4.0.2
-\end{verbatim}
-\clearpage
+Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 6.
+These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
+As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
+\begin{figure}
+\includegraphics[width=1.1\linewidth]{thesis_files/figure-latex/unnamed-chunk-17-1} \caption{Forest plot of effects sizes by study and experiment within a study.
+Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study.
+Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-17}
+\end{figure}
+Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal history), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p \textless{} 0.0001).
+Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (df = 139, p \textless{} 0.00001).
+Organization level was also important in the model, as a statistically significant covariate (df= 139, p \textless{} 0.05).
+\begin{figure}
+\includegraphics[width=1.05\linewidth]{thesis_files/figure-latex/unnamed-chunk-18-1} \caption{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.}\label{fig:unnamed-chunk-18}
+\end{figure}
+Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 7).The most negative estimate (organization level) was significantly influential in our model at a significance level of p \textless{} 0.05.
+Surprisingly, the most negative estimates of our model coefficients were not statistically significant.
+In our analysis of thermal history via a separate random effects model, we found thermal history was not statistically significant in explaining the effect size (Table 3).
+\begin{table}
 
-The example we show here does the following:
-\begin{itemize}
-\item
-  Selects only the \texttt{carrier\_name} and \texttt{arr\_delay} from the \texttt{flights} dataset and then assigns this subset to a new variable called \texttt{flights2}.
-\item
-  Using \texttt{flights2}, we determine the largest arrival delay for each of the carriers.
-\end{itemize}
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{flights2 <-}\StringTok{ }\NormalTok{flights }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(carrier_name, arr_delay)}
-\NormalTok{max_delays <-}\StringTok{ }\NormalTok{flights2 }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{group_by}\NormalTok{(carrier_name) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{max_arr_delay =} \KeywordTok{max}\NormalTok{(arr_delay, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-A useful function in the \texttt{knitr} package for making nice tables in \emph{R Markdown} is called \texttt{kable}.
-It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.
-This again goes to show how nice reproducible documents can be!
-(Note the use of \texttt{results="asis"}, which will produce the table instead of the code to create the table.) The \texttt{caption.short} argument is used to include a shorter title to appear in the List of Tables.
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{kable}\NormalTok{(max_delays,}
-  \DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"Airline"}\NormalTok{, }\StringTok{"Max Arrival Delay"}\NormalTok{),}
-  \DataTypeTok{caption =} \StringTok{"Maximum Delays by Airline"}\NormalTok{,}
-  \DataTypeTok{caption.short =} \StringTok{"Max Delays by Airline"}\NormalTok{,}
-  \DataTypeTok{longtable =} \OtherTok{TRUE}\NormalTok{,}
-  \DataTypeTok{booktabs =} \OtherTok{TRUE}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-\begin{longtable}[t]{lr}
-\caption[Max Delays by Airline]{\label{tab:maxdelays}Maximum Delays by Airline}\\
+\caption{\label{tab:unnamed-chunk-19}Model summary statistics for random effects model using thermal history as sole modifier}
+\centering
+\begin{tabular}[t]{lrrrrrr}
 \toprule
-Airline & Max Arrival Delay\\
+\textbf{Term} & \textbf{estimate} & \textbf{se} & \textbf{zval} & \textbf{pval} & \textbf{ci.lb} & \textbf{ci.ub}\\
 \midrule
-Alaska Airlines Inc. & 338\\
-American Airlines Inc. & 1539\\
-Delta Air Lines Inc. & 371\\
-Frontier Airlines Inc. & 166\\
-Hawaiian Airlines Inc. & 116\\
-\addlinespace
-JetBlue Airways & 256\\
-SkyWest Airlines Inc. & 321\\
-Southwest Airlines Co. & 315\\
-United Air Lines Inc. & 319\\
-US Airways Inc. & 347\\
-\addlinespace
-Virgin America & 366\\
+intercept & 0.2050 & 0.1846 & 1.1104 & 0.2668 & -0.1568 & 0.5669\\
+stressful & -0.0489 & 0.0702 & -0.6975 & 0.4855 & -0.1864 & 0.0886\\
 \bottomrule
-\end{longtable}
-The last two options make the table a little easier-to-read.
+\end{tabular}
+\end{table}
+The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 8.
+At lower mean temperatures, higher fluctuation ranges have a more positive effect on\ldots{}, though that trend starkly ends at about 24O C.
+Beyond 24O C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 8).
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-20-1} \caption{Scatterplot describing the relationship between mean temperature, temperature range, and effect size.
+Temperature fluctuation range from 0-20C, Mean temperature range from 7-33OC.
+Effect size range from -2 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)}\label{fig:unnamed-chunk-20}
+\end{figure}
+The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 9).
+Interestingly, there were not any significant differences between thermal history (Figure 10), life history metric (Figure 11), body size (Figure 12), or age (Figure 13).
 
-We can further look into the properties of the largest value here for American Airlines Inc.~To do so, we can isolate the row corresponding to the arrival delay of 1539 minutes for American in our original \texttt{flights} dataset.
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{flights }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(}
-\NormalTok{    arr_delay }\OperatorTok{==}\StringTok{ }\DecValTok{1539}\NormalTok{,}
-\NormalTok{    carrier_name }\OperatorTok{==}\StringTok{ "American Airlines Inc."}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}
-\NormalTok{    month, day, carrier, dest_name, hour,}
-\NormalTok{    minute, carrier_name, arr_delay}
-\NormalTok{  ))}
-\end{Highlighting}
-\end{Shaded}
-\begin{verbatim}
-  dep_time dep_delay arr_time tailnum flight dest air_time distance
-1     1403      1553     1934  N595AA   1568  DFW      182     1616
-\end{verbatim}
-We see that the flight occurred on March 3rd and departed a little after 2 PM on its way to Dallas/Fort Worth.
-Lastly, we show how we can visualize the arrival delay of all departing flights from Portland on March 3rd against time of departure.
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{flights }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(month }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{, day }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ dep_time, }\DataTypeTok{y =}\NormalTok{ arr_delay)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-\includegraphics{thesis_files/figure-latex/march3plot-1.pdf}
+As noted in the methods, when we excluded one highly influential study and included thermal history (exclusion model), size became much more important in the thermal stress model (df=123, p \textless{} 0.001) as did mean temperature.
+Surprisingly, organization level, fluctuation range, and the interaction between range and mean temperature became insignificant in the thermal stress model (df=123, p \textless{} 0.1).
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-21-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-21}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-22-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal history: stressful or not stressful.}\label{fig:unnamed-chunk-22}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-23-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-23}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-24-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-24}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-25-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-25}
+\end{figure}
+\hypertarget{discussion}{%
+\chapter*{Discussion}\label{discussion}}
+\addcontentsline{toc}{chapter}{Discussion}
 
-\hypertarget{additional-resources}{%
-\section{Additional resources}\label{additional-resources}}
-\begin{itemize}
-\item
-  \emph{Markdown} Cheatsheet - \url{https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet}
-\item
-  \emph{R Markdown}
-  \begin{itemize}
-  \tightlist
-  \item
-    Reference Guide - \url{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}
-  \item
-    Cheatsheet - \url{https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf}
-  \end{itemize}
-\item
-  \emph{RStudio IDE}
-  \begin{itemize}
-  \tightlist
-  \item
-    Cheatsheet - \url{https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf}
-  \item
-    Official website - \url{https://rstudio.com/products/rstudio/}
-  \end{itemize}
-\item
-  Introduction to \texttt{dplyr} - \url{https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html}
-\item
-  \texttt{ggplot2}
-  \begin{itemize}
-  \tightlist
-  \item
-    Documentation - \url{https://ggplot2.tidyverse.org/}
-  \item
-    Cheatsheet - \url{https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf}
-  \end{itemize}
-\end{itemize}
-\hypertarget{math-sci}{%
-\chapter{Mathematics and Science}\label{math-sci}}
+The results of this meta analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level.
+There were not overall differences in the direction o,
 
-Placeholder
+These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e.~higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures.
+The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change (Vasseur et al 2014, Sunday et al 2012).
+With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed ``the point of no return'' (Russill and Nyssa 2009), understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades (West et al 2002, Cheung et al 2008).
+In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature (Vasseur et al 2014).
 
-\hypertarget{math}{%
-\section{Math}\label{math}}
+These conclusions further drive home the decadal conclusion that Jensen's Inequality has demonstrated: averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance (Ruel and Ayres 1999, Bernhardt et al 2018).
+The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive.
 
-\hypertarget{chemistry-101-symbols}{%
-\section{Chemistry 101: Symbols}\label{chemistry-101-symbols}}
+Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal history, life history traits, and experimental age.
+Body size has been identified as important in the allometric scaling of MTE and it is interesting in these subsets of studies, body size did not explain the differences in responses across studies.
+This could simply be because we had a small sample size of studies (n=15), though multiple responses in studies contributed to a large sample of effect sizes (n=139) and there were only a subset of relative body sizes included.
+It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms (Clark and Johnston 1999, Clarke et al 2006).
+This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns variability in temperature drives.
 
-\hypertarget{typesetting-reactions}{%
-\subsection{Typesetting reactions}\label{typesetting-reactions}}
+An important side note is that our full model and thermal history model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
+This study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding Garcia-Ruiz et al (2011) from our full model.
+We decided to include this study in our analysis because the data were extracted from a table, therefore there were no inaccuracies in obtaining the data.
+It provides an important counter and discussion for how and why responses from different studies may be so different from others.
 
-\hypertarget{other-examples-of-reactions}{%
-\subsection{Other examples of reactions}\label{other-examples-of-reactions}}
+While body size and thermal history emerged as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship thermal variability and MTE.
+If we were able to obtain additional information about the CTmax, CTmin, thermal breadth, and tolerance range for each of the species included in the analysis, we may also be able to better understand the trends and the significance of variability patterns in dictating responses.
 
-\hypertarget{physics}{%
-\section{Physics}\label{physics}}
+Both the type of response (i.e.~whether it was a rate or trait) and experimental age were also not important predictors in our full model.
+We expected rates and traits to have different effects, as traits may be more heavily affected by variability because they are not as temporally dynamic as rates.
+The lack of difference between rates and traits may be because variability is not differentially acting on rate or trait based responses, simply dampening their overall effects.
+Again, this may be an artifact of having a small sample size, with mainly trait based responses as opposed to rate based responses.
 
-\hypertarget{biology}{%
-\section{Biology}\label{biology}}
+We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018).
+This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile.
+However, there have been meta analyses, both qualitative (Massey et al 2020), and quantitative (Noble et al 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al 2018).
+These results align with patterns of juvenile organisms (i.e.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Portner and Farrell 2008, Moyano et al 2017).
 
-\hypertarget{ref-labels}{%
-\chapter{Graphics, References, and Labels}\label{ref-labels}}
+Beyond the data included in our meta analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
+As Figure 4 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures.
+It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated in fact.
+Are these diurnal temperature fluctuations accurate for what is expected under natural conditions?
+This lack of variety in pattern may be due to a lack of consensus on the language used to classify and discuss thermal variability.
+Though there is a large body of literature that investigates temperature variation, there appears to be dissonance in the scientific literature about the language used to discuss such variation.
+Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures (Amarasekare et al 2012) as well as different ranges of temperature, both of which are different concepts.
+For instance, Amarasekare et al 2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits.
+Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
+Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes (Colinet et al 2018, Colinet et al 2015).
+Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.
 
-Placeholder
+Many of these studies failed to justify why certain temperatures were chosen as mean temperatures as well as the range of temperatures studied.
+In order to compile information on the thermal history of each of the organisms included in the studies of this meta analysis, additional resources were needed (see Appendix Table S1).
+Understanding where these organisms' performance falls in relation to temperature on its TPC is crucial for drawing reliable conclusions of how thermal variability affects performance.
+Without it, only broad conclusions can be drawn, as have been drawn here, that variability and range affects thermal performance.
 
-\hypertarget{figures}{%
-\section{Figures}\label{figures}}
+Many of the studies made little to no mention of generation time with respect to duration of their experiments or study organisms used.
+Understanding how organisms' life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt et al 2020).
+This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta analysis.
+Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times.
+The converse is true for seasonal or decadal thermal fluctuations.
 
-\hypertarget{footnotes-and-endnotes}{%
-\section{Footnotes and Endnotes}\label{footnotes-and-endnotes}}
+The results of this analysis differ from previous attempts to understand variability as they quantitatively address how variability affects organisms across taxa, not simply one group.
+Understanding how allometric scaling and thermal variability patterns interact is important for predicting realistic organismal responses in the face of variable conditions in the environment.
 
-\hypertarget{bibliographies}{%
-\section{Bibliographies}\label{bibliographies}}
+\hypertarget{conclusions-and-future-directions}{%
+\chapter*{Conclusions and Future Directions}\label{conclusions-and-future-directions}}
+\addcontentsline{toc}{chapter}{Conclusions and Future Directions}
 
-\hypertarget{anything-else}{%
-\section{Anything else?}\label{anything-else}}
+While there are limitations to the reach of this meta analysis and its results, it is the first cross-taxa quantitative summary of how environmental variability has been studied in the literature.
+Our results suggest that thermal variability influences responses across several studies, positively as lower mean temperatures with higher fluctuation ranges, and less so at high mean temperatures with lower fluctuation ranges.
 
-\hypertarget{conclusion}{%
-\chapter*{Conclusion}\label{conclusion}}
-\addcontentsline{toc}{chapter}{Conclusion}
-
-If we don't want Conclusion to have a chapter number next to it, we can add the \texttt{\{-\}} attribute.
-
-\textbf{More info}
-
-And here's some other random info: the first paragraph after a chapter title or section head \emph{shouldn't be} indented, because indents are to tell the reader that you're starting a new paragraph.
-Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.
+However, most importantly, our results highlight the importance of learning from previous research to push how environmental variability is studied further.
+It has been noted for years the importance of including thermal variability in ecological research, yet there are still an overwhelming number of studies that employ the same patterns of variability with the same lack of reference to thermal performance.
+Going forward, experiments that explicitly account for and justify variation patterns with respect to generation time, organization level, duration, and natural relevance are needed.
+In order to draw conclusions that are relevant to natural conditions, patterns of variability across colors of noise (red, white, etc.) and across lengths of time must be investigated.
+A larger sample size of studies to draw more robust conclusions with respect to support for or against allometric scaling and MTE with respect to thermal variability would be useful.
+There is more to be studied on how additional experimental designs, i.e.~acclimation or heat-wave simulations, and biological performance respond to environmental variability, however, this is an important first step in quantitatively summarizing progress so far.
 
@@
 \textbf{In the main Rmd file}
-
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# This chunk ensures that the thesisdown package is}
+\CommentTok{# installed and loaded. This thesisdown package includes}
+\CommentTok{# the template files for the thesis.}
+\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
+  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
+    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \KeywordTok{stop}\NormalTok{(}
+      \KeywordTok{paste}\NormalTok{(}\StringTok{'You need to run install.packages("remotes")",}
+\StringTok{            "first in the Console.'}\NormalTok{)}
+\NormalTok{    )}
+\NormalTok{  \}}
+\NormalTok{\}}
+\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
+  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
+\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \KeywordTok{stop}\NormalTok{(}
+      \KeywordTok{paste}\NormalTok{(}
+        \StringTok{"You need to run"}\NormalTok{,}
+        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
+        \StringTok{"first in the Console."}
+\NormalTok{      )}
+\NormalTok{    )}
+\NormalTok{  \}}
+\NormalTok{\}}
+\KeywordTok{library}\NormalTok{(thesisdown)}
+\CommentTok{# Set how wide the R output will go}
+\KeywordTok{options}\NormalTok{(}\DataTypeTok{width =} \DecValTok{70}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 \textbf{In Chapter \ref{ref-labels}:}
-
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# This chunk ensures that the thesisdown package is}
+\CommentTok{# installed and loaded. This thesisdown package includes}
+\CommentTok{# the template files for the thesis and also two functions}
+\CommentTok{# used for labeling and referencing}
+\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
+  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
+    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \KeywordTok{stop}\NormalTok{(}
+      \KeywordTok{paste}\NormalTok{(}
+        \StringTok{'You need to run install.packages("remotes")'}\NormalTok{,}
+        \StringTok{"first in the Console."}
+\NormalTok{      )}
+\NormalTok{    )}
+\NormalTok{  \}}
+\NormalTok{\}}
+\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(dplyr)) \{}
+  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
+    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \KeywordTok{stop}\NormalTok{(}
+      \KeywordTok{paste}\NormalTok{(}
+        \StringTok{'You need to run install.packages("dplyr")'}\NormalTok{,}
+        \StringTok{"first in the Console."}
+\NormalTok{      )}
+\NormalTok{    )}
+\NormalTok{  \}}
+\NormalTok{\}}
+\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(ggplot2)) \{}
+  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
+    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \KeywordTok{stop}\NormalTok{(}
+      \KeywordTok{paste}\NormalTok{(}
+        \StringTok{'You need to run install.packages("ggplot2")'}\NormalTok{,}
+        \StringTok{"first in the Console."}
+\NormalTok{      )}
+\NormalTok{    )}
+\NormalTok{  \}}
+\NormalTok{\}}
+\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(bookdown)) \{}
+  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
+    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"bookdown"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \KeywordTok{stop}\NormalTok{(}
+      \KeywordTok{paste}\NormalTok{(}
+        \StringTok{'You need to run install.packages("bookdown")'}\NormalTok{,}
+        \StringTok{"first in the Console."}
+\NormalTok{      )}
+\NormalTok{    )}
+\NormalTok{  \}}
+\NormalTok{\}}
+\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
+  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
+\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \KeywordTok{stop}\NormalTok{(}
+      \KeywordTok{paste}\NormalTok{(}
+        \StringTok{"You need to run"}\NormalTok{,}
+        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
+        \StringTok{"first in the Console."}
+\NormalTok{      )}
+\NormalTok{    )}
+\NormalTok{  \}}
+\NormalTok{\}}
+\KeywordTok{library}\NormalTok{(thesisdown)}
+\KeywordTok{library}\NormalTok{(dplyr)}
+\KeywordTok{library}\NormalTok{(ggplot2)}
+\KeywordTok{library}\NormalTok{(knitr)}
+\NormalTok{flights <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data/flights.csv"}\NormalTok{, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 \hypertarget{the-second-appendix-for-fun}{%
 \chapter{The Second Appendix, for Fun}\label{the-second-appendix-for-fun}}
+
+\backmatter
 
@@
 
-Placeholder
+\markboth{References}{References}
+
+\noindent
+
+\setlength{\parindent}{-0.20in}
+\setlength{\leftskip}{0.20in}
+\setlength{\parskip}{8pt}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-angel2000}{}%
+Angel, E. (2000). \emph{Interactive computer graphics : A top-down approach with opengl}. Boston, MA: Addison Wesley Longman.
+
+\leavevmode\hypertarget{ref-angel2001}{}%
+Angel, E. (2001a). \emph{Batch-file computer graphics : A bottom-up approach with quicktime}. Boston, MA: Wesley Addison Longman.
+
+\leavevmode\hypertarget{ref-angel2002a}{}%
+Angel, E. (2001b). \emph{Test second book by angel}. Boston, MA: Wesley Addison Longman.
 

### 2021-04-28 writing /thesis_draft/index.Rmd
an ugly but somewhat formatted draft

@@
 ---
-title: 'My Final College Paper'
+title: 'Mean temperature, fluctuation range, and biological organization affect biological responses'
 author: 'Margaret A. Slein'
@@
   thesisdown::thesis_pdf: default 
-abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
+abstract: 'Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns.
+Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
+Here, we present a quantitative meta analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
+Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model.
+Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.'
 acknowledgements: |
@@
 
-dedication: |
-  You can have a dedication here if you wish. 
 bibliography: bib/thesis.bib
@@
 ## The historical significance of variability 
-It is no secret that the natural world has relied on resilience and persistence in the face of environmental change for millenia (cite).
-Cyanobacteria can trace their origin back to a bacteria that evolved the ability to perform oxygenic photosynthesis to persist in a changing world (Soo et al 2017, science).
-Similarly, vascular plants have evolved different photosynthetic systems (C3, CAM) to manage a variety of thermal and spatial conditions (cite).
-While cyanobacteria and vascular plants evolved new strategies and structures to persist in the environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity may become even more relevant (cite).
-Understanding how organisms respond to the environment and its dynamic nature is significant for how to capitulate ecosystems currently and in the future.
+It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bouchard 2014, Bambach 1993).
+Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Soo et al 2017, Blankenship 2017).
+Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll 2008).
+In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnik et al 2007).
+Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Reed et al 2010, Burggren 2018).
+Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.
 
@@
 In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
-Ecosystems and organisms inhabiting those environments have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.
-ENSO).
-The scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (cite).
+Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.
+El Nino Southern Oscillation).
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (Khelifa et al 2019, Stewart et al 2013).
+How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
 As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.
-hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important (cite).
+hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Helmuth et al 2010, Bernardo 2014, Huey et al 2012).
 
 ## Why this study on variability matters 
-While there is a large body of literature on environmental variability including its relevance for X, y, and Z (include citation), there is a lack of consensus as to which scales of biological organization will be most severely affected by different shifts in environmental conditions (problem statement).
-We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as whether different patterns of variation differentially affect different levels of organization (aims).
+There is a large body of literature on environmental variability including its relevance for positive effects for populations and communities, i.e.
+inflationary effects in sink-source systems (Holt et al 2003, Vasseur and Fox 2009), such that variability aids population density when organisms emigrate.
+However, there is a lack of consensus as to which scales of biological organization, i.e.
+individuals, populations, or communities of organisms will be most severely affected by different shifts in environmental conditions (problem statement).
+We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response (aims).
 To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.

### 2021-04-29 writing /thesis_draft/01-Background.Rmd
fully formatted draft :)

--- 2021-04-28 writing /thesis_draft/01-chap1.Rmd
@@
 The chunk above is called intro and the one below is called chapter1.
-Feel free to change the name of the Rmd file as you wish, but don't forget to change it here from chap1.Rmd.
+Feel free to change the name of the Rmd file as you wish, but don't forget to change it here from 01-Background.Rmd.
 -->
@@
 
-# Background {-}
+# Background {}
 ## Environmental variability
@@
 ```{r, echo=FALSE, fig.cap="An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
-Shown is a comparison of how duration of variability compares across the same environmental factor", out.width = '105%'}
+Shown is a comparison of how duration of variability compares within the same environmental factor.
+Terms highlighted in each panel are described below in the main text", out.width = '105%', fig.align="center", fig.scap="Metrics to describe and explain environmental variability"}
 knitr::include_graphics("figures/figure1.png")
@@
 
-Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms’ response to environmental conditions (Parepa et al 2013, Gudmundson et al 2015).
-Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1).
+Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms’ response to environmental conditions [@parepa_environmental_2013; @gudmundson_environmental_2015].
+Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1.1).
 
@@
 We can expand additional information adjacent to the mean (i.e.
-range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014).
+range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature [@vasseur_increased_2014].
 We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
@@
 The Amazon Rainforest experiences less variation in light availability during the year than Denali National Park in Alaska, which experiences high variation in light availability between the summer and winter seasons.
-The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare et al 2013, 2012), as well as additional life history traits (size, speed, etc.).
-An organism’s ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible (Amarasekare 2012).
-If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale their adaptive evolution can manage.
+The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature [@amarasekare_framework_2012; @amarasekare_intrinsic_2013], as well as additional life history traits (size, speed, etc.).
+An organism’s ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible [@amarasekare_framework_2012].
+If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.
 
@@
 The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability.
-Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1).
-Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet et al 2018).
-Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment (Husak and Lailvaux 2017, Amarasekare and Savage 2012).While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1.1).
+Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance [@colinet_mechanisms_2018].
+Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment [@husak_how_2017; @amarasekare_framework_2012].
+While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
 However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.
@@
 Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
-For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Delmas et al 2007, Bowden and Paitz 2018).
-Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (Lawson et al 2015) depending on the magnitude of variance.
-When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Romanuk and Kolasa 2002, Dobramys and Tauber 2013, Bastille-Rousseau et al 2018).
+For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development [@delmas_mechanistic_2007; @bowden_temperature_2018].
+Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth [@lawson_environmental_2015] depending on the magnitude of variance.
+When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics [@romanuk_environmental_2002; @dobramysl_environmental_2013; @bastille-rousseau_climate_2018].
 
@@
 Predictability can also attempt to describe stochastic vs.
-autocorrelated variation, however, that will be discussed in a different context below (Figure 4).
+autocorrelated variation, however, that will be discussed in a different context below (Figure 1.3).
 
@@
 While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact.
-Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Horne and Schneider 1995, Farhang-Sardroodi et al 2019).
+Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space [@horne_spatial_1995; @farhang-sardroodi_environmental_2019].
 Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e.
 Diel vertical migration (DVM).
-In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (Lampert 1989, Brierly 2014).
+In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability [@lampert_adaptive_1989; @brierley_diel_2014].
 However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.
-an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey 1974, Kearney et al 2011).
+an organism moving from sun to shade during the day to regulate body temperature and metabolic processes [@huey_behavioral_1974; @kearney_potential_2009].
 
-Though environmental variability can include many different forms, from fluctuations in resource availability (Dempster and Pollard 1981, Sommer 1984) to light (Ruel and Ayres 1999, Morison 2020), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (Huey et al 1990, Colinet et al 2015).
+Though environmental variability can include many different forms, from fluctuations in resource availability 
+[@dempster_fluctuations_1981; @sommer_paradox_1984] to light [@ruel_jensens_1999; @morison_light_2020], one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations [@huey_physiological_1990; @colinet_insects_2015].
 
@@
 Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as “a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis” (2018).
-Most generally, temperature variability describes changes in environmental temperature through time (Holmes et al 2016, Colinet et al 2018).
+Most generally, temperature variability describes changes in environmental temperature through time [@holmes_robust_2016; @colinet_mechanisms_2018].
 
@@
 
-Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey et al 1990).
-Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of  ecology (MTE), though highly debated (Clarke et al 2006), which serves as an origin point for understanding how organisms respond to environmental conditions.
+Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment [@huey_physiological_1990].
+Brown et al highlighted the importance of both temperature and body size in the metabolic theory of  ecology (MTE) [@brown_toward_2004].
+Though highly debated [@clarke_temperature_2006], which serves as an origin point for understanding how organisms respond to environmental conditions.
 MTE is described by the following equation:
@@
 $$
+
 Where $B$ is organismal metabolic rate, $b_0$ is a normalization constant unrelated to temperate or body size, $M$ is organismal mass, and $e^-E/kT$ is referred to as the “Boltzman factor”, where $E$ is activation energy, $k$ is the Boltzman constant and $T$ is temperature.
@@
 We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates should be the same.
-However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clark and Johnston 1999, Clarke et al 2006).
-Others believe that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al 2006).
+However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish [@clarke_scaling_1999; @clarke_temperature_2006].
+Others believe that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups [@gillooly_response_2006].
 Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.
@@
 MTE supports temperature impacting metabolic rate, which can have downstream effects on fitness characteristics.
-This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 2).
+This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
 
-```{r, echo=FALSE, fig.cap="Figure 2. Schematic of thermal performance curve and it’s important components for understanding thermal performance. Adapted from (cite, cite)", out.width = '90%'}
+```{r, echo=FALSE, fig.cap="Schematic of thermal performance curve and it’s important components for understanding thermal performance.
+Adapted from (Krenek et al 2012, Tuff et al 2016)", fig.align="center", out.width = '90%',fig.scap="Thermal performance curve components"}
 knitr::include_graphics("figures/figure2.png")
@@
 The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms’ performance is negatively impacted (Brown et al 2004, Dowd et al 2015).
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms’ performance is negatively impacted [@brown_toward_2004; @dowd_thermal_2015].
 Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
@@
 TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (i.e.
-metabolism) (Amarasekare et al 2012).
+metabolism) [@amarasekare_framework_2012].
 If only temperature’s effect on metabolism were considered, the shape of the relationship may be an exponential one.
@@
 
-TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt et al 2018).
+TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment [@bernhardt_nonlinear_2018].
 However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.
@@
 Given the characteristically non-linear shape of TPCs, Jensen’s inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
-Jensen’s inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999).
-For instance, Ruel and Ayres (1999) provide the development rate of ectotherms as an example to demonstrate Jensen’s inequality.
-The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel and Ayres, 1999).
-With Jensen’s inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel and Ayres, 1999).
+Jensen’s inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions [@ruel_jensens_1999].
+For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen’s inequality [@ruel_jensens_1999].
+The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures [@ruel_jensens_1999].
+With Jensen’s inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures [@ruel_jensens_1999].
 Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time.
-It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019).
+It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies [@bernhardt_nonlinear_2018; @khelifa_usefulness_2019].
 
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare et al 2012, Paajimans et al 2013).
-Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare et al 2012).
+For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns [@amarasekare_framework_2012; @paaijmans_temperature_2013].
+Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature [@amarasekare_framework_2012].
 
-### The relevancy of temperature variability in the modern context
+### Relevancy of thermal variability in the modern context
 
 Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
-Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
-Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
+Variation in temperature is predicted to affect performance more significantly than increasing temperatures [@vasseur_increased_2014]; thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
+Khelifa et al demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance [@khelifa_usefulness_2019].
 
-However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 3).
-Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (2019).
+However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 1.3).
+Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable [@khelifa_usefulness_2019].
 Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
-Vasseur and Yodzis (2004) outline how the frequency (1/period) of environmental fluctuations describes the “color of environmental noise.” In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
-Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (Wigley et al 1998, Tabari et al 2012).
+Vasseur and Yodzis outline how the frequency (1/period) of environmental fluctuations describes the “color of environmental noise” [@vasseur_color_2004].
+In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
+Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable [@wigley_anthropogenic_1998-1; @tabari_shift_2012].
 Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
@@
 ```{r, echo=FALSE, fig.cap="Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
-Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.", out.width = '90%'}
+Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.", out.width = '90%', fig.align="center", fig.scap="Patterns of variability and spectral density plots"}
 knitr::include_graphics("figures/figure4.png")
 ```
+
+\clearpage
 
@@
 Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
-* At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve (Bartheld et al 2017) to changes in TSD ratios (Bowden and Paitz 2018)
-* At the population level, this usually comes in the form of decreasing population densities (Anders and Post 2006)
-* At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration (Descamps-Julien and Gonzalez 2005), inflationary effects (Holt et al 2003)
+
+ * At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve [@bartheld_thermal_2017-2] to changes in TSD ratios [@bowden_temperature_2018]
+
+ * At the population level, this usually comes in the form of decreasing population densities [@anders_distribution-wide_2006]
+
+ * At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration [@descamps-julien_stable_2005], inflationary effects [@holt_impacts_2003]
 It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
-Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (Wang et al 2016, Petchey et al 2000).
+Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses [@petchey_environmental_2000].
 
-There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al 2019, Massey et al 2020).
-To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions:
-1) Does variability pattern drive differences in response magnitude?
+There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability [@kroeker_ecological_2020; @massey_measurement_2019].
+To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
 (i.e.

### 2021-04-29 writing /thesis_draft/02-chap2.Rmd
fully formatted draft :)

@@
-# Methods {-}
+# Methods {}
 ```{r, include=FALSE}
@@
 ```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
-Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%'}
+Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%', fig.align="center", fig.scap="Meta analysis paper selection process"}
 knitr::include_graphics("figures/figure5.png")
@@
 
-We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 3).
+We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1).
 After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*) ) AND SU=(Life Sciences & Biomedicine).
@@
 We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1).
-Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 5).
+Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 4).
 Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
@@
 
+We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.
+treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
+We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response [@brown_toward_2004].
+The final number of data points included in the analysis is broken down by study (Table 2.1).
+
+\clearpage
+
 ```{r, echo = FALSE}
@@
                          "Rolandi et al", "Saxon et al", 
-                         "Semenov et al", "Simonici et al", 
-                         "Triedel et al"), 
+                         "Semenov et al", "Simoncini et al", 
+                         "Treidel et al"), 
                  Year = c("2016", "2009", "2011", "2009", "2015", "2014",
@@
 kable(table1, "latex", booktabs = T, col.names = c("Study", "Year", "Extracted data points (n)"), escape = F, caption = "Studies included in analysis of thermal variability (n=15)") %>%
-   row_spec(0,bold=TRUE)
+   row_spec(0,bold=TRUE)%>%
+   kable_styling(latex_options = c("striped", "hold_position"))
 ```
-
-We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.
-treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
-We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al 2004).
-The final number of data points included in the analysis is broken down by study (Table 1).
 
@@
 We extracted mean values for response variables, any measure of variance (Standard Deviation (SD) or Standard Error of the Mean (SEM)), and sample size.
-For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM (see code appendix).
-We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal history, and aggregated these values (Table 2).
+For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM.
+We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal stress, and aggregated these values (Table 2.2).
+
+If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation [@higgins_cochrane_2011].
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
+
+\clearpage
 
@@
                                      exceeded Topt, classified as thermally stressful, else not
-                                     thermally stressful. See appendix for references in which
-                                     organisms’ thermal history was included"))
+                                     thermally stressful. See Appendix Table 1 for references in which
+                                     organisms’ thermal stress metrics were included"))
 
-kable(table2, caption = "All covariates included in the model", "latex", booktabs = T) %>%
+kable(table2, caption = "Covariates included in models", "latex", booktabs = T) %>%
   column_spec(2, width = "8cm") %>%
-   row_spec(0,bold=TRUE)
+   row_spec(0,bold=TRUE)%>%
+   kable_styling(latex_options = c("striped", "hold_position"))
 ```
 
-If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Cochrane Handbook for Systematic Reviews of Interventions, Version 5.1.0, 2011).
-Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
+\clearpage
 
@@
 We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
-Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer et al 2010).
+Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) [@viechtbauer_conducting_2010].
 In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
@@
 kable(table3, "latex", booktabs = T, col.names = c("Column name", "Purpose"), escape = F, caption = "Column names expanded to wider format for data analysis in metafor") %>%
-   row_spec(0,bold=TRUE)
+   row_spec(0,bold=TRUE) %>%
+   kable_styling(latex_options = c("striped", "hold_position"))
 ```
+
+\linebreak
 
@@
 ```
-We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head (Viechtbauer 2010).
+We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head [@viechtbauer_conducting_2010].
 It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer 2010).
 Each of the subsequent arguments of the escalc function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
-The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 6).
+The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
 
-```{r, echo= FALSE, fig.cap= "Workflow diagram of how data were extracted, analyzed, and processed for downstream analysis in metafor.", out.width="100%"}
+```{r, echo= FALSE, fig.cap= "Workflow diagram of how data were extracted, analyzed, and processed for downstream analysis in metafor.", out.width="100%", fig.scap="Workflow of analysis in metafor"}
 knitr::include_graphics("figures/figure6.png")
@@
 
-Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response (see insert of pseudo code).
+Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
+We first ran a multivariate random effects models without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
 
@@
 
-The arguments of the rma.mv function include SMD (yi), sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
+The arguments of the rma.mv function include SMD, sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
 We specified a nested structure to account for the colinearity of data points from within the same study, experiment, and response metric.
-We used Restricted Maximum-Likelihood (REML) to specify the fit of the model, as REML is best at handling small samples sizes compared to Maximum Likelihood (ML) (Viechtbauer 2010).
-We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population (Viechtbauer 2010).
-
-We then ran two multivariate random effects models, one with modifiers and one without (see insert above).
-This was to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies.
-The full model we ran included:
+We used Restricted Maximum-Likelihood (REML) to specify the fit of the model, as REML is best at handling small samples sizes compared to Maximum Likelihood (ML) [@viechtbauer_conducting_2010].
+We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population [@viechtbauer_conducting_2010].
+The random effects model structure also allowed us to account for the nested structure of responses, experiment, and study to address the dependence of observations from within each study.
+The multivariate random effects model with all modifiers (full model) we ran included:
 
@@
 
-We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal history data for one study.
+The model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
+If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed an organisms’ Topt and become thermally stressful.
+
+We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal stress data for one study.
 That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz 2011, Table 1).
-We ran a separate model without this study and included the thermal history covariate separately from all the other modifiers:
+The model we ran without this study, but with all the covariates in the full model as well as thermal stress (exclusion model):
+
+```{r, eval = FALSE}
+rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant 
+       + exp_age + size + org_level + resp_type + stressful, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+```
+
+We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly and important factor or if it was simply excluding those influential data points:
 

### 2021-04-29 writing /thesis_draft/03-chap3.Rmd
fully formatted draft :)

@@
 
-# Results {-}
+# Results {}
 
@@
 
-Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 6.
-These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
-As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
-
 ```{r, echo= FALSE, message= FALSE, fig.cap="Forest plot of effects sizes by study and experiment within a study.
 Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study.
-Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", out.width="110%"}
+Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", fig.scap="Forest plot of effect sizes", out.width="95%", fig.align="center"}
 forest <- dat_MA_ES %>%
@@
   unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
-forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2")
+forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2", 
+       header=TRUE, cex=0.77)
 ```
 
-Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal history), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p < 0.0001).
+Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 3.1.
+These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
+As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
+
+Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p < 0.0001).
 Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (df = 139, p < 0.00001).
@@
 
-```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.", out.width="105%"}
+```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
 full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
@@
 
-Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 7).The most negative estimate (organization level) was significantly influential in our model at a significance level of p < 0.05.
+\clearpage
+
+Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2).The most negative estimate (organization level) was significantly influential in our model at a significance level of p < 0.05.
 Surprisingly, the most negative estimates of our model coefficients were not statistically significant.
-In our analysis of thermal history via a separate random effects model, we found thermal history was not statistically significant in explaining the effect size (Table 3).
+In our analysis of thermal stress  via a separate random effects model, we found thermal stress was not statistically significant in explaining the effect size (Table 3.1).
 
@@
                  ci.ub = c(0.5669, 0.0886))
-kable(table3, "latex", booktabs = T, col.names = c("Term", "estimate", "se", "zval", "pval", "ci.lb", "ci.ub"), escape = F, caption = "Model summary statistics for random effects model using thermal history as sole modifier") %>%
-   row_spec(0,bold=TRUE)
+kable(table3, "latex", booktabs = T, col.names = c("Term", "estimate", "se", "zval", "pval", "ci.lb", "ci.ub"), escape = F, caption = "Model summary statistics for random effects model using thermal stress as sole modifier",
+      caption.short ="Thermal stress model summary statistics") %>%
+   row_spec(0,bold=TRUE)%>%
+   kable_styling(position = "center", latex_options = c("striped", "hold_position"))
 ```
 
-The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 8.
-At lower mean temperatures, higher fluctuation ranges have a more positive effect on..., though that trend starkly ends at about 24O C.
-Beyond 24O C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 8).
+\clearpage
+
+The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
+At lower mean temperatures, higher fluctuation ranges have a more positive effect on..., though that trend starkly ends at about $24^{\circ}$C.
+Beyond $24^{\circ}$C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
 
@@
 Temperature fluctuation range from 0-20C, Mean temperature range from 7-33OC.
-Effect size range from -2 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)", out.width = '90%'}
+Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and effect size"}
 normalized %>%
@@
 
-The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 9).
-Interestingly, there were not any significant differences between thermal history (Figure 10), life history metric (Figure 11), body size (Figure 12), or age (Figure 13).
+The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4).
+Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).
 
-As noted in the methods, when we excluded one highly influential study and included thermal history (exclusion model), size became much more important in the thermal stress model (df=123, p < 0.001) as did mean temperature.
-Surprisingly, organization level, fluctuation range, and the interaction between range and mean temperature became insignificant in the thermal stress model (df=123, p < 0.1).
+As noted in the methods, when we excluded one highly influential study and included thermal stress (exclusion model), size became much more important in the thermal stress model (df=123, p < 0.001) as did mean temperature (df=123, p < 0.001).
+Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the thermal stress model (df=123, p < 0.1).
 
-```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.", out.width = '90%'}
+```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by organization level"}
 # scatterplot of standardized mean response vs flux range colored and lm fit by org level
@@
 
-```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal history: stressful or not stressful.", out.width = '90%'}
+```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal stress.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by thermal stress"}
 normalized %>%
@@
 
-```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.", out.width = '90%'}
+```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by response type"}
 normalized %>%
@@
 
-```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.", out.width = '90%'}
+```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by body size"}
 ggplot(normalized, aes(x=flux_range, y=yi, color = factor(size)))+
@@
 
-```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.", out.width = '90%'}
+```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by experimental age"}
 ggplot(normalized, aes(x=flux_range, y=yi, color = factor(exp_age)))+
@@
 
-# Discussion {-}
+# Discussion {}
 
 The results of this meta analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level.
-There were not overall differences in the direction o,
-
+There were not overall differences in the direction of thermally variable environments positively or negatively impacting performance relative to constant environments.
+However, a considerable amount of variation exists in whether thermally variably environments positively, negatively, or minimally impact performance across studies.
+Our model results highlight the importance of fluctuation range, mean temperature, and organization level to better understand the variety of effects thermal variation has on organismal responses.
 These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e.
 higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures.
-The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change (Vasseur et al 2014, Sunday et al 2012).
-With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed “the point of no return” (Russill and Nyssa 2009), understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades (West et al 2002, Cheung et al 2008).
-In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature (Vasseur et al 2014).
+The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change [@vasseur_increased_2014; @sunday_thermal_2012].
+With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed “the point of no return” [@russill_tipping_2009], understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades [@west_allometric_2002; @cheung_application_2008].
+In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature [@vasseur_increased_2014].
 
-These conclusions further drive home the decadal conclusion that Jensen’s Inequality has demonstrated: averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance (Ruel and Ayres 1999, Bernhardt et al 2018).
+These findings underscore the significance of better understanding why differences across studies and experimental designs occurred.
+Jensen’s inequality is one way to conceptualize why responses differ, as averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance [@ruel_jensens_1999; @bernhardt_nonlinear_2018].
 The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive.
+Acclimation may also explain differences in response across studies as the scale on which variation occurs influences how organisms may respond.
+If organisms’ non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (i.e.
+inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies [@kremer_gradual_2018].
+In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance.
+Rapid evolution could further explain differences in responses across studies, as if organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics [@yoshida_rapid_2003].
+It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.
 
-Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal history, life history traits, and experimental age.
+Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal stress, life history traits, and experimental age.
 Body size has been identified as important in the allometric scaling of MTE and it is interesting in these subsets of studies, body size did not explain the differences in responses across studies.
-This could simply be because we had a small sample size of studies (n=15), though multiple  responses in studies contributed to a large sample of effect sizes (n=139) and there were only a subset of relative body sizes included.
-It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms (Clark and Johnston 1999, Clarke et al 2006).
+This could simply be because we had a small sample size of studies (n=15), though multiple  responses in studies contributed to a large sample of effect sizes (n=139), there were only a subset of relative body sizes included.
+It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms [@clarke_scaling_1999; @clarke_temperature_2006].
 This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns variability in temperature drives.
 
-An important side note is that our full model and thermal history model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
+An important side note is that our full model and thermal stress model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
 This study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding Garcia-Ruiz et al (2011) from our full model.
@@
 
-While body size and thermal history emerged as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship thermal variability and MTE.
+While body size and thermal stress emerged as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship thermal variability and MTE.
 If we were able to obtain additional information about the CTmax, CTmin, thermal breadth, and tolerance range for each of the species included in the analysis, we may also be able to better understand the trends and the significance of variability patterns in dictating responses.
@@
 This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile.
-However, there have been meta analyses, both qualitative (Massey et al 2020), and quantitative (Noble et al 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al 2018).
+However, there have been meta analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018].
 These results align with patterns of juvenile organisms (i.e.
-fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Portner and Farrell 2008, Moyano et al 2017).
+fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance [@portner_ecology_2008; @moyano_effects_2017].
 
@@
 Though there is a large body of literature that investigates temperature variation, there appears to be dissonance in the scientific literature about the language used to discuss such variation.
-Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures (Amarasekare et al 2012) as well as different ranges of temperature, both of which are different concepts.
+Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures [@amarasekare_framework_2012] as well as different ranges of temperature, both of which are different concepts.
 For instance, Amarasekare et al 2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits.
 Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
-Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes (Colinet et al 2018, Colinet et al 2015).
+Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes [@colinet_mechanisms_2018; @colinet_insects_2015].
 Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.
@@
 Many of these studies failed to justify why certain temperatures were chosen as mean temperatures as well as the range of temperatures studied.
-In order to compile information on the thermal history of each of the organisms included in the studies of this meta analysis, additional resources were needed (see Appendix Table S1).
+In order to compile information on the thermal stress of each of the organisms included in the studies of this meta analysis, additional resources were needed (see Appendix Table 1).
 Understanding where these organisms' performance falls in relation to temperature on its TPC is crucial for drawing reliable conclusions of how thermal variability affects performance.
@@
 Many of the studies made little to no mention of generation time with respect to duration of their experiments or study organisms used.
-Understanding how organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt et al 2020).
+Understanding how organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020].
 This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta analysis.

### 2021-04-29 writing /thesis_draft/05-appendix.Rmd
fully formatted draft :)

`r if(knitr:::is_latex_output()) '\\appendix'`

# Appendix {-}

```{r, echo=FALSE, results = 'hide', message=FALSE}
pacman::p_load(metafor, tidyverse, viridis, visreg, forcats, kableExtra)
dat_MA_0<- read_csv("metafor.csv") %>%
  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))

#make standardized SD columns based on N and SE vs SD
dat_MA <-dat_MA_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_MA_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_MA, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) 
count(dat_MA_ES, study_id)

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -4) 
```

```{r, echo = FALSE}
#funnel plot 
### set up 2x2 array for plotting
par(mfrow=c(2,2))
funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
```

```{r, echo=FALSE, message = FALSE, fig.cap="Funnel plot displaying the extracted data points and the distribution of their residuals centered around zero.", fig.align="center", out.width = '90%', fig.scap="Funnel plot"}
### draw funnel plots
#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
```

\clearpage

```{r, echo= FALSE, fig.cap="Distribution of effect sizes (n=140) across data from all 15 studies.", fig.align="center", out.width = '90%', fig.scap="Histogram of effect sizes"}
dat_MA_ES %>%
filter(yi > -20) %>%
  ggplot(aes(x=yi))+
  geom_histogram(binwidth = 1)+
  theme_bw()+
  scale_fill_viridis()
```

```{r, echo= FALSE}
table4 <- data.frame(Species = c("Drosophila melanogaster", "Leptopilina boulardi", 
                         "Mauremys reevesii", "Gryllus firmus", 
                         "Limnodynastes peronii", "Limnodynastes tasmaniensis", 
                         "Platyplectrum ornatum ", "Natrix natrix", 
                         "Coturnix japoncia", "Plestiodon chinensis", 
                         "Rhodnis prolixus", "Escherichia coli", 
                         "Salmonella", "Caiman latirostris", 
                         "Trachemys scripta", "Esteya vermicola"), 
                 Topt = c("24-26.6 C", "25 C", "37.97 +/- 0.64", "< 30 C", "30 C", "30-33 C",
                          "33 C", " </= 30 C", "26 C", "33 C", "30-35 C", "37 C, 35-36 C",
                          "35-36 C", "32-33 C", "32-33 C", "26 C"), 
                 `Reference` = c("Klepsatel 2013, David 1988", "De lava et al 2016", 
                                 "Dang et al 2019", "Singh et al 2020", "Seebacher et al 2014",
                                 "Whitehead et al 1989", "Kern et al 2014", "Issac 1997",
                                 "Alagawany et al 2017", "Baojun et al 2010", "Fresquet and Lazzari 2011",
                                 "Doyle and Schoeni 1984, Bronikowski et al 2001", 
                                 "Bronikowski et al 2001", 
                                 "Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002", 
                                 "Dang et al 2019", "Wang et al 2015"))
kable(table4, "latex", booktabs = T, 
      col.names = c("Species", "Topt", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X. arvicola, used in Garcia-Ruiz et al 2011)", caption.short ="Thermal stress data by species") %>%
   row_spec(0,bold=TRUE) %>%
  column_spec(3, width = "6cm") %>%
   kable_styling(latex_options = c("striped", "hold_position"))
```

### 2021-04-29 writing /thesis_draft/_book/thesis.tex
fully formatted draft :)

@@
 % To pass between YAML and LaTeX the dollar signs are added by CII
-\title{Mean temperature, fluctuation range, and biological organization affect biological responses}
-\author{Margaret A. Slein}
+\title{Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis}
+\author{Margaret Anne Slein}
 % The month and year that you submit your FINAL draft TO THE LIBRARY (May or December)
@@
 
+	\usepackage{setspace}\onehalfspacing
 	\usepackage{booktabs}
@@
 
-It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bouchard 2014, Bambach 1993).
-Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Soo et al 2017, Blankenship 2017).
-Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll 2008).
+It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bambach, 1993; Bouchard, 2014).
+Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Soo, Hemp, Parks, Fischer, \& Hugenholtz, 2017, @blankenship\_how\_2017).
+Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll, 2008).
 In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnik et al 2007).
-Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Reed et al 2010, Burggren 2018).
+Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Burggren, 2018; Reed, Waples, Schindler, Hard, \& Kinnison, 2010).
 Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.
@@
 Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Nino Southern Oscillation).
-As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (Khelifa et al 2019, Stewart et al 2013).
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (khelifa\_usefulness\_2019; Stewart et al., 2013).
 How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
-As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Helmuth et al 2010, Bernardo 2014, Huey et al 2012).
+As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).
 
@@
 
-There is a large body of literature on environmental variability including its relevance for positive effects for populations and communities, i.e.~inflationary effects in sink-source systems (Holt et al 2003, Vasseur and Fox 2009), such that variability aids population density when organisms emigrate.
+There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, i.e.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate.
 However, there is a lack of consensus as to which scales of biological organization, i.e.~individuals, populations, or communities of organisms will be most severely affected by different shifts in environmental conditions (problem statement).
@@
 \hypertarget{background}{%
-\chapter*{Background}\label{background}}
-\addcontentsline{toc}{chapter}{Background}
+\chapter{Background}\label{background}}
 
@@
 \begin{figure}
-\includegraphics[width=1.05\linewidth]{figures/figure1} \caption{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
-Shown is a comparison of how duration of variability compares across the same environmental factor}\label{fig:unnamed-chunk-3}
+
+{\centering \includegraphics[width=1.05\linewidth]{figures/figure1}
+
+}
+
+\caption[Metrics to describe and explain environmental variability]{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
+Shown is a comparison of how duration of variability compares within the same environmental factor.
+Terms highlighted in each panel are described below in the main text}\label{fig:unnamed-chunk-3}
 \end{figure}
-Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms' response to environmental conditions (Parepa et al 2013, Gudmundson et al 2015).
-Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1).
+Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms' response to environmental conditions (Gudmundson, Eklöf, \& Wennergren, 2015; Parepa, Fischer, \& Bossdorf, 2013).
+Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1.1).
 
@@
 
-We can expand additional information adjacent to the mean (i.e.~range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014).
+We can expand additional information adjacent to the mean (i.e.~range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (Vasseur et al., 2014).
 We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
@@
 The Amazon Rainforest experiences less variation in light availability during the year than Denali National Park in Alaska, which experiences high variation in light availability between the summer and winter seasons.
-The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare et al 2013, 2012), as well as additional life history traits (size, speed, etc.).
-An organism's ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible (Amarasekare 2012).
-If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale their adaptive evolution can manage.
+The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare \& Coutinho, 2013; Amarasekare \& Savage, 2012), as well as additional life history traits (size, speed, etc.).
+An organism's ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible (Amarasekare \& Savage, 2012).
+If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.
 
@@
 The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability.
-Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1).
-Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet et al 2018).
-Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Husak and Lailvaux 2017, Amarasekare and Savage 2012).While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1.1).
+Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018).
+Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
+While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
 However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.~heat waves, cold fronts etc.) in which the benefits of FTR in mitigated chill injuries may be helpful.
 Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
-For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Delmas et al 2007, Bowden and Paitz 2018).
-Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (Lawson et al 2015) depending on the magnitude of variance.
-When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Romanuk and Kolasa 2002, Dobramys and Tauber 2013, Bastille-Rousseau et al 2018).
+For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Bowden \& Paitz, 2018; Delmas, Prevot-Julliard, Pieau, \& Girondot, 2007).
+Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (Lawson, Vindenes, Bailey, \& Pol, 2015) depending on the magnitude of variance.
+When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Bastille-Rousseau et al., 2018; Dobramysl \& Tauber, 2013; Romanuk \& Kolasa, 2002).
 
@@
 Variability may not have evenly spaced intervals (i.e.~heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
-Predictability can also attempt to describe stochastic vs.~autocorrelated variation, however, that will be discussed in a different context below (Figure 4).
+Predictability can also attempt to describe stochastic vs.~autocorrelated variation, however, that will be discussed in a different context below (Figure 1.3).
 
@@
 While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact.
-Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Horne and Schneider 1995, Farhang-Sardroodi et al 2019).
+Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Farhang-Sardroodi, Darooneh, Kohandel, \& Komarova, 2019; Horne \& Schneider, 1995).
 Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e.~Diel vertical migration (DVM).
-In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (Lampert 1989, Brierly 2014).
-However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey 1974, Kearney et al 2011).
-
-Though environmental variability can include many different forms, from fluctuations in resource availability (Dempster and Pollard 1981, Sommer 1984) to light (Ruel and Ayres 1999, Morison 2020), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (Huey et al 1990, Colinet et al 2015).
+In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (Brierley, 2014; Lampert, 1989).
+However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).
+
+Though environmental variability can include many different forms, from fluctuations in resource availability
+(Dempster \& Pollard, 1981; Sommer, 1984) to light (Morison, Franzè, Harvey, \& Menden‐Deuer, 2020; Ruel \& Ayres, 1999), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (Colinet, Sinclair, Vernon, \& Renault, 2015; Huey \& Bennett, 1990).
 
@@
 Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as ``a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis'' (2018).
-Most generally, temperature variability describes changes in environmental temperature through time (Holmes et al 2016, Colinet et al 2018).
+Most generally, temperature variability describes changes in environmental temperature through time (Colinet et al., 2018; Holmes, Woollings, Hawkins, \& Vries, 2016).
 
@@
 
-Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey et al 1990).
-Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE), though highly debated (Clarke et al 2006), which serves as an origin point for understanding how organisms respond to environmental conditions.
+Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey \& Bennett, 1990).
+Brown et al highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE) (Brown, Gillooly, Allen, Savage, \& West, 2004).
+Though highly debated (Clarke, 2006), which serves as an origin point for understanding how organisms respond to environmental conditions.
 MTE is described by the following equation:
@@
 \]
+
 Where \(B\) is organismal metabolic rate, \(b_0\) is a normalization constant unrelated to temperate or body size, \(M\) is organismal mass, and \(e^-E/kT\) is referred to as the ``Boltzman factor'', where \(E\) is activation energy, \(k\) is the Boltzman constant and \(T\) is temperature.
@@
 We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates should be the same.
-However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clark and Johnston 1999, Clarke et al 2006).
-Others believe that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al 2006).
+However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clarke, 2006; Clarke \& Johnston, 1999).
+Others believe that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al., 2006).
 Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.
@@
 MTE supports temperature impacting metabolic rate, which can have downstream effects on fitness characteristics.
-This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 2).
+This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{figures/figure2} \caption{Figure 2.
-Schematic of thermal performance curve and it’s important components for understanding thermal performance.
-Adapted from (cite, cite)}\label{fig:unnamed-chunk-4}
+
+{\centering \includegraphics[width=0.9\linewidth]{figures/figure2}
+
+}
+
+\caption[Thermal performance curve components]{Schematic of thermal performance curve and it’s important components for understanding thermal performance.
+Adapted from (Krenek et al 2012, Tuff et al 2016)}\label{fig:unnamed-chunk-4}
 \end{figure}
 The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms' performance is negatively impacted (Brown et al 2004, Dowd et al 2015).
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms' performance is negatively impacted (Brown et al., 2004; Dowd, King, \& Denny, 2015).
 Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
 
-TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (i.e.~metabolism) (Amarasekare et al 2012).
+TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (i.e.~metabolism) (Amarasekare \& Savage, 2012).
 If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one.
@@
 
-TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt et al 2018).
+TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018).
 However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.~15-35 C in 5 C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
@@
 Given the characteristically non-linear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
-Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel and Ayres, 1999).
-For instance, Ruel and Ayres (1999) provide the development rate of ectotherms as an example to demonstrate Jensen's inequality.
-The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel and Ayres, 1999).
-With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel and Ayres, 1999).
+Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel \& Ayres, 1999).
+For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen's inequality (Ruel \& Ayres, 1999).
+The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel \& Ayres, 1999).
+With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel \& Ayres, 1999).
 Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time.
-It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al 2018, Khelifa et al 2019).
+It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al., 2018; Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019).
 
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare et al 2012, Paajimans et al 2013).
-Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare et al 2012).
-
-\hypertarget{the-relevancy-of-temperature-variability-in-the-modern-context}{%
-\subsection{The relevancy of temperature variability in the modern context}\label{the-relevancy-of-temperature-variability-in-the-modern-context}}
+For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013).
+Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare \& Savage, 2012).
+
+\hypertarget{relevancy-of-thermal-variability-in-the-modern-context}{%
+\subsection{Relevancy of thermal variability in the modern context}\label{relevancy-of-thermal-variability-in-the-modern-context}}
 
 Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
-Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
-Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
-
-However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 3).
-Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (2019).
+Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al., 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
+Khelifa et al demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance (Khelifa et al., 2019).
+
+However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 1.3).
+Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (Khelifa et al., 2019).
 Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
-Vasseur and Yodzis (2004) outline how the frequency (1/period) of environmental fluctuations describes the ``color of environmental noise.'' In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
-Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (Wigley et al 1998, Tabari et al 2012).
+Vasseur and Yodzis outline how the frequency (1/period) of environmental fluctuations describes the ``color of environmental noise'' (Vasseur \& Yodzis, 2004).
+In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
+Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (Tabari, Hosseinzadeh Talaee, Ezani, \& Shifteh Some'e, 2012; Wigley, 1998).
 Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{figures/figure4} \caption{Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
+
+{\centering \includegraphics[width=0.9\linewidth]{figures/figure4}
+
+}
+
+\caption[Patterns of variability and spectral density plots]{Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
 Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.}\label{fig:unnamed-chunk-5}
 \end{figure}
+\clearpage
+
 \hypertarget{objectives-driving-questions-hypotheses}{%
@@
 Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
-* At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve (Bartheld et al 2017) to changes in TSD ratios (Bowden and Paitz 2018)
-* At the population level, this usually comes in the form of decreasing population densities (Anders and Post 2006)
-* At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration (Descamps-Julien and Gonzalez 2005), inflationary effects (Holt et al 2003)
+\begin{itemize}
+\item
+  At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve (Bartheld, Artacho, \& Bacigalupe, 2017) to changes in TSD ratios (Bowden \& Paitz, 2018)
+\item
+  At the population level, this usually comes in the form of decreasing population densities (Anders \& Post, 2006)
+\item
+  At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration (Descamps-Julien \& Gonzalez, 2005), inflationary effects (Holt et al., 2003)
   It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
-Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (Wang et al 2016, Petchey et al 2000).
-
-There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al 2019, Massey et al 2020).
-To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions:
-1) Does variability pattern drive differences in response magnitude?
+Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).
+\end{itemize}
+There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al., 2020; Massey, Holt, Brooks, \& Rollinson, 2019).
+To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
 (i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc)?
@@
 \hypertarget{methods}{%
-\chapter*{Methods}\label{methods}}
-\addcontentsline{toc}{chapter}{Methods}
+\chapter{Methods}\label{methods}}
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{figures/figure5} \caption{Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
+
+{\centering \includegraphics[width=0.9\linewidth]{figures/figure5}
+
+}
+
+\caption[Meta analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
 Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
 \end{figure}
-We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 3).
+We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1).
 After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*) ) AND SU=(Life Sciences \& Biomedicine).
@@
 \end{itemize}
-\begin{table}
+We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
+We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004).
+The final number of data points included in the analysis is broken down by study (Table 2.1).
+
+\clearpage
+\begin{table}[!h]
 
@@
 \midrule
-Delava et al & 2016 & 16\\
+\cellcolor{gray!6}{Delava et al} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{16}\\
 Du et al & 2009 & 32\\
-Garcia-Ruiz et al & 2011 & 16\\
+\cellcolor{gray!6}{Garcia-Ruiz et al} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{16}\\
 Glass et al & 2009 & 16\\
-Kern et al & 2015 & 24\\
+\cellcolor{gray!6}{Kern et al} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{24}\\
 \addlinespace
 Kern et al & 2014 & 4\\
-Lowenborg et al & 2012 & 10\\
+\cellcolor{gray!6}{Lowenborg et al} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{10}\\
 Maneti et al & 2014 & 24\\
-Pendlebury et al & 2004 & 16\\
+\cellcolor{gray!6}{Pendlebury et al} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{16}\\
 Qu et al & 2014 & 40\\
 \addlinespace
-Rolandi et al & 2018 & 4\\
+\cellcolor{gray!6}{Rolandi et al} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{4}\\
 Saxon et al & 2017 & 12\\
-Semenov et al & 2007 & 32\\
-Simonici et al & 2019 & 30\\
-Triedel et al & 2015 & 4\\
+\cellcolor{gray!6}{Semenov et al} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{32}\\
+Simoncini et al & 2019 & 30\\
+\cellcolor{gray!6}{Treidel et al} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{4}\\
 \bottomrule
@@
 \end{table}
-We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
-We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al 2004).
-The final number of data points included in the analysis is broken down by study (Table 1).
-
 \hypertarget{data-extraction-and-analysis}{%
@@
 We extracted mean values for response variables, any measure of variance (Standard Deviation (SD) or Standard Error of the Mean (SEM)), and sample size.
-For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM (see code appendix).
-We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal history, and aggregated these values (Table 2).
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-9}All covariates included in the model}
+For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM.
+We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal stress, and aggregated these values (Table 2.2).
+
+If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Higgins et al., 2011).
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
+
+\clearpage
+\begin{table}[!h]
+
+\caption{\label{tab:unnamed-chunk-9}Covariates included in models}
 \centering
@@
 \midrule
-mean\_temp\_constant & Reported mean temperature in study\\
+\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature in study}\\
 flux\_range & Reported range (difference between minimum 
@@
                                      as either 0 for individual or 1 for population level
-                                     responses\\
+\cellcolor{gray!6}{                                     responses}\\
 size & Size estimate of study organism, 0 for extra-small
@@
                                      categorized as 0 for larval, 1 for juvenile, and 
-                                     2 for adult\\
+\cellcolor{gray!6}{                                     2 for adult}\\
 \addlinespace
@@
                                      exceeded Topt, classified as thermally stressful, else not
-                                     thermally stressful. See appendix for references in which
-                                     organisms’ thermal history was included\\
+                                     thermally stressful. See Appendix Table 1 for references in which
+\cellcolor{gray!6}{                                     organisms’ thermal stress metrics were included}\\
 \bottomrule
@@
 \end{table}
-If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Cochrane Handbook for Systematic Reviews of Interventions, Version 5.1.0, 2011).
-Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
+\clearpage
 
@@
 We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
-Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer et al 2010).
+Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer, 2010).
 In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
-\begin{table}
+\begin{table}[!h]
 
@@
 \midrule
-Mf & Mean of experimental group\\
+\cellcolor{gray!6}{Mf} & \cellcolor{gray!6}{Mean of experimental group}\\
 SDf & Standard deviation of the experimental group\\
-Mc & Mean of the control group\\
+\cellcolor{gray!6}{Mc} & \cellcolor{gray!6}{Mean of the control group}\\
 SDc & Standard deviation of the control group\\
-Nf & Sample size (n) of the experimental group\\
+\cellcolor{gray!6}{Nf} & \cellcolor{gray!6}{Sample size (n) of the experimental group}\\
 \addlinespace
@@
 \end{table}
+\linebreak
+
 Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the escalc function in metafor to calculate effect size and sampling variance:
@@
 \end{Shaded}
-We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head (Viechtbauer 2010).
+We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head (Viechtbauer, 2010).
 It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer 2010).
 Each of the subsequent arguments of the escalc function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
-The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 6).
+The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/figure6} \caption{Workflow diagram of how data were extracted, analyzed, and processed for downstream analysis in metafor.}\label{fig:unnamed-chunk-12}
+\includegraphics[width=1\linewidth]{figures/figure6} \caption[Workflow of analysis in metafor]{Workflow diagram of how data were extracted, analyzed, and processed for downstream analysis in metafor.}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 
-Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response (see insert of pseudo code).
+Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
+We first ran a multivariate random effects models without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
 \begin{Shaded}
@@
 \end{Shaded}
-The arguments of the rma.mv function include SMD (yi), sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
+The arguments of the rma.mv function include SMD, sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
 We specified a nested structure to account for the colinearity of data points from within the same study, experiment, and response metric.
-We used Restricted Maximum-Likelihood (REML) to specify the fit of the model, as REML is best at handling small samples sizes compared to Maximum Likelihood (ML) (Viechtbauer 2010).
-We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population (Viechtbauer 2010).
-
-We then ran two multivariate random effects models, one with modifiers and one without (see insert above).
-This was to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies.
-The full model we ran included:
+We used Restricted Maximum-Likelihood (REML) to specify the fit of the model, as REML is best at handling small samples sizes compared to Maximum Likelihood (ML) (Viechtbauer, 2010).
+We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population (Viechtbauer, 2010).
+The random effects model structure also allowed us to account for the nested structure of responses, experiment, and study to address the dependence of observations from within each study.
+The multivariate random effects model with all modifiers (full model) we ran included:
 \begin{Shaded}
@@
 \end{Shaded}
-We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal history data for one study.
+The model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
+If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed an organisms' Topt and become thermally stressful.
+
+We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal stress data for one study.
 That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz 2011, Table 1).
-We ran a separate model without this study and included the thermal history covariate separately from all the other modifiers:
+The model we ran without this study, but with all the covariates in the full model as well as thermal stress (exclusion model):
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{rma.mv}\NormalTok{(yi, vi, }\DataTypeTok{data=}\NormalTok{dat_MA_ES, }\DataTypeTok{mods =} \OperatorTok{~}\NormalTok{flux_range }\OperatorTok{*}\StringTok{ }\NormalTok{mean_temp_constant }
+       \OperatorTok{+}\StringTok{ }\NormalTok{exp_age }\OperatorTok{+}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{org_level }\OperatorTok{+}\StringTok{ }\NormalTok{resp_type }\OperatorTok{+}\StringTok{ }\NormalTok{stressful, }
+               \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{  }\NormalTok{study_id}\OperatorTok{/}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id,}
+                 \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
+\end{Highlighting}
+\end{Shaded}
+We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly and important factor or if it was simply excluding those influential data points:
 \begin{Shaded}
@@
 \hypertarget{results}{%
-\chapter*{Results}\label{results}}
-\addcontentsline{toc}{chapter}{Results}
-
-Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 6.
+\chapter{Results}\label{results}}
+\begin{figure}
+
+{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-18-1}
+
+}
+
+\caption[Forest plot of effect sizes]{Forest plot of effects sizes by study and experiment within a study.
+Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study.
+Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-18}
+\end{figure}
+Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1.
 These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
 As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
-\begin{figure}
-\includegraphics[width=1.1\linewidth]{thesis_files/figure-latex/unnamed-chunk-17-1} \caption{Forest plot of effects sizes by study and experiment within a study.
-Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study.
-Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-17}
-\end{figure}
-Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal history), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p \textless{} 0.0001).
+
+Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p \textless{} 0.0001).
 Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (df = 139, p \textless{} 0.00001).
@@
 \begin{figure}
-\includegraphics[width=1.05\linewidth]{thesis_files/figure-latex/unnamed-chunk-18-1} \caption{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.}\label{fig:unnamed-chunk-18}
+
+{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-19-1}
+
+}
+
+\caption[Coefficient estimates for full model]{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.}\label{fig:unnamed-chunk-19}
 \end{figure}
-Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 7).The most negative estimate (organization level) was significantly influential in our model at a significance level of p \textless{} 0.05.
+\clearpage
+
+Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).The most negative estimate (organization level) was significantly influential in our model at a significance level of p \textless{} 0.05.
 Surprisingly, the most negative estimates of our model coefficients were not statistically significant.
-In our analysis of thermal history via a separate random effects model, we found thermal history was not statistically significant in explaining the effect size (Table 3).
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-19}Model summary statistics for random effects model using thermal history as sole modifier}
+In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant in explaining the effect size (Table 3.1).
+\begin{table}[!h]
+
+\caption[Thermal stress model summary statistics]{\label{tab:unnamed-chunk-20}Model summary statistics for random effects model using thermal stress as sole modifier}
 \centering
@@
 \midrule
-intercept & 0.2050 & 0.1846 & 1.1104 & 0.2668 & -0.1568 & 0.5669\\
+\cellcolor{gray!6}{intercept} & \cellcolor{gray!6}{0.2050} & \cellcolor{gray!6}{0.1846} & \cellcolor{gray!6}{1.1104} & \cellcolor{gray!6}{0.2668} & \cellcolor{gray!6}{-0.1568} & \cellcolor{gray!6}{0.5669}\\
 stressful & -0.0489 & 0.0702 & -0.6975 & 0.4855 & -0.1864 & 0.0886\\
@@
 \end{table}
-The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 8.
-At lower mean temperatures, higher fluctuation ranges have a more positive effect on\ldots{}, though that trend starkly ends at about 24O C.
-Beyond 24O C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 8).
+\clearpage
+
+The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
+At lower mean temperatures, higher fluctuation ranges have a more positive effect on\ldots{}, though that trend starkly ends at about \(24^{\circ}\)C.
+Beyond \(24^{\circ}\)C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-20-1} \caption{Scatterplot describing the relationship between mean temperature, temperature range, and effect size.
+
+{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-21-1}
+
+}
+
+\caption[Scatterplot of relationship between range, mean, and effect size]{Scatterplot describing the relationship between mean temperature, temperature range, and effect size.
 Temperature fluctuation range from 0-20C, Mean temperature range from 7-33OC.
-Effect size range from -2 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)}\label{fig:unnamed-chunk-20}
+Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)}\label{fig:unnamed-chunk-21}
 \end{figure}
-The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 9).
-Interestingly, there were not any significant differences between thermal history (Figure 10), life history metric (Figure 11), body size (Figure 12), or age (Figure 13).
-
-As noted in the methods, when we excluded one highly influential study and included thermal history (exclusion model), size became much more important in the thermal stress model (df=123, p \textless{} 0.001) as did mean temperature.
-Surprisingly, organization level, fluctuation range, and the interaction between range and mean temperature became insignificant in the thermal stress model (df=123, p \textless{} 0.1).
+The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4).
+Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).
+
+As noted in the methods, when we excluded one highly influential study and included thermal stress (exclusion model), size became much more important in the thermal stress model (df=123, p \textless{} 0.001) as did mean temperature (df=123, p \textless{} 0.001).
+Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the thermal stress model (df=123, p \textless{} 0.1).
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-21-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-21}
+
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-22-1}
+
+}
+
+\caption[Effect sizes across temperature range by organization level]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-22}
 \end{figure}
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-22-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal history: stressful or not stressful.}\label{fig:unnamed-chunk-22}
+
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-23-1}
+
+}
+
+\caption[Effect sizes across temperature range by thermal stress]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal stress.}\label{fig:unnamed-chunk-23}
 \end{figure}
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-23-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-23}
+
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-24-1}
+
+}
+
+\caption[Effect sizes across temperature range by response type]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-24}
 \end{figure}
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-24-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-24}
+
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-25-1}
+
+}
+
+\caption[Effect sizes across temperature range by body size]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-25}
 \end{figure}
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-25-1} \caption{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-25}
+
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-26-1}
+
+}
+
+\caption[Effect sizes across temperature range by experimental age]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-26}
 \end{figure}
 \hypertarget{discussion}{%
-\chapter*{Discussion}\label{discussion}}
-\addcontentsline{toc}{chapter}{Discussion}
+\chapter{Discussion}\label{discussion}}
 
 The results of this meta analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level.
-There were not overall differences in the direction o,
-
+There were not overall differences in the direction of thermally variable environments positively or negatively impacting performance relative to constant environments.
+However, a considerable amount of variation exists in whether thermally variably environments positively, negatively, or minimally impact performance across studies.
+Our model results highlight the importance of fluctuation range, mean temperature, and organization level to better understand the variety of effects thermal variation has on organismal responses.
 These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e.~higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures.
-The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change (Vasseur et al 2014, Sunday et al 2012).
-With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed ``the point of no return'' (Russill and Nyssa 2009), understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades (West et al 2002, Cheung et al 2008).
-In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature (Vasseur et al 2014).
-
-These conclusions further drive home the decadal conclusion that Jensen's Inequality has demonstrated: averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance (Ruel and Ayres 1999, Bernhardt et al 2018).
+The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change (Sunday, Bates, \& Dulvy, 2012; Vasseur et al., 2014).
+With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed ``the point of no return'' (Russill \& Nyssa, 2009), understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades (Cheung, Close, Lam, Watson, \& Pauly, 2008; West, Woodruff, \& Brown, 2002).
+In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature (Vasseur et al., 2014).
+
+These findings underscore the significance of better understanding why differences across studies and experimental designs occurred.
+Jensen's inequality is one way to conceptualize why responses differ, as averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance (Bernhardt et al., 2018; Ruel \& Ayres, 1999).
 The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive.
-
-Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal history, life history traits, and experimental age.
+Acclimation may also explain differences in response across studies as the scale on which variation occurs influences how organisms may respond.
+If organisms' non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (i.e.~inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies (Kremer, Fey, Arellano, \& Vasseur, 2018).
+In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance.
+Rapid evolution could further explain differences in responses across studies, as if organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics (Yoshida, Jones, Ellner, Fussmann, \& Hairston, 2003).
+It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.
+
+Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal stress, life history traits, and experimental age.
 Body size has been identified as important in the allometric scaling of MTE and it is interesting in these subsets of studies, body size did not explain the differences in responses across studies.
-This could simply be because we had a small sample size of studies (n=15), though multiple responses in studies contributed to a large sample of effect sizes (n=139) and there were only a subset of relative body sizes included.
-It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms (Clark and Johnston 1999, Clarke et al 2006).
+This could simply be because we had a small sample size of studies (n=15), though multiple responses in studies contributed to a large sample of effect sizes (n=139), there were only a subset of relative body sizes included.
+It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms (Clarke, 2006; Clarke \& Johnston, 1999).
 This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns variability in temperature drives.
 
-An important side note is that our full model and thermal history model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
+An important side note is that our full model and thermal stress model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
 This study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding Garcia-Ruiz et al (2011) from our full model.
@@
 This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile.
-However, there have been meta analyses, both qualitative (Massey et al 2020), and quantitative (Noble et al 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al 2018).
-These results align with patterns of juvenile organisms (i.e.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Portner and Farrell 2008, Moyano et al 2017).
+However, there have been meta analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
+These results align with patterns of juvenile organisms (i.e.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Moyano et al., 2017; Portner \& Farrell, 2008).
 
@@
 Though there is a large body of literature that investigates temperature variation, there appears to be dissonance in the scientific literature about the language used to discuss such variation.
-Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures (Amarasekare et al 2012) as well as different ranges of temperature, both of which are different concepts.
+Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures (Amarasekare \& Savage, 2012) as well as different ranges of temperature, both of which are different concepts.
 For instance, Amarasekare et al 2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits.
 Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
-Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes (Colinet et al 2018, Colinet et al 2015).
+Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes (Colinet et al., 2018, 2015).
 Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.
@@
 Many of the studies made little to no mention of generation time with respect to duration of their experiments or study organisms used.
-Understanding how organisms' life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt et al 2020).
+Understanding how organisms' life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt, O'Connor, Sunday, \& Gonzalez, 2020).
 This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta analysis.
@@
 
-\hypertarget{the-first-appendix}{%
-\chapter{The First Appendix}\label{the-first-appendix}}
-
-This first appendix includes all of the R chunks of code that were hidden throughout the document (using the \texttt{include\ =\ FALSE} chunk tag) to help with readibility and/or setup.
-
-\textbf{In the main Rmd file}
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# This chunk ensures that the thesisdown package is}
-\CommentTok{# installed and loaded. This thesisdown package includes}
-\CommentTok{# the template files for the thesis.}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}\StringTok{'You need to run install.packages("remotes")",}
-\StringTok{            "first in the Console.'}\NormalTok{)}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{"You need to run"}\NormalTok{,}
-        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\KeywordTok{library}\NormalTok{(thesisdown)}
-\CommentTok{# Set how wide the R output will go}
-\KeywordTok{options}\NormalTok{(}\DataTypeTok{width =} \DecValTok{70}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-\textbf{In Chapter \ref{ref-labels}:}
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# This chunk ensures that the thesisdown package is}
-\CommentTok{# installed and loaded. This thesisdown package includes}
-\CommentTok{# the template files for the thesis and also two functions}
-\CommentTok{# used for labeling and referencing}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(remotes)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"remotes"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("remotes")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(dplyr)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("dplyr")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(ggplot2)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("ggplot2")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(bookdown)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-    \KeywordTok{install.packages}\NormalTok{(}\StringTok{"bookdown"}\NormalTok{, }\DataTypeTok{repos =} \StringTok{"https://cran.rstudio.com"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{'You need to run install.packages("bookdown")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(thesisdown)) \{}
-  \ControlFlowTok{if}\NormalTok{ (params}\OperatorTok{$}\StringTok{`}\DataTypeTok{Install needed packages for \{thesisdown\}}\StringTok{`}\NormalTok{) \{}
-\NormalTok{    remotes}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ismayc/thesisdown"}\NormalTok{)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \KeywordTok{stop}\NormalTok{(}
-      \KeywordTok{paste}\NormalTok{(}
-        \StringTok{"You need to run"}\NormalTok{,}
-        \StringTok{'remotes::install_github("ismayc/thesisdown")'}\NormalTok{,}
-        \StringTok{"first in the Console."}
-\NormalTok{      )}
-\NormalTok{    )}
-\NormalTok{  \}}
-\NormalTok{\}}
-\KeywordTok{library}\NormalTok{(thesisdown)}
-\KeywordTok{library}\NormalTok{(dplyr)}
-\KeywordTok{library}\NormalTok{(ggplot2)}
-\KeywordTok{library}\NormalTok{(knitr)}
-\NormalTok{flights <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data/flights.csv"}\NormalTok{, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-\hypertarget{the-second-appendix-for-fun}{%
-\chapter{The Second Appendix, for Fun}\label{the-second-appendix-for-fun}}
-
+\hypertarget{appendix}{%
+\chapter*{Appendix}\label{appendix}}
+\addcontentsline{toc}{chapter}{Appendix}
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-29-1}
+
+}
+
+\caption[Funnel plot]{Funnel plot displaying the extracted data points and the distribution of their residuals centered around zero.}\label{fig:unnamed-chunk-29}
+\end{figure}
+\clearpage
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-30-1}
+
+}
+
+\caption[Histogram of effect sizes]{Distribution of effect sizes (n=140) across data from all 15 studies.}\label{fig:unnamed-chunk-30}
+\end{figure}
+\begin{table}[!h]
+
+\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X.
+arvicola, used in Garcia-Ruiz et al 2011)}
+\centering
+\begin{tabular}[t]{ll>{\raggedright\arraybackslash}p{6cm}}
+\toprule
+\textbf{Species} & \textbf{Topt} & \textbf{Reference}\\
+\midrule
+\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6 C} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
+Leptopilina boulardi & 25 C & De lava et al 2016\\
+\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 +/- 0.64} & \cellcolor{gray!6}{Dang et al 2019}\\
+Gryllus firmus & < 30 C & Singh et al 2020\\
+\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30 C} & \cellcolor{gray!6}{Seebacher et al 2014}\\
+\addlinespace
+Limnodynastes tasmaniensis & 30-33 C & Whitehead et al 1989\\
+\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33 C} & \cellcolor{gray!6}{Kern et al 2014}\\
+Natrix natrix & </= 30 C & Issac 1997\\
+\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26 C} & \cellcolor{gray!6}{Alagawany et al 2017}\\
+Plestiodon chinensis & 33 C & Baojun et al 2010\\
+\addlinespace
+\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35 C} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
+Escherichia coli & 37 C, 35-36 C & Doyle and Schoeni 1984, Bronikowski et al 2001\\
+\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36 C} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
+Caiman latirostris & 32-33 C & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
+\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33 C} & \cellcolor{gray!6}{Dang et al 2019}\\
+\addlinespace
+Esteya vermicola & 26 C & Wang et al 2015\\
+\bottomrule
+\end{tabular}
+\end{table}
 \backmatter
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-angel2000}{}%
-Angel, E. (2000). \emph{Interactive computer graphics : A top-down approach with opengl}. Boston, MA: Addison Wesley Longman.
-
-\leavevmode\hypertarget{ref-angel2001}{}%
-Angel, E. (2001a). \emph{Batch-file computer graphics : A bottom-up approach with quicktime}. Boston, MA: Wesley Addison Longman.
-
-\leavevmode\hypertarget{ref-angel2002a}{}%
-Angel, E. (2001b). \emph{Test second book by angel}. Boston, MA: Wesley Addison Longman.
+\leavevmode\hypertarget{ref-alagawany_heat_2017}{}%
+Alagawany, M., Farag, M., Abd El-Hack, M., \& Patra, A.
+(2017).
+Heat stress: Effects on productive and reproductive performance of quail.
+\emph{World's Poultry Science Journal}, \emph{73}(4), 747--756.
+\url{http://doi.org/10.1017/S0043933917000782}
+
+\leavevmode\hypertarget{ref-amarasekare_intrinsic_2013}{}%
+Amarasekare, P., \& Coutinho, R.
+M.
+(2013).
+The intrinsic growth rate as a predictor of population viability under climate warming.
+\emph{J Anim Ecol}, \emph{82}(6), 1240--1253.
+\url{http://doi.org/10.1111/1365-2656.12112}
+
+\leavevmode\hypertarget{ref-amarasekare_framework_2012}{}%
+Amarasekare, P., \& Savage, V. (2012). A framework for elucidating the temperature dependence of fitness. \emph{The American Naturalist}, \emph{179}(2), 178--191. \url{http://doi.org/10.1086/663677}
+
+\leavevmode\hypertarget{ref-anders_distribution-wide_2006}{}%
+Anders, A.
+D., \& Post, E.
+(2006).
+Distribution-wide effects of climate on population densities of a declining migratory landbird.
+\emph{J Anim Ecology}, \emph{75}(1), 221--227.
+\url{http://doi.org/10.1111/j.1365-2656.2006.01034.x}
+
+\leavevmode\hypertarget{ref-bambach_seafood_1993}{}%
+Bambach, R.
+K.
+(1993).
+Seafood through time: Changes in biomass, energetics, and productivity in the marine ecosystem.
+\emph{Paleobiology}, \emph{19}(3), 372--397.
+\url{http://doi.org/10.1017/S0094837300000336}
+
+\leavevmode\hypertarget{ref-baojun_seasonal_2014}{}%
+Baojun, S., Wenqi, T., Zhigao, Z., \& Weiguo, D.
+(2014).
+The seasonal acclimatisation of locomotion in a terrestrial reptile, plestiodon chinensis(Scincidae).
+\emph{Asian Herpetological Research}, \emph{5}(3), 197.
+\url{http://doi.org/10.3724/SP.J.1245.2014.00197}
+
+\leavevmode\hypertarget{ref-bartheld_thermal_2017-2}{}%
+Bartheld, J.
+L., Artacho, P., \& Bacigalupe, L.
+(2017).
+Thermal performance curves under daily thermal fluctuation: A study in helmeted water toad tadpoles.
+\emph{Journal of Thermal Biology}, \emph{70}, 80--85.
+\url{http://doi.org/10.1016/j.jtherbio.2017.09.008}
+
+\leavevmode\hypertarget{ref-bastille-rousseau_climate_2018}{}%
+Bastille-Rousseau, G., Schaefer, J.
+A., Peers, M.
+J.
+L., Ellington, E.
+H., Mumma, M.
+A., Rayl, N.
+D., \ldots{} Murray, D.
+L.
+(2018).
+Climate change can alter predator--prey dynamics and population viability of prey.
+\emph{Oecologia}, \emph{186}(1), 141--150.
+\url{http://doi.org/10.1007/s00442-017-4017-y}
+
+\leavevmode\hypertarget{ref-bernardo_biologically_2014}{}%
+Bernardo, J.
+(2014).
+Biologically grounded predictions of species resistance and resilience to climate change.
+\emph{Proceedings of the National Academy of Sciences}, \emph{111}(15), 5450--5451.
+\url{http://doi.org/10.1073/pnas.1404505111}
+
+\leavevmode\hypertarget{ref-bernhardt_life_2020}{}%
+Bernhardt, J.
+R., O'Connor, M.
+I., Sunday, J.
+M., \& Gonzalez, A.
+(2020).
+Life in fluctuating environments.
+\emph{Phil.
+Trans.
+R.
+Soc.
+B}, \emph{375}(1814), 20190454.
+\url{http://doi.org/10.1098/rstb.2019.0454}
+
+\leavevmode\hypertarget{ref-bernhardt_nonlinear_2018}{}%
+Bernhardt, J. R., Sunday, J. M., Thompson, P. L., \& O'Connor, M. I. (2018). Nonlinear averaging of thermal experience predicts population growth rates in a thermally variable environment, 10.
+
+\leavevmode\hypertarget{ref-blankenship_how_2017}{}%
+Blankenship, R. E. (2017). How cyanobacteria went green. \emph{Science}, \emph{355}(6332), 1372--1373. \url{http://doi.org/10.1126/science.aam9365}
+
+\leavevmode\hypertarget{ref-bouchard_ecosystem_2014}{}%
+Bouchard, F. (2014). Ecosystem evolution is about variation and persistence, not populations and reproduction. \emph{Biol Theory}, \emph{9}(4), 382--391. \url{http://doi.org/10.1007/s13752-014-0171-1}
+
+\leavevmode\hypertarget{ref-bowden_temperature_2018}{}%
+Bowden, R.
+M., \& Paitz, R.
+T.
+(2018).
+Temperature fluctuations and maternal estrogens as critical factors for understanding temperature-dependent sex determination in nature.
+\emph{J.
+Exp.
+Zool.}, \emph{329}(4), 177--184.
+\url{http://doi.org/10.1002/jez.2183}
+
+\leavevmode\hypertarget{ref-brierley_diel_2014}{}%
+Brierley, A. S. (2014). Diel vertical migration. \emph{Current Biology}, \emph{24}(22), R1074--R1076. \url{http://doi.org/10.1016/j.cub.2014.08.054}
+
+\leavevmode\hypertarget{ref-bronikowski_evolutionary_2001}{}%
+Bronikowski, A.
+M., Bennett, A.
+F., \& Lenski, R.
+E.
+(2001).
+EVOLUTIONARY ADAPTATION TO TEMPERATURE.
+VIII.
+EFFECTS OF TEMPERATURE ON GROWTH RATE IN NATURAL ISOLATES OF ESCHERICHIA COLI AND SALMONELLA ENTERICA FROM DIFFERENT THERMAL ENVIRONMENTS.
+\emph{Evolution}, \emph{55}(1), 33--40.
+\url{http://doi.org/10.1111/j.0014-3820.2001.tb01270.x}
+
+\leavevmode\hypertarget{ref-brown_toward_2004}{}%
+Brown, J. H., Gillooly, J. F., Allen, A. P., Savage, V. M., \& West, G. B. (2004). TOWARD a METABOLIC THEORY OF ECOLOGY. \emph{Ecology}, \emph{85}(7), 1771--1789. \url{http://doi.org/10.1890/03-9000}
+
+\leavevmode\hypertarget{ref-burggren_developmental_2018}{}%
+Burggren, W.
+(2018).
+Developmental phenotypic plasticity helps bridge stochastic weather events associated with climate change.
+\emph{Journal of Experimental Biology}, \emph{221}(9), jeb161984.
+\url{http://doi.org/10.1242/jeb.161984}
+
+\leavevmode\hypertarget{ref-cheung_application_2008}{}%
+Cheung, W., Close, C., Lam, V., Watson, R., \& Pauly, D.
+(2008).
+Application of macroecological theory to predict effects of climate change on global fisheries potential.
+\emph{Mar.
+Ecol.
+Prog.
+Ser.}, \emph{365}, 187--197.
+\url{http://doi.org/10.3354/meps07414}
+
+\leavevmode\hypertarget{ref-clarke_temperature_2006}{}%
+Clarke, A. (2006). Temperature and the metabolic theory of ecology. \emph{Funct Ecology}, \emph{20}(2), 405--412. \url{http://doi.org/10.1111/j.1365-2435.2006.01109.x}
+
+\leavevmode\hypertarget{ref-clarke_scaling_1999}{}%
+Clarke, A., \& Johnston, N.
+M.
+(1999).
+Scaling of metabolic rate with body mass and temperature in teleost fish.
+\emph{J Anim Ecology}, \emph{68}(5), 893--905.
+\url{http://doi.org/10.1046/j.1365-2656.1999.00337.x}
+
+\leavevmode\hypertarget{ref-colinet_mechanisms_2018}{}%
+Colinet, H., Rinehart, J.
+P., Yocum, G.
+D., \& Greenlee, K.
+J.
+(2018).
+Mechanisms underpinning the beneficial effects of fluctuating thermal regimes in insect cold tolerance.
+\emph{J Exp Biol}, \emph{221}(14), jeb164806.
+\url{http://doi.org/10.1242/jeb.164806}
+
+\leavevmode\hypertarget{ref-colinet_insects_2015}{}%
+Colinet, H., Sinclair, B.
+J., Vernon, P., \& Renault, D.
+(2015).
+Insects in fluctuating thermal environments.
+\emph{Annu.
+Rev.
+Entomol.}, \emph{60}(1), 123--140.
+\url{http://doi.org/10.1146/annurev-ento-010814-021017}
+
+\leavevmode\hypertarget{ref-dang_thermal_2019}{}%
+Dang, W., Hu, Y.-C., Geng, J., Wang, J., \& Lu, H.-L.
+(2019).
+Thermal physiological performance of two freshwater turtles acclimated to different temperatures.
+\emph{J Comp Physiol B}, \emph{189}(1), 121--130.
+\url{http://doi.org/10.1007/s00360-018-1194-x}
+
+\leavevmode\hypertarget{ref-delava_effects_2016}{}%
+Delava, E., Fleury, F., \& Gibert, P.
+(2016).
+Effects of daily fluctuating temperatures on the drosophila--leptopilina boulardi parasitoid association.
+\emph{Journal of Thermal Biology}, \emph{60}, 95--102.
+\url{http://doi.org/10.1016/j.jtherbio.2016.06.012}
+
+\leavevmode\hypertarget{ref-delmas_mechanistic_2007}{}%
+Delmas, V., Prevot-Julliard, A.-C., Pieau, C., \& Girondot, M.
+(2007).
+A mechanistic model of temperature-dependent sex determination in a chelonian: The european pond turtle.
+\emph{Funct Ecology}, \emph{0}(0), 071027215958001--???
+\url{http://doi.org/10.1111/j.1365-2435.2007.01349.x}
+
+\leavevmode\hypertarget{ref-dempster_fluctuations_1981}{}%
+Dempster, J. P., \& Pollard, E. (1981). Fluctuations in resource availability and insect populations. \emph{Oecologia}, \emph{50}(3), 412--416. \url{http://doi.org/10.1007/BF00344984}
+
+\leavevmode\hypertarget{ref-descamps-julien_stable_2005}{}%
+Descamps-Julien, B., \& Gonzalez, A.
+(2005).
+STABLE COEXISTENCE IN a FLUCTUATING ENVIRONMENT: AN EXPERIMENTAL DEMONSTRATION.
+\emph{Ecology}, \emph{86}(10), 2815--2824.
+\url{http://doi.org/10.1890/04-1700}
+
+\leavevmode\hypertarget{ref-dobramysl_environmental_2013}{}%
+Dobramysl, U., \& Tauber, U.
+C.
+(2013).
+Environmental vs.
+Demographic variability in two-species predator-prey models.
+\emph{Phys.
+Rev.
+Lett.}, \emph{110}(4), 048105.
+\url{http://doi.org/10.1103/PhysRevLett.110.048105}
+
+\leavevmode\hypertarget{ref-dowd_thermal_2015}{}%
+Dowd, W.
+W., King, F.
+A., \& Denny, M.
+W.
+(2015).
+Thermal variation, thermal extremes and the physiological performance of individuals.
+\emph{Journal of Experimental Biology}, \emph{218}(12), 1956--1967.
+\url{http://doi.org/10.1242/jeb.114926}
+
+\leavevmode\hypertarget{ref-doyle_survival_1984}{}%
+Doyle, M.
+P., \& Schoeni, J.
+L.
+(1984).
+Survival and growth characteristics of escherichia coli associated with hemorrhagic colitis.
+\emph{Applied and Environmental Microbiology}, \emph{48}(4), 855--856.
+\url{http://doi.org/10.1128/AEM.48.4.855-856.1984}
+
+\leavevmode\hypertarget{ref-du_embryonic_2009}{}%
+Du, W.-G., Shen, J.-W., \& Wang, L.
+(2009).
+Embryonic development rate and hatchling phenotypes in the chinese three-keeled pond turtle (chinemys reevesii): The influence of fluctuating temperature versus constant temperature.
+\emph{Journal of Thermal Biology}, \emph{34}(5), 250--255.
+\url{http://doi.org/10.1016/j.jtherbio.2009.03.002}
+
+\leavevmode\hypertarget{ref-farhang-sardroodi_environmental_2019}{}%
+Farhang-Sardroodi, S., Darooneh, A.
+H., Kohandel, M., \& Komarova, N.
+L.
+(2019).
+Environmental spatial and temporal variability and its role in non-favoured mutant dynamics.
+\emph{J.
+R.
+Soc.
+Interface.}, \emph{16}(157), 20180781.
+\url{http://doi.org/10.1098/rsif.2018.0781}
+
+\leavevmode\hypertarget{ref-fresquet_response_2011}{}%
+Fresquet, N., \& Lazzari, C.
+R.
+(2011).
+Response to heat in rhodnius prolixus: The role of the thermal background.
+\emph{Journal of Insect Physiology}, \emph{57}(10), 1446--1449.
+\url{http://doi.org/10.1016/j.jinsphys.2011.07.012}
+
+\leavevmode\hypertarget{ref-garcia-ruiz_effects_2011}{}%
+García-Ruiz, E., Marco, V., \& Pérez-Moreno, I.
+(2011).
+Effects of variable and constant temperatures on the embryonic development and survival of a new grape pest, xylotrechus arvicola (coleoptera: Cerambycidae).
+\emph{Environ Entomol}, \emph{40}(4), 939--947.
+\url{http://doi.org/10.1603/EN11080}
+
+\leavevmode\hypertarget{ref-gillooly_response_2006}{}%
+Gillooly, J.
+F., Allen, A.
+P., Savage, V.
+M., Charnov, E.
+L., West, G.
+B., \& Brown, J.
+H.
+(2006).
+Response to clarke and fraser: Effects of temperature on metabolic rate.
+\emph{Funct Ecology}, \emph{20}(2), 400--404.
+\url{http://doi.org/10.1111/j.1365-2435.2006.01110.x}
+
+\leavevmode\hypertarget{ref-glass_should_2019}{}%
+Glass, J.
+R., \& Stahlschmidt, Z.
+R.
+(2019).
+Should i stay or should i go?
+Complex environments influence the developmental plasticity of flight capacity and flight-related trade-offs.
+\emph{Biological Journal of the Linnean Society}, \emph{128}(1), 59--69.
+\url{http://doi.org/10.1093/biolinnean/blz073}
+
+\leavevmode\hypertarget{ref-gudmundson_environmental_2015}{}%
+Gudmundson, S., Eklöf, A., \& Wennergren, U.
+(2015).
+Environmental variability uncovers disruptive effects of species' interactions on population dynamics.
+\emph{Proc.
+R.
+Soc.
+B.}, \emph{282}(1812), 20151126.
+\url{http://doi.org/10.1098/rspb.2015.1126}
+
+\leavevmode\hypertarget{ref-helmuth_organismal_2010}{}%
+Helmuth, B., Broitman, B.
+R., Yamane, L., Gilman, S.
+E., Mach, K., Mislan, K.
+A.
+S., \& Denny, M.
+W.
+(2010).
+Organismal climatology: Analyzing environmental variability at scales relevant to physiological stress.
+\emph{Journal of Experimental Biology}, \emph{213}(6), 995--1003.
+\url{http://doi.org/10.1242/jeb.038463}
+
+\leavevmode\hypertarget{ref-higgins_cochrane_2011}{}%
+Higgins, J., Thomas, J., Chandler, J., Li, T., Page, M., \& Welch, V.
+(2011).
+\emph{Cochrane handbook for systematic reviews of interventions} (Vol.
+5.1.0).
+Retrieved from \url{www.training.cochrane.org/handbook}
+
+\leavevmode\hypertarget{ref-holmes_robust_2016}{}%
+Holmes, C.
+R., Woollings, T., Hawkins, E., \& Vries, H.
+de.
+(2016).
+Robust future changes in temperature variability under greenhouse gas forcing and the relationship with thermal advection.
+\emph{Journal of Climate}, \emph{29}(6), 2221--2236.
+\url{http://doi.org/10.1175/JCLI-D-14-00735.1}
+
+\leavevmode\hypertarget{ref-holt_impacts_2003}{}%
+Holt, R.
+D., Barfield, M., \& Gonzalez, A.
+(2003).
+Impacts of environmental variability in open populations and communities: ``Inflation'' in sink environments.
+\emph{Theoretical Population Biology}, \emph{64}(3), 315--330.
+\url{http://doi.org/10.1016/S0040-5809(03)00087-X}
+
+\leavevmode\hypertarget{ref-horne_spatial_1995}{}%
+Horne, J. K., \& Schneider, D. C. (1995). Spatial variance in ecology. \emph{Oikos}, \emph{74}(1), 18. \url{http://doi.org/10.2307/3545670}
+
+\leavevmode\hypertarget{ref-huey_behavioral_1974}{}%
+Huey, R. B. (1974). Behavioral thermoregulation in lizards: Importance of associated costs. \emph{Science}, \emph{184}(4140), 1001--1003. \url{http://doi.org/10.1126/science.184.4140.1001}
+
+\leavevmode\hypertarget{ref-huey_physiological_1990}{}%
+Huey, R. B., \& Bennett, A. F. (1990). Physiological adjustments to fluctuating thermal environments: An ecological and evolutionary perspective. \emph{Stress Proteins in Biology and Medicine}, 23.
+
+\leavevmode\hypertarget{ref-huey_predicting_2012}{}%
+Huey, R.
+B., Kearney, M.
+R., Krockenberger, A., Holtum, J.
+A.
+M., Jess, M., \& Williams, S.
+E.
+(2012).
+Predicting organismal vulnerability to climate warming: Roles of behaviour, physiology and adaptation.
+\emph{Phil.
+Trans.
+R.
+Soc.
+B}, \emph{367}(1596), 1665--1679.
+\url{http://doi.org/10.1098/rstb.2012.0005}
+
+\leavevmode\hypertarget{ref-husak_how_2017}{}%
+Husak, J.
+F., \& Lailvaux, S.
+P.
+(2017).
+How do we measure the cost of whole-organism performance traits?
+\emph{Integrative and Comparative Biology}, \emph{57}(2), 333--343.
+\url{http://doi.org/10.1093/icb/icx048}
+
+\leavevmode\hypertarget{ref-isaac_isaac_2003pdf_2003}{}%
+Isaac, L. A. (2003). Isaac\_2003.pdf. University of Victoria. Retrieved from \url{https://dspace.library.uvic.ca/handle/1828/356}
+
+\leavevmode\hypertarget{ref-kearney_potential_2009}{}%
+Kearney, M., Shine, R., \& Porter, W.
+P.
+(2009).
+The potential for behavioral thermoregulation to buffer "cold-blooded" animals against climate warming.
+\emph{Proceedings of the National Academy of Sciences}, \emph{106}(10), 3835--3840.
+\url{http://doi.org/10.1073/pnas.0808913106}
+
+\leavevmode\hypertarget{ref-kern_temperature_2014}{}%
+Kern, P., Cramp, R.
+L., \& Franklin, C.
+E.
+(2014).
+Temperature and UV-b-insensitive performance in tadpoles of the ornate burrowing frog: An ephemeral pond specialist.
+\emph{Journal of Experimental Biology}, \emph{217}(8), 1246--1252.
+\url{http://doi.org/10.1242/jeb.097006}
+
+\leavevmode\hypertarget{ref-kern_physiological_2015-3}{}%
+Kern, P., Cramp, R.
+L., \& Franklin, C.
+E.
+(2015).
+Physiological responses of ectotherms to daily temperature variation.
+\emph{Journal of Experimental Biology}, \emph{218}(19), 3068--3076.
+\url{http://doi.org/10.1242/jeb.123166}
+
+\leavevmode\hypertarget{ref-khelifa_usefulness_2019}{}%
+Khelifa, R., Blanckenhorn, W.
+U., Roy, J., Rohner, P.
+T., \& Mahdjoub, H.
+(2019).
+Usefulness and limitations of thermal performance curves in predicting ectotherm development under climatic variability.
+\emph{J Anim Ecol}, \emph{88}(12), 1901--1912.
+\url{http://doi.org/10.1111/1365-2656.13077}
+
+\leavevmode\hypertarget{ref-klepsatel_variation_2013}{}%
+Klepsatel, P., Gáliková, M., De Maio, N., Huber, C.
+D., Schlötterer, C., \& Flatt, T.
+(2013).
+VARIATION IN THERMAL PERFORMANCE AND REACTION NORMS AMONG POPULATIONS OF \emph{DROSOPHILA MELANOGASTER}: THERMAL PERFORMANCE IN DROSOPHILA.
+\emph{Evolution}, \emph{67}(12), 3573--3587.
+\url{http://doi.org/10.1111/evo.12221}
+
+\leavevmode\hypertarget{ref-knoll_chapter_2008}{}%
+Knoll, A. (2008). Chapter 1. In \emph{Cyanobacteria and earth history}.
+
+\leavevmode\hypertarget{ref-kremer_gradual_2018}{}%
+Kremer, C.
+T., Fey, S.
+B., Arellano, A.
+A., \& Vasseur, D.
+A.
+(2018).
+Gradual plasticity alters population dynamics in variable environments: Thermal acclimation in the green alga \emph{chlamydomonas reinhartdii}.
+\emph{Proc.
+R.
+Soc.
+B.}, \emph{285}(1870), 20171942.
+\url{http://doi.org/10.1098/rspb.2017.1942}
+
+\leavevmode\hypertarget{ref-krenek_coping_2012}{}%
+Krenek, S., Petzoldt, T., \& Berendonk, T.
+U.
+(2012).
+Coping with temperature at the warm edge -- patterns of thermal adaptation in the microbial eukaryote paramecium caudatum.
+\emph{PLoS ONE}, \emph{7}(3), e30598.
+\url{http://doi.org/10.1371/journal.pone.0030598}
+
+\leavevmode\hypertarget{ref-kroeker_ecological_2020}{}%
+Kroeker, K.
+J., Bell, L.
+E., Donham, E.
+M., Hoshijima, U., Lummis, S., Toy, J.
+A., \& Willis‐Norton, E.
+(2020).
+Ecological change in dynamic environments: Accounting for temporal environmental variability in studies of ocean change biology.
+\emph{Glob Change Biol}, \emph{26}(1), 54--67.
+\url{http://doi.org/10.1111/gcb.14868}
+
+\leavevmode\hypertarget{ref-lampert_adaptive_1989}{}%
+Lampert, W. (1989). The adaptive significance of diel vertical migration of zooplankton. \emph{Functional Ecology}, \emph{3}(1), 21. \url{http://doi.org/10.2307/2389671}
+
+\leavevmode\hypertarget{ref-lawson_environmental_2015}{}%
+Lawson, C.
+R., Vindenes, Y., Bailey, L., \& Pol, M.
+van de.
+(2015).
+Environmental variation and population responses to global change.
+\emph{Ecol Lett}, \emph{18}(7), 724--736.
+\url{http://doi.org/10.1111/ele.12437}
+
+\leavevmode\hypertarget{ref-lowenborg_how_2012}{}%
+Löwenborg, K., Gotthard, K., \& Hagman, M.
+(2012).
+How a thermal dichotomy in nesting environments influences offspring of the world's most northerly oviparous snake, \emph{natrix natrix} (colubridae): Grass snake nest-site dichotomy.
+\emph{Biol J Linn Soc Lond}, \emph{107}(4), 833--844.
+\url{http://doi.org/10.1111/j.1095-8312.2012.01972.x}
+
+\leavevmode\hypertarget{ref-manenti_predictability_2014-2}{}%
+Manenti, T., Sørensen, J.
+G., Moghadam, N.
+N., \& Loeschcke, V.
+(2014).
+Predictability rather than amplitude of temperature fluctuations determines stress resistance in a natural population of \emph{drosophila simulans}.
+\emph{J.
+Evol.
+Biol.}, \emph{27}(10), 2113--2122.
+\url{http://doi.org/10.1111/jeb.12463}
+
+\leavevmode\hypertarget{ref-massey_measurement_2019}{}%
+Massey, M.
+D., Holt, S.
+M., Brooks, R.
+J., \& Rollinson, N.
+(2019).
+Measurement and modelling of primary sex ratios for species with temperature-dependent sex determination.
+\emph{J Exp Biol}, \emph{222}(1), jeb190215.
+\url{http://doi.org/10.1242/jeb.190215}
+
+\leavevmode\hypertarget{ref-morison_light_2020}{}%
+Morison, F., Franzè, G., Harvey, E., \& Menden‐Deuer, S.
+(2020).
+Light fluctuations are key in modulating plankton trophic dynamics and their impact on primary production.
+\emph{Limnol Oceanogr}, \emph{5}(5), 346--353.
+\url{http://doi.org/10.1002/lol2.10156}
+
+\leavevmode\hypertarget{ref-moyano_effects_2017}{}%
+Moyano, M., Candebat, C., Ruhbaum, Y., Álvarez-Fernández, S., Claireaux, G., Zambonino-Infante, J.-L., \& Peck, M.
+A.
+(2017).
+Effects of warming rate, acclimation temperature and ontogeny on the critical thermal maximum of temperate marine fish larvae.
+\emph{PLoS ONE}, \emph{12}(7), e0179928.
+\url{http://doi.org/10.1371/journal.pone.0179928}
+
+\leavevmode\hypertarget{ref-noble_developmental_2018}{}%
+Noble, D.
+W.
+A., Stenhouse, V., \& Schwanz, L.
+E.
+(2018).
+Developmental temperatures and phenotypic plasticity in reptiles: A systematic review and meta-analysis: Incubation temperature and plasticity.
+\emph{Biol Rev}, \emph{93}(1), 72--97.
+\url{http://doi.org/10.1111/brv.12333}
+
+\leavevmode\hypertarget{ref-paaijmans_temperature_2013}{}%
+Paaijmans, K.
+P., Heinig, R.
+L., Seliga, R.
+A., Blanford, J.
+I., Blanford, S., Murdock, C.
+C., \& Thomas, M.
+B.
+(2013).
+Temperature variation makes ectotherms more sensitive to climate change.
+\emph{Glob Change Biol}, \emph{19}(8), 2373--2380.
+\url{http://doi.org/10.1111/gcb.12240}
+
+\leavevmode\hypertarget{ref-parachu_marco_new_2017}{}%
+Parachu Marco, M.
+V., Leiva, P., Iungman, J.
+L., Simonici, M.
+S., \& Pina, C.
+I.
+(2017).
+New evidence characterizing temperature-dependent sex determination in broad-snouted caiman, caiman latirostris.
+\emph{Herpetological Conservation and Biology}, (12), 78--84.
+
+\leavevmode\hypertarget{ref-parepa_environmental_2013}{}%
+Parepa, M., Fischer, M., \& Bossdorf, O. (2013). Environmental variability promotes plant invasion. \emph{Nat Commun}, \emph{4}(1), 1604. \url{http://doi.org/10.1038/ncomms2632}
+
+\leavevmode\hypertarget{ref-pendlebury_variation_2004-1}{}%
+Pendlebury, C. J. (2004). Variation in temperature increases the cost of living in birds. \emph{Journal of Experimental Biology}, \emph{207}(12), 2065--2070. \url{http://doi.org/10.1242/jeb.00999}
+
+\leavevmode\hypertarget{ref-petchey_environmental_2000}{}%
+Petchey, O.
+L.
+(2000).
+Environmental colour affects aspects of single--species population dynamics.
+\emph{Proc.
+R.
+Soc.
+Lond.
+B}, \emph{267}(1445), 747--754.
+\url{http://doi.org/10.1098/rspb.2000.1066}
+
+\leavevmode\hypertarget{ref-pina_effect_2003}{}%
+Piña, C.
+I., Larriera, A., \& Cabrera, M.
+R.
+(2003).
+Effect of incubation temperature on incubation period, sex ratio, hatching success, and survivorship in caiman latirostris (crocodylia, alligatoridae).
+\emph{Journal of Herpetology}, \emph{37}(1), 199--202.
+\href{http://doi.org/10.1670/0022-1511(2003)037\%5B0199:EOITOI\%5D2.0.CO;2}{http://doi.org/10.1670/0022-1511(2003)037{[}0199:EOITOI{]}2.0.CO;2}
+
+\leavevmode\hypertarget{ref-portner_ecology_2008}{}%
+Portner, H. O., \& Farrell, A. P. (2008). ECOLOGY: Physiology and climate change. \emph{Science}, \emph{322}(5902), 690--692. \url{http://doi.org/10.1126/science.1163156}
+
+\leavevmode\hypertarget{ref-qu_incubation_2014}{}%
+Qu, Y.-F., Lu, H.-L., Li, H., \& Ji, X.
+(2014).
+Incubation temperature fluctuation does not affect incubation length and hatchling phenotype in the chinese skink plestiodon chinensis.
+\emph{Journal of Thermal Biology}, \emph{46}, 10--15.
+\url{http://doi.org/10.1016/j.jtherbio.2014.09.008}
+
+\leavevmode\hypertarget{ref-reed_phenotypic_2010}{}%
+Reed, T.
+E., Waples, R.
+S., Schindler, D.
+E., Hard, J.
+J., \& Kinnison, M.
+T.
+(2010).
+Phenotypic plasticity and population viability: The importance of environmental predictability.
+\emph{Proc.
+R.
+Soc.
+B.}, \emph{277}(1699), 3391--3400.
+\url{http://doi.org/10.1098/rspb.2010.0771}
+
+\leavevmode\hypertarget{ref-rolandi_costs_2018}{}%
+Rolandi, C., \& Schilman, P.
+E.
+(2018).
+The costs of living in a thermal fluctuating environment for the tropical haematophagous bug, rhodnius prolixus.
+\emph{Journal of Thermal Biology}, \emph{74}, 92--99.
+\url{http://doi.org/10.1016/j.jtherbio.2018.03.022}
+
+\leavevmode\hypertarget{ref-romanuk_environmental_2002}{}%
+Romanuk, T.
+N., \& Kolasa, J.
+(2002).
+Environmental variability alters the relationship between richness and variability of community abundances in aquatic rock pool microcosms.
+\emph{Écoscience}, \emph{9}(1), 55--62.
+\url{http://doi.org/10.1080/11956860.2002.11682690}
+
+\leavevmode\hypertarget{ref-ruel_jensens_1999}{}%
+Ruel, J.
+J., \& Ayres, M.
+P.
+(1999).
+Jensen's inequality predicts effects of environmental variation.
+\emph{Trends in Ecology \& Evolution}, \emph{14}(9), 361--366.
+\url{http://doi.org/10.1016/S0169-5347(99)01664-X}
+
+\leavevmode\hypertarget{ref-russill_tipping_2009}{}%
+Russill, C., \& Nyssa, Z.
+(2009).
+The tipping point trend in climate change communication.
+\emph{Global Environmental Change}, \emph{19}(3), 336--344.
+\url{http://doi.org/10.1016/j.gloenvcha.2009.04.001}
+
+\leavevmode\hypertarget{ref-saxon_temperature_2018-1}{}%
+Saxon, A.
+D., O'Brien, E.
+K., \& Bridle, J.
+R.
+(2018).
+Temperature fluctuations during development reduce male fitness and may limit adaptive potential in tropical rainforest \emph{drosophila}.
+\emph{J.
+Evol.
+Biol.}, \emph{31}(3), 405--415.
+\url{http://doi.org/10.1111/jeb.13231}
+
+\leavevmode\hypertarget{ref-seebacher_embryonic_2014}{}%
+Seebacher, F., \& Grigaltchik, V.
+S.
+(2014).
+Embryonic developmental temperatures modulate thermal acclimation of performance curves in tadpoles of the frog limnodynastes peronii.
+\emph{PLoS ONE}, \emph{9}(9), e106492.
+\url{http://doi.org/10.1371/journal.pone.0106492}
+
+\leavevmode\hypertarget{ref-semenov_influence_2007}{}%
+Semenov, A.
+V., Van Bruggen, A.
+H., Van Overbeek, L., Termorshuizen, A.
+J., \& Semenov, A.
+M.
+(2007).
+Influence of temperature fluctuations on escherichia coli o157:H7 and salmonella enterica serovar typhimurium in cow manure: Effect of temperature fluctuation on human pathogens.
+\emph{FEMS Microbiology Ecology}, \emph{60}(3), 419--428.
+\url{http://doi.org/10.1111/j.1574-6941.2007.00306.x}
+
+\leavevmode\hypertarget{ref-simoncini_influence_2019}{}%
+Simoncini, M.
+S., Leiva, P.
+M., Piña, C.
+I., \& Cruz, F.
+B.
+(2019).
+Influence of temperature variation on incubation period, hatching success, sex ratio, and phenotypes in \emph{caiman latirostris}: SIMONCINI \textless{}span style="font-variant:Small-caps;"\textgreater{}et al.\textless{}/span\textgreater{}.
+\emph{J.
+Exp.
+Zool.}, \emph{331}(5), 299--307.
+\url{http://doi.org/10.1002/jez.2265}
+
+\leavevmode\hypertarget{ref-singh_effect_2020}{}%
+Singh, R., Prathibha, P., \& Jain, M.
+(2020).
+\emph{Effect of temperature on life-history traits and mating calls of a field cricket, \textup{acanthogryllus asiaticus}} (preprint).
+Ecology.
+Retrieved from \url{http://biorxiv.org/lookup/doi/10.1101/2020.06.06.137869}
+
+\leavevmode\hypertarget{ref-sommer_paradox_1984}{}%
+Sommer, U.
+(1984).
+The paradox of the plankton: Fluctuations of phosphorus availability maintain diversity of phytoplankton in flow-through cultures1.
+\emph{Limnol.
+Oceanogr.}, \emph{29}(3), 633--636.
+\url{http://doi.org/10.4319/lo.1984.29.3.0633}
+
+\leavevmode\hypertarget{ref-soo_origins_2017}{}%
+Soo, R.
+M., Hemp, J., Parks, D.
+H., Fischer, W.
+W., \& Hugenholtz, P.
+(2017).
+On the origins of oxygenic photosynthesis and aerobic respiration in cyanobacteria.
+\emph{Science}, \emph{355}(6332), 1436--1440.
+\url{http://doi.org/10.1126/science.aal3794}
+
+\leavevmode\hypertarget{ref-stewart_mesocosm_2013}{}%
+Stewart, R.
+I., Dossena, M., Bohan, D.
+A., Jeppesen, E., Kordas, R.
+L., Ledger, M.
+E., \ldots{} Woodward, G.
+(2013).
+Mesocosm experiments as a tool for ecological climate-change research.
+In \emph{Advances in ecological research} (Vol.
+48, pp.
+71--181).
+Elsevier.
+\url{http://doi.org/10.1016/B978-0-12-417199-2.00002-1}
+
+\leavevmode\hypertarget{ref-sunday_thermal_2012}{}%
+Sunday, J.
+M., Bates, A.
+E., \& Dulvy, N.
+K.
+(2012).
+Thermal tolerance and the global redistribution of animals.
+\emph{Nature Clim Change}, \emph{2}(9), 686--690.
+\url{http://doi.org/10.1038/nclimate1539}
+
+\leavevmode\hypertarget{ref-tabari_shift_2012}{}%
+Tabari, H., Hosseinzadeh Talaee, P., Ezani, A., \& Shifteh Some'e, B.
+(2012).
+Shift changes and monotonic trends in autocorrelated temperature series over iran.
+\emph{Theor Appl Climatol}, \emph{109}(1), 95--108.
+\url{http://doi.org/10.1007/s00704-011-0568-8}
+
+\leavevmode\hypertarget{ref-treidel_temperature_2016-1}{}%
+Treidel, L.
+A., Carter, A.
+W., \& Bowden, R.
+M.
+(2016).
+Temperature experienced during incubation affects antioxidant capacity but not oxidative damage in hatchling red-eared slider turtles ( \emph{trachemys scripta elegans} ).
+\emph{J Exp Biol}, \emph{219}(4), 561--570.
+\url{http://doi.org/10.1242/jeb.128843}
+
+\leavevmode\hypertarget{ref-tuff_framework_2016}{}%
+Tuff, K.
+T., Tuff, T., \& Davies, K.
+F.
+(2016).
+A framework for integrating thermal biology into fragmentation research.
+\emph{Ecol Lett}, \emph{19}(4), 361--374.
+\url{http://doi.org/10.1111/ele.12579}
+
+\leavevmode\hypertarget{ref-vasseur_increased_2014}{}%
+Vasseur, D.
+A., DeLong, J.
+P., Gilbert, B., Greig, H.
+S., Harley, C.
+D.
+G., McCann, K.
+S., \ldots{} O'Connor, M.
+I.
+(2014).
+Increased temperature variation poses a greater risk to species than climate warming.
+\emph{Proc.
+R.
+Soc.
+B.}, \emph{281}(1779), 20132612.
+\url{http://doi.org/10.1098/rspb.2013.2612}
+
+\leavevmode\hypertarget{ref-vasseur_phase-locking_2009}{}%
+Vasseur, D.
+A., \& Fox, J.
+W.
+(2009).
+Phase-locking and environmental fluctuations generate synchrony in a predator--prey community.
+\emph{Nature}, \emph{460}(7258), 1007--1010.
+\url{http://doi.org/10.1038/nature08208}
+
+\leavevmode\hypertarget{ref-vasseur_color_2004}{}%
+Vasseur, D. A., \& Yodzis, P. (2004). THE COLOR OF ENVIRONMENTAL NOISE. \emph{Ecology}, \emph{85}(4), 1146--1152. \url{http://doi.org/10.1890/02-3122}
+
+\leavevmode\hypertarget{ref-viechtbauer_conducting_2010}{}%
+Viechtbauer, W. (2010). Conducting meta-analyses in \emph{r} with the \textbf{metafor} package. \emph{J. Stat. Soft.}, \emph{36}(3). \url{http://doi.org/10.18637/jss.v036.i03}
+
+\leavevmode\hypertarget{ref-wang_effects_2015}{}%
+Wang, Y.-b., Pang, W.-x., Yv, X.-n., Li, J.-j., Zhang, Y.-a., \& Sung, C.-k.
+(2015).
+The effects of fluctuating culture temperature on stress tolerance and antioxidase expression in esteya vermicola.
+\emph{J Microbiol.}, \emph{53}(2), 122--126.
+\url{http://doi.org/10.1007/s12275-015-4529-2}
+
+\leavevmode\hypertarget{ref-west_allometric_2002}{}%
+West, G.
+B., Woodruff, W.
+H., \& Brown, J.
+H.
+(2002).
+Allometric scaling of metabolic rate from molecules and mitochondria to cells and mammals.
+\emph{Proceedings of the National Academy of Sciences}, \emph{99}(Supplement 1), 2473--2478.
+\url{http://doi.org/10.1073/pnas.012579799}
+
+\leavevmode\hypertarget{ref-whitehead_effect_1989}{}%
+Whitehead, P.
+J., Puckridge, J.
+T., Leigh, C.
+M., \& Seymour, R.
+S.
+(1989).
+Effect of temperature on jump performance of the frog \emph{limnodynastes tasmaniensis}.
+\emph{Physiological Zoology}, \emph{62}(4), 937--949.
+\url{http://doi.org/10.1086/physzool.62.4.30157938}
+
+\leavevmode\hypertarget{ref-wigley_anthropogenic_1998-1}{}%
+Wigley, T.
+M.
+(1998).
+Anthropogenic influence on the autocorrelation structure of hemispheric-mean temperatures.
+\emph{Science}, \emph{282}(5394), 1676--1679.
+\url{http://doi.org/10.1126/science.282.5394.1676}
+
+\leavevmode\hypertarget{ref-yoshida_rapid_2003}{}%
+Yoshida, T., Jones, L.
+E., Ellner, S.
+P., Fussmann, G.
+F., \& Hairston, N.
+G.
+(2003).
+Rapid evolution drives ecological dynamics in a predator--prey system.
+\emph{Nature}, \emph{424}(6946), 303--306.
+\url{http://doi.org/10.1038/nature01767}
 

### 2021-04-29 writing /thesis_draft/index.Rmd
fully formatted draft :)

@@
 ---
-title: 'Mean temperature, fluctuation range, and biological organization affect biological responses'
-author: 'Margaret A. Slein'
+title: 'Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis'
+author: 'Margaret Anne Slein'
 date: 'May 2021'
@@
 degree: 'Bachelor of Arts'
+header-includes:
+    - \usepackage{setspace}\onehalfspacing
 knit: bookdown::render_book
@@
 lof: true
+nocite : |
+
+  @alagawany_heat_2017, @delava_effects_2016, @dang_thermal_2019, 
+  @singh_effect_2020, @seebacher_embryonic_2014, @whitehead_effect_1989, 
+  @kern_temperature_2014, @isaac_isaac_2003pdf_2003, @sunday_thermal_2012, 
+  @baojun_seasonal_2014, @fresquet_response_2011, @doyle_survival_1984, 
+  @bronikowski_evolutionary_2001, @parachu_marco_new_2017,
+  @pina_effect_2003, @simoncini_influence_2019, @dang_thermal_2019, 
+  @wang_effects_2015, @du_embryonic_2009, @garcia-ruiz_effects_2011, 
+  @glass_should_2019, @kern_physiological_2015-3, @lowenborg_how_2012, 
+  @manenti_predictability_2014-2, @pendlebury_variation_2004-1,
+  @qu_incubation_2014, @rolandi_costs_2018, @saxon_temperature_2018-1, 
+  @semenov_influence_2007, @treidel_temperature_2016-1, @klepsatel_variation_2013,
+  @krenek_coping_2012, @tuff_framework_2016
+
 ---
@@
 ## The historical significance of variability 
-It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bouchard 2014, Bambach 1993).
-Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Soo et al 2017, Blankenship 2017).
-Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll 2008).
+It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia [@bouchard_ecosystem_2014; @bambach_seafood_1993].
+Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever [@soo_origins_2017, @blankenship_how_2017].
+Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time [@knoll_chapter_2008].
 In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnik et al 2007).
-Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Reed et al 2010, Burggren 2018).
+Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant [@reed_phenotypic_2010; @burggren_developmental_2018].
 Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.
@@
 El Nino Southern Oscillation).
-As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (Khelifa et al 2019, Stewart et al 2013).
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems [khelifa_usefulness_2019; @stewart_mesocosm_2013].
 How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
 As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.
-hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Helmuth et al 2010, Bernardo 2014, Huey et al 2012).
+hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses [@helmuth_organismal_2010; @bernardo_biologically_2014; @huey_predicting_2012].
 
 ## Why this study on variability matters 
-There is a large body of literature on environmental variability including its relevance for positive effects for populations and communities, i.e.
-inflationary effects in sink-source systems (Holt et al 2003, Vasseur and Fox 2009), such that variability aids population density when organisms emigrate.
+There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, i.e.
+inflationary effects in sink-source systems [@holt_impacts_2003; @vasseur_phase-locking_2009], such that variability aids population density when organisms emigrate.
 However, there is a lack of consensus as to which scales of biological organization, i.e.

### 2021-04-29 writing /thesis_draft/_book/thesis.tex
fixing the damn title

@@
 % To pass between YAML and LaTeX the dollar signs are added by CII
-\title{Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis}
+\title{Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation range and biological organization: A Meta Analysis}
 \author{Margaret Anne Slein}

### 2021-04-29 writing /thesis_draft/index.Rmd
fixing the damn title

@@
 ---
-title: 'Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis'
+title: 'Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation range and biological organization: A Meta Analysis'
 author: 'Margaret Anne Slein'

### 2021-04-29 writing /thesis_draft/02-chap2.Rmd
all formatting is in, just need to read for grammar

@@
                          "stressful"),
-                     Description = c("Reported mean temperature in study",
+                     Description = c("Reported mean temperature (°C) in study",
                                      "Reported range (difference between minimum 
-                                     and maximum) of temperature fluctuation experienced", 
+                                     and maximum) of temperature fluctuation (°C) experienced", 
                                      "Level of biological organization, categorized

### 2021-04-29 writing /thesis_draft/03-chap3.Rmd
all formatting is in, just need to read for grammar

@@
 
-Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 3.1.
+Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 3.1.
 These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
@@
 
-Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p < 0.0001).
-Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (df = 139, p < 0.00001).
-Organization level was also important in the model, as a statistically significant covariate (df= 139, p < 0.05).
+Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001).
+Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p < 0.00001).
+Organization level (n=140) was also important in the model, as a statistically significant covariate (df= 7, p < 0.05).
 
-```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
+```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
+Orange error bars represent 95 percent confidence intervals", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
 full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
@@
                                        "size", "experimental age","organization level" 
-                                      ))
+                                      ), 
+         lb = c(-0.9781, -0.1371, 0.1794, -0.0184, -0.0177, -0.8898, -1.3401, -1.2954), 
+         ub= c(2.7433, 1.3568, 0.3711, 0.0536, -0.0097, 0.2348, 0.2367, 0.0041))
 
 label.df <- data.frame(coefficients = c("fluctuation range", "fluctuation range:mean temperature", "organization level"),
-                       value = c(0.35, -0.07, -0.69))
+                       value = c(0.5, -0.25, -1.4))
 
@@
              y = value,
-             fill = value > 0)) +
-  geom_col(show.legend = FALSE) +
+             color = value > 0)) +
+  geom_errorbar(aes(ymin=lb, ymax=ub), width=.2,
+                 position=position_dodge(.95), color = "black")+ 
+  geom_point(size = 4, show.legend = FALSE) +
   labs(x = NULL, y = "Coefficient estimate") +
-  scale_fill_viridis(discrete = TRUE)+
+  scale_color_viridis(discrete = TRUE)+
   theme_bw()+
   coord_flip()+
-  geom_text(data=label.df, label = "*")
+  geom_text(data=label.df, label = "*", color = "black", size = 6)
 ```
@@
 
-Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2).The most negative estimate (organization level) was significantly influential in our model at a significance level of p < 0.05.
-Surprisingly, the most negative estimates of our model coefficients were not statistically significant.
-In our analysis of thermal stress  via a separate random effects model, we found thermal stress was not statistically significant in explaining the effect size (Table 3.1).
+Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2).The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p < 0.05.
+Surprisingly, the remaining negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
+In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).
 
@@
 The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
-At lower mean temperatures, higher fluctuation ranges have a more positive effect on..., though that trend starkly ends at about $24^{\circ}$C.
+At lower mean temperatures, higher fluctuation ranges have a more positive effect on organism responses and performance, though that trend starkly ends at about $24^{\circ}$C.
 Beyond $24^{\circ}$C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
@@
 ```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature range, and effect size.
-Temperature fluctuation range from 0-20C, Mean temperature range from 7-33OC.
+Temperature fluctuation range from 0-20°C, Mean temperature range from 7-33°C.
 Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and effect size"}
@@
   geom_jitter( alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
-  scale_fill_viridis('Temperature range (C)', option = "C")+
+  scale_fill_viridis('Temperature range (°C)', option = "C")+
   theme_classic()+
-  xlab("Mean Temperature (C)")+
+  xlab("Mean Temperature (°C)")+
   ylab("Effect Size")+
@@
 
-As noted in the methods, when we excluded one highly influential study and included thermal stress (exclusion model), size became much more important in the thermal stress model (df=123, p < 0.001) as did mean temperature (df=123, p < 0.001).
-Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the thermal stress model (df=123, p < 0.1).
+As noted in the methods, when we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p < 0.001) as did mean temperature (df=8, p < 0.001).
+Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p < 0.1).
 
-```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by organization level"}
+```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by organization level: individual or population level responses reported in studies.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by organization level"}
 # scatterplot of standardized mean response vs flux range colored and lm fit by org level
@@
   geom_point()+
-  labs(x = "Temperature fluctuation range (C)",
+  labs(x = "Temperature fluctuation range (°C)",
        y ="SMD",
@@
 
-```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal stress.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by thermal stress"}
+```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by thermal stress.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by thermal stress"}
 normalized %>%
@@
   geom_jitter()+
-  labs(x = "Temperature fluctuation range (C)",
+  labs(x = "Temperature fluctuation range (°C)",
        y ="SMD",
@@
 
-```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by response type"}
+```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by life history metric: responses categorized as rate or trait.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by response type"}
 normalized %>%
@@
 
-```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by body size"}
+```{r, echo=FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by body size: extra-small, small, medium, or large.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by body size"}
 ggplot(normalized, aes(x=flux_range, y=yi, color = factor(size)))+
   geom_point()+
-  labs(x = "Temperature fluctuation range (C)",
+  labs(x = "Temperature fluctuation range (°C)",
        y ="SMD",
@@
 
-```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by experimental age"}
+```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by experimental age: larval, juvenile, or adult.", fig.align="center", out.width = '85%', fig.scap="Effect sizes across temperature range by experimental age"}
 ggplot(normalized, aes(x=flux_range, y=yi, color = factor(exp_age)))+
   geom_point()+
-  labs(x = "Temperature fluctuation range (C)",
+  labs(x = "Temperature fluctuation range (°C)",
        y ="SMD",

### 2021-04-29 writing /thesis_draft/05-appendix.Rmd
all formatting is in, just need to read for grammar

@@
                          "Trachemys scripta", "Esteya vermicola"), 
-                 Topt = c("24-26.6 C", "25 C", "37.97 +/- 0.64", "< 30 C", "30 C", "30-33 C",
-                          "33 C", " </= 30 C", "26 C", "33 C", "30-35 C", "37 C, 35-36 C",
-                          "35-36 C", "32-33 C", "32-33 C", "26 C"), 
+                 Topt = c("24-26.6°C", "25°C", "37.97 ± 0.64 °C", "~30°C", "30°C", "30-33°C",
+                          "33°C", "~30°C", "26°C", "33°C", "30-35°C", "37°C, 35-36°C",
+                          "35-36°C", "32-33°C", "32-33°C", "26°C"), 
                  `Reference` = c("Klepsatel 2013, David 1988", "De lava et al 2016", 
@@
 kable(table4, "latex", booktabs = T, 
-      col.names = c("Species", "Topt", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X.
+      col.names = c("Species", "Topt (°C)", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X.
 arvicola, used in Garcia-Ruiz et al 2011)", caption.short ="Thermal stress data by species") %>%

### 2021-04-29 writing /thesis_draft/_book/thesis.tex
all formatting is in, just need to read for grammar

--- 2021-04-29 writing /thesis_draft/_book/thesis.tex
@@
 % To pass between YAML and LaTeX the dollar signs are added by CII
-\title{Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis}
+\title{Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta analysis}
 \author{Margaret Anne Slein}
@@
 In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
-Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Nino Southern Oscillation).
+Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Niño Southern Oscillation).
 As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (khelifa\_usefulness\_2019; Stewart et al., 2013).
@@
 \midrule
-\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature in study}\\
+\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature (°C) in study}\\
 flux\_range & Reported range (difference between minimum 
-                                     and maximum) of temperature fluctuation experienced\\
+                                     and maximum) of temperature fluctuation (°C) experienced\\
 org\_level & Level of biological organization, categorized
@@
 \end{figure}
-Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1.
+Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1.
 These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
@@
 
-Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p \textless{} 0.0001).
-Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (df = 139, p \textless{} 0.00001).
-Organization level was also important in the model, as a statistically significant covariate (df= 139, p \textless{} 0.05).
+Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p \textless{} 0.0001).
+Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p \textless{} 0.00001).
+Organization level (n=140) was also important in the model, as a statistically significant covariate (df= 7, p \textless{} 0.05).
 \begin{figure}
@@
 
-\caption[Coefficient estimates for full model]{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.}\label{fig:unnamed-chunk-19}
+\caption[Coefficient estimates for full model]{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
+Orange error bars represent 95 percent confidence intervals}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 
-Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).The most negative estimate (organization level) was significantly influential in our model at a significance level of p \textless{} 0.05.
-Surprisingly, the most negative estimates of our model coefficients were not statistically significant.
-In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant in explaining the effect size (Table 3.1).
+Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p \textless{} 0.05.
+Surprisingly, the remaining negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
+In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).
 \begin{table}[!h]
@@
 The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
-At lower mean temperatures, higher fluctuation ranges have a more positive effect on\ldots{}, though that trend starkly ends at about \(24^{\circ}\)C.
+At lower mean temperatures, higher fluctuation ranges have a more positive effect on organism responses and performance, though that trend starkly ends at about \(24^{\circ}\)C.
 Beyond \(24^{\circ}\)C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
@@
 \caption[Scatterplot of relationship between range, mean, and effect size]{Scatterplot describing the relationship between mean temperature, temperature range, and effect size.
-Temperature fluctuation range from 0-20C, Mean temperature range from 7-33OC.
+Temperature fluctuation range from 0-20°C, Mean temperature range from 7-33°C.
 Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)}\label{fig:unnamed-chunk-21}
@@
 
-As noted in the methods, when we excluded one highly influential study and included thermal stress (exclusion model), size became much more important in the thermal stress model (df=123, p \textless{} 0.001) as did mean temperature (df=123, p \textless{} 0.001).
-Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the thermal stress model (df=123, p \textless{} 0.1).
+As noted in the methods, when we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p \textless{} 0.001) as did mean temperature (df=8, p \textless{} 0.001).
+Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p \textless{} 0.1).
 \begin{figure}
@@
 
-\caption[Effect sizes across temperature range by organization level]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-22}
+\caption[Effect sizes across temperature range by organization level]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-22}
 \end{figure}
@@
 
-\caption[Effect sizes across temperature range by thermal stress]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal stress.}\label{fig:unnamed-chunk-23}
+\caption[Effect sizes across temperature range by thermal stress]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by thermal stress.}\label{fig:unnamed-chunk-23}
 \end{figure}
@@
 
-\caption[Effect sizes across temperature range by response type]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-24}
+\caption[Effect sizes across temperature range by response type]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-24}
 \end{figure}
@@
 
-\caption[Effect sizes across temperature range by body size]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-25}
+\caption[Effect sizes across temperature range by body size]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-25}
 \end{figure}
@@
 
-\caption[Effect sizes across temperature range by experimental age]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-26}
+\caption[Effect sizes across temperature range by experimental age]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-26}
 \end{figure}
@@
 \toprule
-\textbf{Species} & \textbf{Topt} & \textbf{Reference}\\
+\textbf{Species} & \textbf{Topt (°C)} & \textbf{Reference}\\
 \midrule
-\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6 C} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
-Leptopilina boulardi & 25 C & De lava et al 2016\\
-\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 +/- 0.64} & \cellcolor{gray!6}{Dang et al 2019}\\
-Gryllus firmus & < 30 C & Singh et al 2020\\
-\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30 C} & \cellcolor{gray!6}{Seebacher et al 2014}\\
+\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6°C} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
+Leptopilina boulardi & 25°C & De lava et al 2016\\
+\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 ± 0.64 °C} & \cellcolor{gray!6}{Dang et al 2019}\\
+Gryllus firmus & ~30°C & Singh et al 2020\\
+\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30°C} & \cellcolor{gray!6}{Seebacher et al 2014}\\
 \addlinespace
-Limnodynastes tasmaniensis & 30-33 C & Whitehead et al 1989\\
-\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33 C} & \cellcolor{gray!6}{Kern et al 2014}\\
-Natrix natrix & </= 30 C & Issac 1997\\
-\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26 C} & \cellcolor{gray!6}{Alagawany et al 2017}\\
-Plestiodon chinensis & 33 C & Baojun et al 2010\\
+Limnodynastes tasmaniensis & 30-33°C & Whitehead et al 1989\\
+\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33°C} & \cellcolor{gray!6}{Kern et al 2014}\\
+Natrix natrix & ~30°C & Issac 1997\\
+\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26°C} & \cellcolor{gray!6}{Alagawany et al 2017}\\
+Plestiodon chinensis & 33°C & Baojun et al 2010\\
 \addlinespace
-\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35 C} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
-Escherichia coli & 37 C, 35-36 C & Doyle and Schoeni 1984, Bronikowski et al 2001\\
-\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36 C} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
-Caiman latirostris & 32-33 C & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
-\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33 C} & \cellcolor{gray!6}{Dang et al 2019}\\
+\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35°C} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
+Escherichia coli & 37°C, 35-36°C & Doyle and Schoeni 1984, Bronikowski et al 2001\\
+\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36°C} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
+Caiman latirostris & 32-33°C & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
+\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33°C} & \cellcolor{gray!6}{Dang et al 2019}\\
 \addlinespace
-Esteya vermicola & 26 C & Wang et al 2015\\
+Esteya vermicola & 26°C & Wang et al 2015\\
 \bottomrule

### 2021-04-29 writing /thesis_draft/index.Rmd
all formatting is in, just need to read for grammar

--- 2021-04-29 writing /thesis_draft/index.Rmd
@@
 ---
-title: 'Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis'
+title: 'Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta analysis'
 author: 'Margaret Anne Slein'
@@
 Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.
-El Nino Southern Oscillation).
+El Niño Southern Oscillation).
 As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems [khelifa_usefulness_2019; @stewart_mesocosm_2013].

### 2021-04-29 writing /thesis_draft/01-Background.Rmd
more clean up and fixes

@@
 There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability [@kroeker_ecological_2020; @massey_measurement_2019].
-To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
+To understand how environmental variability affects organisms, populations, and communities, we conducted a meta-analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
 (i.e.

### 2021-04-29 writing /thesis_draft/02-chap2.Rmd
more clean up and fixes

@@
 ```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
-Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%', fig.align="center", fig.scap="Meta analysis paper selection process"}
+Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%', fig.align="center", fig.scap="Meta-analysis paper selection process"}
 knitr::include_graphics("figures/figure5.png")

### 2021-04-29 writing /thesis_draft/03-chap3.Rmd
more clean up and fixes

@@
              y = value,
-             color = value > 0)) +
+             fill = value > 0)) +
   geom_errorbar(aes(ymin=lb, ymax=ub), width=.2,
                  position=position_dodge(.95), color = "black")+ 
-  geom_point(size = 4, show.legend = FALSE) +
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black") +
   labs(x = NULL, y = "Coefficient estimate") +
-  scale_color_viridis(discrete = TRUE)+
+  scale_fill_viridis(discrete = TRUE)+
   theme_bw()+
@@
 
+\clearpage
+
 ```{r, echo= FALSE, fig.cap="Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by thermal stress.", out.width = '85%', fig.align="center", fig.scap="Effect sizes across temperature range by thermal stress"}
@@
 
-The results of this meta analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level.
+The results of this meta-analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level.
 There were not overall differences in the direction of thermally variable environments positively or negatively impacting performance relative to constant environments.
@@
 This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile.
-However, there have been meta analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018].
+However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018].
 These results align with patterns of juvenile organisms (i.e.
@@
 
-Beyond the data included in our meta analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
+Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
 As Figure 4 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures.
@@
 Many of these studies failed to justify why certain temperatures were chosen as mean temperatures as well as the range of temperatures studied.
-In order to compile information on the thermal stress of each of the organisms included in the studies of this meta analysis, additional resources were needed (see Appendix Table 1).
+In order to compile information on the thermal stress of each of the organisms included in the studies of this meta-analysis, additional resources were needed (see Appendix Table 1).
 Understanding where these organisms' performance falls in relation to temperature on its TPC is crucial for drawing reliable conclusions of how thermal variability affects performance.
@@
 Understanding how organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020].
-This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta analysis.
+This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.
 Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times.

### 2021-04-29 writing /thesis_draft/04-conclusion.Rmd
fully formatted draft :)
more clean up and fixes

@@
 
-While there are limitations to the reach of this meta analysis and its results, it is the first cross-taxa quantitative summary of how environmental variability has been studied in the literature.
+While there are limitations to the reach of this meta-analysis and its results, it is the first cross-taxa quantitative summary of how environmental variability has been studied in the literature.
 Our results suggest that thermal variability influences responses across several studies, positively as lower mean temperatures with higher fluctuation ranges, and less so at high mean temperatures with lower fluctuation ranges.

### 2021-04-29 writing /thesis_draft/05-appendix.Rmd
more clean up and fixes

--- 2021-04-29 writing /thesis_draft/05-appendix.Rmd
@@
 kable(table4, "latex", booktabs = T, 
-      col.names = c("Species", "Topt", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X.
+      col.names = c("Species", "Topt (°C)", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis (missing data on species X.
 arvicola, used in Garcia-Ruiz et al 2011)", caption.short ="Thermal stress data by species") %>%

### 2021-04-29 writing /thesis_draft/_book/thesis.tex
more clean up and fixes

@@
 % To pass between YAML and LaTeX the dollar signs are added by CII
-\title{Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta analysis}
+\title{Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta-analysis}
 \author{Margaret Anne Slein}
@@
 Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
-Here, we present a quantitative meta analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
+Here, we present a quantitative meta-analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
 Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model.
@@
 There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al., 2020; Massey, Holt, Brooks, \& Rollinson, 2019).
-To understand how environmental variability affects organisms, populations, and communities, we conducted a meta analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
+To understand how environmental variability affects organisms, populations, and communities, we conducted a meta-analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
 (i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc)?
@@
 
-\caption[Meta analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
+\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
 Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
@@
 \end{figure}
+\clearpage
 \begin{figure}
@@
 This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile.
-However, there have been meta analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
+However, there have been meta-analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
 These results align with patterns of juvenile organisms (i.e.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Moyano et al., 2017; Portner \& Farrell, 2008).
 
-Beyond the data included in our meta analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
+Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
 As Figure 4 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures.
@@
 Understanding how organisms' life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt, O'Connor, Sunday, \& Gonzalez, 2020).
-This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta analysis.
+This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.
 Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times.
@@
 
-\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X.
+\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis (missing data on species X.
 arvicola, used in Garcia-Ruiz et al 2011)}

### 2021-04-29 writing /thesis_draft/index.Rmd
more clean up and fixes

@@
 ---
-title: 'Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta analysis'
+title: 'Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta-analysis'
 author: 'Margaret Anne Slein'

### 2021-04-30 writing /thesis_draft/01-Background.Rmd
final edits and submitted :)

@@
 Shown is a comparison of how duration of variability compares within the same environmental factor.
-Terms highlighted in each panel are described below in the main text", out.width = '105%', fig.align="center", fig.scap="Metrics to describe and explain environmental variability"}
+Terms highlighted in each panel are described below in the main text", out.width = '105%', fig.align="center", fig.scap="Metrics to describe and explain environmental variability."}
 knitr::include_graphics("figures/figure1.png")
@@
 Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance [@colinet_mechanisms_2018].
-Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment [@husak_how_2017; @amarasekare_framework_2012].
+Performance is a set of whole organism traits (i.e.
+assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment [@husak_how_2017; @amarasekare_framework_2012].
 While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
 However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.
-heat waves, cold fronts etc.) in which the benefits of FTR in mitigated chill injuries may be helpful.
+heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful.
 Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
 For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development [@delmas_mechanistic_2007; @bowden_temperature_2018].
-Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth [@lawson_environmental_2015] depending on the magnitude of variance.
+Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth depending on the magnitude of variance [@lawson_environmental_2015].
 When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics [@romanuk_environmental_2002; @dobramysl_environmental_2013; @bastille-rousseau_climate_2018].
@@
 Brown et al highlighted the importance of both temperature and body size in the metabolic theory of  ecology (MTE) [@brown_toward_2004].
-Though highly debated [@clarke_temperature_2006], which serves as an origin point for understanding how organisms respond to environmental conditions.
+Though highly debated [@clarke_temperature_2006], MTE serves as an origin point for understanding how organisms respond to environmental conditions.
 MTE is described by the following equation:
@@
 ```{r, echo=FALSE, fig.cap="Schematic of thermal performance curve and it’s important components for understanding thermal performance.
-Adapted from (Krenek et al 2012, Tuff et al 2016)", fig.align="center", out.width = '90%',fig.scap="Thermal performance curve components"}
+Adapted from (Krenek et al 2012, Tuff et al 2016).", fig.align="center", out.width = '90%',fig.scap="Thermal performance curve components"}
 knitr::include_graphics("figures/figure2.png")
@@
 However, there are several other constraining factors (i.e.
-energy acquisition, nutrient availability etc.) that derive the characteristically nonlinear pattern of TPCs.
+energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs.
 TPCs strive to represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
@@
 However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.
-15-35 C in 5 C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
+15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
 These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.
 
-Given the characteristically non-linear shape of TPCs, Jensen’s inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
+Given the characteristically nonlinear shape of TPCs, Jensen’s inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
 Jensen’s inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions [@ruel_jensens_1999].
@@
 
- * At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration [@descamps-julien_stable_2005], inflationary effects [@holt_impacts_2003]
+ * At the community level, this usually comes in the form of promoting coexistence in the face of emigration [@descamps-julien_stable_2005], or population success via inflationary effects [@holt_impacts_2003]
+
 It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
@@
 does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.
-periodicity, age, size, thermal history, etc)?
+periodicity, age, size, thermal history, etc.)?
 We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
 
-We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and larger organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).
+We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and smaller organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).

### 2021-04-30 writing /thesis_draft/02-chap2.Rmd
final edits and submitted :)

@@
 
-```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
+```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 19 studies included in the analysis.
 Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%', fig.align="center", fig.scap="Meta-analysis paper selection process"}
@@
 We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1).
-After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*) ) AND SU=(Life Sciences & Biomedicine).
+After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*)) AND SU=(Life Sciences & Biomedicine).
 This search returned a total of 176 results.
 We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1).
-Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 4).
+Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 2.1).
 Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
@@
 * We excluded studies that did not report measurements of error
-* We excluded studies in which the response variables are comparable or extrable (i.e. one study performed Principal component analysis for their responses)
+* We excluded studies in which the response variables are comparable or extractable (i.e. one study performed Principal component analysis for their responses)
 
 We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.
-treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
+treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
 We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response [@brown_toward_2004].
@@
                                                  "12", "32", "30", "4"))
-kable(table1, "latex", booktabs = T, col.names = c("Study", "Year", "Extracted data points (n)"), escape = F, caption = "Studies included in analysis of thermal variability (n=15)") %>%
+kable(table1, "latex", booktabs = T, col.names = c("Study", "Year", "Extracted data points (n)"), escape = F, caption = "Studies included in analysis of thermal variability (n=15)", 
+      caption.short ="Studies included in analysis of thermal variability") %>%
    row_spec(0,bold=TRUE)%>%
@@
 From the studies that met our inclusion criteria, we extracted data from tables and figures.
-Figures were extracted via png files imported into WebPlotDigitizer to obtain the raw data.
+Figures were extracted via .png files imported into WebPlotDigitizer to obtain the raw data.
 The extracted data was then imported into a master spreadsheet of data from all studies for analysis.
@@
 
-If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation [@higgins_cochrane_2011].
-Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
+If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and a subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation [@higgins_cochrane_2011].
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, metafor.
 
@@
                                      responses", 
-                                     "Size estimate of study organism, 0 for extra-small
-                                     (microscopic), 1 for small (macroscopic), 2 for medium 
-                                     (less than a foot), and 3 for large (more than a foot)", 
+                                     "Size estimate of study organism, 0 for extra-small (less than 1mm),
+                                     1 for small (less than 5 cm), 2 for medium (less than 0.3 m), and 3
+                                     for large (greater than 0.3 m)", 
                                      "Life stage at which experiment was conducted, 
-                                     categorized as 0 for larval, 1 for juvenile, and 
-                                     2 for adult",
-                                     "Response variables units classified as rates to traits, 
+                                     categorized as 0 for larval, 1 for juvenile (not sexually mature),
+                                     and 2 for adult",
+                                     "Response variable units classified as rates or traits, 
                                      such that rates included any response variable units with a
-                                     “per day” clarifier were classified as rates, all else as
-                                     traits",
+                                     “per day” clarifier, all else as traits",
                                      "Whether or not the organisms experienced thermally stressful 
@@
 We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
-Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) [@viechtbauer_conducting_2010].
+Metafor builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) [@viechtbauer_conducting_2010].
 In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
@@
 
-Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the escalc function in metafor to calculate effect size and sampling variance:
+Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the `escalc` function in metafor to calculate effect size and sampling variance:
 ```{r, eval= FALSE}
@@
 We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head [@viechtbauer_conducting_2010].
-It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer 2010).
-Each of the subsequent arguments of the escalc function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
-The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
+It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias [@viechtbauer_conducting_2010].
+Each of the subsequent arguments of the `escalc` function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
+The `escalc` function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
 
@@
 
-Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
-We first ran a multivariate random effects models without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
+Both the SMD and sampling variance are then input into the `rma.mv` function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
+We first ran a multivariate random effects model without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
 
@@
 
-The arguments of the rma.mv function include SMD, sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
+The arguments of the `rma.mv` function include SMD (yi), sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
 We specified a nested structure to account for the colinearity of data points from within the same study, experiment, and response metric.
@@
 
-The model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
+The full model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
 If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed an organisms’ Topt and become thermally stressful.
@@
 We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal stress data for one study.
-That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz 2011, Table 1).
-The model we ran without this study, but with all the covariates in the full model as well as thermal stress (exclusion model):
+That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz et al 2011, Table 1).
+The  multivariate random effects model with all the covariates in the full model as well as thermal stress (exclusion model), which excluded Garcia-Ruiz et al 2011 entirely, included:
 
@@
 
-We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly and important factor or if it was simply excluding those influential data points:
+We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly an important factor or if it was simply excluding those influential data points:
 

### 2021-04-30 writing /thesis_draft/03-chap3.Rmd
final edits and submitted :)

@@
 ```{r, echo= FALSE, message= FALSE, fig.cap="Forest plot of effects sizes by study and experiment within a study.
-Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study.
+Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study.
 Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", fig.scap="Forest plot of effect sizes", out.width="95%", fig.align="center"}
@@
 Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 3.1.
-These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
+These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1).
 As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
 
-Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001).
+Our full model, including 7 covariates (all, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001).
 Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p < 0.00001).
@@
 
-```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
-Orange error bars represent 95 percent confidence intervals", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
+```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficient estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
+Error bars represent 95 percent confidence intervals.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
 full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
@@
 
-Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2).The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p < 0.05.
-Surprisingly, the remaining negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
+Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2).
+The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p < 0.05.
+Surprisingly, the additional negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
 In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).
@@
 
-```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature range, and effect size.
-Temperature fluctuation range from 0-20°C, Mean temperature range from 7-33°C.
-Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and effect size"}
+```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and effect size.
+Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C.
+Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure 2).", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and effect size"}
 normalized %>%
@@
 In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance.
-Rapid evolution could further explain differences in responses across studies, as if organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics [@yoshida_rapid_2003].
+Rapid evolution could further explain differences in responses across studies.
+If organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics [@yoshida_rapid_2003].
 It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.
@@
 Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal stress, life history traits, and experimental age.
-Body size has been identified as important in the allometric scaling of MTE and it is interesting in these subsets of studies, body size did not explain the differences in responses across studies.
-This could simply be because we had a small sample size of studies (n=15), though multiple  responses in studies contributed to a large sample of effect sizes (n=139), there were only a subset of relative body sizes included.
+Body size has been identified as important in the allometric scaling of MTE and it is interesting that in these subsets of studies, body size did not explain the differences in responses across studies.
+This could simply be because we had a small sample size of studies (n=15), though multiple responses within each study contributed to a large sample of effect sizes (n=140), there were only a subset of relative body sizes included.
 It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms [@clarke_scaling_1999; @clarke_temperature_2006].
-This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns variability in temperature drives.
+This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns that variability in temperature drives.
 
-An important side note is that our full model and thermal stress model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
-This study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding Garcia-Ruiz et al (2011) from our full model.
-We decided to include this study in our analysis because the data were extracted from a table, therefore there were no inaccuracies in obtaining the data.
+An important side note is that our full model and thermal stress model differed in the number of studies, as one study, @garcia-ruiz_effects_2011, did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
+Further, this study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding @garcia-ruiz_effects_2011 from our full model.
+We decided to include this study in our analysis because data were extracted from a table, therefore there were no inaccuracies in obtaining the data.
 It provides an important counter and discussion for how and why responses from different studies may be so different from others.
 
-While body size and thermal stress emerged as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship thermal variability and MTE.
+While body size and thermal stress emerging as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship between thermal variability and MTE.
 If we were able to obtain additional information about the CTmax, CTmin, thermal breadth, and tolerance range for each of the species included in the analysis, we may also be able to better understand the trends and the significance of variability patterns in dictating responses.
@@
 We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018).
-This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile.
+This may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile stage instead.
 However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018].
@@
 Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
-As Figure 4 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures.
-It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated in fact.
+As Figure 1.3 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures.
+It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated.
 Are these diurnal temperature fluctuations accurate for what is expected under natural conditions?
@@
 Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures [@amarasekare_framework_2012] as well as different ranges of temperature, both of which are different concepts.
-For instance, Amarasekare et al 2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits.
+For instance, @amarasekare_framework_2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits.
 Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
@@
 
-Many of the studies made little to no mention of generation time with respect to duration of their experiments or study organisms used.
-Understanding how organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020].
+Several studies made little to no mention of generation time with respect to duration of their experiments or study organisms used.
+How organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020].
 This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.

### 2021-04-30 writing /thesis_draft/04-conclusion.Rmd
final edits and submitted :)

@@
 It has been noted for years the importance of including thermal variability in ecological research, yet there are still an overwhelming number of studies that employ the same patterns of variability with the same lack of reference to thermal performance.
-Going forward, experiments that explicitly account for and justify variation patterns with respect to generation time, organization level, duration, and natural relevance are needed.
+Going forward, experiments that explicitly account for and justify variation patterns with respect to generation time, organization level, duration, and relevance to natural conditions are needed.
 In order to draw conclusions that are relevant to natural conditions, patterns of variability across colors of noise (red, white, etc.) and across lengths of time must be investigated.
@@
 There is more to be studied on how additional experimental designs, i.e.
-acclimation or heat-wave simulations, and biological performance respond to environmental variability, however, this is an important first step in quantitatively summarizing progress so far.
+acclimation or heat-wave simulations, as well as how biological performance responds to environmental variability.
+However, this is an important first step in quantitatively summarizing progress so far.

### 2021-04-30 writing /thesis_draft/_book/thesis.tex
final edits and submitted :)

@@
 It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bambach, 1993; Bouchard, 2014).
-Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Soo, Hemp, Parks, Fischer, \& Hugenholtz, 2017, @blankenship\_how\_2017).
+Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Blankenship, 2017; Soo, Hemp, Parks, Fischer, \& Hugenholtz, 2017).
 Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll, 2008).
-In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnik et al 2007).
+In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnick, Cline, Smith, \& Sage, 2007).
 Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Burggren, 2018; Reed, Waples, Schindler, Hard, \& Kinnison, 2010).
@@
 Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Niño Southern Oscillation).
-As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (khelifa\_usefulness\_2019; Stewart et al., 2013).
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019; Stewart et al., 2013).
 How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
@@
 
-\caption[Metrics to describe and explain environmental variability]{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
+\caption[Metrics to describe and explain environmental variability.]{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
 Shown is a comparison of how duration of variability compares within the same environmental factor.
@@
 Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018).
-Performance is a set of whole organism traits (i.e.assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
+Performance is a set of whole organism traits (i.e.~assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
 While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
-However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.~heat waves, cold fronts etc.) in which the benefits of FTR in mitigated chill injuries may be helpful.
+However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful.
 Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
 For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Bowden \& Paitz, 2018; Delmas, Prevot-Julliard, Pieau, \& Girondot, 2007).
-Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth (Lawson, Vindenes, Bailey, \& Pol, 2015) depending on the magnitude of variance.
+Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth depending on the magnitude of variance (Lawson, Vindenes, Bailey, \& Pol, 2015).
 When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Bastille-Rousseau et al., 2018; Dobramysl \& Tauber, 2013; Romanuk \& Kolasa, 2002).
@@
 Brown et al highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE) (Brown, Gillooly, Allen, Savage, \& West, 2004).
-Though highly debated (Clarke, 2006), which serves as an origin point for understanding how organisms respond to environmental conditions.
+Though highly debated (Clarke, 2006), MTE serves as an origin point for understanding how organisms respond to environmental conditions.
 MTE is described by the following equation:
@@
 \caption[Thermal performance curve components]{Schematic of thermal performance curve and it’s important components for understanding thermal performance.
-Adapted from (Krenek et al 2012, Tuff et al 2016)}\label{fig:unnamed-chunk-4}
+Adapted from (Krenek et al 2012, Tuff et al 2016).}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one.
-However, there are several other constraining factors (i.e.~energy acquisition, nutrient availability etc.) that derive the characteristically nonlinear pattern of TPCs.
+However, there are several other constraining factors (i.e.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs.
 TPCs strive to represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
@@
 TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018).
-However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.~15-35 C in 5 C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
+However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
 These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.
 
-Given the characteristically non-linear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
+Given the characteristically nonlinear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
 Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel \& Ayres, 1999).
@@
 Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time.
-It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al., 2018; Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019).
+It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al., 2018; Khelifa et al., 2019).
 
@@
 \item
-  At the community level, this usually comes in the form of promoting coexistence or population success in the face of emigration (Descamps-Julien \& Gonzalez, 2005), inflationary effects (Holt et al., 2003)
+  At the community level, this usually comes in the form of promoting coexistence in the face of emigration (Descamps-Julien \& Gonzalez, 2005), or population success via inflationary effects (Holt et al., 2003)
+\end{itemize}
 It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
 Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).
-\end{itemize}
+
 There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al., 2020; Massey, Holt, Brooks, \& Rollinson, 2019).
 To understand how environmental variability affects organisms, populations, and communities, we conducted a meta-analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
-(i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc)?
+(i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc.)?
 We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
 
-We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and larger organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).
+We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and smaller organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).
 
@@
 
-\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science to the 19 studies included in the analysis.
+\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 19 studies included in the analysis.
 Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
@@
 We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1).
-After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*) ) AND SU=(Life Sciences \& Biomedicine).
+After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*)) AND SU=(Life Sciences \& Biomedicine).
 This search returned a total of 176 results.
 We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1).
-Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 4).
+Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 2.1).
 Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
@@
 \item
-  We excluded studies in which the response variables are comparable or extrable (i.e.~one study performed Principal component analysis for their responses)
+  We excluded studies in which the response variables are comparable or extractable (i.e.~one study performed Principal component analysis for their responses)
 \end{itemize}
-We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1 C of the two treatments.
+We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
 We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004).
@@
 
-\caption{\label{tab:unnamed-chunk-8}Studies included in analysis of thermal variability (n=15)}
+\caption[Studies included in analysis of thermal variability]{\label{tab:unnamed-chunk-8}Studies included in analysis of thermal variability (n=15)}
 \centering
@@
 
-If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Higgins et al., 2011).
-Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via data analysis software metafor.
+If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and a subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Higgins et al., 2011).
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, metafor.
 
@@
 \cellcolor{gray!6}{                                     responses}\\
-size & Size estimate of study organism, 0 for extra-small
-                                     (microscopic), 1 for small (macroscopic), 2 for medium 
-                                     (less than a foot), and 3 for large (more than a foot)\\
+size & Size estimate of study organism, 0 for extra-small (less than 1mm),
+                                     1 for small (less than 5 cm), 2 for medium (less than 0.3 m), and 3
+                                     for large (greater than 0.3 m)\\
 exp\_age & Life stage at which experiment was conducted, 
-                                     categorized as 0 for larval, 1 for juvenile, and 
-\cellcolor{gray!6}{                                     2 for adult}\\
+                                     categorized as 0 for larval, 1 for juvenile (not sexually mature),
+\cellcolor{gray!6}{                                     and 2 for adult}\\
 \addlinespace
-resp\_type & Response variables units classified as rates to traits, 
+resp\_type & Response variable units classified as rates or traits, 
                                      such that rates included any response variable units with a
-                                     “per day” clarifier were classified as rates, all else as
-                                     traits\\
+                                     “per day” clarifier, all else as traits\\
 stressful & Whether or not the organisms experienced thermally stressful 
@@
 We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
-Metafor builds on the functionality of older meta-analysis packages (meta, etc) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer, 2010).
+Metafor builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer, 2010).
 In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
@@
 
-Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the escalc function in metafor to calculate effect size and sampling variance:
+Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the \texttt{escalc} function in metafor to calculate effect size and sampling variance:
 \begin{Shaded}
@@
 We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head (Viechtbauer, 2010).
-It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer 2010).
-Each of the subsequent arguments of the escalc function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
-The escalc function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
+It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer, 2010).
+Each of the subsequent arguments of the \texttt{escalc} function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
+The \texttt{escalc} function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
 \begin{figure}
@@
 
-Both the SMD and sampling variance are then input into the rma.mv function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
-We first ran a multivariate random effects models without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
+Both the SMD and sampling variance are then input into the \texttt{rma.mv} function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
+We first ran a multivariate random effects model without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
 \begin{Shaded}
@@
 \end{Shaded}
-The arguments of the rma.mv function include SMD, sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
+The arguments of the \texttt{rma.mv} function include SMD (yi), sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
 We specified a nested structure to account for the colinearity of data points from within the same study, experiment, and response metric.
@@
 \end{Shaded}
-The model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
+The full model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
 If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed an organisms' Topt and become thermally stressful.
@@
 We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal stress data for one study.
-That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz 2011, Table 1).
-The model we ran without this study, but with all the covariates in the full model as well as thermal stress (exclusion model):
+That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz et al 2011, Table 1).
+The multivariate random effects model with all the covariates in the full model as well as thermal stress (exclusion model), which excluded Garcia-Ruiz et al 2011 entirely, included:
 \begin{Shaded}
@@
 \end{Shaded}
-We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly and important factor or if it was simply excluding those influential data points:
+We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly an important factor or if it was simply excluding those influential data points:
 \begin{Shaded}
@@
 \caption[Forest plot of effect sizes]{Forest plot of effects sizes by study and experiment within a study.
-Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study.
+Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study.
 Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-18}
@@
 Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1.
-These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis.
+These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1).
 As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
 
-Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p \textless{} 0.0001).
+Our full model, including 7 covariates (all, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p \textless{} 0.0001).
 Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p \textless{} 0.00001).
@@
 
-\caption[Coefficient estimates for full model]{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
-Orange error bars represent 95 percent confidence intervals}\label{fig:unnamed-chunk-19}
+\caption[Coefficient estimates for full model]{Coefficient estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
+Error bars represent 95 percent confidence intervals.}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 
-Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p \textless{} 0.05.
-Surprisingly, the remaining negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
+Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).
+The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p \textless{} 0.05.
+Surprisingly, the additional negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
 In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).
@@
 
-\caption[Scatterplot of relationship between range, mean, and effect size]{Scatterplot describing the relationship between mean temperature, temperature range, and effect size.
-Temperature fluctuation range from 0-20°C, Mean temperature range from 7-33°C.
-Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)}\label{fig:unnamed-chunk-21}
+\caption[Scatterplot of relationship between range, mean, and effect size]{Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and effect size.
+Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C.
+Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-21}
 \end{figure}
@@
 In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance.
-Rapid evolution could further explain differences in responses across studies, as if organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics (Yoshida, Jones, Ellner, Fussmann, \& Hairston, 2003).
+Rapid evolution could further explain differences in responses across studies.
+If organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics (Yoshida, Jones, Ellner, Fussmann, \& Hairston, 2003).
 It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.
@@
 Equally as interesting as the significant model coefficients were the insignificant model coefficients: body size, thermal stress, life history traits, and experimental age.
-Body size has been identified as important in the allometric scaling of MTE and it is interesting in these subsets of studies, body size did not explain the differences in responses across studies.
-This could simply be because we had a small sample size of studies (n=15), though multiple responses in studies contributed to a large sample of effect sizes (n=139), there were only a subset of relative body sizes included.
+Body size has been identified as important in the allometric scaling of MTE and it is interesting that in these subsets of studies, body size did not explain the differences in responses across studies.
+This could simply be because we had a small sample size of studies (n=15), though multiple responses within each study contributed to a large sample of effect sizes (n=140), there were only a subset of relative body sizes included.
 It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms (Clarke, 2006; Clarke \& Johnston, 1999).
-This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns variability in temperature drives.
-
-An important side note is that our full model and thermal stress model differed in the number of studies, as one study, Garcia-Ruiz et al (2011) did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
-This study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding Garcia-Ruiz et al (2011) from our full model.
-We decided to include this study in our analysis because the data were extracted from a table, therefore there were no inaccuracies in obtaining the data.
+This theory may also explain why thermal stress was not a significant predictor in our model, if MTE does not aptly describe the patterns that variability in temperature drives.
+
+An important side note is that our full model and thermal stress model differed in the number of studies, as one study, García-Ruiz, Marco, \& Pérez-Moreno (2011), did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
+Further, this study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding García-Ruiz et al.
+(2011) from our full model.
+We decided to include this study in our analysis because data were extracted from a table, therefore there were no inaccuracies in obtaining the data.
 It provides an important counter and discussion for how and why responses from different studies may be so different from others.
 
-While body size and thermal stress emerged as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship thermal variability and MTE.
+While body size and thermal stress emerging as not statistically significant coefficients in our model may contradict MTE and its allometric scaling component, an additional analysis, with explicit masses for each of these organisms, would better explore the relationship between thermal variability and MTE.
 If we were able to obtain additional information about the CTmax, CTmin, thermal breadth, and tolerance range for each of the species included in the analysis, we may also be able to better understand the trends and the significance of variability patterns in dictating responses.
@@
 We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018).
-This result may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile.
+This may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile stage instead.
 However, there have been meta-analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
@@
 Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures (Amarasekare \& Savage, 2012) as well as different ranges of temperature, both of which are different concepts.
-For instance, Amarasekare et al 2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits.
+For instance, Amarasekare \& Savage (2012) addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits.
 Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
@@
 
-Many of the studies made little to no mention of generation time with respect to duration of their experiments or study organisms used.
-Understanding how organisms' life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt, O'Connor, Sunday, \& Gonzalez, 2020).
+Several studies made little to no mention of generation time with respect to duration of their experiments or study organisms used.
+How organisms' life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt, O'Connor, Sunday, \& Gonzalez, 2020).
 This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.
@@
 It has been noted for years the importance of including thermal variability in ecological research, yet there are still an overwhelming number of studies that employ the same patterns of variability with the same lack of reference to thermal performance.
-Going forward, experiments that explicitly account for and justify variation patterns with respect to generation time, organization level, duration, and natural relevance are needed.
+Going forward, experiments that explicitly account for and justify variation patterns with respect to generation time, organization level, duration, and relevance to natural conditions are needed.
 In order to draw conclusions that are relevant to natural conditions, patterns of variability across colors of noise (red, white, etc.) and across lengths of time must be investigated.
 A larger sample size of studies to draw more robust conclusions with respect to support for or against allometric scaling and MTE with respect to thermal variability would be useful.
-There is more to be studied on how additional experimental designs, i.e.~acclimation or heat-wave simulations, and biological performance respond to environmental variability, however, this is an important first step in quantitatively summarizing progress so far.
+There is more to be studied on how additional experimental designs, i.e.~acclimation or heat-wave simulations, as well as how biological performance responds to environmental variability.
+However, this is an important first step in quantitatively summarizing progress so far.
 
@@
 
+\leavevmode\hypertarget{ref-guralnick_evolutionary_2007}{}%
+Guralnick, L.
+J., Cline, A., Smith, M., \& Sage, R.
+F.
+(2007).
+Evolutionary physiology: The extent of c4 and CAM photosynthesis in the genera anacampseros and grahamia of the portulacaceae.
+\emph{Journal of Experimental Botany}, \emph{59}(7), 1735--1742.
+\url{http://doi.org/10.1093/jxb/ern081}
+
 \leavevmode\hypertarget{ref-helmuth_organismal_2010}{}%

### 2021-04-30 writing /thesis_draft/index.Rmd
final edits and submitted :)

@@
 It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia [@bouchard_ecosystem_2014; @bambach_seafood_1993].
-Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever [@soo_origins_2017, @blankenship_how_2017].
+Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever [@soo_origins_2017; @blankenship_how_2017].
 Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time [@knoll_chapter_2008].
-In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnik et al 2007).
+In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions [@guralnick_evolutionary_2007].
 Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant [@reed_phenotypic_2010; @burggren_developmental_2018].
@@
 El Niño Southern Oscillation).
-As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems [khelifa_usefulness_2019; @stewart_mesocosm_2013].
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems [@khelifa_usefulness_2019; @stewart_mesocosm_2013].
 How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.

### 2021-05-11 writing /thesis_draft/01-Background.Rmd
putting in everyone's edits from orals

@@
 It provides a departure point for what we expect the average value to be for all time.
-In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, i.e.
+In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, e.g.
 averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations.
@@
 ### Amplitude/Range 
-We can expand additional information adjacent to the mean (i.e.
-range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature [@vasseur_increased_2014].
+We can expand additional information adjacent to the mean (e.g.
+range, duration, predictability) for a finer scale picture of environmental conditions.
+Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014).
 We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
@@
 The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature [@amarasekare_framework_2012; @amarasekare_intrinsic_2013], as well as additional life history traits (size, speed, etc.).
-An organism’s ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible [@amarasekare_framework_2012].
-If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.
+An organism’s ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature relative to the temperature at which survival is no longer possible and may be significantly impacted by increases in temperature range [@amarasekare_framework_2012].
+If an organism has an optimal temperature close to their upper limit of survival, a larger range of variation may push the organism to spend more time beyond the optimal temperature.
+This would negatively impact their ability to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.
 
@@
 Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance [@colinet_mechanisms_2018].
-Performance is a set of whole organism traits (i.e.
+Performance is a set of whole organism traits (e.g.
 assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment [@husak_how_2017; @amarasekare_framework_2012].
-While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
-However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.
+Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.
 heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful.
@@
 FTR can also describe the predictability of the fluctuation period.
-Variability may not have evenly spaced intervals (i.e.
+Variability may not have evenly spaced intervals (e.g.
 heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
@@
 Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space [@horne_spatial_1995; @farhang-sardroodi_environmental_2019].
-Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e.
+Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, e.g.
 Diel vertical migration (DVM).
-In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability [@lampert_adaptive_1989; @brierley_diel_2014].
-However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.
+In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability [@lampert_adaptive_1989; @brierley_diel_2014].
+However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, e.g.
 an organism moving from sun to shade during the day to regulate body temperature and metabolic processes [@huey_behavioral_1974; @kearney_potential_2009].
@@
 
-## Thermal variability
+## Temperature variability
 
@@
 
-Changes in temperature can be grouped into a couple different orientations, the first being the difference between the terms “fluctuating” and “variation” with respect to temperature.
-Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as “a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis” (2018).
-Most generally, temperature variability describes changes in environmental temperature through time [@holmes_robust_2016; @colinet_mechanisms_2018].
+Changes in temperature can be grouped into a couple different definitions, the first being the difference between the terms “fluctuating” and “variation” with respect to temperature.
+Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al (2018) explicitly defined fluctuating temperature as “a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis.” Most generally, temperature variability describes changes in environmental temperature through time [@holmes_robust_2016; @colinet_mechanisms_2018].
 
-### Why is it important to ecological processes?
+### Why is temperature variability important to ecological processes?
 
@@
 Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate.
-We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates should be the same.
+We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and converse for smaller organisms.
 However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish [@clarke_scaling_1999; @clarke_temperature_2006].
-Others believe that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups [@gillooly_response_2006].
+Others highlight that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups [@gillooly_response_2006].
 Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.
 
-MTE supports temperature impacting metabolic rate, which can have downstream effects on fitness characteristics.
+MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics.
 This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
 
-```{r, echo=FALSE, fig.cap="Schematic of thermal performance curve and it’s important components for understanding thermal performance.
+```{r, echo=FALSE, fig.cap="Schematic of thermal performance curve and its important components for understanding thermal performance.
+CTmin denotes Critical Thermal Minimum, CTmax denotes Critical Thermal Maximum, and Topt denotes Thermal Optimum.
 Adapted from (Krenek et al 2012, Tuff et al 2016).", fig.align="center", out.width = '90%',fig.scap="Thermal performance curve components"}
@@
 The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms’ performance is negatively impacted [@brown_toward_2004; @dowd_thermal_2015].
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive [@brown_toward_2004; @dowd_thermal_2015].
 Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
 
-TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (i.e.
+TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (e.g.
 metabolism) [@amarasekare_framework_2012].
-If only temperature’s effect on metabolism were considered, the shape of the relationship may be an exponential one.
-However, there are several other constraining factors (i.e.
+If only temperature’s effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after whiich it would no longer increase and the organism would fail to survive.
+However, there are several other constraining factors (e.g.
 energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs.
-TPCs strive to represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
+TPCs represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
 
@@
 TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment [@bernhardt_nonlinear_2018].
-However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.
+However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.
 15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
@@
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns [@amarasekare_framework_2012; @paaijmans_temperature_2013].
+For instance, Amarasekare et al (2012) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns [@amarasekare_framework_2012; @paaijmans_temperature_2013].
 Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature [@amarasekare_framework_2012].
@@
 Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
-Variation in temperature is predicted to affect performance more significantly than increasing temperatures [@vasseur_increased_2014]; thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
-Khelifa et al demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance [@khelifa_usefulness_2019].
+Variation in temperature is predicted to affect performance more significantly than increasing temperatures [@vasseur_increased_2014]; thus, if metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
+Will increased variation have differing effects on different levels of biological organization?
+Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
 
 However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 1.3).
-Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable [@khelifa_usefulness_2019].
+Khelifa et al (2019) demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable.
 Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
@@
 It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
-Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses [@petchey_environmental_2000].
+Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas studies at higher levels of biological organization have explored how colored noise elicits these characteristic responses [@petchey_environmental_2000].
 
-There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability [@kroeker_ecological_2020; @massey_measurement_2019].
+There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at the significance of variability on ecosystem dynamics [@kroeker_ecological_2020; @massey_measurement_2019].
 To understand how environmental variability affects organisms, populations, and communities, we conducted a meta-analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
 (i.e.
-does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.
+does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (e.g.
 periodicity, age, size, thermal history, etc.)?
@@
 
-We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and smaller organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).
+We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response, 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will exhibit increased performance and smaller organisms will exhibit decreased performance with increased thermal variability, 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics, 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted.

### 2021-05-11 writing /thesis_draft/02-chap2.Rmd
putting in everyone's edits from orals

@@
 ```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 19 studies included in the analysis.
-Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%', fig.align="center", fig.scap="Meta-analysis paper selection process"}
+Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%', fig.align="center", fig.scap="Meta-analysis paper selection process"}
 knitr::include_graphics("figures/figure5.png")
@@
 
-* We excluded studies that did not feature both a constant and fluctuating treatment (i.e.
+* We excluded studies that did not feature both a constant and fluctuating treatment (e.g.
 studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
-* We excluded studies that were not completed in a laboratory (i.e. modeling predictions or results from models, mesocosm experiments, in situ experiments)
-* We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (i.e.
+* We excluded studies that were not completed in a laboratory (e.g. modeling predictions or results from models, mesocosm experiments, *in situ* experiments)
+* We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.
 organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
@@
 * We excluded studies that did not report measurements of error
-* We excluded studies in which the response variables are comparable or extractable (i.e. one study performed Principal component analysis for their responses)
+* We excluded studies in which the response variables were not comparable or extractable (e.g. one study performed Principal component analysis for their responses)
 
-We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.
+We excluded studies that did not have the same mean for constant and fluctuating treatments (e.g.
 treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
@@
 If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and a subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation [@higgins_cochrane_2011].
-Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, metafor.
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, *metafor*.
 
@@
                                      "Level of biological organization, categorized
-                                     as either 0 for individual or 1 for population level
-                                     responses", 
+                                     as either 0 for individual, 1 for population level
+                                     responses, 2 for community level responses", 
                                      "Size estimate of study organism, 0 for extra-small (less than 1mm),
@@
                                      such that rates included any response variable units with a
-                                     “per day” clarifier, all else as traits",
+                                     “per day” clarifier (e.g. growth rate), all else as traits (e.g. body
+                                     width)",
                                      "Whether or not the organisms experienced thermally stressful 
@@
 
-We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
-Metafor builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) [@viechtbauer_conducting_2010].
-In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
+We used *metafor* (version 2.4.0) in R to analyze the extracted, aggregated data.
+*Metafor* builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) [@viechtbauer_conducting_2010].
+In order to use *metafor*, the package requires the data to be in a wider format, in which data are input into the following columns:
 
@@
 
-First, it calculates the pooled standard deviation (Pooled SD) from the sample sizes of both studies.
+First, it calculated the pooled standard deviation (Pooled SD) from the sample sizes of both studies.
 Our data had studies with a wide range of sample sizes, so we wanted to include separate sample sizes between the constant and fluctuating treatments.
-The Pooled SD, Cohen’s d, which is one of the most common ways to measure effect size, is used to normalize the mean differences between constant and fluctuating treatments.
-Then, using a correction factor J, which again normalizes the data further, we obtain the final SMD.
-The SMD is then ultimately used to calculate the sampling variance, which creates a unitless metric for variance.
+The Pooled SD, Cohen’s d, which is one of the most common ways to measure effect size, was used to normalize the mean differences between constant and fluctuating treatments.
+Then, using a correction factor J, which again normalized the data further, we obtained the final SMD.
+The SMD was then ultimately used to calculate the sampling variance, which created a unitless metric for variance.
 
-Both the SMD and sampling variance are then input into the `rma.mv` function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
-We first ran a multivariate random effects model without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
+Both the SMD and sampling variance were then input into the `rma.mv` function in *metafor* to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
+We first ran a multivariate random effects model without modifiers (simple model) to establish baseline heterogeneity between studies, and then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
 
@@
 
-We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly an important factor or if it was simply excluding those influential data points:
+We also ran a separate model without Garcia-Ruiz et al 2011 and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly an important factor or if it was simply excluding those influential data points:
 

### 2021-05-11 writing /thesis_draft/03-chap3.Rmd
putting in everyone's edits from orals

@@
 ```{r, echo= FALSE, message= FALSE, fig.cap="Forest plot of effects sizes by study and experiment within a study.
-Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study.
+Data indicate that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study.
 Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", fig.scap="Forest plot of effect sizes", out.width="95%", fig.align="center"}
@@
 
-Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 3.1.
-These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1).
-As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
+Overall, even our simple model demonstrated that the intercept was not significant but there was significant heterogeneity across studies (n=140, z= 0.79, p < 0.0001), as demonstrated by Figure 3.1.
+These results allowed us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1).
+This then allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
 
-Our full model, including 7 covariates (all, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001).
-Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p < 0.00001).
-Organization level (n=140) was also important in the model, as a statistically significant covariate (df= 7, p < 0.05).
+Our full model, including 7 covariates (all covariates in Table 2.2, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001).
+Further, fluctuation range  and the interaction term between fluctuation range and mean temperature  were statistically significant in the model (n=140, df = 7, p < 0.00001).
+Organization level was also important in the model, as a statistically significant covariate (n=140, df= 7, p < 0.05).
 
-```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficient estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
-Error bars represent 95 percent confidence intervals.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
+```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficient estimates for each covariate in the full model, where * signify statistically significant model predictors.
+Error bars represent 95 percent confidence intervals.
+Yellow indicates a positive coefficient estimate, purple indicates a negative coefficient estimate.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
 full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
@@
 
-Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2).
-The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p < 0.05.
-Surprisingly, the additional negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
-In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).
+Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant (n=140, df=7, p < 0.001 level (Figure 3.2).
+The most negative estimate, organization level (n=140), was also significantly influential in our model (n=140, df=7, p < 0.05).The additional negative estimates of our model coefficients, experimental age and size, were not statistically significant.
+In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant (n=132, df=1, p = 0.4855) in
+explaining the effect size (Table 3.1).
 
@@
 The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
-At lower mean temperatures, higher fluctuation ranges have a more positive effect on organism responses and performance, though that trend starkly ends at about $24^{\circ}$C.
-Beyond $24^{\circ}$C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
+At lower mean temperatures, higher fluctuation ranges generally have a more positive effect on organism responses and performance, though that trend starkly ends at about $24^{\circ}$C.
+However, at higher mean temperatures, lower fluctuation ranges appears to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
 
-```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and effect size.
+```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD.
 Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C.
-Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure 2).", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and effect size"}
+SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and SMD"}
 normalized %>%
 ggplot(aes(x=mean_temp_constant, y = yi, fill=flux_range))+
-  geom_jitter( alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
+  geom_jitter(alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
   scale_fill_viridis('Temperature range (°C)', option = "C")+
@@
   xlab("Mean Temperature (°C)")+
-  ylab("Effect Size")+
+  ylab("SMD")+
   geom_line(y=0, linetype = "dashed")+
@@
 
-As noted in the methods, when we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p < 0.001) as did mean temperature (df=8, p < 0.001).
-Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p < 0.1).
+When we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p < 0.001) as did mean temperature (df=8, p < 0.001).
+Organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p < 0.1).
 
@@
 These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e.
-higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures.
+higher fluctuation ranges at lower mean temperatures generally have higher effect sizes compared to higher and lower effects sizes of lower fluctuation ranges at higher mean temperatures.
 The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change [@vasseur_increased_2014; @sunday_thermal_2012].
@@
 Acclimation may also explain differences in response across studies as the scale on which variation occurs influences how organisms may respond.
-If organisms’ non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (i.e.
+If organisms’ non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (e.g.
 inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies [@kremer_gradual_2018].
@@
 If organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics [@yoshida_rapid_2003].
-It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.
+It is important to note that rapid evolution will only be relevant to population levels responses and not individual level responses.
 
@@
 
-An important side note is that our full model and thermal stress model differed in the number of studies, as one study, @garcia-ruiz_effects_2011, did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
+An important side note is that our full model and thermal stress model differed in the number of studies, as one study, @garcia-ruiz_effects_2011, did not have any supplemental information on thermal performance curve points for the taxa used in that study (*Xylotrechus arvicola*).
 Further, this study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding @garcia-ruiz_effects_2011 from our full model.
@@
 We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018).
-This may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile stage instead.
+This may again be the result of studies not focused on the larval stage of development, but instead on the juvenile stage.
 However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018].
-These results align with patterns of juvenile organisms (i.e.
+These results align with patterns of juvenile organisms (e.g.
 fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance [@portner_ecology_2008; @moyano_effects_2017].
@@
 Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
-As Figure 1.3 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures.
-It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated.
+Diurnal cycles are more correlated than reddened cycles, like seasonal temperatures (Figure 1.3).
+It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, longer-term temperature patterns are more stable than diurnal cycles, and less autocorrelated.
 Are these diurnal temperature fluctuations accurate for what is expected under natural conditions?
@@
 Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
-Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes [@colinet_mechanisms_2018; @colinet_insects_2015].
+Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly ecologically irrelevant thermal regimes [@colinet_mechanisms_2018; @colinet_insects_2015].
 Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.
@@
 How organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020].
-This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.
+This is also pertinent to the scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.
 Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times.

### 2021-05-11 writing /thesis_draft/05-appendix.Rmd
putting in everyone's edits from orals

--- 2021-04-29 writing /thesis_draft/05-appendix.Rmd
@@
                          "Salmonella", "Caiman latirostris", 
-                         "Trachemys scripta", "Esteya vermicola"), 
-                 Topt = c("24-26.6 C", "25 C", "37.97 +/- 0.64", "< 30 C", "30 C", "30-33 C",
-                          "33 C", " </= 30 C", "26 C", "33 C", "30-35 C", "37 C, 35-36 C",
-                          "35-36 C", "32-33 C", "32-33 C", "26 C"), 
+                         "Trachemys scripta", "Esteya vermicola", "Xylotrechus arvicola"), 
+                 Topt = c("24-26.6°C", "25°C", "37.97 ± 0.64 °C", "30°C", "30°C", "30-33°C",
+                          "33°C", "30°C", "26°C", "33°C", "30-35°C", "37°C, 35-36°C",
+                          "35-36°C", "32-33°C", "32-33°C", "26°C", "N/A"), 
+                 Class = c("Insecta", "Insecta", "Reptilia", "Insecta", "Amphibia", "Amphibia", "Amphibia",
+                           "Reptilia", "Aves", "Reptilia", "Insecta", "Gammaproteobacteria",
+                           "Gammaproteobacteria","Reptilia", "Reptilia", "Not assigned; Fungus", "Insecta"),
                  `Reference` = c("Klepsatel 2013, David 1988", "De lava et al 2016", 
@@
                                  "Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002", 
-                                 "Dang et al 2019", "Wang et al 2015"))
+                                 "Dang et al 2019", "Wang et al 2015", "Garcia-Ruiz et al 2011"))
 kable(table4, "latex", booktabs = T, 
-      col.names = c("Species", "Topt", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X.
-arvicola, used in Garcia-Ruiz et al 2011)", caption.short ="Thermal stress data by species") %>%
+      col.names = c("Species", "Topt (°C)", "Class", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis (missing data on species X.
+arvicola, used in Garcia-Ruiz et al 2011) and scientific classification at the class level", caption.short ="Thermal stress data by species") %>%
    row_spec(0,bold=TRUE) %>%
-  column_spec(3, width = "6cm") %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(4, width = "3cm") %>%
    kable_styling(latex_options = c("striped", "hold_position"))

### 2021-05-11 writing /thesis_draft/_book/thesis.tex
putting in everyone's edits from orals

@@
 \date{May 2021}
-\division{Mathematics and Natural Sciences}
+\division{Mathematical and Natural Sciences}
 \advisor{Sam Fey}
@@
 \Abstract{
-Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns.
-Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
-Here, we present a quantitative meta-analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
-Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model.
+Ecosystems and organisms have experienced variation on many temporal scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns.
+Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the persistence and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
+Here, we present a quantitative meta-analysis of how thermal variability impacts biological responses across multiple taxa, organism sizes, and levels of biological organization.
+Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model of biological responsiveness to temperature variability.
 Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.
@@
   \begin{abstract}
-    Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns.
-Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
-Here, we present a quantitative meta-analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
-Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model.
+    Ecosystems and organisms have experienced variation on many temporal scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns.
+Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the persistence and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
+Here, we present a quantitative meta-analysis of how thermal variability impacts biological responses across multiple taxa, organism sizes, and levels of biological organization.
+Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model of biological responsiveness to temperature variability.
 Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.
@@
 Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll, 2008).
-In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions (Guralnick, Cline, Smith, \& Sage, 2007).
-Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Burggren, 2018; Reed, Waples, Schindler, Hard, \& Kinnison, 2010).
+Vascular plants have also evolved different photosynthetic systems (e.g.~C4, CAM) to manage a variety of environmental conditions (Guralnick, Cline, Smith, \& Sage, 2007).
+Though cyanobacteria and vascular plants evolved new strategies and structures to persist as the environment changed, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Burggren, 2018; Reed, Waples, Schindler, Hard, \& Kinnison, 2010).
 Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.
@@
 In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
-Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Niño Southern Oscillation).
-As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019; Stewart et al., 2013).
+Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.~El Niño Southern Oscillation).
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems (Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019; Stewart et al., 2013).
 How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
-As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).
+As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).
 
@@
 
-There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, i.e.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate.
-However, there is a lack of consensus as to which scales of biological organization, i.e.~individuals, populations, or communities of organisms will be most severely affected by different shifts in environmental conditions (problem statement).
-We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response (aims).
+There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate.
+However, there is a lack of consensus as to which scales of biological organization, e.g.~individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions.
+We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response.
 To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.
@@
 It provides a departure point for what we expect the average value to be for all time.
-In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, i.e.~averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations.
+In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, e.g.~averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations.
 However, in other cases, the mean is less useful compared to the variability.
@@
 
-We can expand additional information adjacent to the mean (i.e.~range, duration, predictability) for a finer scale picture of environmental conditions, as Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (Vasseur et al., 2014).
+We can expand additional information adjacent to the mean (e.g.~range, duration, predictability) for a finer scale picture of environmental conditions.
+Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014).
 We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
@@
 The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare \& Coutinho, 2013; Amarasekare \& Savage, 2012), as well as additional life history traits (size, speed, etc.).
-An organism's ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature and the temperature at which survival is no longer possible (Amarasekare \& Savage, 2012).
-If an organism has an optimal temperature that is very close to that survival threshold, they are less able to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.
+An organism's ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature relative to the temperature at which survival is no longer possible and may be significantly impacted by increases in temperature range (Amarasekare \& Savage, 2012).
+If an organism has an optimal temperature close to their upper limit of survival, a larger range of variation may push the organism to spend more time beyond the optimal temperature.
+This would negatively impact their ability to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.
 
@@
 Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018).
-Performance is a set of whole organism traits (i.e.~assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
-While studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
-However, there are instances in nature where organisms do experience short periods of extreme temperature changes (i.e.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful.
+Performance is a set of whole organism traits (e.g.~assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
+Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful.
 Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
@@
 FTR can also describe the predictability of the fluctuation period.
-Variability may not have evenly spaced intervals (i.e.~heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
+Variability may not have evenly spaced intervals (e.g.~heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
 Predictability can also attempt to describe stochastic vs.~autocorrelated variation, however, that will be discussed in a different context below (Figure 1.3).
@@
 Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Farhang-Sardroodi, Darooneh, Kohandel, \& Komarova, 2019; Horne \& Schneider, 1995).
-Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, i.e.~Diel vertical migration (DVM).
-In DVM, organisms migrate to the surface and bottom of lakes based mainly on cues of light availability (Brierley, 2014; Lampert, 1989).
-However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, i.e.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).
+Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, e.g.~Diel vertical migration (DVM).
+In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability (Brierley, 2014; Lampert, 1989).
+However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, e.g.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).
 
@@
 
-\hypertarget{thermal-variability}{%
-\section{Thermal variability}\label{thermal-variability}}
+\hypertarget{temperature-variability}{%
+\section{Temperature variability}\label{temperature-variability}}
 
@@
 
-Changes in temperature can be grouped into a couple different orientations, the first being the difference between the terms ``fluctuating'' and ``variation'' with respect to temperature.
-Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al explicitly defined fluctuating temperature as ``a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis'' (2018).
-Most generally, temperature variability describes changes in environmental temperature through time (Colinet et al., 2018; Holmes, Woollings, Hawkins, \& Vries, 2016).
-
-\hypertarget{why-is-it-important-to-ecological-processes}{%
-\subsection{Why is it important to ecological processes?}\label{why-is-it-important-to-ecological-processes}}
+Changes in temperature can be grouped into a couple different definitions, the first being the difference between the terms ``fluctuating'' and ``variation'' with respect to temperature.
+Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al (2018) explicitly defined fluctuating temperature as ``a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis.'' Most generally, temperature variability describes changes in environmental temperature through time (Colinet et al., 2018; Holmes, Woollings, Hawkins, \& Vries, 2016).
+
+\hypertarget{why-is-temperature-variability-important-to-ecological-processes}{%
+\subsection{Why is temperature variability important to ecological processes?}\label{why-is-temperature-variability-important-to-ecological-processes}}
 
@@
 Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate.
-We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates should be the same.
+We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and converse for smaller organisms.
 However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clarke, 2006; Clarke \& Johnston, 1999).
-Others believe that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al., 2006).
+Others highlight that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al., 2006).
 Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.
 
-MTE supports temperature impacting metabolic rate, which can have downstream effects on fitness characteristics.
+MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics.
 This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
@@
 
-\caption[Thermal performance curve components]{Schematic of thermal performance curve and it’s important components for understanding thermal performance.
+\caption[Thermal performance curve components]{Schematic of thermal performance curve and its important components for understanding thermal performance.
+CTmin denotes Critical Thermal Minimum, CTmax denotes Critical Thermal Maximum, and Topt denotes Thermal Optimum.
 Adapted from (Krenek et al 2012, Tuff et al 2016).}\label{fig:unnamed-chunk-4}
@@
 The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, organisms' performance is negatively impacted (Brown et al., 2004; Dowd, King, \& Denny, 2015).
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive (Brown et al., 2004; Dowd, King, \& Denny, 2015).
 Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
 
-TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (i.e.~metabolism) (Amarasekare \& Savage, 2012).
-If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one.
-However, there are several other constraining factors (i.e.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs.
-TPCs strive to represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
+TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (e.g.~metabolism) (Amarasekare \& Savage, 2012).
+If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after whiich it would no longer increase and the organism would fail to survive.
+However, there are several other constraining factors (e.g.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs.
+TPCs represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
 
@@
 TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018).
-However, TPCs usually consist of measurements taken across an array of constant temperatures (i.e.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
+However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
 These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.
@@
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al 2012 demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013).
+For instance, Amarasekare et al (2012) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013).
 Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare \& Savage, 2012).
@@
 Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
-Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al., 2014); thus, if there is disagreement about how metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
-Khelifa et al demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance (Khelifa et al., 2019).
+Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al., 2014); thus, if metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
+Will increased variation have differing effects on different levels of biological organization?
+Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
 
 However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 1.3).
-Khelifa et al demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable (Khelifa et al., 2019).
+Khelifa et al (2019) demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable.
 Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
@@
 It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
-Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas higher levels of organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).
-
-There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at variability (Kroeker et al., 2020; Massey, Holt, Brooks, \& Rollinson, 2019).
+Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas studies at higher levels of biological organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).
+
+There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at the significance of variability on ecosystem dynamics (Kroeker et al., 2020; Massey, Holt, Brooks, \& Rollinson, 2019).
 To understand how environmental variability affects organisms, populations, and communities, we conducted a meta-analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
-(i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (i.e.~periodicity, age, size, thermal history, etc.)?
+(i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (e.g.~periodicity, age, size, thermal history, etc.)?
 We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
 
-We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response (full model), 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will cope better and smaller organisms will cope worse with increased thermal variability (full model), 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics (full model), 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted (thermal stress model).
+We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response, 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will exhibit increased performance and smaller organisms will exhibit decreased performance with increased thermal variability, 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics, 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted.
 
@@
 \caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 19 studies included in the analysis.
-Red indicates no usable data, purple indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
+Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 \item
-  We excluded studies that did not feature both a constant and fluctuating treatment (i.e.~studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
+  We excluded studies that did not feature both a constant and fluctuating treatment (e.g.~studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
 \item
-  We excluded studies that were not completed in a laboratory (i.e.~modeling predictions or results from models, mesocosm experiments, in situ experiments)
+  We excluded studies that were not completed in a laboratory (e.g.~modeling predictions or results from models, mesocosm experiments, \emph{in situ} experiments)
 \item
-  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (i.e.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
+  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
 \item
@@
 \item
-  We excluded studies in which the response variables are comparable or extractable (i.e.~one study performed Principal component analysis for their responses)
+  We excluded studies in which the response variables were not comparable or extractable (e.g.~one study performed Principal component analysis for their responses)
 \end{itemize}
-We excluded studies that did not have the same mean for constant and fluctuating treatments (i.e.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
+We excluded studies that did not have the same mean for constant and fluctuating treatments (e.g.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
 We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004).
@@
 If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and a subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Higgins et al., 2011).
-Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, metafor.
+Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, \emph{metafor}.
 
@@
 org\_level & Level of biological organization, categorized
-                                     as either 0 for individual or 1 for population level
-\cellcolor{gray!6}{                                     responses}\\
+                                     as either 0 for individual, 1 for population level
+\cellcolor{gray!6}{                                     responses, 2 for community level responses}\\
 size & Size estimate of study organism, 0 for extra-small (less than 1mm),
@@
                                      such that rates included any response variable units with a
-                                     “per day” clarifier, all else as traits\\
+                                     “per day” clarifier (e.g. growth rate), all else as traits (e.g. body
+                                     width)\\
 stressful & Whether or not the organisms experienced thermally stressful 
@@
 
-We utilized metafor (version 2.4.0) in R to analyze the extracted, aggregated data.
-Metafor builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer, 2010).
-In order to use metafor, the package requires the data to be in a wider format, in which data are input into the following columns:
+We used \emph{metafor} (version 2.4.0) in R to analyze the extracted, aggregated data.
+\emph{Metafor} builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer, 2010).
+In order to use \emph{metafor}, the package requires the data to be in a wider format, in which data are input into the following columns:
 \begin{table}[!h]
@@
 \end{figure}
-First, it calculates the pooled standard deviation (Pooled SD) from the sample sizes of both studies.
+First, it calculated the pooled standard deviation (Pooled SD) from the sample sizes of both studies.
 Our data had studies with a wide range of sample sizes, so we wanted to include separate sample sizes between the constant and fluctuating treatments.
-The Pooled SD, Cohen's d, which is one of the most common ways to measure effect size, is used to normalize the mean differences between constant and fluctuating treatments.
-Then, using a correction factor J, which again normalizes the data further, we obtain the final SMD.
-The SMD is then ultimately used to calculate the sampling variance, which creates a unitless metric for variance.
-
-Both the SMD and sampling variance are then input into the \texttt{rma.mv} function in metafor to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
-We first ran a multivariate random effects model without modifiers (simple model) to establish baseline heterogeneity between studies to then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
+The Pooled SD, Cohen's d, which is one of the most common ways to measure effect size, was used to normalize the mean differences between constant and fluctuating treatments.
+Then, using a correction factor J, which again normalized the data further, we obtained the final SMD.
+The SMD was then ultimately used to calculate the sampling variance, which created a unitless metric for variance.
+
+Both the SMD and sampling variance were then input into the \texttt{rma.mv} function in \emph{metafor} to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
+We first ran a multivariate random effects model without modifiers (simple model) to establish baseline heterogeneity between studies, and then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
 \begin{Shaded}
@@
 \end{Shaded}
-We also ran a separate model without this study and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly an important factor or if it was simply excluding those influential data points:
+We also ran a separate model without Garcia-Ruiz et al 2011 and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly an important factor or if it was simply excluding those influential data points:
 \begin{Shaded}
@@
 \caption[Forest plot of effect sizes]{Forest plot of effects sizes by study and experiment within a study.
-Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study.
+Data indicate that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study.
 Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-18}
 \end{figure}
-Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1.
-These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1).
-As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
-
-Our full model, including 7 covariates (all, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p \textless{} 0.0001).
-Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p \textless{} 0.00001).
-Organization level (n=140) was also important in the model, as a statistically significant covariate (df= 7, p \textless{} 0.05).
+Overall, even our simple model demonstrated that the intercept was not significant but there was significant heterogeneity across studies (n=140, z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1.
+These results allowed us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1).
+This then allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.
+
+Our full model, including 7 covariates (all covariates in Table 2.2, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p \textless{} 0.0001).
+Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (n=140, df = 7, p \textless{} 0.00001).
+Organization level was also important in the model, as a statistically significant covariate (n=140, df= 7, p \textless{} 0.05).
 \begin{figure}
@@
 
-\caption[Coefficient estimates for full model]{Coefficient estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.
-Error bars represent 95 percent confidence intervals.}\label{fig:unnamed-chunk-19}
+\caption[Coefficient estimates for full model]{Coefficient estimates for each covariate in the full model, where * signify statistically significant model predictors.
+Error bars represent 95 percent confidence intervals.
+Yellow indicates a positive coefficient estimate, purple indicates a negative coefficient estimate.}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 
-Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).
-The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p \textless{} 0.05.
-Surprisingly, the additional negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant.
-In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).
+Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant (n=140, df=7, p \textless{} 0.001 level (Figure 3.2).
+The most negative estimate, organization level (n=140), was also significantly influential in our model (n=140, df=7, p \textless{} 0.05).The additional negative estimates of our model coefficients, experimental age and size, were not statistically significant.
+In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant (n=132, df=1, p = 0.4855) in
+explaining the effect size (Table 3.1).
 \begin{table}[!h]
@@
 The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
-At lower mean temperatures, higher fluctuation ranges have a more positive effect on organism responses and performance, though that trend starkly ends at about \(24^{\circ}\)C.
-Beyond \(24^{\circ}\)C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
+At lower mean temperatures, higher fluctuation ranges generally have a more positive effect on organism responses and performance, though that trend starkly ends at about \(24^{\circ}\)C.
+However, at higher mean temperatures, lower fluctuation ranges appears to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
 \begin{figure}
@@
 
-\caption[Scatterplot of relationship between range, mean, and effect size]{Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and effect size.
+\caption[Scatterplot of relationship between range, mean, and SMD]{Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD.
 Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C.
-Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-21}
+SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-21}
 \end{figure}
@@
 
-As noted in the methods, when we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p \textless{} 0.001) as did mean temperature (df=8, p \textless{} 0.001).
-Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p \textless{} 0.1).
+When we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p \textless{} 0.001) as did mean temperature (df=8, p \textless{} 0.001).
+Organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p \textless{} 0.1).
 \begin{figure}
@@
 Our model results highlight the importance of fluctuation range, mean temperature, and organization level to better understand the variety of effects thermal variation has on organismal responses.
-These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e.~higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures.
+These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e.~higher fluctuation ranges at lower mean temperatures generally have higher effect sizes compared to higher and lower effects sizes of lower fluctuation ranges at higher mean temperatures.
 The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change (Sunday, Bates, \& Dulvy, 2012; Vasseur et al., 2014).
@@
 Acclimation may also explain differences in response across studies as the scale on which variation occurs influences how organisms may respond.
-If organisms' non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (i.e.~inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies (Kremer, Fey, Arellano, \& Vasseur, 2018).
+If organisms' non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (e.g.~inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies (Kremer, Fey, Arellano, \& Vasseur, 2018).
 In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance.
@@
 If organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics (Yoshida, Jones, Ellner, Fussmann, \& Hairston, 2003).
-It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.
+It is important to note that rapid evolution will only be relevant to population levels responses and not individual level responses.
 
@@
 
-An important side note is that our full model and thermal stress model differed in the number of studies, as one study, García-Ruiz, Marco, \& Pérez-Moreno (2011), did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola).
+An important side note is that our full model and thermal stress model differed in the number of studies, as one study, García-Ruiz, Marco, \& Pérez-Moreno (2011), did not have any supplemental information on thermal performance curve points for the taxa used in that study (\emph{Xylotrechus arvicola}).
 Further, this study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding García-Ruiz et al.
@@
 We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018).
-This may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile stage instead.
+This may again be the result of studies not focused on the larval stage of development, but instead on the juvenile stage.
 However, there have been meta-analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
-These results align with patterns of juvenile organisms (i.e.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Moyano et al., 2017; Portner \& Farrell, 2008).
+These results align with patterns of juvenile organisms (e.g.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Moyano et al., 2017; Portner \& Farrell, 2008).
 
 Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project.
-As Figure 1.3 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures.
-It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated.
+Diurnal cycles are more correlated than reddened cycles, like seasonal temperatures (Figure 1.3).
+It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, longer-term temperature patterns are more stable than diurnal cycles, and less autocorrelated.
 Are these diurnal temperature fluctuations accurate for what is expected under natural conditions?
@@
 Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
-Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes (Colinet et al., 2018, 2015).
+Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly ecologically irrelevant thermal regimes (Colinet et al., 2018, 2015).
 Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.
@@
 How organisms' life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond (Bernhardt, O'Connor, Sunday, \& Gonzalez, 2020).
-This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.
+This is also pertinent to the scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis.
 Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times.
@@
 A larger sample size of studies to draw more robust conclusions with respect to support for or against allometric scaling and MTE with respect to thermal variability would be useful.
-There is more to be studied on how additional experimental designs, i.e.~acclimation or heat-wave simulations, as well as how biological performance responds to environmental variability.
+There is more to be studied on how additional experimental designs, e.g.~acclimation or heat-wave simulations, as well as how biological performance responds to environmental variability.
 However, this is an important first step in quantitatively summarizing progress so far.
@@
 \caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis (missing data on species X.
-arvicola, used in Garcia-Ruiz et al 2011)}
+arvicola, used in Garcia-Ruiz et al 2011) and scientific classification at the class level}
 \centering
-\begin{tabular}[t]{ll>{\raggedright\arraybackslash}p{6cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}ll>{\raggedright\arraybackslash}p{3cm}}
 \toprule
-\textbf{Species} & \textbf{Topt (°C)} & \textbf{Reference}\\
+\textbf{Species} & \textbf{Topt (°C)} & \textbf{Class} & \textbf{Reference}\\
 \midrule
-\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6°C} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
-Leptopilina boulardi & 25°C & De lava et al 2016\\
-\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 ± 0.64 °C} & \cellcolor{gray!6}{Dang et al 2019}\\
-Gryllus firmus & ~30°C & Singh et al 2020\\
-\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30°C} & \cellcolor{gray!6}{Seebacher et al 2014}\\
+\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6°C} & \cellcolor{gray!6}{Insecta} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
+Leptopilina boulardi & 25°C & Insecta & De lava et al 2016\\
+\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 ± 0.64 °C} & \cellcolor{gray!6}{Reptilia} & \cellcolor{gray!6}{Dang et al 2019}\\
+Gryllus firmus & 30°C & Insecta & Singh et al 2020\\
+\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30°C} & \cellcolor{gray!6}{Amphibia} & \cellcolor{gray!6}{Seebacher et al 2014}\\
 \addlinespace
-Limnodynastes tasmaniensis & 30-33°C & Whitehead et al 1989\\
-\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33°C} & \cellcolor{gray!6}{Kern et al 2014}\\
-Natrix natrix & ~30°C & Issac 1997\\
-\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26°C} & \cellcolor{gray!6}{Alagawany et al 2017}\\
-Plestiodon chinensis & 33°C & Baojun et al 2010\\
+Limnodynastes tasmaniensis & 30-33°C & Amphibia & Whitehead et al 1989\\
+\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33°C} & \cellcolor{gray!6}{Amphibia} & \cellcolor{gray!6}{Kern et al 2014}\\
+Natrix natrix & 30°C & Reptilia & Issac 1997\\
+\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26°C} & \cellcolor{gray!6}{Aves} & \cellcolor{gray!6}{Alagawany et al 2017}\\
+Plestiodon chinensis & 33°C & Reptilia & Baojun et al 2010\\
 \addlinespace
-\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35°C} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
-Escherichia coli & 37°C, 35-36°C & Doyle and Schoeni 1984, Bronikowski et al 2001\\
-\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36°C} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
-Caiman latirostris & 32-33°C & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
-\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33°C} & \cellcolor{gray!6}{Dang et al 2019}\\
+\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35°C} & \cellcolor{gray!6}{Insecta} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
+Escherichia coli & 37°C, 35-36°C & Gammaproteobacteria & Doyle and Schoeni 1984, Bronikowski et al 2001\\
+\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36°C} & \cellcolor{gray!6}{Gammaproteobacteria} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
+Caiman latirostris & 32-33°C & Reptilia & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
+\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33°C} & \cellcolor{gray!6}{Reptilia} & \cellcolor{gray!6}{Dang et al 2019}\\
 \addlinespace
-Esteya vermicola & 26°C & Wang et al 2015\\
+Esteya vermicola & 26°C & Not assigned; Fungus & Wang et al 2015\\
+\cellcolor{gray!6}{Xylotrechus arvicola} & \cellcolor{gray!6}{N/A} & \cellcolor{gray!6}{Insecta} & \cellcolor{gray!6}{Garcia-Ruiz et al 2011}\\
 \bottomrule

### 2021-05-11 writing /thesis_draft/index.Rmd
putting in everyone's edits from orals

@@
 institution: 'Reed College'
-division: 'Mathematics and Natural Sciences'
+division: 'Mathematical and Natural Sciences'
 advisor: 'Sam Fey'
@@
   thesisdown::thesis_pdf: default 
-abstract: 'Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns.
-Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
-Here, we present a quantitative meta-analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization.
-Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model.
+abstract: 'Ecosystems and organisms have experienced variation on many temporal scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns.
+Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the persistence and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization.
+Here, we present a quantitative meta-analysis of how thermal variability impacts biological responses across multiple taxa, organism sizes, and levels of biological organization.
+Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model of biological responsiveness to temperature variability.
 Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.'
@@
   @semenov_influence_2007, @treidel_temperature_2016-1, @klepsatel_variation_2013,
-  @krenek_coping_2012, @tuff_framework_2016
+  @krenek_coping_2012, @tuff_framework_2016, @vasseur_increased_2014, @khelifa_usefulness_2019
 
@@
 Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time [@knoll_chapter_2008].
-In contrast, vascular plants have evolved different photosynthetic systems (C4, CAM) to manage a variety of environmental conditions [@guralnick_evolutionary_2007].
-Though cyanobacteria and vascular plants evolved new strategies and structures to persist in a new environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant [@reed_phenotypic_2010; @burggren_developmental_2018].
+Vascular plants have also evolved different photosynthetic systems (e.g.
+C4, CAM) to manage a variety of environmental conditions [@guralnick_evolutionary_2007].
+Though cyanobacteria and vascular plants evolved new strategies and structures to persist as the environment changed, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant [@reed_phenotypic_2010; @burggren_developmental_2018].
 Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.
@@
 In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
-Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.
+Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.
 El Niño Southern Oscillation).
-As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems [@khelifa_usefulness_2019; @stewart_mesocosm_2013].
+As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems [@khelifa_usefulness_2019; @stewart_mesocosm_2013].
 How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
-As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.
+As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g.
 hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses [@helmuth_organismal_2010; @bernardo_biologically_2014; @huey_predicting_2012].
@@
 ## Why this study on variability matters 
-There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, i.e.
+There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g.
 inflationary effects in sink-source systems [@holt_impacts_2003; @vasseur_phase-locking_2009], such that variability aids population density when organisms emigrate.
-However, there is a lack of consensus as to which scales of biological organization, i.e.
-individuals, populations, or communities of organisms will be most severely affected by different shifts in environmental conditions (problem statement).
-We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response (aims).
+However, there is a lack of consensus as to which scales of biological organization, e.g.
+individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions.
+We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response.
 To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.

### 2021-05-12 writing /thesis_draft/01-Background.Rmd
final copy submitted to the library

@@
 Shown is a comparison of how duration of variability compares within the same environmental factor.
-Terms highlighted in each panel are described below in the main text", out.width = '105%', fig.align="center", fig.scap="Metrics to describe and explain environmental variability."}
+Terms highlighted in each panel are described below in the main text.", out.width = '105%', fig.align="center", fig.scap="Metrics to describe and explain environmental variability."}
 knitr::include_graphics("figures/figure1.png")
@@
 ### Mean
-Aggregating values together, whether it be temperature, light availability, precipitation, allows for a better understanding of trends and patterns as opposed to random events.
+Aggregating values together, whether it be temperature, light availability, or precipitation, allows for a better understanding of trends and patterns as opposed to random events.
 The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time.
@@
 Performance is a set of whole organism traits (e.g.
-assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment [@husak_how_2017; @amarasekare_framework_2012].
-Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+assimilation, development, metabolic rate) with the availability to perform dynamic actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms’ persistence in the environment [@husak_how_2017; @amarasekare_framework_2012].
+Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience multiple warming exposures during a single day.
 However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.
@@
 In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability [@lampert_adaptive_1989; @brierley_diel_2014].
-However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, e.g.
+However, there are also instances in which the environment is not temporally variable, yet an organism utilizes spatial variability to regulate its behavior, e.g.
 an organism moving from sun to shade during the day to regulate body temperature and metabolic processes [@huey_behavioral_1974; @kearney_potential_2009].
@@
 Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment [@huey_physiological_1990].
-Brown et al highlighted the importance of both temperature and body size in the metabolic theory of  ecology (MTE) [@brown_toward_2004].
+Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of  ecology (MTE).
 Though highly debated [@clarke_temperature_2006], MTE serves as an origin point for understanding how organisms respond to environmental conditions.
@@
 Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate.
-We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and converse for smaller organisms.
+We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and for smaller organisms should be lower.
 However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish [@clarke_scaling_1999; @clarke_temperature_2006].
@@
 MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics.
-This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
+This is typically described by a characteristic unimodal, left skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
 
@@
 metabolism) [@amarasekare_framework_2012].
-If only temperature’s effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after whiich it would no longer increase and the organism would fail to survive.
+If only temperature’s effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after which it would no longer increase and the organism would fail to survive.
 However, there are several other constraining factors (e.g.
@@
 
-TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment [@bernhardt_nonlinear_2018].
+TPCs have been crucial in predicting and understanding how organisms respond to their environment [@bernhardt_nonlinear_2018].
 However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.
 15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
-These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.
+These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their natural environments.
 
@@
 Jensen’s inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions [@ruel_jensens_1999].
-For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen’s inequality [@ruel_jensens_1999].
-The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures [@ruel_jensens_1999].
+For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen’s inequality (1999).
+The development rate of ectotherms has a characteristically nonlinear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures [@ruel_jensens_1999].
 With Jensen’s inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures [@ruel_jensens_1999].
@@
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al (2012) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns [@amarasekare_framework_2012; @paaijmans_temperature_2013].
+For instance, Amarasekare et al (2012) demonstrated the importance of where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns [@amarasekare_framework_2012; @paaijmans_temperature_2013].
 Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature [@amarasekare_framework_2012].
@@
 
- * At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve [@bartheld_thermal_2017-2] to changes in TSD ratios [@bowden_temperature_2018]
+ * At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, from flattening the thermal performance curve [@bartheld_thermal_2017-2] to changes in TSD ratios [@bowden_temperature_2018]
 

### 2021-05-12 writing /thesis_draft/02-chap2.Rmd
final copy submitted to the library

@@
 
-```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 19 studies included in the analysis.
+```{r, echo=FALSE, fig.cap="Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 15 studies included in the analysis.
 Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.", out.width = '90%', fig.align="center", fig.scap="Meta-analysis paper selection process"}
@@
 * We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.
-organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
+organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a TPC)
 * We excluded any articles not written in English 
 * We excluded studies that did not report measurements of error
-* We excluded studies in which the response variables were not comparable or extractable (e.g. one study performed Principal component analysis for their responses)
+* We excluded studies in which the response variables were not comparable or extractable (e.g. one study performed Principal Component Analysis for their responses)
 
@@
 treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
-We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response [@brown_toward_2004].
+We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude most simply to variation, as changes in mean temperature significantly influence the magnitude of response [@brown_toward_2004].
 The final number of data points included in the analysis is broken down by study (Table 2.1).
@@
                                      as either 0 for individual, 1 for population level
-                                     responses, 2 for community level responses", 
+                                     responses", 
                                      "Size estimate of study organism, 0 for extra-small (less than 1mm),
@@
 
-## Metafor
+## *Metafor*
 
@@
 Each of the subsequent arguments of the `escalc` function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
-The `escalc` function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
+The `escalc` function uses multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
 
@@
 We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population [@viechtbauer_conducting_2010].
-The random effects model structure also allowed us to account for the nested structure of responses, experiment, and study to address the dependence of observations from within each study.
+The random effects model structure also allowed us to account for the nested structure of responses, experiments, and studies to address the dependence of observations from within each study.
 The multivariate random effects model with all modifiers (full model) we ran included:
@@
 The full model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
-If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed an organisms’ Topt and become thermally stressful.
+If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed the organisms’ Topt and become thermally stressful.
 

### 2021-05-12 writing /thesis_draft/_book/thesis.tex
final copy submitted to the library

@@
 
-In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
 Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.~El Niño Southern Oscillation).
@@
 There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate.
-However, there is a lack of consensus as to which scales of biological organization, e.g.~individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions.
+However, there is a lack of consensus as to which scales of biological organization, i.e.~individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions.
 We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response.
@@
 Shown is a comparison of how duration of variability compares within the same environmental factor.
-Terms highlighted in each panel are described below in the main text}\label{fig:unnamed-chunk-3}
+Terms highlighted in each panel are described below in the main text.}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 
-Aggregating values together, whether it be temperature, light availability, precipitation, allows for a better understanding of trends and patterns as opposed to random events.
+Aggregating values together, whether it be temperature, light availability, or precipitation, allows for a better understanding of trends and patterns as opposed to random events.
 The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time.
@@
 Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018).
-Performance is a set of whole organism traits (e.g.~assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
-Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day.
+Performance is a set of whole organism traits (e.g.~assimilation, development, metabolic rate) with the availability to perform dynamic actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
+Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience multiple warming exposures during a single day.
 However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful.
@@
 In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability (Brierley, 2014; Lampert, 1989).
-However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, e.g.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).
+However, there are also instances in which the environment is not temporally variable, yet an organism utilizes spatial variability to regulate its behavior, e.g.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).
 
@@
 Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey \& Bennett, 1990).
-Brown et al highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE) (Brown, Gillooly, Allen, Savage, \& West, 2004).
+Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE).
 Though highly debated (Clarke, 2006), MTE serves as an origin point for understanding how organisms respond to environmental conditions.
@@
 Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate.
-We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and converse for smaller organisms.
+We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and for smaller organisms should be lower.
 However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clarke, 2006; Clarke \& Johnston, 1999).
@@
 MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics.
-This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
+This is typically described by a characteristic unimodal, left skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
 \begin{figure}
@@
 The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive (Brown et al., 2004; Dowd, King, \& Denny, 2015).
+CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive (Brown, Gillooly, Allen, Savage, \& West, 2004; Dowd, King, \& Denny, 2015).
 Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
@@
 TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (e.g.~metabolism) (Amarasekare \& Savage, 2012).
-If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after whiich it would no longer increase and the organism would fail to survive.
+If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after which it would no longer increase and the organism would fail to survive.
 However, there are several other constraining factors (e.g.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs.
@@
 
-TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018).
+TPCs have been crucial in predicting and understanding how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018).
 However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
-These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.
+These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their natural environments.
 
@@
 Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel \& Ayres, 1999).
-For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen's inequality (Ruel \& Ayres, 1999).
-The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel \& Ayres, 1999).
+For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen's inequality (1999).
+The development rate of ectotherms has a characteristically nonlinear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel \& Ayres, 1999).
 With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel \& Ayres, 1999).
@@
 Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al (2012) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013).
+For instance, Amarasekare et al (2012) demonstrated the importance of where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013).
 Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare \& Savage, 2012).
@@
 \item
-  At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve (Bartheld, Artacho, \& Bacigalupe, 2017) to changes in TSD ratios (Bowden \& Paitz, 2018)
+  At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, from flattening the thermal performance curve (Bartheld, Artacho, \& Bacigalupe, 2017) to changes in TSD ratios (Bowden \& Paitz, 2018)
 \item
@@
 
-\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 19 studies included in the analysis.
+\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 15 studies included in the analysis.
 Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
@@
 \item
-  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
+  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a TPC)
 \item
@@
 \item
-  We excluded studies in which the response variables were not comparable or extractable (e.g.~one study performed Principal component analysis for their responses)
+  We excluded studies in which the response variables were not comparable or extractable (e.g.~one study performed Principal Component Analysis for their responses)
 \end{itemize}
 We excluded studies that did not have the same mean for constant and fluctuating treatments (e.g.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
-We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004).
+We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude most simply to variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004).
 The final number of data points included in the analysis is broken down by study (Table 2.1).
@@
                                      as either 0 for individual, 1 for population level
-\cellcolor{gray!6}{                                     responses, 2 for community level responses}\\
+\cellcolor{gray!6}{                                     responses}\\
 size & Size estimate of study organism, 0 for extra-small (less than 1mm),
@@
 \hypertarget{metafor}{%
-\section{Metafor}\label{metafor}}
+\section{\texorpdfstring{\emph{Metafor}}{Metafor}}\label{metafor}}
 
@@
 Each of the subsequent arguments of the \texttt{escalc} function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
-The \texttt{escalc} function is using multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
+The \texttt{escalc} function uses multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
 \begin{figure}
@@
 We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population (Viechtbauer, 2010).
-The random effects model structure also allowed us to account for the nested structure of responses, experiment, and study to address the dependence of observations from within each study.
+The random effects model structure also allowed us to account for the nested structure of responses, experiments, and studies to address the dependence of observations from within each study.
 The multivariate random effects model with all modifiers (full model) we ran included:
@@
 The full model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
-If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed an organisms' Topt and become thermally stressful.
+If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed the organisms' Topt and become thermally stressful.
 
@@
 The most negative estimate, organization level (n=140), was also significantly influential in our model (n=140, df=7, p \textless{} 0.05).The additional negative estimates of our model coefficients, experimental age and size, were not statistically significant.
-In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant (n=132, df=1, p = 0.4855) in
-explaining the effect size (Table 3.1).
+In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant (n=132, df=1, p = 0.4855) in explaining the effect size (Table 3.1).
 \begin{table}[!h]
@@
 
-The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
+The interaction between fluctuation range and mean temperature as well as effect size is best displayed in Figure 3.3.
 At lower mean temperatures, higher fluctuation ranges generally have a more positive effect on organism responses and performance, though that trend starkly ends at about \(24^{\circ}\)C.
-However, at higher mean temperatures, lower fluctuation ranges appears to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
+However, at higher mean temperatures, lower fluctuation ranges appear to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
 \begin{figure}
@@
 \caption[Scatterplot of relationship between range, mean, and SMD]{Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD.
-Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C.
+Temperature fluctuation ranged from 0-20°C, mean temperature ranged from 7-33°C.
 SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-21}
@@
 Body size has been identified as important in the allometric scaling of MTE and it is interesting that in these subsets of studies, body size did not explain the differences in responses across studies.
-This could simply be because we had a small sample size of studies (n=15), though multiple responses within each study contributed to a large sample of effect sizes (n=140), there were only a subset of relative body sizes included.
+This could simply be because we had a small sample size of studies (n=15).
+Though multiple responses within each study contributed to a large sample of effect sizes (n=140), there were only a subset of relative body sizes included.
 It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms (Clarke, 2006; Clarke \& Johnston, 1999).
@@
 This may again be the result of studies not focused on the larval stage of development, but instead on the juvenile stage.
-However, there have been meta-analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
+However, there have been meta-analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptile development, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
 These results align with patterns of juvenile organisms (e.g.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Moyano et al., 2017; Portner \& Farrell, 2008).
@@
 
-\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis (missing data on species X.
-arvicola, used in Garcia-Ruiz et al 2011) and scientific classification at the class level}
+\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis and scientific classification at the class level}
 \centering

### 2021-05-12 writing /thesis_draft/index.Rmd
final copy submitted to the library

@@
 ## Relevant significance of variability today 
-In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales.
 Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.
@@
 inflationary effects in sink-source systems [@holt_impacts_2003; @vasseur_phase-locking_2009], such that variability aids population density when organisms emigrate.
-However, there is a lack of consensus as to which scales of biological organization, e.g.
+However, there is a lack of consensus as to which scales of biological organization, i.e.
 individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions.

### 2021-06-22 shared analysis files /manuscript_datavis_22jun21.Rmd
reorganizing repo and creating data vis file for manuscript data analysis

---
title: "manuscript_data_wrangling"
author: "Maggie Slein"
date: "6/14/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, results = 'hide', message=FALSE}
## loading libraries 
pacman::p_load(metafor, tidyverse, viridis, visreg, forcats)

#load data
dat_acclim_0<- read_csv("acclimation_full_2june21.csv")
dat_full_var_0 <- read_csv("metafor.csv")
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
##wrangling calculations for acclimation data
#make standardized SD columns based on N and SE vs SD
dat_acclim <-dat_acclim_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_acclim_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_acclim, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_acclim_ES<-dat_acclim_ES %>% 
  filter (!is.na(yi)) 
```

```{r, echo= FALSE, results = 'hide', message=FALSE}
##wrangling calculations for full variability data
#make standardized SD columns based on N and SE vs SD
dat_full_var <-dat_full_var_0 %>% 
  mutate ( SD_constant= if_else ( variance_type== 0, constant_variance * sqrt(constant_samp), constant_variance))  %>%
  mutate ( SD_variable= if_else ( variance_type== 0, flux_variance * sqrt(flux_variance) , flux_variance))

#calculate ES for data
dat_full_var_ES <-escalc(measure="SMD", m1i=constant_resp, m2i=flux_resp, 
            sd1i=`SD_constant`, sd2i= `SD_variable`, n1i=constant_samp, n2i=flux_samp, 
            data=dat_full_var, slab=paste(study_id, experiment_id, response_id, sep=", "))

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_full_var_ES<-dat_full_var_ES %>% 
  filter (!is.na(yi)) 
```

# Acclimation Model

This data and analysis accounts for the temperatures at which organisms are reared and how their performance compares when exposed to different temperature. The results from this model has very similar results from the model I originally ran for my thesis...

```{r}
simple_acclimation_model <- rma.mv(yi, vi, data=dat_acclim_ES, 
               random = ~1 | study_id / experiment_id/ response_id,
                 method="REML") 
simple_acclimation_model
```

```{r}
acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, mods = ~flux_range * mean_temp_reared +
                        exp_age + size + org_level + exposure_temp, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML") 
acclimation_model
```

# Neglecting Acclimation

This uses the full dataset from my thesis (with more data from the additional search). Interesting that this time, mean temp is the only significant predictor?

```{r}
simple_full_var_model <- rma.mv(yi, vi, data=dat_full_var_ES, 
               random = ~1 | study_id / experiment_id/ response_id,
                 method="REML") 
simple_full_var_model
```

```{r}
full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML") 
full_var_model
```

```{r}
dat_acclim_ES %>%
  filter(yi > -5) %>%
ggplot(aes(x=mean_temp_reared, y = yi, fill=flux_range))+
  geom_point(alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Temperature range (°C)', option = "C")+
  theme_classic()+
  xlab("Mean Temperature (°C)")+
  ylab("SMD")+
  geom_line(y=0, linetype = "dashed")+
  theme(legend.position="bottom")+
  ggtitle("acclimation")
```

```{r}
dat_full_var_ES %>%
  filter(yi > -5) %>%
ggplot(aes(x=mean_temp_constant, y = yi, fill=flux_range))+
  geom_point(alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Temperature range (°C)', option = "C")+
  theme_classic()+
  xlab("Mean Temperature (°C)")+
  ylab("SMD")+
  geom_line(y=0, linetype = "dashed")+
  theme(legend.position="bottom")+
  ggtitle("neglecting acclimation (full dataset)")
```

### 2021-06-22 thesis writing/data_wrangling_plot.Rmd
reorganizing repo and creating data vis file for manuscript data analysis

--- 2021-04-27 shared analysis files /data_wrangling_plot.Rmd
@@
 full_rf_model
+
+#full model plus thermal stress
+full_stress_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
+                        exp_age + size + org_level + resp_type + stressful, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+full_stress_model
+
 ```
@@
   rename(value = "coef(full_rf_model)") %>%
-  arrange(desc(value))
+  arrange(desc(value)) %>%
+  mutate(coefficients = c("intercept", "response trait type", 
+                          "fluctuation range", "mean temperature",
+                          "fluctuation range:mean temperature",
+                                       "size", "experimental age","organization level" 
+                                      ))
 
-label.df <- data.frame(covariates = c("flux_range", "flux_range:mean_constant_temp", "org_level"),
-                       value = c(0.3, -0.01, -0.69))
+label.df <- data.frame(coefficients = c("fluctuation range", "fluctuation range:mean temperature", "organization level"),
+                       value = c(0.35, -0.07, -0.69))
 
-ggplot(data=full_model_coefficients, aes(x = reorder(covariates, value),
+ggplot(data=full_model_coefficients, aes(x = reorder(coefficients, value),
              y = value,

### 2021-06-22 thesis writing/thesis_draft/03-chap3.Rmd
reorganizing repo and creating data vis file for manuscript data analysis

--- 2021-05-11 writing /thesis_draft/03-chap3.Rmd
@@
 The most negative estimate, organization level (n=140), was also significantly influential in our model (n=140, df=7, p < 0.05).The additional negative estimates of our model coefficients, experimental age and size, were not statistically significant.
-In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant (n=132, df=1, p = 0.4855) in
-explaining the effect size (Table 3.1).
+In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant (n=132, df=1, p = 0.4855) in explaining the effect size (Table 3.1).
 
@@
 
-The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3.
+The interaction between fluctuation range and mean temperature as well as effect size is best displayed in Figure 3.3.
 At lower mean temperatures, higher fluctuation ranges generally have a more positive effect on organism responses and performance, though that trend starkly ends at about $24^{\circ}$C.
-However, at higher mean temperatures, lower fluctuation ranges appears to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
+However, at higher mean temperatures, lower fluctuation ranges appear to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
 
 ```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD.
-Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C.
+Temperature fluctuation ranged from 0-20°C, mean temperature ranged from 7-33°C.
 SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and SMD"}
@@
 Body size has been identified as important in the allometric scaling of MTE and it is interesting that in these subsets of studies, body size did not explain the differences in responses across studies.
-This could simply be because we had a small sample size of studies (n=15), though multiple responses within each study contributed to a large sample of effect sizes (n=140), there were only a subset of relative body sizes included.
+This could simply be because we had a small sample size of studies (n=15).
+Though multiple responses within each study contributed to a large sample of effect sizes (n=140), there were only a subset of relative body sizes included.
 It could also provide support for the argument against allometric scaling in MTE, such that there only individualized trends amongst taxonomic groups as opposed to general trends across individuals, populations, and communities of organisms [@clarke_scaling_1999; @clarke_temperature_2006].
@@
 This may again be the result of studies not focused on the larval stage of development, but instead on the juvenile stage.
-However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018].
+However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptile development, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018].
 These results align with patterns of juvenile organisms (e.g.

### 2021-06-22 thesis writing/thesis_draft/05-appendix.Rmd
reorganizing repo and creating data vis file for manuscript data analysis

--- 2021-04-29 writing /thesis_draft/05-appendix.Rmd
@@
                          "Salmonella", "Caiman latirostris", 
-                         "Trachemys scripta", "Esteya vermicola"), 
-                 Topt = c("24-26.6°C", "25°C", "37.97 ± 0.64 °C", "~30°C", "30°C", "30-33°C",
-                          "33°C", "~30°C", "26°C", "33°C", "30-35°C", "37°C, 35-36°C",
-                          "35-36°C", "32-33°C", "32-33°C", "26°C"), 
+                         "Trachemys scripta", "Esteya vermicola", "Xylotrechus arvicola"), 
+                 Topt = c("24-26.6°C", "25°C", "37.97 ± 0.64 °C", "30°C", "30°C", "30-33°C",
+                          "33°C", "30°C", "26°C", "33°C", "30-35°C", "37°C, 35-36°C",
+                          "35-36°C", "32-33°C", "32-33°C", "26°C", "N/A"), 
+                 Class = c("Insecta", "Insecta", "Reptilia", "Insecta", "Amphibia", "Amphibia", "Amphibia",
+                           "Reptilia", "Aves", "Reptilia", "Insecta", "Gammaproteobacteria",
+                           "Gammaproteobacteria","Reptilia", "Reptilia", "Not assigned; Fungus", "Insecta"),
                  `Reference` = c("Klepsatel 2013, David 1988", "De lava et al 2016", 
@@
                                  "Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002", 
-                                 "Dang et al 2019", "Wang et al 2015"))
+                                 "Dang et al 2019", "Wang et al 2015", "Garcia-Ruiz et al 2011"))
 kable(table4, "latex", booktabs = T, 
-      col.names = c("Species", "Topt (°C)", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis (missing data on species X.
-arvicola, used in Garcia-Ruiz et al 2011)", caption.short ="Thermal stress data by species") %>%
+      col.names = c("Species", "Topt (°C)", "Class", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis and scientific classification at the class level", caption.short ="Thermal stress data by species") %>%
    row_spec(0,bold=TRUE) %>%
-  column_spec(3, width = "6cm") %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(4, width = "3cm") %>%
    kable_styling(latex_options = c("striped", "hold_position"))

### 2021-06-22 thesis writing/thesis_draft/_book/thesis.tex
reorganizing repo and creating data vis file for manuscript data analysis

--- 2021-05-12 writing /thesis_draft/_book/thesis.tex
@@
 % Syntax highlighting #22
-  \usepackage{color}
-  \usepackage{fancyvrb}
-  \newcommand{\VerbBar}{|}
-  \newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-  % Add ',fontsize=\small' for more characters per line
-  \usepackage{framed}
-  \definecolor{shadecolor}{RGB}{248,248,248}
-  \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-  \newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-  \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-  \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-  \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-  \newcommand{\BuiltInTok}[1]{#1}
-  \newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-  \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-  \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-  \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-  \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-  \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-  \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-  \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-  \newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-  \newcommand{\ExtensionTok}[1]{#1}
-  \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-  \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-  \newcommand{\ImportTok}[1]{#1}
-  \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-  \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-  \newcommand{\NormalTok}[1]{#1}
-  \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-  \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-  \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-  \newcommand{\RegionMarkerTok}[1]{#1}
-  \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-  \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-  \newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-  \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-  \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-  \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 
+Placeholder
+
 \hypertarget{the-historical-significance-of-variability}{%
@@
 
-It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bambach, 1993; Bouchard, 2014).
-Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Blankenship, 2017; Soo, Hemp, Parks, Fischer, \& Hugenholtz, 2017).
-Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll, 2008).
-Vascular plants have also evolved different photosynthetic systems (e.g.~C4, CAM) to manage a variety of environmental conditions (Guralnick, Cline, Smith, \& Sage, 2007).
-Though cyanobacteria and vascular plants evolved new strategies and structures to persist as the environment changed, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Burggren, 2018; Reed, Waples, Schindler, Hard, \& Kinnison, 2010).
-Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.
-
 \hypertarget{relevant-significance-of-variability-today}{%
@@
 
-Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.~El Niño Southern Oscillation).
-As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems (Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019; Stewart et al., 2013).
-How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics.
-As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).
-
 \hypertarget{why-this-study-on-variability-matters}{%
@@
 
-There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate.
-However, there is a lack of consensus as to which scales of biological organization, i.e.~individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions.
-We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response.
-To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.
-
 \hypertarget{background}{%
@@
 
+Placeholder
+
 \hypertarget{environmental-variability}{%
 \section{Environmental variability}\label{environmental-variability}}
-\begin{figure}
-
-{\centering \includegraphics[width=1.05\linewidth]{figures/figure1}
-
-}
-
-\caption[Metrics to describe and explain environmental variability.]{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability.
-Shown is a comparison of how duration of variability compares within the same environmental factor.
-Terms highlighted in each panel are described below in the main text.}\label{fig:unnamed-chunk-3}
-\end{figure}
-Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms' response to environmental conditions (Gudmundson, Eklöf, \& Wennergren, 2015; Parepa, Fischer, \& Bossdorf, 2013).
-Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1.1).
 
@@
 
-Aggregating values together, whether it be temperature, light availability, or precipitation, allows for a better understanding of trends and patterns as opposed to random events.
-The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time.
-It provides a departure point for what we expect the average value to be for all time.
-In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, e.g.~averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations.
-However, in other cases, the mean is less useful compared to the variability.
-For instance, if there are key factors contributing to population growth rates from specific locations, comparing the amount of variability in additional variables across locations (individual respiration rates, temperature, etc.) is more fruitful for the task at this scale.
-At some point however, all data of interest are averages.
-The important distinction is the scale at which we average the data and the level of detail provided at each scale.
-Means are more useful in describing variation amongst normally distributed data, however, other statistical metrics like the median or mode may be more appropriate when the data is not normally distributed.
-
 \hypertarget{amplituderange}{%
@@
 
-We can expand additional information adjacent to the mean (e.g.~range, duration, predictability) for a finer scale picture of environmental conditions.
-Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014).
-We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience.
-This range can be short or long, for example, geographic location drastically influences the amount of sunlight an ecosystem receives.
-The Amazon Rainforest experiences less variation in light availability during the year than Denali National Park in Alaska, which experiences high variation in light availability between the summer and winter seasons.
-The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare \& Coutinho, 2013; Amarasekare \& Savage, 2012), as well as additional life history traits (size, speed, etc.).
-An organism's ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature relative to the temperature at which survival is no longer possible and may be significantly impacted by increases in temperature range (Amarasekare \& Savage, 2012).
-If an organism has an optimal temperature close to their upper limit of survival, a larger range of variation may push the organism to spend more time beyond the optimal temperature.
-This would negatively impact their ability to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.
-
 \hypertarget{duration}{%
@@
 
-The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability.
-Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1.1).
-Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018).
-Performance is a set of whole organism traits (e.g.~assimilation, development, metabolic rate) with the availability to perform dynamic actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017).
-Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience multiple warming exposures during a single day.
-However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful.
-Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes.
-For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Bowden \& Paitz, 2018; Delmas, Prevot-Julliard, Pieau, \& Girondot, 2007).
-Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth depending on the magnitude of variance (Lawson, Vindenes, Bailey, \& Pol, 2015).
-When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Bastille-Rousseau et al., 2018; Dobramysl \& Tauber, 2013; Romanuk \& Kolasa, 2002).
-
 \hypertarget{predictability}{%
@@
 
-FTR can also describe the predictability of the fluctuation period.
-Variability may not have evenly spaced intervals (e.g.~heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time.
-Predictability can also attempt to describe stochastic vs.~autocorrelated variation, however, that will be discussed in a different context below (Figure 1.3).
-
 \hypertarget{link-between-temporal-and-spatial-variability}{%
@@
 
-While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact.
-Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Farhang-Sardroodi, Darooneh, Kohandel, \& Komarova, 2019; Horne \& Schneider, 1995).
-Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, e.g.~Diel vertical migration (DVM).
-In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability (Brierley, 2014; Lampert, 1989).
-However, there are also instances in which the environment is not temporally variable, yet an organism utilizes spatial variability to regulate its behavior, e.g.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).
-
-Though environmental variability can include many different forms, from fluctuations in resource availability
-(Dempster \& Pollard, 1981; Sommer, 1984) to light (Morison, Franzè, Harvey, \& Menden‐Deuer, 2020; Ruel \& Ayres, 1999), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (Colinet, Sinclair, Vernon, \& Renault, 2015; Huey \& Bennett, 1990).
-
 \hypertarget{temperature-variability}{%
@@
 
-Changes in temperature can be grouped into a couple different definitions, the first being the difference between the terms ``fluctuating'' and ``variation'' with respect to temperature.
-Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al (2018) explicitly defined fluctuating temperature as ``a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis.'' Most generally, temperature variability describes changes in environmental temperature through time (Colinet et al., 2018; Holmes, Woollings, Hawkins, \& Vries, 2016).
-
 \hypertarget{why-is-temperature-variability-important-to-ecological-processes}{%
@@
 
-Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey \& Bennett, 1990).
-Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE).
-Though highly debated (Clarke, 2006), MTE serves as an origin point for understanding how organisms respond to environmental conditions.
-MTE is described by the following equation:
-
-\[
-B = b_0M^{3/4}e^{-E/kT}
-\]
-
-Where \(B\) is organismal metabolic rate, \(b_0\) is a normalization constant unrelated to temperate or body size, \(M\) is organismal mass, and \(e^-E/kT\) is referred to as the ``Boltzman factor'', where \(E\) is activation energy, \(k\) is the Boltzman constant and \(T\) is temperature.
-Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate.
-We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and for smaller organisms should be lower.
-However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clarke, 2006; Clarke \& Johnston, 1999).
-Others highlight that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al., 2006).
-Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.
-
-MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics.
-This is typically described by a characteristic unimodal, left skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{figures/figure2}
-
-}
-
-\caption[Thermal performance curve components]{Schematic of thermal performance curve and its important components for understanding thermal performance.
-CTmin denotes Critical Thermal Minimum, CTmax denotes Critical Thermal Maximum, and Topt denotes Thermal Optimum.
-Adapted from (Krenek et al 2012, Tuff et al 2016).}\label{fig:unnamed-chunk-4}
-\end{figure}
-The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range.
-CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive (Brown, Gillooly, Allen, Savage, \& West, 2004; Dowd, King, \& Denny, 2015).
-Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.
-
-TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (e.g.~metabolism) (Amarasekare \& Savage, 2012).
-If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after which it would no longer increase and the organism would fail to survive.
-However, there are several other constraining factors (e.g.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs.
-TPCs represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.
-
 \hypertarget{jensens-inequality-and-its-impacts-on-metabolic-rate}{%
@@
 
-TPCs have been crucial in predicting and understanding how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018).
-However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance).
-These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their natural environments.
-
-Given the characteristically nonlinear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions.
-Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel \& Ayres, 1999).
-For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen's inequality (1999).
-The development rate of ectotherms has a characteristically nonlinear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel \& Ayres, 1999).
-With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel \& Ayres, 1999).
-Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time.
-It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al., 2018; Khelifa et al., 2019).
-
-Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms.
-For instance, Amarasekare et al (2012) demonstrated the importance of where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013).
-Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare \& Savage, 2012).
-
 \hypertarget{relevancy-of-thermal-variability-in-the-modern-context}{%
@@
 
-Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate.
-Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al., 2014); thus, if metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance?
-Will increased variation have differing effects on different levels of biological organization?
-Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.
-
-However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 1.3).
-Khelifa et al (2019) demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable.
-Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature.
-Vasseur and Yodzis outline how the frequency (1/period) of environmental fluctuations describes the ``color of environmental noise'' (Vasseur \& Yodzis, 2004).
-In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods.
-Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (Tabari, Hosseinzadeh Talaee, Ezani, \& Shifteh Some'e, 2012; Wigley, 1998).
-Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{figures/figure4}
-
-}
-
-\caption[Patterns of variability and spectral density plots]{Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F).
-Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.}\label{fig:unnamed-chunk-5}
-\end{figure}
-\clearpage
-
 \hypertarget{objectives-driving-questions-hypotheses}{%
@@
 
-Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
-\begin{itemize}
-\item
-  At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, from flattening the thermal performance curve (Bartheld, Artacho, \& Bacigalupe, 2017) to changes in TSD ratios (Bowden \& Paitz, 2018)
-\item
-  At the population level, this usually comes in the form of decreasing population densities (Anders \& Post, 2006)
-\item
-  At the community level, this usually comes in the form of promoting coexistence in the face of emigration (Descamps-Julien \& Gonzalez, 2005), or population success via inflationary effects (Holt et al., 2003)
-\end{itemize}
-It is, however, less understood how variation patterns influence these characteristic responses at each level of organization.
-Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas studies at higher levels of biological organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).
-
-There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at the significance of variability on ecosystem dynamics (Kroeker et al., 2020; Massey, Holt, Brooks, \& Rollinson, 2019).
-To understand how environmental variability affects organisms, populations, and communities, we conducted a meta-analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude?
-(i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (e.g.~periodicity, age, size, thermal history, etc.)?
-We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.
-
-We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response, 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will exhibit increased performance and smaller organisms will exhibit decreased performance with increased thermal variability, 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics, 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted.
-
 \hypertarget{methods}{%
 \chapter{Methods}\label{methods}}
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{figures/figure5}
-
-}
-
-\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 15 studies included in the analysis.
-Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
-\end{figure}
-We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1).
-After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*)) AND SU=(Life Sciences \& Biomedicine).
-This search returned a total of 176 results.
-We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1).
-Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 2.1).
-Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
-\begin{itemize}
-\tightlist
-\item
-  We excluded studies that did not feature both a constant and fluctuating treatment (e.g.~studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
-\item
-  We excluded studies that were not completed in a laboratory (e.g.~modeling predictions or results from models, mesocosm experiments, \emph{in situ} experiments)
-\item
-  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a TPC)
-\item
-  We excluded any articles not written in English
-\item
-  We excluded studies that did not report measurements of error
-\item
-  We excluded studies in which the response variables were not comparable or extractable (e.g.~one study performed Principal Component Analysis for their responses)
-\end{itemize}
-We excluded studies that did not have the same mean for constant and fluctuating treatments (e.g.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments.
-We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude most simply to variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004).
-The final number of data points included in the analysis is broken down by study (Table 2.1).
-
-\clearpage
-\begin{table}[!h]
-
-\caption[Studies included in analysis of thermal variability]{\label{tab:unnamed-chunk-8}Studies included in analysis of thermal variability (n=15)}
-\centering
-\begin{tabular}[t]{lll}
-\toprule
-\textbf{Study} & \textbf{Year} & \textbf{Extracted data points (n)}\\
-\midrule
-\cellcolor{gray!6}{Delava et al} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{16}\\
-Du et al & 2009 & 32\\
-\cellcolor{gray!6}{Garcia-Ruiz et al} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{16}\\
-Glass et al & 2009 & 16\\
-\cellcolor{gray!6}{Kern et al} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{24}\\
-\addlinespace
-Kern et al & 2014 & 4\\
-\cellcolor{gray!6}{Lowenborg et al} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{10}\\
-Maneti et al & 2014 & 24\\
-\cellcolor{gray!6}{Pendlebury et al} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{16}\\
-Qu et al & 2014 & 40\\
-\addlinespace
-\cellcolor{gray!6}{Rolandi et al} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{4}\\
-Saxon et al & 2017 & 12\\
-\cellcolor{gray!6}{Semenov et al} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{32}\\
-Simoncini et al & 2019 & 30\\
-\cellcolor{gray!6}{Treidel et al} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{4}\\
-\bottomrule
-\end{tabular}
-\end{table}
+
+Placeholder
+
 \hypertarget{data-extraction-and-analysis}{%
@@
 
-From the studies that met our inclusion criteria, we extracted data from tables and figures.
-Figures were extracted via .png files imported into WebPlotDigitizer to obtain the raw data.
-The extracted data was then imported into a master spreadsheet of data from all studies for analysis.
-We extracted mean values for response variables, any measure of variance (Standard Deviation (SD) or Standard Error of the Mean (SEM)), and sample size.
-For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM.
-We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal stress, and aggregated these values (Table 2.2).
-
-If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and a subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Higgins et al., 2011).
-Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, \emph{metafor}.
-
-\clearpage
-\begin{table}[!h]
-
-\caption{\label{tab:unnamed-chunk-9}Covariates included in models}
-\centering
-\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-\textbf{Variable} & \textbf{Description}\\
-\midrule
-\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature (°C) in study}\\
-flux\_range & Reported range (difference between minimum 
-                                     and maximum) of temperature fluctuation (°C) experienced\\
-org\_level & Level of biological organization, categorized
-                                     as either 0 for individual, 1 for population level
-\cellcolor{gray!6}{                                     responses}\\
-size & Size estimate of study organism, 0 for extra-small (less than 1mm),
-                                     1 for small (less than 5 cm), 2 for medium (less than 0.3 m), and 3
-                                     for large (greater than 0.3 m)\\
-exp\_age & Life stage at which experiment was conducted, 
-                                     categorized as 0 for larval, 1 for juvenile (not sexually mature),
-\cellcolor{gray!6}{                                     and 2 for adult}\\
-\addlinespace
-resp\_type & Response variable units classified as rates or traits, 
-                                     such that rates included any response variable units with a
-                                     “per day” clarifier (e.g. growth rate), all else as traits (e.g. body
-                                     width)\\
-stressful & Whether or not the organisms experienced thermally stressful 
-                                     temperatures throughout study duration, if temperature range 
-                                     exceeded Topt, classified as thermally stressful, else not
-                                     thermally stressful. See Appendix Table 1 for references in which
-\cellcolor{gray!6}{                                     organisms’ thermal stress metrics were included}\\
-\bottomrule
-\end{tabular}
-\end{table}
-\clearpage
-
 \hypertarget{metafor}{%
@@
 
-We used \emph{metafor} (version 2.4.0) in R to analyze the extracted, aggregated data.
-\emph{Metafor} builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer, 2010).
-In order to use \emph{metafor}, the package requires the data to be in a wider format, in which data are input into the following columns:
-\begin{table}[!h]
-
-\caption{\label{tab:unnamed-chunk-10}Column names expanded to wider format for data analysis in metafor}
-\centering
-\begin{tabular}[t]{ll}
-\toprule
-\textbf{Column name} & \textbf{Purpose}\\
-\midrule
-\cellcolor{gray!6}{Mf} & \cellcolor{gray!6}{Mean of experimental group}\\
-SDf & Standard deviation of the experimental group\\
-\cellcolor{gray!6}{Mc} & \cellcolor{gray!6}{Mean of the control group}\\
-SDc & Standard deviation of the control group\\
-\cellcolor{gray!6}{Nf} & \cellcolor{gray!6}{Sample size (n) of the experimental group}\\
-\addlinespace
-Nc & Sample size (n) of the control group\\
-\bottomrule
-\end{tabular}
-\end{table}
-\linebreak
-
-Formatting the data in this way allows for metafor to calculate an effect size and sampling variance for each row in the dataset, such that a constant treatment and corresponding fluctuating treatment will be paired together using the \texttt{escalc} function in metafor to calculate effect size and sampling variance:
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{escalc}\NormalTok{(}\DataTypeTok{measure=}\StringTok{"SMD"}\NormalTok{, }\DataTypeTok{m1i=}\StringTok{`}\DataTypeTok{constant_resp}\StringTok{`}\NormalTok{, }\DataTypeTok{m2i=}\NormalTok{flux_resp, }
-            \DataTypeTok{sd1i=}\StringTok{`}\DataTypeTok{SD_constant}\StringTok{`}\NormalTok{, }\DataTypeTok{sd2i=} \StringTok{`}\DataTypeTok{SD_variable}\StringTok{`}\NormalTok{, }\DataTypeTok{n1i=}\NormalTok{constant_samp, }
-       \DataTypeTok{n2i=}\NormalTok{flux_samp, }\DataTypeTok{data=}\NormalTok{dat_MA, }\DataTypeTok{slab=}\KeywordTok{paste}\NormalTok{(study_id, experiment_id, }
-\NormalTok{                                              response_id, }\DataTypeTok{sep=}\StringTok{", "}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-We specified the effect size metric of interest to be the standardized mean difference (SMD), as this is a common calculation used to compare two groups head to head (Viechtbauer, 2010).
-It is also a powerful metric to use in a meta-analysis because it standardizes the responses across studies to reduce heterogeneity and bias (Viechtbauer, 2010).
-Each of the subsequent arguments of the \texttt{escalc} function specify the mean, standard deviation, and sample size for each paired constant and fluctuating treatment.
-The \texttt{escalc} function uses multiple equations to ultimately output the SMD and sampling variance, all of which are below (Figure 2.2).
-\begin{figure}
-\includegraphics[width=1\linewidth]{figures/figure6} \caption[Workflow of analysis in metafor]{Workflow diagram of how data were extracted, analyzed, and processed for downstream analysis in metafor.}\label{fig:unnamed-chunk-12}
-\end{figure}
-First, it calculated the pooled standard deviation (Pooled SD) from the sample sizes of both studies.
-Our data had studies with a wide range of sample sizes, so we wanted to include separate sample sizes between the constant and fluctuating treatments.
-The Pooled SD, Cohen's d, which is one of the most common ways to measure effect size, was used to normalize the mean differences between constant and fluctuating treatments.
-Then, using a correction factor J, which again normalized the data further, we obtained the final SMD.
-The SMD was then ultimately used to calculate the sampling variance, which created a unitless metric for variance.
-
-Both the SMD and sampling variance were then input into the \texttt{rma.mv} function in \emph{metafor} to generate a multivariate random effects model to understand how variability compares across studies and what factors may influence the magnitude of response.
-We first ran a multivariate random effects model without modifiers (simple model) to establish baseline heterogeneity between studies, and then examine modifiers that could better explain the heterogeneity in effect sizes across studies:
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{rma.mv}\NormalTok{(yi, vi, }
-       \DataTypeTok{data=}\NormalTok{dat_MA_ES, }
-       \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{ }\NormalTok{study_id }\OperatorTok{/}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id, }
-       \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-The arguments of the \texttt{rma.mv} function include SMD (yi), sampling variance (vi), the datafile, as well the random argument to specify the structure of the model.
-We specified a nested structure to account for the colinearity of data points from within the same study, experiment, and response metric.
-We used Restricted Maximum-Likelihood (REML) to specify the fit of the model, as REML is best at handling small samples sizes compared to Maximum Likelihood (ML) (Viechtbauer, 2010).
-We chose a random effects model as opposed to a fixed effects model because we were interested in better understanding the patterns of the larger population of environmental variability literature, as we assumed the studies that were included in this analysis to be a random selection from the true population (Viechtbauer, 2010).
-The random effects model structure also allowed us to account for the nested structure of responses, experiments, and studies to address the dependence of observations from within each study.
-The multivariate random effects model with all modifiers (full model) we ran included:
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{rma.mv}\NormalTok{(yi, vi, }\DataTypeTok{data=}\NormalTok{dat_MA_ES, }\DataTypeTok{mods =} \OperatorTok{~}\NormalTok{flux_range }\OperatorTok{*}\StringTok{ }\NormalTok{mean_temp_constant }
-       \OperatorTok{+}\StringTok{ }\NormalTok{exp_age }\OperatorTok{+}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{org_level }\OperatorTok{+}\StringTok{ }\NormalTok{resp_type, }
-               \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{  }\NormalTok{study_id}\OperatorTok{/}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id,}
-                 \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-The full model also included an interaction term between mean temperature and fluctuation range to understand if mean temperature coupled with fluctuation range was an important component in explaining patterns of variation.
-If an organism experiences large fluctuations in temperature range at a high mean temperature, it is more likely to exceed the organisms' Topt and become thermally stressful.
-
-We decided to exclude the thermally stressful covariate as an argument in the full model because we lacked thermal stress data for one study.
-That particular study happened to be very influential in our model but the data was not misrepresented, as the data was extracted from a table (Garcia-Ruiz et al 2011, Table 1).
-The multivariate random effects model with all the covariates in the full model as well as thermal stress (exclusion model), which excluded Garcia-Ruiz et al 2011 entirely, included:
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{rma.mv}\NormalTok{(yi, vi, }\DataTypeTok{data=}\NormalTok{dat_MA_ES, }\DataTypeTok{mods =} \OperatorTok{~}\NormalTok{flux_range }\OperatorTok{*}\StringTok{ }\NormalTok{mean_temp_constant }
-       \OperatorTok{+}\StringTok{ }\NormalTok{exp_age }\OperatorTok{+}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{org_level }\OperatorTok{+}\StringTok{ }\NormalTok{resp_type }\OperatorTok{+}\StringTok{ }\NormalTok{stressful, }
-               \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{  }\NormalTok{study_id}\OperatorTok{/}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id,}
-                 \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-We also ran a separate model without Garcia-Ruiz et al 2011 and included the thermal stress covariate separately from all the other modifiers (thermal stress model) to see if thermal stress was truly an important factor or if it was simply excluding those influential data points:
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{rma.mv}\NormalTok{(yi, vi, }\DataTypeTok{data=}\NormalTok{dat_MA_ES, }\DataTypeTok{mods =}  \OperatorTok{~}\StringTok{ }\NormalTok{stressful, }
-               \DataTypeTok{random =} \OperatorTok{~}\DecValTok{1} \OperatorTok{|}\StringTok{ }\NormalTok{experiment_id}\OperatorTok{/}\StringTok{ }\NormalTok{study_id}\OperatorTok{/}\StringTok{ }\NormalTok{response_id,}
-                 \DataTypeTok{method=}\StringTok{"REML"}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
 \hypertarget{results}{%
@@
 
-{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-18-1}
+{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-2-1}
 
@@
 Data indicate that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study.
-Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-18}
+Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-19-1}
+{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-3-1}
 
@@
 Error bars represent 95 percent confidence intervals.
-Yellow indicates a positive coefficient estimate, purple indicates a negative coefficient estimate.}\label{fig:unnamed-chunk-19}
+Yellow indicates a positive coefficient estimate, purple indicates a negative coefficient estimate.}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 
-\caption[Thermal stress model summary statistics]{\label{tab:unnamed-chunk-20}Model summary statistics for random effects model using thermal stress as sole modifier}
+\caption[Thermal stress model summary statistics]{\label{tab:unnamed-chunk-4}Model summary statistics for random effects model using thermal stress as sole modifier}
 \centering
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-21-1}
+{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-5-1}
 
@@
 Temperature fluctuation ranged from 0-20°C, mean temperature ranged from 7-33°C.
-SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-21}
+SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-22-1}
-
-}
-
-\caption[Effect sizes across temperature range by organization level]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-22}
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-6-1}
+
+}
+
+\caption[Effect sizes across temperature range by organization level]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-23-1}
-
-}
-
-\caption[Effect sizes across temperature range by thermal stress]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by thermal stress.}\label{fig:unnamed-chunk-23}
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-7-1}
+
+}
+
+\caption[Effect sizes across temperature range by thermal stress]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by thermal stress.}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-24-1}
-
-}
-
-\caption[Effect sizes across temperature range by response type]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-24}
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-8-1}
+
+}
+
+\caption[Effect sizes across temperature range by response type]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-25-1}
-
-}
-
-\caption[Effect sizes across temperature range by body size]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-25}
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-9-1}
+
+}
+
+\caption[Effect sizes across temperature range by body size]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-26-1}
-
-}
-
-\caption[Effect sizes across temperature range by experimental age]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-26}
+{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-10-1}
+
+}
+
+\caption[Effect sizes across temperature range by experimental age]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 These findings underscore the significance of better understanding why differences across studies and experimental designs occurred.
-Jensen's inequality is one way to conceptualize why responses differ, as averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance (Bernhardt et al., 2018; Ruel \& Ayres, 1999).
+Jensen's inequality is one way to conceptualize why responses differ, as averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance (Bernhardt, Sunday, Thompson, \& O'Connor, 2018; Ruel \& Ayres, 1999).
 The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive.
@@
 This may again be the result of studies not focused on the larval stage of development, but instead on the juvenile stage.
-However, there have been meta-analyses, both qualitative (Massey et al., 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptile development, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
+However, there have been meta-analyses, both qualitative (Massey, Holt, Brooks, \& Rollinson, 2019), and quantitative (Noble, Stenhouse, \& Schwanz, 2018), that explicitly looked at the effects of incubation temperature on reptile development, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response (Noble et al., 2018).
 These results align with patterns of juvenile organisms (e.g.~fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance (Moyano et al., 2017; Portner \& Farrell, 2008).
@@
 Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern.
-Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly ecologically irrelevant thermal regimes (Colinet et al., 2018, 2015).
+Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly ecologically irrelevant thermal regimes (Colinet, Rinehart, Yocum, \& Greenlee, 2018; Colinet, Sinclair, Vernon, \& Renault, 2015).
 Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.
@@
 
-\appendix
-
 \hypertarget{appendix}{%
@@
 \addcontentsline{toc}{chapter}{Appendix}
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-29-1}
-
-}
-
-\caption[Funnel plot]{Funnel plot displaying the extracted data points and the distribution of their residuals centered around zero.}\label{fig:unnamed-chunk-29}
-\end{figure}
-\clearpage
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-30-1}
-
-}
-
-\caption[Histogram of effect sizes]{Distribution of effect sizes (n=140) across data from all 15 studies.}\label{fig:unnamed-chunk-30}
-\end{figure}
-\begin{table}[!h]
-
-\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis and scientific classification at the class level}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}ll>{\raggedright\arraybackslash}p{3cm}}
-\toprule
-\textbf{Species} & \textbf{Topt (°C)} & \textbf{Class} & \textbf{Reference}\\
-\midrule
-\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6°C} & \cellcolor{gray!6}{Insecta} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
-Leptopilina boulardi & 25°C & Insecta & De lava et al 2016\\
-\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 ± 0.64 °C} & \cellcolor{gray!6}{Reptilia} & \cellcolor{gray!6}{Dang et al 2019}\\
-Gryllus firmus & 30°C & Insecta & Singh et al 2020\\
-\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30°C} & \cellcolor{gray!6}{Amphibia} & \cellcolor{gray!6}{Seebacher et al 2014}\\
-\addlinespace
-Limnodynastes tasmaniensis & 30-33°C & Amphibia & Whitehead et al 1989\\
-\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33°C} & \cellcolor{gray!6}{Amphibia} & \cellcolor{gray!6}{Kern et al 2014}\\
-Natrix natrix & 30°C & Reptilia & Issac 1997\\
-\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26°C} & \cellcolor{gray!6}{Aves} & \cellcolor{gray!6}{Alagawany et al 2017}\\
-Plestiodon chinensis & 33°C & Reptilia & Baojun et al 2010\\
-\addlinespace
-\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35°C} & \cellcolor{gray!6}{Insecta} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
-Escherichia coli & 37°C, 35-36°C & Gammaproteobacteria & Doyle and Schoeni 1984, Bronikowski et al 2001\\
-\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36°C} & \cellcolor{gray!6}{Gammaproteobacteria} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
-Caiman latirostris & 32-33°C & Reptilia & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
-\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33°C} & \cellcolor{gray!6}{Reptilia} & \cellcolor{gray!6}{Dang et al 2019}\\
-\addlinespace
-Esteya vermicola & 26°C & Not assigned; Fungus & Wang et al 2015\\
-\cellcolor{gray!6}{Xylotrechus arvicola} & \cellcolor{gray!6}{N/A} & \cellcolor{gray!6}{Insecta} & \cellcolor{gray!6}{Garcia-Ruiz et al 2011}\\
-\bottomrule
-\end{tabular}
-\end{table}
-\backmatter
+
+Placeholder
 
@@
 
-\markboth{References}{References}
-
-\noindent
-
-\setlength{\parindent}{-0.20in}
-\setlength{\leftskip}{0.20in}
-\setlength{\parskip}{8pt}
+Placeholder
 
@@
 
-\leavevmode\hypertarget{ref-amarasekare_intrinsic_2013}{}%
-Amarasekare, P., \& Coutinho, R.
-M.
-(2013).
-The intrinsic growth rate as a predictor of population viability under climate warming.
-\emph{J Anim Ecol}, \emph{82}(6), 1240--1253.
-\url{http://doi.org/10.1111/1365-2656.12112}
-
 \leavevmode\hypertarget{ref-amarasekare_framework_2012}{}%
 Amarasekare, P., \& Savage, V. (2012). A framework for elucidating the temperature dependence of fitness. \emph{The American Naturalist}, \emph{179}(2), 178--191. \url{http://doi.org/10.1086/663677}
-
-\leavevmode\hypertarget{ref-anders_distribution-wide_2006}{}%
-Anders, A.
-D., \& Post, E.
-(2006).
-Distribution-wide effects of climate on population densities of a declining migratory landbird.
-\emph{J Anim Ecology}, \emph{75}(1), 221--227.
-\url{http://doi.org/10.1111/j.1365-2656.2006.01034.x}
-
-\leavevmode\hypertarget{ref-bambach_seafood_1993}{}%
-Bambach, R.
-K.
-(1993).
-Seafood through time: Changes in biomass, energetics, and productivity in the marine ecosystem.
-\emph{Paleobiology}, \emph{19}(3), 372--397.
-\url{http://doi.org/10.1017/S0094837300000336}
 
@@
 \url{http://doi.org/10.3724/SP.J.1245.2014.00197}
-
-\leavevmode\hypertarget{ref-bartheld_thermal_2017-2}{}%
-Bartheld, J.
-L., Artacho, P., \& Bacigalupe, L.
-(2017).
-Thermal performance curves under daily thermal fluctuation: A study in helmeted water toad tadpoles.
-\emph{Journal of Thermal Biology}, \emph{70}, 80--85.
-\url{http://doi.org/10.1016/j.jtherbio.2017.09.008}
-
-\leavevmode\hypertarget{ref-bastille-rousseau_climate_2018}{}%
-Bastille-Rousseau, G., Schaefer, J.
-A., Peers, M.
-J.
-L., Ellington, E.
-H., Mumma, M.
-A., Rayl, N.
-D., \ldots{} Murray, D.
-L.
-(2018).
-Climate change can alter predator--prey dynamics and population viability of prey.
-\emph{Oecologia}, \emph{186}(1), 141--150.
-\url{http://doi.org/10.1007/s00442-017-4017-y}
-
-\leavevmode\hypertarget{ref-bernardo_biologically_2014}{}%
-Bernardo, J.
-(2014).
-Biologically grounded predictions of species resistance and resilience to climate change.
-\emph{Proceedings of the National Academy of Sciences}, \emph{111}(15), 5450--5451.
-\url{http://doi.org/10.1073/pnas.1404505111}
 
@@
 
-\leavevmode\hypertarget{ref-blankenship_how_2017}{}%
-Blankenship, R. E. (2017). How cyanobacteria went green. \emph{Science}, \emph{355}(6332), 1372--1373. \url{http://doi.org/10.1126/science.aam9365}
-
-\leavevmode\hypertarget{ref-bouchard_ecosystem_2014}{}%
-Bouchard, F. (2014). Ecosystem evolution is about variation and persistence, not populations and reproduction. \emph{Biol Theory}, \emph{9}(4), 382--391. \url{http://doi.org/10.1007/s13752-014-0171-1}
-
-\leavevmode\hypertarget{ref-bowden_temperature_2018}{}%
-Bowden, R.
-M., \& Paitz, R.
-T.
-(2018).
-Temperature fluctuations and maternal estrogens as critical factors for understanding temperature-dependent sex determination in nature.
-\emph{J.
-Exp.
-Zool.}, \emph{329}(4), 177--184.
-\url{http://doi.org/10.1002/jez.2183}
-
-\leavevmode\hypertarget{ref-brierley_diel_2014}{}%
-Brierley, A. S. (2014). Diel vertical migration. \emph{Current Biology}, \emph{24}(22), R1074--R1076. \url{http://doi.org/10.1016/j.cub.2014.08.054}
-
 \leavevmode\hypertarget{ref-bronikowski_evolutionary_2001}{}%
@@
 
-\leavevmode\hypertarget{ref-brown_toward_2004}{}%
-Brown, J. H., Gillooly, J. F., Allen, A. P., Savage, V. M., \& West, G. B. (2004). TOWARD a METABOLIC THEORY OF ECOLOGY. \emph{Ecology}, \emph{85}(7), 1771--1789. \url{http://doi.org/10.1890/03-9000}
-
-\leavevmode\hypertarget{ref-burggren_developmental_2018}{}%
-Burggren, W.
-(2018).
-Developmental phenotypic plasticity helps bridge stochastic weather events associated with climate change.
-\emph{Journal of Experimental Biology}, \emph{221}(9), jeb161984.
-\url{http://doi.org/10.1242/jeb.161984}
-
 \leavevmode\hypertarget{ref-cheung_application_2008}{}%
@@
 
-\leavevmode\hypertarget{ref-delmas_mechanistic_2007}{}%
-Delmas, V., Prevot-Julliard, A.-C., Pieau, C., \& Girondot, M.
-(2007).
-A mechanistic model of temperature-dependent sex determination in a chelonian: The european pond turtle.
-\emph{Funct Ecology}, \emph{0}(0), 071027215958001--???
-\url{http://doi.org/10.1111/j.1365-2435.2007.01349.x}
-
-\leavevmode\hypertarget{ref-dempster_fluctuations_1981}{}%
-Dempster, J. P., \& Pollard, E. (1981). Fluctuations in resource availability and insect populations. \emph{Oecologia}, \emph{50}(3), 412--416. \url{http://doi.org/10.1007/BF00344984}
-
-\leavevmode\hypertarget{ref-descamps-julien_stable_2005}{}%
-Descamps-Julien, B., \& Gonzalez, A.
-(2005).
-STABLE COEXISTENCE IN a FLUCTUATING ENVIRONMENT: AN EXPERIMENTAL DEMONSTRATION.
-\emph{Ecology}, \emph{86}(10), 2815--2824.
-\url{http://doi.org/10.1890/04-1700}
-
-\leavevmode\hypertarget{ref-dobramysl_environmental_2013}{}%
-Dobramysl, U., \& Tauber, U.
-C.
-(2013).
-Environmental vs.
-Demographic variability in two-species predator-prey models.
-\emph{Phys.
-Rev.
-Lett.}, \emph{110}(4), 048105.
-\url{http://doi.org/10.1103/PhysRevLett.110.048105}
-
-\leavevmode\hypertarget{ref-dowd_thermal_2015}{}%
-Dowd, W.
-W., King, F.
-A., \& Denny, M.
-W.
-(2015).
-Thermal variation, thermal extremes and the physiological performance of individuals.
-\emph{Journal of Experimental Biology}, \emph{218}(12), 1956--1967.
-\url{http://doi.org/10.1242/jeb.114926}
-
 \leavevmode\hypertarget{ref-doyle_survival_1984}{}%
@@
 
-\leavevmode\hypertarget{ref-farhang-sardroodi_environmental_2019}{}%
-Farhang-Sardroodi, S., Darooneh, A.
-H., Kohandel, M., \& Komarova, N.
-L.
-(2019).
-Environmental spatial and temporal variability and its role in non-favoured mutant dynamics.
-\emph{J.
-R.
-Soc.
-Interface.}, \emph{16}(157), 20180781.
-\url{http://doi.org/10.1098/rsif.2018.0781}
-
 \leavevmode\hypertarget{ref-fresquet_response_2011}{}%
@@
 \url{http://doi.org/10.1603/EN11080}
-
-\leavevmode\hypertarget{ref-gillooly_response_2006}{}%
-Gillooly, J.
-F., Allen, A.
-P., Savage, V.
-M., Charnov, E.
-L., West, G.
-B., \& Brown, J.
-H.
-(2006).
-Response to clarke and fraser: Effects of temperature on metabolic rate.
-\emph{Funct Ecology}, \emph{20}(2), 400--404.
-\url{http://doi.org/10.1111/j.1365-2435.2006.01110.x}
 
@@
 
-\leavevmode\hypertarget{ref-gudmundson_environmental_2015}{}%
-Gudmundson, S., Eklöf, A., \& Wennergren, U.
-(2015).
-Environmental variability uncovers disruptive effects of species' interactions on population dynamics.
-\emph{Proc.
-R.
-Soc.
-B.}, \emph{282}(1812), 20151126.
-\url{http://doi.org/10.1098/rspb.2015.1126}
-
-\leavevmode\hypertarget{ref-guralnick_evolutionary_2007}{}%
-Guralnick, L.
-J., Cline, A., Smith, M., \& Sage, R.
-F.
-(2007).
-Evolutionary physiology: The extent of c4 and CAM photosynthesis in the genera anacampseros and grahamia of the portulacaceae.
-\emph{Journal of Experimental Botany}, \emph{59}(7), 1735--1742.
-\url{http://doi.org/10.1093/jxb/ern081}
-
-\leavevmode\hypertarget{ref-helmuth_organismal_2010}{}%
-Helmuth, B., Broitman, B.
-R., Yamane, L., Gilman, S.
-E., Mach, K., Mislan, K.
-A.
-S., \& Denny, M.
-W.
-(2010).
-Organismal climatology: Analyzing environmental variability at scales relevant to physiological stress.
-\emph{Journal of Experimental Biology}, \emph{213}(6), 995--1003.
-\url{http://doi.org/10.1242/jeb.038463}
-
-\leavevmode\hypertarget{ref-higgins_cochrane_2011}{}%
-Higgins, J., Thomas, J., Chandler, J., Li, T., Page, M., \& Welch, V.
-(2011).
-\emph{Cochrane handbook for systematic reviews of interventions} (Vol.
-5.1.0).
-Retrieved from \url{www.training.cochrane.org/handbook}
-
-\leavevmode\hypertarget{ref-holmes_robust_2016}{}%
-Holmes, C.
-R., Woollings, T., Hawkins, E., \& Vries, H.
-de.
-(2016).
-Robust future changes in temperature variability under greenhouse gas forcing and the relationship with thermal advection.
-\emph{Journal of Climate}, \emph{29}(6), 2221--2236.
-\url{http://doi.org/10.1175/JCLI-D-14-00735.1}
-
-\leavevmode\hypertarget{ref-holt_impacts_2003}{}%
-Holt, R.
-D., Barfield, M., \& Gonzalez, A.
-(2003).
-Impacts of environmental variability in open populations and communities: ``Inflation'' in sink environments.
-\emph{Theoretical Population Biology}, \emph{64}(3), 315--330.
-\url{http://doi.org/10.1016/S0040-5809(03)00087-X}
-
-\leavevmode\hypertarget{ref-horne_spatial_1995}{}%
-Horne, J. K., \& Schneider, D. C. (1995). Spatial variance in ecology. \emph{Oikos}, \emph{74}(1), 18. \url{http://doi.org/10.2307/3545670}
-
-\leavevmode\hypertarget{ref-huey_behavioral_1974}{}%
-Huey, R. B. (1974). Behavioral thermoregulation in lizards: Importance of associated costs. \emph{Science}, \emph{184}(4140), 1001--1003. \url{http://doi.org/10.1126/science.184.4140.1001}
-
-\leavevmode\hypertarget{ref-huey_physiological_1990}{}%
-Huey, R. B., \& Bennett, A. F. (1990). Physiological adjustments to fluctuating thermal environments: An ecological and evolutionary perspective. \emph{Stress Proteins in Biology and Medicine}, 23.
-
-\leavevmode\hypertarget{ref-huey_predicting_2012}{}%
-Huey, R.
-B., Kearney, M.
-R., Krockenberger, A., Holtum, J.
-A.
-M., Jess, M., \& Williams, S.
-E.
-(2012).
-Predicting organismal vulnerability to climate warming: Roles of behaviour, physiology and adaptation.
-\emph{Phil.
-Trans.
-R.
-Soc.
-B}, \emph{367}(1596), 1665--1679.
-\url{http://doi.org/10.1098/rstb.2012.0005}
-
-\leavevmode\hypertarget{ref-husak_how_2017}{}%
-Husak, J.
-F., \& Lailvaux, S.
-P.
-(2017).
-How do we measure the cost of whole-organism performance traits?
-\emph{Integrative and Comparative Biology}, \emph{57}(2), 333--343.
-\url{http://doi.org/10.1093/icb/icx048}
-
 \leavevmode\hypertarget{ref-isaac_isaac_2003pdf_2003}{}%
 Isaac, L. A. (2003). Isaac\_2003.pdf. University of Victoria. Retrieved from \url{https://dspace.library.uvic.ca/handle/1828/356}
-
-\leavevmode\hypertarget{ref-kearney_potential_2009}{}%
-Kearney, M., Shine, R., \& Porter, W.
-P.
-(2009).
-The potential for behavioral thermoregulation to buffer "cold-blooded" animals against climate warming.
-\emph{Proceedings of the National Academy of Sciences}, \emph{106}(10), 3835--3840.
-\url{http://doi.org/10.1073/pnas.0808913106}
 
@@
 \url{http://doi.org/10.1111/evo.12221}
-
-\leavevmode\hypertarget{ref-knoll_chapter_2008}{}%
-Knoll, A. (2008). Chapter 1. In \emph{Cyanobacteria and earth history}.
 
@@
 
-\leavevmode\hypertarget{ref-kroeker_ecological_2020}{}%
-Kroeker, K.
-J., Bell, L.
-E., Donham, E.
-M., Hoshijima, U., Lummis, S., Toy, J.
-A., \& Willis‐Norton, E.
-(2020).
-Ecological change in dynamic environments: Accounting for temporal environmental variability in studies of ocean change biology.
-\emph{Glob Change Biol}, \emph{26}(1), 54--67.
-\url{http://doi.org/10.1111/gcb.14868}
-
-\leavevmode\hypertarget{ref-lampert_adaptive_1989}{}%
-Lampert, W. (1989). The adaptive significance of diel vertical migration of zooplankton. \emph{Functional Ecology}, \emph{3}(1), 21. \url{http://doi.org/10.2307/2389671}
-
-\leavevmode\hypertarget{ref-lawson_environmental_2015}{}%
-Lawson, C.
-R., Vindenes, Y., Bailey, L., \& Pol, M.
-van de.
-(2015).
-Environmental variation and population responses to global change.
-\emph{Ecol Lett}, \emph{18}(7), 724--736.
-\url{http://doi.org/10.1111/ele.12437}
-
 \leavevmode\hypertarget{ref-lowenborg_how_2012}{}%
@@
 
-\leavevmode\hypertarget{ref-morison_light_2020}{}%
-Morison, F., Franzè, G., Harvey, E., \& Menden‐Deuer, S.
-(2020).
-Light fluctuations are key in modulating plankton trophic dynamics and their impact on primary production.
-\emph{Limnol Oceanogr}, \emph{5}(5), 346--353.
-\url{http://doi.org/10.1002/lol2.10156}
-
 \leavevmode\hypertarget{ref-moyano_effects_2017}{}%
@@
 
-\leavevmode\hypertarget{ref-paaijmans_temperature_2013}{}%
-Paaijmans, K.
-P., Heinig, R.
-L., Seliga, R.
-A., Blanford, J.
-I., Blanford, S., Murdock, C.
-C., \& Thomas, M.
-B.
-(2013).
-Temperature variation makes ectotherms more sensitive to climate change.
-\emph{Glob Change Biol}, \emph{19}(8), 2373--2380.
-\url{http://doi.org/10.1111/gcb.12240}
-
 \leavevmode\hypertarget{ref-parachu_marco_new_2017}{}%
@@
 
-\leavevmode\hypertarget{ref-parepa_environmental_2013}{}%
-Parepa, M., Fischer, M., \& Bossdorf, O. (2013). Environmental variability promotes plant invasion. \emph{Nat Commun}, \emph{4}(1), 1604. \url{http://doi.org/10.1038/ncomms2632}
-
 \leavevmode\hypertarget{ref-pendlebury_variation_2004-1}{}%
 Pendlebury, C. J. (2004). Variation in temperature increases the cost of living in birds. \emph{Journal of Experimental Biology}, \emph{207}(12), 2065--2070. \url{http://doi.org/10.1242/jeb.00999}
-
-\leavevmode\hypertarget{ref-petchey_environmental_2000}{}%
-Petchey, O.
-L.
-(2000).
-Environmental colour affects aspects of single--species population dynamics.
-\emph{Proc.
-R.
-Soc.
-Lond.
-B}, \emph{267}(1445), 747--754.
-\url{http://doi.org/10.1098/rspb.2000.1066}
 
@@
 
-\leavevmode\hypertarget{ref-reed_phenotypic_2010}{}%
-Reed, T.
-E., Waples, R.
-S., Schindler, D.
-E., Hard, J.
-J., \& Kinnison, M.
-T.
-(2010).
-Phenotypic plasticity and population viability: The importance of environmental predictability.
-\emph{Proc.
-R.
-Soc.
-B.}, \emph{277}(1699), 3391--3400.
-\url{http://doi.org/10.1098/rspb.2010.0771}
-
 \leavevmode\hypertarget{ref-rolandi_costs_2018}{}%
@@
 \url{http://doi.org/10.1016/j.jtherbio.2018.03.022}
-
-\leavevmode\hypertarget{ref-romanuk_environmental_2002}{}%
-Romanuk, T.
-N., \& Kolasa, J.
-(2002).
-Environmental variability alters the relationship between richness and variability of community abundances in aquatic rock pool microcosms.
-\emph{Écoscience}, \emph{9}(1), 55--62.
-\url{http://doi.org/10.1080/11956860.2002.11682690}
 
@@
 
-\leavevmode\hypertarget{ref-sommer_paradox_1984}{}%
-Sommer, U.
-(1984).
-The paradox of the plankton: Fluctuations of phosphorus availability maintain diversity of phytoplankton in flow-through cultures1.
-\emph{Limnol.
-Oceanogr.}, \emph{29}(3), 633--636.
-\url{http://doi.org/10.4319/lo.1984.29.3.0633}
-
-\leavevmode\hypertarget{ref-soo_origins_2017}{}%
-Soo, R.
-M., Hemp, J., Parks, D.
-H., Fischer, W.
-W., \& Hugenholtz, P.
-(2017).
-On the origins of oxygenic photosynthesis and aerobic respiration in cyanobacteria.
-\emph{Science}, \emph{355}(6332), 1436--1440.
-\url{http://doi.org/10.1126/science.aal3794}
-
-\leavevmode\hypertarget{ref-stewart_mesocosm_2013}{}%
-Stewart, R.
-I., Dossena, M., Bohan, D.
-A., Jeppesen, E., Kordas, R.
-L., Ledger, M.
-E., \ldots{} Woodward, G.
-(2013).
-Mesocosm experiments as a tool for ecological climate-change research.
-In \emph{Advances in ecological research} (Vol.
-48, pp.
-71--181).
-Elsevier.
-\url{http://doi.org/10.1016/B978-0-12-417199-2.00002-1}
-
 \leavevmode\hypertarget{ref-sunday_thermal_2012}{}%
@@
 \url{http://doi.org/10.1038/nclimate1539}
-
-\leavevmode\hypertarget{ref-tabari_shift_2012}{}%
-Tabari, H., Hosseinzadeh Talaee, P., Ezani, A., \& Shifteh Some'e, B.
-(2012).
-Shift changes and monotonic trends in autocorrelated temperature series over iran.
-\emph{Theor Appl Climatol}, \emph{109}(1), 95--108.
-\url{http://doi.org/10.1007/s00704-011-0568-8}
 
@@
 
-\leavevmode\hypertarget{ref-vasseur_phase-locking_2009}{}%
-Vasseur, D.
-A., \& Fox, J.
-W.
-(2009).
-Phase-locking and environmental fluctuations generate synchrony in a predator--prey community.
-\emph{Nature}, \emph{460}(7258), 1007--1010.
-\url{http://doi.org/10.1038/nature08208}
-
-\leavevmode\hypertarget{ref-vasseur_color_2004}{}%
-Vasseur, D. A., \& Yodzis, P. (2004). THE COLOR OF ENVIRONMENTAL NOISE. \emph{Ecology}, \emph{85}(4), 1146--1152. \url{http://doi.org/10.1890/02-3122}
-
-\leavevmode\hypertarget{ref-viechtbauer_conducting_2010}{}%
-Viechtbauer, W. (2010). Conducting meta-analyses in \emph{r} with the \textbf{metafor} package. \emph{J. Stat. Soft.}, \emph{36}(3). \url{http://doi.org/10.18637/jss.v036.i03}
-
 \leavevmode\hypertarget{ref-wang_effects_2015}{}%
@@
 
-\leavevmode\hypertarget{ref-wigley_anthropogenic_1998-1}{}%
-Wigley, T.
-M.
-(1998).
-Anthropogenic influence on the autocorrelation structure of hemispheric-mean temperatures.
-\emph{Science}, \emph{282}(5394), 1676--1679.
-\url{http://doi.org/10.1126/science.282.5394.1676}
-
 \leavevmode\hypertarget{ref-yoshida_rapid_2003}{}%

### 2021-06-30 data extraction /scripts /scopus405_screening.R
recloning repo to try and fix things....

@@ -1,8 +1,0 @@
-library(revtools)
-
-scopus_405<-read_bibliography("scopus_405.bib")
-screen_abstracts(scopus_405)
-
-
-scopus_233<-read_bibliography("scopus_233.bib")
-screen_abstracts(scopus_233)

### 2021-06-30 thesis writing/background figures /generation_time_backgroundfig.R

@@ -1,35 +1,0 @@
-library(tidyverse)
-library(cowplot)
-library(png)
-
-
-#generating generation time data 
-a<-sample(0.5:10, 80, replace = T) 
-b<-sample(10:100, 35, replace = T) 
-c<-sample(100:5000, 10, replace =T)
-generation_time<-c(a, b, c)
-d<-sample(1:10, 100, replace = T)
-e<-sample(10:30, 20, replace = T)
-f<-sample(30:90, 5, replace = T)
-flucuation_period <- c(f, e, d)
-phase<-data.frame(generation_time, flucuation_period)
-p<-ggplot(phase, aes(y=flucuation_period, x=generation_time))+
-  geom_point(alpha=0.8, color ="grey25")+
-  scale_x_log10()+
-  scale_y_log10()+
-  geom_vline(xintercept=1, color = "powderblue", size =2)+
-  geom_vline(xintercept=10, color = "darkseagreen3", size =2)+
-  geom_vline(xintercept=100, color ="lightslateblue", size = 2)+
-  geom_vline(xintercept=4000, color ="hotpink3", size = 2)+
-  geom_hline(yintercept = 1, color = "yellow3", size = 20, alpha=0.4)+
-  geom_hline(yintercept = 10, color ="gray53", size =10, alpha=0.4)+
-  geom_hline(yintercept = 8.5, color ="firebrick3", size =10, alpha=0.4)+
-  xlab("Generation time (days)")+
-  ylab("Fluctuation period (days)")+
-  theme_bw(base_size=18)
-  
- 
-
-  
-
-

### 2021-06-30 meeting materials/thesis meeting 9:21:20.Rmd
i don't even know what i've done...

--- 2020-09-20 thesis meeting sept21.Rmd
@@
 
-```{r, echo=FALSE}
-knitr::include_graphics("images/kroeker_fig1.png")
-```
-
-**Option 2:** Look exclusively at diurnal fluctuations of 3 different environmental variables (temp, pH, salinity, DO, light, etc.) across all levels of organization
+**Option 2:** Look exclusively at diurnal fluctuations of 3 different environmental variables (temp, pH, salinity) across all levels of organization
 
@@
 * Potential outcome: a better understanding of these metrics that covary may not necessarily covary in their responses to variation across different levels of organization?
-```{r, echo=FALSE}
-knitr::include_graphics("images/kroeker_fig1.png")
-```
 

### 2021-06-30 meeting materials/thesis proposal .Rmd
i don't even know what i've done...

--- 2020-09-13 thesis proposal .Rmd
@@
 
-Current search term ideas:
+### Current search term ideas:
+
 * "environmental variability" AND ecology AND temperature (1,281 results)

### 2021-06-30 data extraction /scripts /revtools_screeningfiles.R

@@ -1,0 +1,30 @@
+install.packages("revtools")
+library(revtools)
+
+#wos search with 106 papers
+wos1 <- read_bibliography("thesis_papers106.bib")
+view(wos1)
+screen_titles(wos1)
+
+#wos search with 354 papers 
+wos2<-read_bibliography("354.bib")
+screen_titles(wos2)
+
+###read in abstracts for screening
+as <- read_csv(file = "wos354_included.csv")
+as <- as.data.frame(as)
+screen_abstracts(as)
+
+
+#wos search with 176 papers 
+wos176<-read_bibliography("wos176.bib")
+screen_titles(wos176)
+
+wos_176 <-read_csv(file = "wos176_included.csv")
+wos_176_screen <- as.data.frame(wos_176)
+screen_abstracts(wos_176_screen)
+
+#wos search with 2361 papers (part a of e) 
+wos2361a<-read_bibliography("wos2361_a.bib")
+screen_titles(wos2361a)
+

### 2021-06-30 data extraction /scripts /scopus405_screening.R

@@ -1,0 +1,8 @@
+library(revtools)
+
+scopus_405<-read_bibliography("scopus_405.bib")
+screen_abstracts(scopus_405)
+
+
+scopus_233<-read_bibliography("scopus_233.bib")
+screen_abstracts(scopus_233)

### 2021-06-30 thesis writing/background figures /background_figs_series.R

@@ -1,0 +1,81 @@
+install.packages("colorednoise")
+library(colorednoise)
+library(tidyverse)
+library(cowplot)
+####long time series######
+#generating time series for several years 
+red <- colored_noise(timesteps = 365, mean = 0.5, sd = 0.2, phi = 0.9)
+white <-colored_noise(timesteps = 365, mean = 0.5, sd = 0.2, phi = 0)
+blue<-colored_noise(timesteps = 365, mean = 0.5, sd = 0.2, phi = -0.9)
+t<-seq(0, 364, 1)
+diurnal<-sin(t)
+diurnal<-data.frame(t, diurnal)
+#generating time to go along x axis
+time<-c(1:365)
+#combining time and all noise colors into one data frame
+rainbow<-data.frame(red, white, blue, time, diurnal)
+#data tidying from above generation
+rainbow<-rainbow %>% gather(type, temperature_change, red, white, blue, diurnal)
+#filtering out by type for plots 
+red<-rainbow %>% filter(type == "red")
+white<-rainbow %>% filter(type == "white")
+diurnal<-rainbow %>% filter(type == "diurnal")
+#plotting all noise colors across time
+red_series<-ggplot(red, aes(y=temperature_change, x=time))+
+  geom_line(alpha=0.8, color="firebrick3")+
+  theme_bw()
+white_series<-ggplot(white, aes(y=temperature_change, x=time))+
+  geom_line(alpha=0.8, color="gray53")+
+  theme_bw()
+diurnal_series<-ggplot(diurnal, aes(y=temperature_change, x=time))+
+  geom_line(alpha=0.8, color="yellow3")+
+  theme_bw()
+plot_grid(red_series, white_series, diurnal_series,labels=c("A", "B", "C"), ncol=1, nrow=3)
+###spectral density
+### estimate the spectral density for red first 
+mspect <- spectrum(red$temperature_change, plot=FALSE)
+delta <- 1
+specx <- mspect$freq/delta
+specy <- 2*mspect$spec
+spectral_red <- data.frame(specx = mspect$freq/delta, specy = specy)
+ #then make a plot with these calcs on the x andy 
+spec_red<-spectral_red %>% 
+  ggplot(aes(x = specx, y = specy)) + geom_line(color="firebrick3") +
+  geom_smooth(method = "lm", se = FALSE, color = "black") +
[…]

### 2021-06-30 thesis writing/background figures /generation_time_backgroundfig.R

@@ -1,0 +1,35 @@
+library(tidyverse)
+library(cowplot)
+library(png)
+
+
+#generating generation time data 
+a<-sample(0.5:10, 80, replace = T) 
+b<-sample(10:100, 35, replace = T) 
+c<-sample(100:5000, 10, replace =T)
+generation_time<-c(a, b, c)
+d<-sample(1:10, 100, replace = T)
+e<-sample(10:30, 20, replace = T)
+f<-sample(30:90, 5, replace = T)
+flucuation_period <- c(f, e, d)
+phase<-data.frame(generation_time, flucuation_period)
+p<-ggplot(phase, aes(y=flucuation_period, x=generation_time))+
+  geom_point(alpha=0.8, color ="grey25")+
+  scale_x_log10()+
+  scale_y_log10()+
+  geom_vline(xintercept=1, color = "powderblue", size =2)+
+  geom_vline(xintercept=10, color = "darkseagreen3", size =2)+
+  geom_vline(xintercept=100, color ="lightslateblue", size = 2)+
+  geom_vline(xintercept=4000, color ="hotpink3", size = 2)+
+  geom_hline(yintercept = 1, color = "yellow3", size = 20, alpha=0.4)+
+  geom_hline(yintercept = 10, color ="gray53", size =10, alpha=0.4)+
+  geom_hline(yintercept = 8.5, color ="firebrick3", size =10, alpha=0.4)+
+  xlab("Generation time (days)")+
+  ylab("Fluctuation period (days)")+
+  theme_bw(base_size=18)
+  
+ 
+
+  
+
+

### 2021-06-30 shared analysis files /manuscript_datavis_22jun21.Rmd
data wrangling file
model wrangling for meeting with sam and mary

@@
 ---
-title: "manuscript_data_wrangling"
+title: "Meta-analysis Manuscript Wrangling"
 author: "Maggie Slein"
@@
 #load data
-dat_acclim_0<- read_csv("acclimation_full_2june21.csv")
+dat_acclim_0<- read_csv("acclimation_dummytherm_30jun.csv")
 dat_full_var_0 <- read_csv("metafor.csv")
@@
 ```
+# Motivation and Hypotheses
+
+Variability has been identified as important to organismal success and ecosystem dynamics (Vasseur et al 2014).
+To further understand the impacts of variability of performance, we conducted to analyses on data that explicitly accounted for acclimation (Acclimation model) and studies focuses solely on non-linear averaging (Non-linear averaging model).
+
+H-Acclimation: If reared in fluctuating environments, when exposed to different thermal environments, organisms with larger fluctuation ranges will perform worse than those reared in constant environments.
+Additional covariates, such as age, size, and expsoure temperature will be correlated negative responses.
+
+H-Non-linear averaging: Organisms will perform better in constant environments than fluctuating environments.
+Additional covariates, such as age, size, and large fluctuation range will be correlated negative responses.
+
+```{r pressure, echo=FALSE, fig.cap="Conceptual figure demonstrating how acclimation and non-linear averaging account for different aspects of performance.", out.width = '100%'}
+knitr::include_graphics("conceptfig_30jun.png")
+```
+
+\clearpage
 
@@
 
-```{r}
+## Acclimation without modifiers
+```{r, echo=FALSE, message=FALSE}
 simple_acclimation_model <- rma.mv(yi, vi, data=dat_acclim_ES, 
@@
 
-```{r}
+## Acclimation with modifiers
+```{r, echo=FALSE, message=FALSE}
 acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, mods = ~flux_range * mean_temp_reared +
@@
 
-# Neglecting Acclimation
+# Acclimation data split out by CT data and other thermal performance metrics 
+```{r, echo=FALSE, message=FALSE}
+CT_acclim_data <- dat_acclim_ES %>%
+  filter(resp_def %in% c("CTmax", "CT max", "CTmin", "LTmax", "LTmin", "Topt", "Tpref", "Thermal preference", "ULT50", "Scope of thermal tolerance (CTmax - CTmin)"))
 
+curve_acclim_data <- dat_acclim_ES %>%
+  filter(resp_def %in% 
+           c("CCO activity", "CS activity", "LDH activity", "log CO2 production", 
+             "maximal swimming speed", "Maximum burst swim speed", 
+             "Performance breadth (Tbr)", "RMR", "ULT50", "Speed", "Time to heat knockdown",
+             "Ucrit", "Umax"))
+```
+
+## CT aggregated model 
+```{r, echo=FALSE, message=FALSE}
+CT_model <-rma.mv(yi, vi, data=CT_acclim_data, mods = ~flux_range * mean_temp_reared +
+                        exp_age + size + org_level + exposure_temp, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+CT_model
+```
+## Remaining acclimation metrics model 
+```{r, echo=FALSE, message=FALSE}
+curve_model <- rma.mv(yi, vi, data=curve_acclim_data, mods = ~flux_range * mean_temp_reared +
+                        exp_age + size + org_level + exposure_temp, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+curve_model
+```
+
+# Effects of non-linear averaging
 This uses the full dataset from my thesis (with more data from the additional search). Interesting that this time, mean temp is the only significant predictor?
 
-```{r}
+## Non-linear averaging model with no modifiers
+```{r, echo= FALSE, message=FALSE}
 simple_full_var_model <- rma.mv(yi, vi, data=dat_full_var_ES, 
@@
 
-```{r}
+## Non-linear averaging model with modifiers
+```{r, echo = FALSE, message=FALSE}
 full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~flux_range * mean_temp_constant +
@@
 
-```{r}
+# Figures
+```{r, echo = FALSE, message=FALSE}
 dat_acclim_ES %>%
@@
   theme(legend.position="bottom")+
-  ggtitle("acclimation")
+  ggtitle("Acclimation")
 ```
 
-```{r}
+```{r, echo = FALSE, message=FALSE}
 dat_full_var_ES %>%
@@
   theme(legend.position="bottom")+
-  ggtitle("neglecting acclimation (full dataset)")
+  ggtitle("Non-linear averaging")
 ```

### 2021-07-02 shared analysis files /manuscript_datavis_22jun21.Rmd
model selection and meta-analysis plots

@@
 
-## Acclimation without modifiers
+## Acclimation with temperature modifiers
 ```{r, echo=FALSE, message=FALSE}
-simple_acclimation_model <- rma.mv(yi, vi, data=dat_acclim_ES, 
+temp_acclimation_model <- rma.mv(yi, vi, data=dat_acclim_ES, mods=~flux_range * mean_temp_reared,
                random = ~1 | study_id / experiment_id/ response_id,
                  method="REML") 
-simple_acclimation_model
+temp_acclimation_model
 ```
 
-## Acclimation with modifiers
+## Acclimation with trait modifiers
 ```{r, echo=FALSE, message=FALSE}
-acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, mods = ~flux_range * mean_temp_reared +
-                        exp_age + size + org_level + exposure_temp, 
+trait_acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, 
+                                mods = ~exp_age + size + org_level + exposure_temp, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+trait_acclimation_model
+```
+
+## Acclimation with all modifiers
+```{r, echo=FALSE, message=FALSE}
+acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, 
+            mods = ~flux_range * mean_temp_reared + exp_age + size + org_level + exposure_temp, 
                random = ~1 |  study_id/ experiment_id/ response_id,
@@
 
-# Acclimation data split out by CT data and other thermal performance metrics 
-```{r, echo=FALSE, message=FALSE}
-CT_acclim_data <- dat_acclim_ES %>%
-  filter(resp_def %in% c("CTmax", "CT max", "CTmin", "LTmax", "LTmin", "Topt", "Tpref", "Thermal preference", "ULT50", "Scope of thermal tolerance (CTmax - CTmin)"))
-
-curve_acclim_data <- dat_acclim_ES %>%
-  filter(resp_def %in% 
-           c("CCO activity", "CS activity", "LDH activity", "log CO2 production", 
-             "maximal swimming speed", "Maximum burst swim speed", 
-             "Performance breadth (Tbr)", "RMR", "ULT50", "Speed", "Time to heat knockdown",
-             "Ucrit", "Umax"))
-```
-
-## CT aggregated model 
-```{r, echo=FALSE, message=FALSE}
-CT_model <-rma.mv(yi, vi, data=CT_acclim_data, mods = ~flux_range * mean_temp_reared +
-                        exp_age + size + org_level + exposure_temp, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
-                 method="REML")
-CT_model
-```
-## Remaining acclimation metrics model 
-```{r, echo=FALSE, message=FALSE}
-curve_model <- rma.mv(yi, vi, data=curve_acclim_data, mods = ~flux_range * mean_temp_reared +
-                        exp_age + size + org_level + exposure_temp, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
-                 method="REML")
-curve_model
-```
-
 # Effects of non-linear averaging
@@
 
-## Non-linear averaging model with no modifiers
+## Non-linear averaging model with temperature modifiers
 ```{r, echo= FALSE, message=FALSE}
-simple_full_var_model <- rma.mv(yi, vi, data=dat_full_var_ES, 
+temp_full_var_model <- rma.mv(yi, vi, data=dat_full_var_ES, 
+                                mods = ~flux_range * mean_temp_constant,
                random = ~1 | study_id / experiment_id/ response_id,
                  method="REML") 
-simple_full_var_model
+temp_full_var_model
 ```
 
-## Non-linear averaging model with modifiers
+## Non-linear averaging model with trait modifiers
 ```{r, echo = FALSE, message=FALSE}
-full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~flux_range * mean_temp_constant +
-                        exp_age + size + org_level, 
+trait_full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~exp_age + size + org_level, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+trait_full_var_model
+```
+
+## Non-linear averaging model with all modifiers
+```{r, echo = FALSE, message=FALSE}
+full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, 
+                        mods = ~flux_range * mean_temp_constant + exp_age + size + org_level, 
                random = ~1 |  study_id/ experiment_id/ response_id,
@@
 full_var_model
+```
+
+```{r}
+#Model selection
+
+AIC(temp_acclimation_model, trait_acclimation_model, acclimation_model, temp_full_var_model, trait_full_var_model, full_var_model)
+```
+
+```{r}
+normalized_acclim <- dat_acclim_ES %>% filter(between(yi, -5, 5))
+normalized_var <- dat_full_var_ES  %>% filter(between(yi, -5, 5))
+
+dummy_acclim <- rma.mv(yi, vi, data=normalized_acclim, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+dummy_var <- rma.mv(yi, vi, data=normalized_var, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+```
+
+```{r}
+install.packages("devtools")
+install.packages("tidyverse")
+install.packages("metafor")
+install.packages("patchwork")
+install.packages("R.rsp")
+
+devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", method= 'libcurl', force = TRUE, build_vignettes = TRUE)
+library(orchaRd)
+library(patchwork)
+I2_acclim <- i2_ml(dummy_var)
+I2_var <-i2_ml(dummy_acclim)
+
+p1 <- orchard_plot(dummy_acclim, mod = "Int", xlab = "Standardised mean difference", transfm = "none", )+
+  annotate(geom = "text", x = 0, y = 1.5, label = "Acclimation")+
+  annotate(geom = "text", x = -4.7, y = 1.5, label = paste0("italic(I)^{2} == ",
+round(I2_acclim[1] * 100, 2)), color = "black", parse = TRUE, size = 5) + annotate(geom = "text",
+x = -3.7, y = 1.49, label = "%", color = "black", parse = FALSE, size = 5)
+
+p2 <- orchard_plot(dummy_var, mod = "Int", xlab = "Standardised mean difference", transfm = "none")+
+  annotate(geom = "text", x = 0, y = 1.5, label = "Non-linear averaging and acclimation")+
+  annotate(geom = "text", x = -4.1, y = 1.5, label = paste0("italic(I)^{2} == ",
+round(I2_var[1] * 100, 2)), color = "black", parse = TRUE, size = 5) + annotate(geom = "text",
+x = -3.1, y = 1.49, label = "%", color = "black", parse = FALSE, size = 5)
+p1 / p2
+
+```
+
+```{r}
+normalized_acclim %>%
+  ggplot(aes(y=yi, x=as.factor(mean_temp_reared)))+
+  geom_boxplot()+
+  geom_point()
+
+normalized_acclim %>%
+  group_by(flux_range, mean_temp_reared) %>%
+  summarise(yi=mean(yi)) %>%
+  ggplot(aes(x=flux_range, y=yi, color=mean_temp_reared))+
+  geom_point(size= 4)+
+  geom_smooth(method="lm")
 ```
@@
 ```
+
+# Supplementary code and models
+
+## Acclimation data split out by CT data and other thermal performance metrics 
+```{r, echo=FALSE, message=FALSE, results='hide'}
+CT_acclim_data <- dat_acclim_ES %>%
+  filter(resp_def %in% c("CTmax", "CT max", "CTmin", "LTmax", "LTmin", "Topt", "Tpref", "Thermal preference", "ULT50", "Scope of thermal tolerance (CTmax - CTmin)"))
+
+curve_acclim_data <- dat_acclim_ES %>%
+  filter(resp_def %in% 
+           c("CCO activity", "CS activity", "LDH activity", "log CO2 production", 
+             "maximal swimming speed", "Maximum burst swim speed", 
+             "Performance breadth (Tbr)", "RMR", "ULT50", "Speed", "Time to heat knockdown",
+             "Ucrit", "Umax"))
+```
+
+## CT aggregated model 
+```{r, echo=FALSE, message=FALSE, results='hide'}
+CT_model <-rma.mv(yi, vi, data=CT_acclim_data, mods = ~flux_range * mean_temp_reared +
+                        exp_age + size + org_level + exposure_temp, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+CT_model
+```
+## Remaining acclimation metrics model 
+```{r, echo=FALSE, message=FALSE, results='hide'}
+curve_model <- rma.mv(yi, vi, data=curve_acclim_data, mods = ~flux_range * mean_temp_reared +
+                        exp_age + size + org_level + exposure_temp, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+curve_model
+```

### 2021-07-05 shared analysis files /manuscript_datavis_22jun21.Rmd
figuring out plots of model results

--- 2021-06-30 shared analysis files /manuscript_datavis_22jun21.Rmd
@@
 ## loading libraries 
-pacman::p_load(metafor, tidyverse, viridis, visreg, forcats)
-
+pacman::p_load(metafor, tidyverse, viridis, visreg, forcats, devtools, patchwork, R.rsp)
+#devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", method= 'libcurl', force = TRUE, build_vignettes = TRUE)
+library(orchaRd)
 #load data
@@
 
-## Acclimation without modifiers
-```{r, echo=FALSE, message=FALSE}
-simple_acclimation_model <- rma.mv(yi, vi, data=dat_acclim_ES, 
+## Acclimation with temperature modifiers
+```{r, echo=FALSE}
+temp_acclimation_model <- rma.mv(yi, vi, data=dat_acclim_ES, mods=~flux_range * mean_temp_reared,
                random = ~1 | study_id / experiment_id/ response_id,
                  method="REML") 
-simple_acclimation_model
+temp_acclimation_model
 ```
 
-## Acclimation with modifiers
-```{r, echo=FALSE, message=FALSE}
-acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, mods = ~flux_range * mean_temp_reared +
-                        exp_age + size + org_level + exposure_temp, 
+## Acclimation with trait modifiers
+```{r, echo=FALSE}
+trait_acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, 
+                                mods = ~exp_age + size + org_level + exposure_temp, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+trait_acclimation_model
+```
+
+## Acclimation with all modifiers
+```{r, echo=FALSE}
+acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, 
+            mods = ~flux_range * mean_temp_reared + exp_age + size + org_level + exposure_temp, 
                random = ~1 |  study_id/ experiment_id/ response_id,
@@
 
-# Acclimation data split out by CT data and other thermal performance metrics 
-```{r, echo=FALSE, message=FALSE}
+# Effects of non-linear averaging
+This uses the full dataset from my thesis (with more data from the additional search). Interesting that this time, mean temp is the only significant predictor?
+
+## Non-linear averaging model with temperature modifiers
+```{r, echo= FALSE}
+temp_full_var_model <- rma.mv(yi, vi, data=dat_full_var_ES, 
+                                mods = ~flux_range * mean_temp_constant,
+               random = ~1 | study_id / experiment_id/ response_id,
+                 method="REML") 
+temp_full_var_model
+```
+
+## Non-linear averaging model with trait modifiers
+```{r, echo = FALSE}
+trait_full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~exp_age + size + org_level, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+trait_full_var_model
+```
+
+## Non-linear averaging model with all modifiers
+```{r, echo = FALSE}
+full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, 
+                        mods = ~flux_range * mean_temp_constant + exp_age + size + org_level, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML") 
+full_var_model
+```
+
+```{r}
+#Model selection 
+AIC(temp_acclimation_model, trait_acclimation_model, acclimation_model, temp_full_var_model, trait_full_var_model, full_var_model)
+```
+
+# Figures
+
+```{r, echo=FALSE}
+normalized_acclim <- dat_acclim_ES %>% filter(between(yi, -5, 5))
+normalized_var <- dat_full_var_ES  %>% filter(between(yi, -5, 5))
+
+dummy_acclim <- rma.mv(yi, vi, data=normalized_acclim, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+dummy_var <- rma.mv(yi, vi, data=normalized_var, 
+               random = ~1 |  study_id/ experiment_id/ response_id,
+                 method="REML")
+
+I2_acclim <- i2_ml(dummy_var)
+I2_var <-i2_ml(dummy_acclim)
+
+p1 <- orchard_plot(dummy_acclim, mod = "Int", xlab = "Standardised mean difference", transfm = "none", )+
+  annotate(geom = "text", x = 0, y = 1.5, label = "Acclimation")+
+  annotate(geom = "text", x = -4.7, y = 1.5, label = paste0("italic(I)^{2} == ",
+round(I2_acclim[1] * 100, 2)), color = "black", parse = TRUE, size = 5) + annotate(geom = "text",
+x = -3.7, y = 1.49, label = "%", color = "black", parse = FALSE, size = 5)
+
+p2 <- orchard_plot(dummy_var, mod = "Int", xlab = "Standardised mean difference", transfm = "none")+
+  annotate(geom = "text", x = 0, y = 1.5, label = "Non-linear averaging and acclimation")+
+  annotate(geom = "text", x = -4.1, y = 1.5, label = paste0("italic(I)^{2} == ",
+round(I2_var[1] * 100, 2)), color = "black", parse = TRUE, size = 5) + annotate(geom = "text",
+x = -3.1, y = 1.49, label = "%", color = "black", parse = FALSE, size = 5)
+p1 / p2
+```
+
+```{r, echo=FALSE}
+forest_acclim<- dat_acclim_ES %>%
+  group_by(mean_temp_reared, flux_range) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  unite(group, c(mean_temp_reared, flux_range), sep = " range ", remove = FALSE)%>%
+  arrange(desc(flux_range))
+
+forest_var<- normalized_var %>%
+  group_by(mean_temp_constant, flux_range) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  unite(group, c(mean_temp_constant, flux_range), sep = "+/-", remove = FALSE) %>%
+  arrange(desc(flux_range))
+
+forest(forest_acclim$mean_yi, forest_acclim$mean_vi, slab=forest_acclim$group, pch =19, psize=1, col = "gray2", 
+       header=TRUE, cex=0.77)
+
+forest(forest_var$mean_yi, forest_var$mean_vi, slab=forest_var$group, pch =19, psize=1, col = "gray2", 
+       header=TRUE, cex=0.77)
+
+```
+
+## Funnel plot of acclimation data 
+````{r, echo=FALSE}
+funnel(dummy_acclim, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+```
+## Funnel plot of non-linear averaging + acclimation data 
+````{r, echo=FALSE}
+funnel(dummy_var, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+```
+
+```{r, echo=FALSE}
+p3<-normalized_acclim %>%
+  ggplot(aes(x=flux_range, y=yi, color=as.factor(mean_temp_reared)))+
+  scale_color_viridis(discrete=TRUE)+
+  geom_point(size=3, alpha =0.5)+
+  geom_smooth(method="lm")+
+  theme_bw()
+
+p4<-normalized_var %>%
+  ggplot(aes(x=flux_range, y=yi, color=as.factor(mean_temp_constant)))+
+  scale_color_viridis(discrete=TRUE)+
+  geom_point(size= 3, alpha =0.5)+
+  geom_smooth(method="lm")+
+  theme_bw()
+p3/p4
+
+normalized_acclim %>%
+  ggplot(aes(x=as.factor(flux_range), y=yi, color=as.factor(mean_temp_reared)))+
+  scale_color_viridis(discrete = TRUE)+
+  geom_boxplot()+
+  geom_point(alpha=0.5)+
+  theme_bw()
+
+normalized_var %>%
+  ggplot(aes(x=as.factor(flux_range), y=yi, color=as.factor(mean_temp_constant)))+
+  scale_color_viridis(discrete = TRUE)+
+  geom_boxplot()+
+  geom_point(alpha=0.5)+
+  theme_bw()
+
+```
+
+# Supplementary code and models
+
+## Acclimation data split out by CT data and other thermal performance metrics 
+```{r, echo=FALSE, message=FALSE, results='hide'}
 CT_acclim_data <- dat_acclim_ES %>%
@@
 ## CT aggregated model 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, results='hide'}
 CT_model <-rma.mv(yi, vi, data=CT_acclim_data, mods = ~flux_range * mean_temp_reared +
@@
 ## Remaining acclimation metrics model 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, results='hide'}
 curve_model <- rma.mv(yi, vi, data=curve_acclim_data, mods = ~flux_range * mean_temp_reared +
@@
 ```
-
-# Effects of non-linear averaging
-This uses the full dataset from my thesis (with more data from the additional search). Interesting that this time, mean temp is the only significant predictor?
-
-## Non-linear averaging model with no modifiers
-```{r, echo= FALSE, message=FALSE}
-simple_full_var_model <- rma.mv(yi, vi, data=dat_full_var_ES, 
-               random = ~1 | study_id / experiment_id/ response_id,
-                 method="REML") 
-simple_full_var_model
-```
-
-## Non-linear averaging model with modifiers
-```{r, echo = FALSE, message=FALSE}
-full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~flux_range * mean_temp_constant +
-                        exp_age + size + org_level, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
-                 method="REML") 
-full_var_model
-```
-
-# Figures
-```{r, echo = FALSE, message=FALSE}
-dat_acclim_ES %>%
-  filter(yi > -5) %>%
-ggplot(aes(x=mean_temp_reared, y = yi, fill=flux_range))+
-  geom_point(alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
-  scale_fill_viridis('Temperature range (°C)', option = "C")+
-  theme_classic()+
-  xlab("Mean Temperature (°C)")+
-  ylab("SMD")+
-  geom_line(y=0, linetype = "dashed")+
-  theme(legend.position="bottom")+
-  ggtitle("Acclimation")
-```
-
-```{r, echo = FALSE, message=FALSE}
-dat_full_var_ES %>%
-  filter(yi > -5) %>%
-ggplot(aes(x=mean_temp_constant, y = yi, fill=flux_range))+
-  geom_point(alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
-  scale_fill_viridis('Temperature range (°C)', option = "C")+
-  theme_classic()+
-  xlab("Mean Temperature (°C)")+
-  ylab("SMD")+
-  geom_line(y=0, linetype = "dashed")+
-  theme(legend.position="bottom")+
-  ggtitle("Non-linear averaging")
-```

### 2021-07-05 shared analysis files /manuscript_datavis_22jun21.Rmd
additional figure drafts and stats calcs for heterogeneity

@@
 full_var_model
+
+pacman::p_load(esc, dmetar)
+summary(full_var_model)
+convert_z2r(1.3702) #r=0.8787378
+var.comp(full_var_model)
+
+summary(acclimation_model)
+convert_z2r(6.0093) #r=0.9999879
 ```
@@
 p1 / p2
+```
+
+```{r, echo=FALSE}
+mean_temp_acclim <- normalized_acclim %>%
+  group_by(mean_temp_reared, flux_range) %>%
+  summarise(yi=mean(yi), 
+            vi=mean(vi))
+
+mean_var<-normalized_var %>%
+  group_by(mean_temp_constant, flux_range) %>%
+  summarise(yi=mean(yi), 
+            vi=mean(vi))
+```
+
+```{r, echo=FALSE}
+ggplot(normalized_acclim, aes(y=yi, x=mean_temp_reared, fill=flux_range))+
+  geom_line(y=0, linetype = "dashed")+
+  geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.7, height = 0.7)+
+  geom_errorbar(data=mean_temp_acclim, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_point(data = mean_temp_acclim, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Acclimation and fluctuation range")
+
+ggplot(normalized_var, aes(y=yi, x=mean_temp_constant, fill=flux_range))+
+  geom_jitter(alpha = 0.5, pch=21,size=1)+
+  geom_line(y=0, linetype = "dashed")+
+  geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Non-linear averaging + acclimation and fluctuation range")
+
+ggplot(normalized_var, aes(y=yi, x=mean_temp_constant, fill=mean_temp_constant))+
+  geom_jitter(alpha = 0.5, pch=21,size=1)+
+  geom_line(y=0, linetype = "dashed")+
+  geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Non-linear averaging + acclimation and mean temperature")
+
 ```

### 2021-07-13 shared analysis files /manuscript_datavis_22jun21.Rmd
saving changes and trying to figure out which modle structure to go with

--- 2021-07-05 shared analysis files /manuscript_datavis_22jun21.Rmd
@@
                  method="REML") 
-temp_acclimation_model
+summary(temp_acclimation_model)
 ```
@@
 acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, 
-            mods = ~flux_range * mean_temp_reared + exp_age + size + org_level + exposure_temp, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
+            mods = ~flux_range * mean_temp_reared + exp_age + size + org_level + exposure_temp +environment, 
+               random = ~1 |  study_id/ response_id,
                  method="REML") 
@@
 ```{r, echo = FALSE}
+
 full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, 
-                        mods = ~flux_range * mean_temp_constant + exp_age + size + org_level, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
+                        mods = ~flux_range * mean_temp_constant + exp_age + size + org_level +environment, 
+               random = list(~1 | study_id/ response_id),
                  method="REML") 
 full_var_model
+
+pacman::p_load(esc, dmetar)
+summary(full_var_model)
+convert_z2r(1.3702) #r=0.8787378
+
+summary(acclimation_model)
+convert_z2r(6.0093) #r=0.9999879
+
+acclimation_model
+
+##I-SQUARED CALCS FROM ASSINK + WIBBELINK 2016 (doi:10.20982/tqmp.12.3.p154)
+n <- length(dat_acclim_ES$vi)
+list.inverse.variances <- 1 / (dat_acclim_ES$vi)
+sum.inverse.variances <- sum(list.inverse.variances)
+squared.sum.inverse.variances <- (sum.inverse.variances) ^ 2
+list.inverse.variances.square <- 1 / (dat_acclim_ES$vi^2)
+sum.inverse.variances.square <-
+sum(list.inverse.variances.square)
+numerator <- (n - 1) * sum.inverse.variances
+denominator <- squared.sum.inverse.variances -
+sum.inverse.variances.square
+estimated.sampling.variance <- numerator / denominator
+I2_1 <- (estimated.sampling.variance) / (acclimation_model$sigma2[1]
++ acclimation_model$sigma2[2] + estimated.sampling.variance)
+I2_2 <- (acclimation_model$sigma2[1]) / (acclimation_model$sigma2[1]
++ acclimation_model$sigma2[2] + estimated.sampling.variance)
+I2_3 <- (acclimation_model$sigma2[2]) / (acclimation_model$sigma2[1]
++ acclimation_model$sigma2[2] + estimated.sampling.variance)
+
+amountvariancelevel1 <- I2_1 * 100
+amountvariancelevel2 <- I2_2 * 100
+amountvariancelevel3 <- I2_3 * 100
+amountvariancelevel1 ##16.20628
+amountvariancelevel2 ##1.966446e-06
+amountvariancelevel3 ##83.79732
+
+n_f <- length(dat_full_var_ES$vi)
+list.inverse.variances_f <- 1 / (dat_full_var_ES$vi)
+sum.inverse.variances_f <- sum(list.inverse.variances_f)
+squared.sum.inverse.variances_f <- (sum.inverse.variances_f) ^ 2
+list.inverse.variances.square_f <- 1 / (dat_full_var_ES$vi^2)
+sum.inverse.variances.square_f <-
+sum(list.inverse.variances.square_f)
+numerator_f <- (n_f - 1) * sum.inverse.variances_f
+denominator_f <- squared.sum.inverse.variances_f -
+sum.inverse.variances.square_f
+estimated.sampling.variance_f <- numerator_f / denominator_f
+I2_1f <- (estimated.sampling.variance_f) / (full_var_model$sigma2[1]
++ full_var_model$sigma2[2] + estimated.sampling.variance_f)
+I2_2f <- (full_var_model$sigma2[1]) / (full_var_model$sigma2[1]
++ full_var_model$sigma2[2] + estimated.sampling.variance_f)
+I2_3f <- (full_var_model$sigma2[2]) / (full_var_model$sigma2[1]
++ full_var_model$sigma2[2] + estimated.sampling.variance_f)
+
+amountvariancelevel1f <- I2_1f * 100
+amountvariancelevel2f <- I2_2f * 100
+amountvariancelevel3f <- I2_3f * 100
+amountvariancelevel1f #7.629274
+amountvariancelevel2f #15.61584
+amountvariancelevel3f #76.75488
+
+count(dat_acclim_ES, resp_units)
+count(dat_full_var_ES, resp_units)
+
+dat_acclim_ES %>%
+
+```
+
+```{r}
+#looking at the effects sizes for each categorical moderator not related to temperature...
+# acclimation analysis
+
+a <- dat_acclim_ES %>%
+  group_by(size) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=as.factor(size), y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Body Size")+
+  ylim(-1,3.5)+
+  ggtitle("ACCLIMATION")+
+  coord_flip()
+
+b <- dat_acclim_ES %>%
+  group_by(exp_age) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=as.factor(exp_age), y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Age")+
+  ylim(-1,0.75)+
+  coord_flip()
+
+c <- dat_acclim_ES %>%
+  group_by(environment) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=as.factor(environment), y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Environment")+
+  ylim(-1,0.5)+
+  coord_flip()
+
+# assay analysis
+d <- dat_full_var_ES %>%
+  group_by(size) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=as.factor(size), y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Body Size")+
+  ylim(-1,3.5)+
+  ggtitle("ASSAY")+
+  coord_flip()
+
+e <- dat_full_var_ES %>%
+  group_by(exp_age) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=as.factor(exp_age), y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Age")+
+  ylim(-1,0.75)+
+  coord_flip()
+
+f <- dat_full_var_ES %>%
+  group_by(org_level) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=as.factor(org_level), y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Organization Level")+
+  ggtitle("ASSAY")+
+  coord_flip()
+
+g <- dat_full_var_ES %>%
+  group_by(environment) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=as.factor(environment), y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Environment")+
+  ylim(-1,0.5)+
+  coord_flip()
+
+acclimation_mods<- a + b + c 
+var_mods <- d + e + g
+
+acclimation_mods / var_mods
+
+f
+
+```
+
+```{r}
+full_model_coefficients<- as.data.frame(coef(full_var_model)) %>%
+  tibble::rownames_to_column("covariates") %>%
+  rename(value = "coef(full_var_model)") %>%
+  arrange(desc(value)) %>%
+  mutate(coefficients = c("intercept", "experimental age", "fluctuation range:mean temperature",
+                          "fluctuation range", "environmentterrestrial", "mean temperature",
+                                       "size", "organization level" 
+                                      ), 
+         lb = c(0.4597, -0.0738, -0.0008, -0.0646, -0.7671, -0.0719, -0.5783, -0.8331), 
+         ub= c(2.2806, 0.2150, 0.0031, 0.0227, 0.6876, -0.0214, 0.2861, 0.2738))
+
+label.df_full <- data.frame(coefficients = c("intercept","mean temperature"),
+                       value = c(0.35, -0.2))
+
+ggplot(data=full_model_coefficients, aes(x = reorder(coefficients, value),
+             y = value,
+             fill = value > 0)) +
+  geom_errorbar(aes(ymin=lb, ymax=ub), width=.2, size =0.6,
+                 position=position_dodge(.95), color = "black")+ 
+  geom_point(size = 3, show.legend = FALSE, pch=21,colour="black") +
+  labs(x = NULL, y = "Coefficient estimate") +
+  scale_fill_viridis(discrete = TRUE)+
+  theme_bw()+
+  coord_flip()+
+  geom_text(data=label.df_full, label = "*", color = "black", size = 6)
+
+acclim_model_coefficients<- as.data.frame(coef(acclimation_model)) %>%
+  tibble::rownames_to_column("covariates") %>%
+  rename(value = "coef(acclimation_model)") %>%
+  arrange(desc(value)) %>%
+  mutate(coefficients = c("intercept", "fluctuation range:mean temperature",  "exposure_temp",
+                          "experimental age", "mean temperature", 
+                          "fluctuation range", "environmentterrestrial","size"), 
+         lb = c(3.2855, 0.0140, -0.0030, -0.7078, -0.3457, -0.6700, -2.5526, -1.4671), 
+         ub= c(8.6331, 0.0312, 0.0093, 0.5405, -0.1733, -0.2814, 1.0587, 0.1290))
+
+label.df_acclim <- data.frame(coefficients = c("intercept", "fluctuation range:mean temperature", "mean temperature", "fluctuation range"),
+                       value = c(3, -0.5, -0.7, -1))
+
+ggplot(data=acclim_model_coefficients, aes(x = reorder(coefficients, value),
+             y = value,
+             fill = value > 0)) +
+  geom_errorbar(aes(ymin=lb, ymax=ub), width=.2, size =0.6,
+                 position=position_dodge(.95), color = "black")+ 
+  geom_point(size = 3, show.legend = FALSE, pch=21,colour="black") +
+  labs(x = NULL, y = "Coefficient estimate") +
+  scale_fill_viridis(discrete = TRUE)+
+  theme_bw()+
+  coord_flip()+
+  geom_text(data=label.df_acclim, label = "*", color = "black", size = 6)
 ```
@@
 ```{r, echo=FALSE}
+#creating orchard plot with trimmed/normalized effect sizes from both datasets
 normalized_acclim <- dat_acclim_ES %>% filter(between(yi, -5, 5))
 normalized_var <- dat_full_var_ES  %>% filter(between(yi, -5, 5))
-
+#corresponding models
 dummy_acclim <- rma.mv(yi, vi, data=normalized_acclim, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
+               random = ~1 |  study_id/ response_id,
                  method="REML")
 dummy_var <- rma.mv(yi, vi, data=normalized_var, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
+               random = ~1 |  study_id/ response_id,
                  method="REML")
-
+#calculating I2 for each of the analyses
 I2_acclim <- i2_ml(dummy_var)
 I2_var <-i2_ml(dummy_acclim)
-
+#orchard plots 
 p1 <- orchard_plot(dummy_acclim, mod = "Int", xlab = "Standardised mean difference", transfm = "none", )+
@@
 ```{r, echo=FALSE}
-forest_acclim<- dat_acclim_ES %>%
+#calculating mean effect sizes by mean temperature and flux range for ggplots
+mean_temp_acclim <- normalized_acclim %>%
   group_by(mean_temp_reared, flux_range) %>%
-  summarise(mean_yi = mean(yi),
-            mean_vi = mean(vi)) %>%
-  unite(group, c(mean_temp_reared, flux_range), sep = " range ", remove = FALSE)%>%
-  arrange(desc(flux_range))
+  summarise(yi=mean(yi), 
+            vi=mean(vi))
 
-forest_var<- normalized_var %>%
+mean_var<-normalized_var %>%
   group_by(mean_temp_constant, flux_range) %>%
-  summarise(mean_yi = mean(yi),
-            mean_vi = mean(vi)) %>%
-  unite(group, c(mean_temp_constant, flux_range), sep = "+/-", remove = FALSE) %>%
-  arrange(desc(flux_range))
+  summarise(yi=mean(yi), 
+            vi=mean(vi))
+```
 
-forest(forest_acclim$mean_yi, forest_acclim$mean_vi, slab=forest_acclim$group, pch =19, psize=1, col = "gray2", 
-       header=TRUE, cex=0.77)
+```{r, echo=FALSE}
+ggplot(normalized_acclim, aes(y=yi, x=mean_temp_reared, fill=flux_range))+
+  geom_line(y=0, linetype = "dashed")+
+  geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.7, height = 0.7)+
+  geom_errorbar(data=mean_temp_acclim, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_point(data = mean_temp_acclim, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Acclimation and fluctuation range")+
+  labs(x = "Mean temperature (°C)",
+       y ="SMD",
+       fill = "Temperature fluctuation range (°C)")+
+  theme(legend.position="bottom")
 
-forest(forest_var$mean_yi, forest_var$mean_vi, slab=forest_var$group, pch =19, psize=1, col = "gray2", 
-       header=TRUE, cex=0.77)
+ggplot(normalized_var, aes(y=yi, x=mean_temp_constant, fill=flux_range))+
+  geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.3, height = 0.7)+
+  geom_line(y=0, linetype = "dashed")+
+  geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Non-linear averaging + acclimation and fluctuation range")+
+  labs(x = "Mean temperature (°C)",
+       y ="SMD",
+       fill = "Temperature fluctuation range (°C)")+
+  theme(legend.position="bottom")
 
+ggplot(normalized_var, aes(y=yi, x=mean_temp_constant, fill=mean_temp_constant))+
+  geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.3, height = 0.7)+
+  geom_line(y=0, linetype = "dashed")+
+  geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Non-linear averaging + acclimation and mean temperature")+
+  labs(x = "Mean temperature (°C)",
+       y ="SMD",
+       fill = "Mean temperature (°C)")+
+  theme(legend.position="bottom")
 ```
@@
 ```
+
+# Supplementary code and models
 
@@
 
-# Supplementary code and models
+```{r, echo=FALSE}
+forest_acclim<- dat_acclim_ES %>%
+  group_by(mean_temp_reared, flux_range) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  unite(group, c(mean_temp_reared, flux_range), sep = " range ", remove = FALSE)%>%
+  arrange(desc(flux_range))
+
+forest_var<- normalized_var %>%
+  group_by(mean_temp_constant, flux_range) %>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  unite(group, c(mean_temp_constant, flux_range), sep = "+/-", remove = FALSE) %>%
+  arrange(desc(flux_range))
+
+forest(forest_acclim$mean_yi, forest_acclim$mean_vi, slab=forest_acclim$group, pch =19, psize=1, col = "gray2", 
+       header=TRUE, cex=0.77)
+
+forest(forest_var$mean_yi, forest_var$mean_vi, slab=forest_var$group, pch =19, psize=1, col = "gray2", 
+       header=TRUE, cex=0.77)
+
+```
 

### 2021-07-14 shared analysis files /manuscript_datavis_22jun21.Rmd
adding in broader response columns for additional figures

@@
 dat_acclim_ES<-dat_acclim_ES %>% 
-  filter (!is.na(yi)) 
+  filter (!is.na(yi)) %>%
+  mutate(metric = case_when(resp_def %in% c("log CO2 production", 
+                                            "maximal swimming speed", 
+                                            "Maximum burst swim speed", 
+                                            "Oxygen consumption", 
+                                            "Performance breadth (Tbr)",
+                                            "Time to heat knockdown",
+                                            "Scope of thermal tolerance (CTmax - CTmin)",
+                                            "RMR", 
+                                            "Speed", "Thermal preference", 
+                                            "Topt", "Tpref", 
+                                            "Ucrit", "Umax") ~ "Pmax",
+                            resp_def %in% c("CTmin", "LTmin") ~ "CTmin",
+                            resp_def %in% c("CT max", "CTmax", "LTmax", "ULT50") ~ "CTmax",
+                            resp_def %in% c("LDH activity", "CCO activity", 
+                                            "CS activity") ~ "Biochemical"
+                            ))
 ```
@@
 dat_full_var_ES<-dat_full_var_ES %>% 
-  filter (!is.na(yi)) 
+  filter (!is.na(yi)) %>%
+  mutate(metric = case_when(resp_def %in% c("abdomen length", "adult weight", 
+                                            "blade length/total leaf length", 
+                                            "body length", "body mass", "body size", 
+                                            "body weight", "carapace height", 
+                                            "carapace width", 
+                                            "final length", "final mass",
+                                            "fore-limb length", 
+                                            "group mean body weight", 
+                                            "head length", "head width", 
+                                            "hind-limb length", 
+                                            "hypocotyl length", "mass", 
+                                            "maximal length", 
+                                            "snout-vent length", "tail length",
+                                            "total length", 
+                                            "wing centroid", "body (centroid) size", 
+                                            "egg mass", "ovary mass, dry", 
+                                             "testes mass, dry", 
+                                             "pupal shell weight", "pupal weight") ~ "size", 
+                            resp_def %in%  c("average cumulative number of eggs laid per female",
+                                             "hatching success", 
+                                             "offspring per mating",
+                                             "total offspring", 
+                                             "percent females") ~ "fecundity",
+                            resp_def %in% c("Adult eclosion to death", 
+                                            "dessication tolerance", 
+                                            "egg to adult viability",
+                                            "infestation rate", 
+                                            "Mortaility", 
+                                            "Parasitism success", 
+                                            "startvation tolerance", 
+                                            "success of parasitism", 
+                                            "survival", "Survival", 
+                                            "Larval mortaility", "longevity") ~ "survivorship", 
+                            resp_def %in% c("aquatic speed", 
+                                            "daily energy expenditure", 
+                                            "distance covered", 
+                                            "Energy consumption", 
+                                            "feeding efficiency", 
+                                            "feeding rate", ,
+                                            "RMR", "sprint speed",
+                                          "Sqrt Cellular energy allocation (CEA)", 
+                                            "Sqrt of available energy", 
+                                          "population growth rate","productivity", 
+                                             "rate of change", 
+                                             "specific growth rate") ~ "energetics", 
+                            resp_def %in% c("development time", 
+                                            "development to stages 35-37", 
+                                            "days to first slough", 
+                                            "developmental time", 
+                                            "egg to adult", 
+                                            "egg to pupa", "incubation period", 
+                                            "median days to bolt", 
+                                            "Third instar to adult", 
+                                            "Third instar to pupa") ~ "development", 
+                            resp_def %in% c("Catalase", "cortisol",
+                                             "Hexokinase", "Hsp70", 
+                                            "oxidative damage", 
+                                            "Pyruvate kinase", "TAC", 
+                                            "SOD", "glucose") ~ "biochemistry"))
+
+count(dat_full_var_ES, metric)
 ```
@@
 
-count(dat_acclim_ES, resp_units)
-count(dat_full_var_ES, resp_units)
-
-dat_acclim_ES %>%
-
 ```
@@
 
+```
+
+```{r}
+g <- dat_full_var_ES %>%
+  group_by(metric)%>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=metric, y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Metric")+
+  ylim(-2.6,1.7)+
+  coord_flip()+
+  ggtitle("ASSAY")
+
+h <- dat_acclim_ES %>%
+  group_by(metric)%>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=metric, y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Metric")+
+  ylim(-2.6,1.7)+
+  coord_flip()+
+  ggtitle("ACCLIMATION")
+
+h / g
 ```

### 2021-07-15 shared analysis files /manuscript_datavis_22jun21.Rmd
model troubleshooting and figure generation

@@
                                             "feeding efficiency", 
-                                            "feeding rate", ,
+                                            "feeding rate",
                                             "RMR", "sprint speed",
@@
 
-count(dat_full_var_ES, metric)
 ```
@@
 ```{r, echo=FALSE}
-acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, 
-            mods = ~flux_range * mean_temp_reared + exp_age + size + org_level + exposure_temp +environment, 
-               random = ~1 |  study_id/ response_id,
+acclimation_model<-rma.mv(yi, vi, data=normalized_acclim, 
+            mods = ~flux_range * mean_temp_reared + exp_age + size + org_level + exposure_temp +environment +metric, 
+               random = ~1 |  study_id/response_id,
                  method="REML") 
@@
 
-full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, 
-                        mods = ~flux_range * mean_temp_constant + exp_age + size + org_level +environment, 
-               random = list(~1 | study_id/ response_id),
+full_var_model <-rma.mv(yi, vi, data=normalized_var, 
+                        mods = ~flux_range * mean_temp_constant + exp_age + size  +environment +org_level +metric, 
+               random = ~1 | study_id/response_id,
                  method="REML") 
@@
 amountvariancelevel3f #76.75488
+
+```
+
+```{r}
+#devtools::install_github("MathiasHarrer/dmetar")
+library(dmetar)
+res <- var.comp(full_var_model)
+plot(res)
+
+res_acclim <- var.comp(acclimation_model)
+plot(res_acclim)
 
@@
 
-f
-
 ```
@@
 ```{r}
-g <- dat_full_var_ES %>%
+h <- dat_full_var_ES %>%
   group_by(metric)%>%
@@
   xlab("Metric")+
-  ylim(-2.6,1.7)+
+  ylim(-3,1.7)+
   coord_flip()+
@@
 
-h <- dat_acclim_ES %>%
+i <- dat_acclim_ES %>%
   group_by(metric)%>%
@@
 
-h / g
+i / h
+
+h <- dat_full_var_ES %>%
+  group_by(metric)%>%
+  summarise(mean_yi = mean(yi),
+            mean_vi = mean(vi)) %>%
+  ggplot(aes(x=metric, y=mean_yi, fill = mean_yi > 0))+
+  geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  geom_point(size = 4, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  xlab("Metric")+
+  ylim(-3,1.7)+
+  coord_flip()+
+  ggtitle("ASSAY")
+
+i <- normalized_acclim %>%
+  ggplot(aes(x=yi, y=mean_temp_reared,size = flux_range))+
+  #geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  #geom_boxplot(aes(fill = yi > 0))+
+  geom_point(aes(fill = yi > 0), alpha=0.8, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  coord_flip()+
+  ggtitle("ACCLIMATION")+
+  facet_wrap(~metric)
+
+j <- normalized_var %>%
+  ggplot(aes(x=yi, y=mean_temp_constant, size=flux_range))+
+  #geom_errorbar(aes(ymin=mean_yi-mean_vi, ymax=mean_yi+mean_vi), width=.2)+
+  #geom_boxplot(aes(fill = yi > 0))+
+  geom_point(aes(fill = yi > 0), alpha=0.8, show.legend = FALSE, pch=21,colour="black")+
+  scale_fill_viridis(discrete = TRUE)+
+  coord_flip()+
+  ggtitle("ASSAY")+
+  facet_wrap(~metric)
+
+i + j
+
 ```
@@
 
-ggplot(normalized_var, aes(y=yi, x=mean_temp_constant, fill=flux_range))+
+ggplot(normalized_var, aes(y=yi, fill=mean_temp_constant, x=flux_range))+
   geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.3, height = 0.7)+
@@
   ggtitle("Non-linear averaging + acclimation and fluctuation range")+
-  labs(x = "Mean temperature (°C)",
+  labs(x = "flux range",
        y ="SMD",
-       fill = "Temperature fluctuation range (°C)")+
-  theme(legend.position="bottom")
-
-ggplot(normalized_var, aes(y=yi, x=mean_temp_constant, fill=mean_temp_constant))+
-  geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.3, height = 0.7)+
-  geom_line(y=0, linetype = "dashed")+
-  geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
-  geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
-  theme_bw()+
-  scale_fill_viridis()+
-  ggtitle("Non-linear averaging + acclimation and mean temperature")+
-  labs(x = "Mean temperature (°C)",
-       y ="SMD",
-       fill = "Mean temperature (°C)")+
-  theme(legend.position="bottom")
+       fill = "mean temp")+
+  theme(legend.position="bottom")+ 
+  facet_wrap(~mean_temp_constant)
 ```
@@
 ````{r, echo=FALSE}
-funnel(dummy_acclim, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+funnel(acclimation_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+
+par(mfrow=c(2,2))
+
+### draw funnel plots
+funnel(dummy_acclim, main="Standard Error")
+funnel(dummy_acclim, yaxis="vi", main="Sampling Variance")
+funnel(dummy_acclim, yaxis="seinv", main="Inverse Standard Error")
+funnel(dummy_acclim, yaxis="vinv", main="Inverse Sampling Variance")
+
+funnel(dummy_var, main="Standard Error")
+funnel(dummy_var, yaxis="vi", main="Sampling Variance")
+funnel(dummy_var, yaxis="seinv", main="Inverse Standard Error")
+funnel(dummy_var, yaxis="vinv", main="Inverse Sampling Variance")
+
+ggplot(normalized_acclim, aes(y=yi, x=vi))+
+  geom_point()
+
+ggplot(normalized_var, aes(y=yi, x=vi))+
+  geom_point()
+
 ```
@@
 ````{r, echo=FALSE}
-funnel(dummy_var, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
+funnel(full_var_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
 ```

### 2021-07-17 shared analysis files /manuscript_datavis_22jun21.Rmd
savings edits

@@
                             ))
+
+count(dat_acclim_ES, metric)
 ```
@@
 temp_acclimation_model <- rma.mv(yi, vi, data=dat_acclim_ES, mods=~flux_range * mean_temp_reared,
-               random = ~1 | study_id / experiment_id/ response_id,
-                 method="REML") 
-summary(temp_acclimation_model)
+               random = ~1 | study_id / response_id,
+                 method="ML") 
+temp_acclimation_model
 ```
@@
 trait_acclimation_model<-rma.mv(yi, vi, data=dat_acclim_ES, 
-                                mods = ~exp_age + size + org_level + exposure_temp, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
-                 method="REML") 
+                                mods = ~exp_age + size + org_level + exposure_temp + metric, 
+               random = ~1 |  study_id/  response_id,
+                 method="ML") 
 trait_acclimation_model
@@
                random = ~1 |  study_id/response_id,
-                 method="REML") 
+                 method="ML") 
 acclimation_model
@@
                                 mods = ~flux_range * mean_temp_constant,
-               random = ~1 | study_id / experiment_id/ response_id,
-                 method="REML") 
+               random = ~1 | study_id / response_id,
+                 method="ML") 
 temp_full_var_model
@@
 ```{r, echo = FALSE}
-trait_full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~exp_age + size + org_level, 
-               random = ~1 |  study_id/ experiment_id/ response_id,
-                 method="REML") 
+trait_full_var_model <-rma.mv(yi, vi, data=dat_full_var_ES, mods = ~exp_age + size + org_level +metric +environment, 
+               random = ~1 |  study_id/ response_id,
+                 method="ML") 
 trait_full_var_model
@@
                random = ~1 | study_id/response_id,
-                 method="REML") 
+                 method="ML") 
 full_var_model
+
+full_no_org <- rma.mv(yi, vi, data=normalized_var, 
+                        mods = ~flux_range * mean_temp_constant + exp_age + size  +environment +metric, 
+               random = ~1 | study_id/response_id,
+                 method="ML")
 
@@
 #Model selection 
-AIC(temp_acclimation_model, trait_acclimation_model, acclimation_model, temp_full_var_model, trait_full_var_model, full_var_model)
+AIC(temp_acclimation_model, trait_acclimation_model, acclimation_model)
+
+AIC(temp_full_var_model, trait_full_var_model, full_var_model, full_no_org)
+
 ```
@@
   facet_wrap(~mean_temp_constant)
+
+ggplot(normalized_var, aes(y=flux_range, fill=yi > 0, x=mean_temp_constant))+
+  geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.3, height = 0.7)+
+  geom_line(y=0, linetype = "dashed")+
+  #geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis(discrete =TRUE)+
+  ggtitle("Non-linear averaging + acclimation and fluctuation range")+
+  labs(x = "mean temp",
+       y ="fllux_range",
+       fill = "yi")+
+  theme(legend.position="bottom")+
+  geom_smooth(method=lm)
+
+ggplot(normalized_var, aes(y=flux_range, fill=yi > 0, x=mean_temp_constant))+
+  geom_jitter(alpha = 0.5, pch=21,size=1, width = 0.3, height = 0.7)+
+  geom_line(y=0, linetype = "dashed")+
+  #geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis(discrete =TRUE)+
+  ggtitle("Non-linear averaging + acclimation and fluctuation range")+
+  labs(x = "flux range",
+       y ="SMD",
+       fill = "mean temp")+
+  theme(legend.position="bottom")+
+  geom_smooth(method=lm)
+
+ggplot(normalized_var, aes(y=yi, fill=mean_temp_constant, x=flux_range))+
+  geom_jitter(alpha = 0.5, pch=21,size=3, width = 0.3, height = 0.7)+
+  geom_line(y=0, linetype = "dashed")+
+  #geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  #geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Non-linear averaging + acclimation and fluctuation range")+
+  labs(x = "flux_range",
+       y ="yi",
+       fill = "mean temp")+
+  theme(legend.position="bottom")+
+  geom_smooth(method=lm)+
+  facet_wrap(~mean_temp_constant)
+
+ggplot(normalized_acclim, aes(y=yi, x=flux_range, fill=mean_temp_reared))+
+  geom_jitter(alpha = 0.5, pch=21,size=3, width = 0.3, height = 0.7)+
+  geom_line(y=0, linetype = "dashed")+
+  #geom_errorbar(data=mean_var, aes(ymin=yi-vi, ymax=yi+vi), size=.7, width=.4, color = "black")+
+  #geom_jitter(data = mean_var, size=4, pch=21, colour="black")+
+  theme_bw()+
+  scale_fill_viridis()+
+  ggtitle("Acclimation and fluctuation range")+
+  labs(x = "mean temp",
+       y ="yi",
+       fill = " flux range")+
+  theme(legend.position="bottom")+
+  geom_smooth(method=lm)+
+  facet_wrap(~mean_temp_reared)
+
 ```
